### 2019-11-22 codecheck-paper.Rmd
add first outline of manuscript

---
title: "CODE CHECK"
author:
- affiliation: Department of Applied Mathematics and Theoretical Physics, University of Cambridge, Cambridge, Cambridgeshire, GB
  corresponding: sje30@cam.ac.uk
  email: sje30@cam.ac.uk
  name: Stephen Eglen
  orcid: 0000-0001-8607-8025
- affiliation: Institute for Geoinformatics, University of M"unster, M"unster, Germany
  email: daniel.nuest@uni-muenster.de
  name: Daniel Nüst
  orcid: 0000-0002-0024-5046
date: "`r format(Sys.time(), '%d %B, %Y')`"
abstract: |
          | ...
#author_summary: |
#  TBD
bibliography: bibliography.bib
#output: rticles::plos_article
#csl: plos.csl
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- Let's stick to one sentence per line! -->
<!-- Citations: @Figueredo:2009dg, citep: [@Figueredo:2009dg], multiple citep: [@Figueredo:2009dg; @other_2018] -->
<!-- add " {-}" after headings to remove numbering -->

# Background

- role of data and code, data science
- reproducible resarch
- data enclaves
- CASCAID
- ROpenSci/PyOpenSci

**main contributions**

- a concept for integrating a minimal code review into common scholarly communication processes around peer review
- set of shared principles (recognition value) that allow researchers to quickly understand the level of evaluation across journals/publishers/conferences, and allows helps these stakeholders to establish workflow checking in their domain/solution/product
- a common language and rubric for classifying different levels of code and data evaluation as part of a peer-review

# What is CODE CHECK?

- **Principles**: https://codecheck.org.uk/
- relation to "proper" citation of data and software, and depositing of data and software in suitable repos _besides_ the bundle for CODE CHECK: do it, only provide the concrete analysis script for the check
- **Why is it useful?**
  - reduce the barrier to evaluating non-text parts of research
  - acknowledges different skill sets and existing high load on reviewers
  - CC breaks problem that publishers staff expertise today is not likely to suffice
  - engaging ECRs is a perfet fit (they are up-to-date in methods, are introduced to peer review)
  - relation to research compendia? point it out for authors unsure about how to structure their workflow
  - CrossRef and Publons and ORCID have features around reviews
  - shifting burden from reviewer to author (but not too much)

# CODE CHECK implementation concepts

- need to get wording straight: process/implementation/workflow (DN: suggest to use "workflow" for research-analysis workflow)
- https://codecheck.org.uk/process/
- _dimensions_ along which a check can vary
- independent/pre-submission (Peer_Coding) by author + colleagues
- AGILE
- premier (OA) publisher
- community OA / scholarly society with public peer review
- invited reproducibility paper track, cf. https://www.elsevier.com/journals/information-systems/0306-4379/guide-for-authors

# Annotated Examples

- **the default implementation**
- critically discuss our/the default implementation vs. the principles
- HPC example

# Open problems

- need publishers build up data science experts who can conduct checks?
- integration of software/data citation checks benefitial?
- bot-support and tools (linter)
- Does nudging towards better practices work (highlight the successes) or is any check with opt-in doomed to fail?

# Acknowledgements

Collaborators and potential collaborators:

- eLife
- GigaScience
- AH, SciData
- Mozilla
- PLOS?

# Contribution

..

### 2019-11-27 codecheck-paper.Rmd
add an examples file so we have a place to manage/document those
Update codecheck-paper.Rmd

@@
   - shifting burden from reviewer to author (but not too much)
+- keynote https://doi.org/10.7557/5.4963: in discussion, the speaker mentioned value of peer review for "baseline reassurance", i.e.
+a paper has at least been checked by someone with an understanding, to increase trust especially if looking at papers from other disciplines
 
@@
 # Annotated Examples
+<!-- see examples.md -->
 
@@
 - Does nudging towards better practices work (highlight the successes) or is any check with opt-in doomed to fail?
+- Do we need more than one codechecker? Is there a similar diversity as in a peer review where more opinions are needed?
 

### 2019-12-04 codecheck-paper.Rmd
Update codecheck-paper.Rmd

@@
 
+# Related work
+
+- _Successes and struggles with computational reproducibility: Lessons from the Fragile Families Challenge_ > https://osf.io/preprints/socarxiv/g3pdb/
+
 # CODE CHECK implementation concepts
@@
 - invited reproducibility paper track, cf. https://www.elsevier.com/journals/information-systems/0306-4379/guide-for-authors
+- on giving credit: public databases
+  - https://www.reviewercredits.com/
+  - Publons
+  - ORCID
+- compare with existing processes that are publicly documented
+  - AJPS Verification Policy: https://ajps.org/ajps-verification-policy/
+  - ...
 
@@
 - Do we need more than one codechecker? Is there a similar diversity as in a peer review where more opinions are needed?
+- can CODE CHECK work for preprints? how?
 

### 2019-12-06 codecheck-paper.Rmd
Update codecheck-paper.Rmd

@@
 - _Successes and struggles with computational reproducibility: Lessons from the Fragile Families Challenge_ > https://osf.io/preprints/socarxiv/g3pdb/
+- community standards for software, cf. slide 6 in https://figshare.com/articles/_/5675104
+  - "good enough practice" is "get a colleague to try using it", slide 19 in https://figshare.com/articles/_/5675104
 

### 2019-12-09 codecheck-paper.Rmd
Update codecheck-paper.Rmd

@@
   - "good enough practice" is "get a colleague to try using it", slide 19 in https://figshare.com/articles/_/5675104
+- journals doing code reviews
+  - _"We know of 16 journals that take this valuable step (ie TOP Level 3): AEA as already mentioned, plus others in polisci, psychology, biostats, and org chem: see Data Rep Policies here https://t.co/plOF8j6ADU "_ (https://twitter.com/EvoMellor/status/1202692360456589339?s=09); see also whole thread!
 

### 2019-12-09 codecheck-paper.Rmd
Update codecheck-paper.Rmd

@@
 - ROpenSci/PyOpenSci
+- time capsules > https://twitter.com/DougBlank/status/1135904909663068165?s=09
 
@@
   - _"We know of 16 journals that take this valuable step (ie TOP Level 3): AEA as already mentioned, plus others in polisci, psychology, biostats, and org chem: see Data Rep Policies here https://t.co/plOF8j6ADU "_ (https://twitter.com/EvoMellor/status/1202692360456589339?s=09); see also whole thread!
+- open science platforms/cloud services > https://www.nature.com/articles/d41586-019-03366-x
+- live code in articles > o2r, eLife RDS https://www.nature.com/articles/d41586-019-00724-7, ...
 

### 2020-01-24 codecheck-paper.Rmd
Update codecheck-paper.Rmd

@@
 - live code in articles > o2r, eLife RDS https://www.nature.com/articles/d41586-019-00724-7, ...
+- Artefact Evaluation - https://www.artifact-eval.org/about.html
 
@@
   - ...
+- flexible level of detail? a check can be enhanced to (potentially semi-automatically) check for good practices
+  - "A group of us in the @force11rescomm Software Citation Implementation Working Group produced a "Software Citation Checklist for Authors" https://t.co/4Noe66FsoX providing guidelines on how to cite software and describing why it is important and what software should be cited" via https://twitter.com/alegonbel/status/1207664922932453376?s=09)
 

### 2020-01-25 codecheck-paper.Rmd
Update codecheck-paper.Rmd

@@
 
+- https://medium.com/bits-and-behavior/a-modern-vision-for-peer-review-d5f73f0fae07 - work by Amy J. Ko: https://faculty.washington.edu/ajko/publications
 - _Successes and struggles with computational reproducibility: Lessons from the Fragile Families Challenge_ > https://osf.io/preprints/socarxiv/g3pdb/

### 2020-01-27 codecheck-paper.Rmd
Update codecheck-paper.Rmd

@@
   - _"We know of 16 journals that take this valuable step (ie TOP Level 3): AEA as already mentioned, plus others in polisci, psychology, biostats, and org chem: see Data Rep Policies here https://t.co/plOF8j6ADU "_ (https://twitter.com/EvoMellor/status/1202692360456589339?s=09); see also whole thread!
+  - https://medium.com/@NeurIPSConf/call-for-papers-689294418f43
+  - https://www.journals.elsevier.com/information-systems/editorial-board/
+  - SIGMOD Reproducibility Review?
+  - Biostatistics "AER" (https://academic.oup.com/biostatistics/article/10/3/405/293660)
 - open science platforms/cloud services > https://www.nature.com/articles/d41586-019-03366-x
@@
 - Artefact Evaluation - https://www.artifact-eval.org/about.html
+- continuous analysis - https://www.nature.com/articles/nbt.3780
+-
 

### 2020-01-31 codecheck-paper.Rmd
Update codecheck-paper.Rmd

@@
 a paper has at least been checked by someone with an understanding, to increase trust especially if looking at papers from other disciplines
+- **report** must fulfil some requirements to make clear what is really checked and not mislead - readers will have very different understandings of a green checkmark!
 
@@
 - Do we need more than one codechecker? Is there a similar diversity as in a peer review where more opinions are needed?
-- can CODE CHECK work for preprints? how?
+- can CODE CHECK work for preprints? how? is a re-check required?
+- risk: is the hurdle too low? are "levels" of checks are solution to increase transparency of check's level of detail
 

### 2020-01-31 codecheck-paper.Rmd
Update codecheck-paper.Rmd

@@
 - risk: is the hurdle too low? are "levels" of checks are solution to increase transparency of check's level of detail
+- code check in the context of innovations and disruptions in scientific publishing
+  - https://doi.org/10.1042/ETLS20180172
+  - https://doi.org/10.3390/publications7020034
 

### 2020-01-31 codecheck-paper.Rmd
Update codecheck-paper.Rmd

@@
 - https://medium.com/bits-and-behavior/a-modern-vision-for-peer-review-d5f73f0fae07 - work by Amy J. Ko: https://faculty.washington.edu/ajko/publications
+  - is CODE CHECK's process too small, too incremental a change?
 - _Successes and struggles with computational reproducibility: Lessons from the Fragile Families Challenge_ > https://osf.io/preprints/socarxiv/g3pdb/
@@
 - continuous analysis - https://www.nature.com/articles/nbt.3780
--
 

### 2020-02-06 codecheck-paper.Rmd
Update codecheck-paper.Rmd

@@
 - continuous analysis - https://www.nature.com/articles/nbt.3780
+- "A Universal Identifier for Computational Results" > https://www.sciencedirect.com/science/article/pii/S1877050911001256
+- Discussion about using Docker containers during JOSS reviews > https://github.com/openjournals/joss/issues/498#issuecomment-462046912
 

### 2020-02-06 codecheck-paper.Rmd
Update codecheck-paper.Rmd

@@
 
+- https://paperpile.com/shared/rVNwBS
 - https://medium.com/bits-and-behavior/a-modern-vision-for-peer-review-d5f73f0fae07 - work by Amy J. Ko: https://faculty.washington.edu/ajko/publications
@@
   - https://doi.org/10.3390/publications7020034
+- see https://github.com/codecheckers/discussion
 

### 2020-02-06 codecheck-paper.Rmd
Update codecheck-paper.Rmd

@@
 - **report** must fulfil some requirements to make clear what is really checked and not mislead - readers will have very different understandings of a green checkmark!
+- reproducibility is hard https://twitter.com/annakrystalli/status/1144176149859377152?ref_src=twsrc%5Etfw > code check gives at least some benefits
 
@@
 - Discussion about using Docker containers during JOSS reviews > https://github.com/openjournals/joss/issues/498#issuecomment-462046912
+- https://scigen.report, an independent site for registering reproductions, via https://annakrystalli.me/talks/ro-reprohack.html#32
+  - example https://scigen.report/reviews/get?doi=10.1063/1.1823034
+  - seems to be closed infrastructure
 
@@
 - HPC example
+- if we see the corpus of examles as a scientific dataset, it should be published in a citable way (maybe even submit to a data journal)
 

### 2020-05-11 codecheck-paper.Rmd
Update codecheck-paper.Rmd

@@
 - time capsules > https://twitter.com/DougBlank/status/1135904909663068165?s=09
+- science is changing: https://theconversation.com/what-you-need-to-know-about-how-coronavirus-is-changing-science-137641
 

### 2020-05-11 install.R
make Binder-ready with Makefile magic for venv

@@ -1,0 +1,6 @@
+# Install CRAN packages
+install.packages("rticles")
+
+# Install some common LaTeX packages
+tinytex:::install_yihui_pkgs()
+#tinytex::tlmgr_install("xcolor")

### 2020-05-15 codecheck-paper.Rmd
Update codecheck-paper.Rmd

--- 2020-02-06 codecheck-paper.Rmd
@@
 - CASCAID
-- ROpenSci/PyOpenSci
+- ROpenSci/PyOpenSci (and how their software review differs, though they also collaborate with journals, https://devguide.ropensci.org/softwarereviewintro.html)
 - time capsules > https://twitter.com/DougBlank/status/1135904909663068165?s=09
+- science is changing: https://theconversation.com/what-you-need-to-know-about-how-coronavirus-is-changing-science-137641
 

### 2020-05-19 codecheck-paper.Rmd
Update codecheck-paper.Rmd

@@
   - seems to be closed infrastructure
+- new journal _ReScienceX_ (about experiments) also seems to have an ECR angle, https://www.nature.com/articles/d41586-020-01328-2
 

### 2020-05-26 codecheck-paper.Rmd
add stodden and whitaker references
potentially useful for classification

@@
 - reproducibility is hard https://twitter.com/annakrystalli/status/1144176149859377152?ref_src=twsrc%5Etfw > code check gives at least some benefits
+- what are the requirements / level or reproducibility required or checked?
+  - [Stodden et al., 2013](https://stodden.net/icerm_report.pdf): five-level hierarchy of research - reviewable, replicable, confirmable, auditable, and open or reproducible
+  - [Whitaker 2017](https://figshare.com/articles/Showing_your_working_A_guide_to_reproducible_neuroimaging_analyses/4244996/2): reproducible, generalisable, replicability, robustness
 

### 2020-05-26 codecheck-paper.Rmd
Update codecheck-paper.Rmd

@@
   - is CODE CHECK's process too small, too incremental a change?
+- overview of peer review (if not a reference itself, it gives plenty of references): http://juser.fz-juelich.de/record/865096/files/Peer%20Review%20and%20Bibliometrics.pdf
 - _Successes and struggles with computational reproducibility: Lessons from the Fragile Families Challenge_ > https://osf.io/preprints/socarxiv/g3pdb/

### 2020-05-26 codecheck-paper.Rmd
Update codecheck-paper.Rmd

--- 2020-05-26 codecheck-paper.Rmd
@@
   - is CODE CHECK's process too small, too incremental a change?
+- overview of peer review and bibliometrics (if not a reference itself, it gives plenty of references): http://juser.fz-juelich.de/record/865096/files/Peer%20Review%20and%20Bibliometrics.pdf
 - _Successes and struggles with computational reproducibility: Lessons from the Fragile Families Challenge_ > https://osf.io/preprints/socarxiv/g3pdb/
@@
   - https://doi.org/10.3390/publications7020034
+- should every code be checkd, or should the labour be capped by only checking "important" works or at important stages; if yes, isn't peer review such a stage?
 - see https://github.com/codecheckers/discussion

### 2020-05-27 codecheck-paper.Rmd

@@
 - https://paperpile.com/shared/rVNwBS
+- Are "Replicated Computations Results (RCR) Report" a thing? apparently so (using ACM terminology)
+  - https://dl.acm.org/doi/10.1145/3185337
+  - http://users.iems.northwestern.edu/~nelsonb/Publications/Nelson_RCRgreenSim.pdf
+  - https://dl.acm.org/doi/epdf/10.1145/3341094
+    - citation for replicated workflow lists it as "to appear"
+  - getting proper papers for reproductions != CODECHECK approach where we want credit as reviewers
 - https://medium.com/bits-and-behavior/a-modern-vision-for-peer-review-d5f73f0fae07 - work by Amy J. Ko: https://faculty.washington.edu/ajko/publications

### 2020-05-29 codecheck-paper.Rmd
Update codecheck-paper.Rmd

@@
 - bot-support and tools (linter)
+- how can open platforms be used to enable a review in private (cf.
+https://www.cambridge.org/core/blog/2019/08/19/how-to-make-the-data-and-code-for-your-manuscript-available-to-peer-reviewers-before-making-it-public/), if that is desired
 - Does nudging towards better practices work (highlight the successes) or is any check with opt-in doomed to fail?

### 2020-05-29 codecheck-paper.Rmd
ACM badges = related work

@@
 
+- CODECHECK vs. the "Artifacts Evaluated - Functional" badge of the ACM, see https://www.acm.org/publications/policies/artifact-review-badging
 - https://paperpile.com/shared/rVNwBS

### 2020-05-29 codecheck-paper.Rmd
Update codecheck-paper.Rmd

@@
   - "A group of us in the @force11rescomm Software Citation Implementation Working Group produced a "Software Citation Checklist for Authors" https://t.co/4Noe66FsoX providing guidelines on how to cite software and describing why it is important and what software should be cited" via https://twitter.com/alegonbel/status/1207664922932453376?s=09)
+- focus on ECR as codecheckers
+  - ECR are the one's breaking new ground in schol comm, cf. the percentages of career stages submitting registered reports on slide 20 > https://osf.io/2vjus/
 
@@
 - see https://github.com/codecheckers/discussion
+- can we establish code checks as a type of review within the schol comm metadata (or should we?) > we should not, put them on the same level as reviews!
 
@@
 
-..
+SE ...; DN ...; see [CRediT](https://casrai.org/credit/).

### 2020-06-02 codecheck-paper.Rmd
Update codecheck-paper.Rmd

@@
 - science is changing: https://theconversation.com/what-you-need-to-know-about-how-coronavirus-is-changing-science-137641
+- reproducibility is important, also
+  - effects on well-being/mental health, https://www.nature.com/articles/d41586-020-01642-9
 
@@
 - can we establish code checks as a type of review within the schol comm metadata (or should we?) > we should not, put them on the same level as reviews!
+- how can we turn a CODECHECK into a positive experience?
+  - cf. kindness in peer review in https://www.mdpi.com/2304-6775/8/2/26
 

### 2020-06-03 codecheck-paper.Rmd
Update codecheck-paper.Rmd

@@
 - new journal _ReScienceX_ (about experiments) also seems to have an ECR angle, https://www.nature.com/articles/d41586-020-01328-2
+- really interesting work on reproductions in hydrology, with a neat survey that reproducers filled out which might be interesting to capture from codecheckers, too: https://www.nature.com/articles/sdata201930#Sec9
 

### 2020-06-11 codecheck-paper.Rmd
Update codecheck-paper.Rmd

@@
   - shifting burden from reviewer to author (but not too much)
+  - What issues with (open) peer review does CODECHECK address?
 - keynote https://doi.org/10.7557/5.4963: in discussion, the speaker mentioned value of peer review for "baseline reassurance", i.e.
@@
 - overview of peer review and bibliometrics (if not a reference itself, it gives plenty of references): http://juser.fz-juelich.de/record/865096/files/Peer%20Review%20and%20Bibliometrics.pdf
+- Open Peer Review > https://researchintegrityjournal.biomedcentral.com/articles/10.1186/s41073-019-0063-9
+  - Are OPR lessons transferable to CODECHECK? Does CODECHECK work better with OPR than with (single/double) blind peer review?
+  - similarities to their advice: open reports, give credit
 - _Successes and struggles with computational reproducibility: Lessons from the Fragile Families Challenge_ > https://osf.io/preprints/socarxiv/g3pdb/
@@
   - cf. kindness in peer review in https://www.mdpi.com/2304-6775/8/2/26
+- how can we allow reactions to checks? (commenting on reports)
 

### 2020-06-14 codecheck-paper.Rmd
Update codecheck-paper.Rmd

@@
   - [Stodden et al., 2013](https://stodden.net/icerm_report.pdf): five-level hierarchy of research - reviewable, replicable, confirmable, auditable, and open or reproducible
-  - [Whitaker 2017](https://figshare.com/articles/Showing_your_working_A_guide_to_reproducible_neuroimaging_analyses/4244996/2): reproducible, generalisable, replicability, robustness
+  - [Whitaker 2017](https://figshare.com/articles/Showing_your_working_A_guide_to_reproducible_neuroimaging_analyses/4244996/2): reproducible, generalisable, replicability, robustnes
+- can authors retract their code/data once a CODECHECK has started? In the eLife prepreint review, there seems no "way back", see https://twitter.com/ceptional/status/1271528976763121664?s=09
 

### 2020-06-24 codecheck-paper.Rmd
Update codecheck-paper.Rmd

@@
 
-- need publishers build up data science experts who can conduct checks?
-- integration of software/data citation checks benefitial?
-- bot-support and tools (linter)
-- how can open platforms be used to enable a review in private (cf.
+- Do publishers need to build up data science experts who can conduct checks?
+- Is an integration of software/data citation checks benefitial/realistic?
+- Bot-support and tools (linter, report authoring with R Markdown)
+- How can open platforms be used to enable a review in private (cf.
 https://www.cambridge.org/core/blog/2019/08/19/how-to-make-the-data-and-code-for-your-manuscript-available-to-peer-reviewers-before-making-it-public/), if that is desired
@@
 - Do we need more than one codechecker? Is there a similar diversity as in a peer review where more opinions are needed?
-- can CODE CHECK work for preprints? how? is a re-check required?
-- risk: is the hurdle too low? are "levels" of checks are solution to increase transparency of check's level of detail
-- code check in the context of innovations and disruptions in scientific publishing
+- can CODECHECK work for preprints? how? is a re-check required?
+- Risk: is the hurdle too low? are "levels" of checks are solution to increase transparency of check's level of detail
+- CODECHECK in the context of innovations and disruptions in scientific publishing
   - https://doi.org/10.1042/ETLS20180172
   - https://doi.org/10.3390/publications7020034
-- should every code be checkd, or should the labour be capped by only checking "important" works or at important stages; if yes, isn't peer review such a stage?
+- Should every code be checkd, or should the labour be capped by only checking "important" works or at important stages; if yes, isn't peer review such a stage?
 - see https://github.com/codecheckers/discussion
-- can we establish code checks as a type of review within the schol comm metadata (or should we?) > we should not, put them on the same level as reviews!
-- how can we turn a CODECHECK into a positive experience?
+- Can we establish code checks as a type of review within the schol comm metadata (or should we?) > we should not, put them on the same level as reviews!
+- How can we turn a CODECHECK into a positive experience?
   - cf. kindness in peer review in https://www.mdpi.com/2304-6775/8/2/26
-- how can we allow reactions to checks? (commenting on reports)
+- How can we allow reactions to checks? (commenting on reports)
+- Will bitwise reproducibility become easier than what a CODECHECK does? triggered by https://twitter.com/khinsen/status/1242842759733665799?s=09)
 

### 2020-06-29 codecheck-paper.Rmd
Update codecheck-paper.Rmd

@@
 - Will bitwise reproducibility become easier than what a CODECHECK does? triggered by https://twitter.com/khinsen/status/1242842759733665799?s=09)
+- If a report is not published for failures: Isn't there a bias for the codechecker to have a "successful" reproduction because then a report is published?
 

### 2020-07-28 codecheck-paper.Rmd
Update codecheck-paper.Rmd

@@
 - CODECHECK vs. the "Artifacts Evaluated - Functional" badge of the ACM, see https://www.acm.org/publications/policies/artifact-review-badging
+- Does CODECHECK do internal replication, as defined by https://gatesopenresearch.org/articles/4-17/v2 ?
 - https://paperpile.com/shared/rVNwBS

### 2020-08-21 codecheck-paper.Rmd
Update codecheck-paper.Rmd

@@
 
+- CODECHECK vs. Verification Reports > > https://www.sciencedirect.com/science/article/pii/S0010945220301738#fn3
 - CODECHECK vs. the "Artifacts Evaluated - Functional" badge of the ACM, see https://www.acm.org/publications/policies/artifact-review-badging
@@
   - getting proper papers for reproductions != CODECHECK approach where we want credit as reviewers
+  - see also [Google Search](https://www.google.com/search?q=%22Replicated+Computations+Results+%28RCR%29+Report+for%22)
 - https://medium.com/bits-and-behavior/a-modern-vision-for-peer-review-d5f73f0fae07 - work by Amy J. Ko: https://faculty.washington.edu/ajko/publications

### 2020-09-15 codecheck-paper.Rmd
On a roll...

@@
 
-# Background
+# Open questions
+
+* Can we switch to latex for this?  (Find it slightly easier.)
+* Which journal to target -- maybe worry abotu that once we have a
+  preprint.  F1000R my current favourite?
+
+# Introduction
+
+Many areas of scientific research now use computation to either
+simulate or analyse their data.  These computations are increasingly
+complex, and difficult to explain coherently in a paper. (MARWICK).
+To complement the traditional route of sharing research by writing
+papers, there is a growing trend/demand to share the underlying artifacts,
+notably code and datasets, so that others can inspect, reproduce or
+expand that work (Figure 1).  Some of the earliest proponents were
+Buckheit and Donoho (1995) who commented "advert is scholarship for
+the paper". (get correct quote).
+
+So, assuming that researchers begin to share more artifacts, how
+might they be examined to test that they do what they claim?  For
+example, most scientific journals now require a data sharing statement
+that outlines what data the authors have (or will) share.  The
+implementation status of this varies considerably according to the
+journal.  At one of the spectrum, there are specialist journals  that
+have been created to accept "data papers" (e.g. *Scientific Data*,
+... NAME COUPLE OF OTHERS).  These journals have established rigorous
+procedures by which the data is validated according to standards in
+each field.  At the other end of the spectrum, we still witness today
+the infamous statement "Data available upon reasonable request"
+whereby authors, well-intentioned, often cannot provide the data when
+readers ask for it.
+
+Given that there are no clear standards yet for sharing data, what
+hope might there be for sharing computer programs?  Our anectodal
+experiences on this matter suggest that often a variety of reasons are
+given for why their code cannot be shared, e.g. "there is no
+documentation / I do not want to maintain it / I do not want to give
+away my code to my competitors".  Our view, rooted in open science
+princples, is that wherever possible sharing code is good for the
+community, as outlined ten years ago (Barnes, 2010).  Having the code
+freely available, and suitably archived, provides a valuable resource
+for others to learn from, even when it doesn't run or if there is no
+documentation.  However, with a little effort, we believe that if
+someone independent can run/use the code, this is worth documenting as
+early as possible.  With this in mind, we have developed a set of
+PRINCIPLES and an example workflow that provides what he hope is a
+pragmatic way of checking that code works.  We call this system
+CODECHECK as outlined next.
+
+Figure 1: the inverse problem figure.
+
+# What is a CODECHECK?
+
+CODECHECK is best demonstrated by way of our example workflow, and
+later we expand on the underlying principles.  The workflow involves
+three groups of people (1) the AUTHOR providing the code to be
+checked, (2) the PUBLISHER of a journal intereseted in publishing the
+paper by the AUTHOR and (3) the CODECHECKER who checks that the
+AUTHOR's code works.  The workflow that we have refined is documented
+in Figure 2.
+
+Step 1: the author submit their manuscript along with code and data
+to the publishers.  The code and data need not be openly available at
+this point.
+
+Step 2: the publisher finds a codechecker to check the code.  This is
+analgous to the publisher finding one or more reviewers to evaluating
+the code.  However a cruical difference is that we suggest the
+codechcker and the author can talk freely to each directly (see next).
+
+Step 3: the codechecker runs the code, based on instructions provided
+by the author, and checks if some/all of the results from the paper
+can be reproduced.  (See later for exact nature of "reproduced").  If
+there are any problems running the code, the codechecker asks the
+author for help to resolve the problems.  The codechecker then tries
+to run the code again.  This process iterates until either the
+codechecker is successful, or the codechecker concludes the code does
+not (*what do in this case*).  As part of this process, the
+codechecker could work entirely locally on their own compute resource,
+or in the cloud, e.g. using mybinder infrastructure.  Using such
+cloud-based infrastructure allows for the codechecker and author to
+collaboratively run the code together.
+
+Step 4: The codechecker writes a certificate stating how the code was
+run and includes a copy of outputs (figures, tables) that were
+independently generated.
+
+Step 5: the certificate and code/data get deposited on an open archive
+(currently Zenodo), and the certificate is given to the PUBLISHER.
+
+Step 6: the Publisher can then use the certificate as part of their
+supplementary records for a paper, and give credit to the codechecker
+for their work by depositing appropriate records (*bit hazy here*).
+
+Figure 2: the detective.  Could we number the arrows as steps?
+(Implementing the CODECHECK process)
+
+## Variations
+
+This example workflow leaves sufficient room for different
+implementations of this process.  There are several aspects that can
+be considered:
+
+### When to do a codecheck
+
+preprint (no publisher available), pre-condition for peer review,
+during peer review, after acceptance.  Each of these have their own
+pros/cons.
+
+### Who does the codecheck
+
+in house staff, pool of ECRs, volunteeres,
+
+### Who knows who?
+
+single blind, double blind, private/public communication between author and
+codechecker, collaborative.  (This might need finessing with principle
+2 below).
+
+## Core principles
+
+The workflow and variations just outlined below reflect our current
+views on how a codecheck should be performed.  They are not set in
+stone, but we do believe the following four core principles underlying
+CODECHECK:
+
+**1. Codecheckers record but don’t investigate or fix.**  The
+codechecker follows the instructions provided by the author in running
+the code.  If any steps are unclear, or if code does not run
+correctly, the codechecker records this fact, and then communicates
+this information to the author.  We believe that the job of the
+codechecker is not to fix all these problems, but simply to report
+them, and await a fix from the author.
+
+**2. Communication between humans is key.**  Some code may just work
+without any interaction (e.g. Manuel's certificate); however, more
+often there are hidden dependencies that need to be found, or revised
+software installations.  Allowing the codechecker and human to
+communicate directly and openly with each other should make this
+process as constructive as possible.  We feel that routing this
+conversation (possibly anonymously) through a publisher would slow
+things down and reduce the chances for community bulding.
+
+**3. Credit is given to codecheckers.** (TODO Explain in a few
+sentences what each means, e.g. see 1,2)
+
+**4. Workflows must be auditable.** (TODO Explain in a few sentences
+what each means, e.g. see 1,2)
+
+# Register
+
+We have a curated list of 15 certificates available at
+<https://codecheck.org.uk/register>.  These are a mixture of
+reproductions of papers for a variety of reasons (do we name them
+all?) from interesting historical papers, to papers codechecked as
+part of peer review, or because of public interest (covid).  several
+covid models have been done during the time of peer review of these
+papers, under our initative rather than requested from journal, but
+certificate have then been acknowledged/cited in the paper (LANCET /
+LSHTM).
+
+We would like to briefly highlight three examples here.
+
+1. Gigascience -- our first paper.  Only visualisation, not machine
+   learning (which would have taken sseveral days of compute time).
+2. report 9 -- covid model of UK lockdown model.
+3. Geoscience
+
+## Annotated certificate
+
+Should we annotate a certificate, at least the cover page, to
+demonstrate what we mean by a certificate?  We could print a four page
+certificate 4-up so that it fits onto one page.  Text might be too
+small though to read?  A 2-up, printed sideways would work well
+though.
+
+To run the system, we rely on github to host our work, and use a
+custom R pacakge, codecheck, to help with the generation of
+certificates and depositing work on zenodo.  (Cite zen4R).
+
+# Related work
+
+All the related projectsfrom A-Z... e.g. CASCAID, Rescience, modeldb,
+o2r,
+
+# Limitations
+
+Compute time:  for those papers that take significant compute time
+(think days, not minutes), who will pay for the compute time?
+
+Software: authors can currently provide code that requires
+propieratiry software.  Given the prevalence of software like MATLAB,
+this seems like a pragmatic choice, but then (a) we can't rely on
+open infrastructure like mybinder, (b) codechecker must have that
+software.
+
+Can't someone cheat? yes.  we don't check it is correct code, just
+simply that it runs.  Very low bar.  But by having the code open,
+interested parties can then examine the code, and knowing that the
+code will be open we hope will be an incentive for authors to share.
+
+Who's got time for more peer review?  Agree, but ECRS might be
+attracted to working with groups (recall it is not anonymous)
+
+Who finds the codechecker?  Good point!  The journal.
+
+As a principle, we do not require results to be exactly the same for a
+codecheck to pass, simply that the code runs and generates the output
+files that the author claims.  Stochastic simulations mean that often
+we cannot get exactly the same, or even different versions of
+libraries.  (Cite freesurfer example that generates different outputs
+on different operating systems).
+
+- untilhe co modelDB principle.
+
+# Future work and conclusions
+
+- possibility to allow CI for open papers such that for "nearly free"
+  we can find out when softwave breaks, and if provenance tools are
+  used, when.
+
+- How to embed this in juornals.
+- ???
 
@@
 - PLOS?
+- Richard van Noorden, Dalmeet Singh (paper 9)
+- Iain Davies and Sebastian 
+- SSI for some funding.
 

### 2020-09-15 codecheck-paper.Rmd
push

@@
 
+* Prefer longlines or short?
 * Can we switch to latex for this?  (Find it slightly easier.)
@@
 
+* Check with Giuliano re: "code works".
+
 # Introduction
 
-Many areas of scientific research now use computation to either
-simulate or analyse their data.  These computations are increasingly
-complex, and difficult to explain coherently in a paper. (MARWICK).
-To complement the traditional route of sharing research by writing
-papers, there is a growing trend/demand to share the underlying artifacts,
-notably code and datasets, so that others can inspect, reproduce or
-expand that work (Figure 1).  Some of the earliest proponents were
-Buckheit and Donoho (1995) who commented "advert is scholarship for
-the paper". (get correct quote).
+Many areas of scientific research now use computation to either simulate or analyse their data.
+These computations are increasingly complex, and difficult to explain coherently in a paper.
+(MARWICK).
+To complement the traditional route of sharing research by writing papers, there is a growing trend/demand to share the underlying artifacts, notably code and datasets, so that others can inspect, reproduce or expand that work (Figure 1).
+Some of the earliest proponents were Buckheit and Donoho (1995) who commented "advert is scholarship for the paper".
+(get correct quote).
 
@@
 
-Figure 1: the inverse problem figure.
+\begin{figure}
+  \centering
+  \includegraphics{figs/rr.pdf}
+  \caption{The inverse problem in reproducible research.  The left
+  half of the diagram shows the diverse range of materials used
+  internally within a laboratory.  These materials are often then
+  condensed into a paper for sharing with the outside world via the
+  research paper, astatic PDF document.  Working backwards from the
+  PDF to the underlying materials is impossible.  By sharing the
+  materials on the left, others outside the lab are able to reproduce
+  or build on this work.  (Should we draw a backward arrow showing the
+  impossibility of this backward flow?)}
+  \label{fig:inverse}
+\end{figure}
 
@@
 
-Figure 2: the detective.  Could we number the arrows as steps?
-(Implementing the CODECHECK process)
+\begin{figure}
+  \centering
+  \caption{The detective. Could we number the arrows as steps?
+(Implementing the CODECHECK process).  DANIEL: best source for this figure?}
+  \label{fig:worfklow}
+\end{figure}
 
@@
 All the related projectsfrom A-Z... e.g. CASCAID, Rescience, modeldb,
-o2r,
+o2r.  Community iniatives like reprohack along similar lines of
+checking code. <https://reprohack.github.io/reprohack-hq/>
 
 # Limitations
+
+Handling failure:  we have no established process for this, as so far
+all our codechecks succeeded.  Open question remains as what to do in
+this case -- should we publicly report that the code would not run.
 
@@
 
-- untilhe co modelDB principle.
+- modelDB principle: minimum of one figure should be reproducibe.
+  Exampler certificate 2020-016 (I think) which had only a small
+  subset of the possible figures.  Still useful and worth sharing.
 
@@
 
+# Declaration of interest
+
+SJE is on the senior editorial board at the journal *Scientific Data*.
+
 # Acknowledgements
@@
 
-- eLife
-- GigaScience
-- AH, SciData
-- Mozilla
-- PLOS?
-- Richard van Noorden, Dalmeet Singh (paper 9)
-- Iain Davies and Sebastian 
-- SSI for some funding.
+We are grateful to the following individuals for discussions regarding
+the work presented here: Andy Collings, Melissa Harrison, Giuliano
+Maciocci (eLife), Rebecca Kirk (PLOS Computational Biology), Scott
+Edmunds (Gigascience) Andrew Hufton (Scientific Data). Iain Davies and
+Yuhao (Sebastian) Wang worked on developing code and example
+certificates.  This work was financially supported by the UK Software
+Sustainability Institute and a Mozilla Science mini grant.
 

### 2020-09-15 codecheck-paper.Rmd
rough edits from discussion

@@
 
-* Prefer longlines or short?
-* Can we switch to latex for this?  (Find it slightly easier.)
+* Prefer one sentence per line / short / one para per line.
+* Nust* and Eglen*. contributed equally.
+* latex.
 * Which journal to target -- maybe worry abotu that once we have a
-  preprint.  F1000R my current favourite?
+  preprint.  F1000R my current favourite.
 
-* Check with Giuliano re: "code works".
+* Check with Giuliano re: "code works".  Write to Naomi to apologise / placeholder.
 
@@
 
-So, assuming that researchers begin to share more artifacts, how
-might they be examined to test that they do what they claim?  For
-example, most scientific journals now require a data sharing statement
-that outlines what data the authors have (or will) share.  The
-implementation status of this varies considerably according to the
-journal.  At one of the spectrum, there are specialist journals  that
-have been created to accept "data papers" (e.g. *Scientific Data*,
-... NAME COUPLE OF OTHERS).  These journals have established rigorous
-procedures by which the data is validated according to standards in
-each field.  At the other end of the spectrum, we still witness today
-the infamous statement "Data available upon reasonable request"
-whereby authors, well-intentioned, often cannot provide the data when
-readers ask for it.
+So, assuming that researchers begin to share more artifacts, how might they be examined to test that they do what they claim?
+For example, most scientific journals now require a data sharing statement that outlines what data the authors have (or will) share.
+The implementation status of this varies considerably according to the journal.
+At one of the spectrum, there are specialist journals that have been created to accept "data papers" (e.g.
+*Scientific Data*, ...
+NAME COUPLE OF OTHERS).
+These journals have established rigorous procedures by which the data is validated according to standards in each field.
+At the other end of the spectrum, we still witness today the infamous statement "Data available upon reasonable request" whereby authors, well-intentioned, often cannot provide the data when readers ask for it.
+(ANDREW H: any references to sum up poor/evolving state of data sharing; Vines et al 2014).
 
-Given that there are no clear standards yet for sharing data, what
-hope might there be for sharing computer programs?  Our anectodal
-experiences on this matter suggest that often a variety of reasons are
-given for why their code cannot be shared, e.g. "there is no
-documentation / I do not want to maintain it / I do not want to give
-away my code to my competitors".  Our view, rooted in open science
-princples, is that wherever possible sharing code is good for the
-community, as outlined ten years ago (Barnes, 2010).  Having the code
-freely available, and suitably archived, provides a valuable resource
-for others to learn from, even when it doesn't run or if there is no
-documentation.  However, with a little effort, we believe that if
-someone independent can run/use the code, this is worth documenting as
-early as possible.  With this in mind, we have developed a set of
-PRINCIPLES and an example workflow that provides what he hope is a
-pragmatic way of checking that code works.  We call this system
-CODECHECK as outlined next.
+Given that there are no clear standards yet for sharing data, what hope might there be for sharing computer programs?
+Our anectodal experiences on this matter suggest that often a variety of reasons are given for why their code cannot be shared, e.g.
+"there is no documentation / I do not want to maintain it / I do not want to give away my code to my competitors".
+Our view, rooted in open science princples, is that wherever possible sharing code is good for the community, as outlined ten years ago (Barnes, 2010).
+Having the code freely available, and suitably archived, provides a valuable resource for others to learn from, even when it doesn't run or if there is no documentation.
+However, with a little effort, we believe that if someone independent can run/use the code, this is worth documenting as early as possible.
+With this in mind, we have developed a set of PRINCIPLES and an example workflow that provides what he hope is a pragmatic way of checking that code works.
+We call
+this system CODECHECK as outlined next.
 
@@
 in house staff, pool of ECRs, volunteeres,
+for inhouse staff, credit not an issue.
 
@@
 
-**3. Credit is given to codecheckers.** (TODO Explain in a few
-sentences what each means, e.g. see 1,2)
+**3. Credit is given to codecheckers.** (TODO Explain in a few sentences
+what each means, e.g. see 1,2) Ideally publons, or similar, but can just
+be a citation to the publicly available DOI.  (Community DOIs use
+these.)
 
@@
 what each means, e.g. see 1,2)
+
+Do we want to list them in priority?  Open by default.  Community principles.
 
@@
 2. report 9 -- covid model of UK lockdown model.
-3. Geoscience
+3. Geoscience  (add 6 for Agile).
 
 ## Annotated certificate
+
+What can you expect to find a certificate?  Draw an annotated example.
 
@@
 
+### Tools
 To run the system, we rely on github to host our work, and use a
@@
 
+You can do all with open infrastructure (e.g. automated rendering of
+register). but you can use your own tools if you prefer.
+
 # Related work
+
+TO BE CONTINUED.
 
@@
 
+DELIBERATEY NOT COMPREHENSIVE HERE.
+
 # Limitations
 
-Handling failure:  we have no established process for this, as so far
+**Handling failure**:  we have no established process for this, as so far
 all our codechecks succeeded.  Open question remains as what to do in
@@
 
-Compute time:  for those papers that take significant compute time
+**Compute time**:  for those papers that take significant compute time
 (think days, not minutes), who will pay for the compute time?
 
-Software: authors can currently provide code that requires
+**Software**: authors can currently provide code that requires
 propieratiry software.  Given the prevalence of software like MATLAB,
@@
 
-Can't someone cheat? yes.  we don't check it is correct code, just
+**Can't someone cheat?** yes.  we don't check it is correct code, just
 simply that it runs.  Very low bar.  But by having the code open,
@@
 
-Who's got time for more peer review?  Agree, but ECRS might be
+**Who's got time for more peer review?**  Agree, but ECRS might be
 attracted to working with groups (recall it is not anonymous)
 
-Who finds the codechecker?  Good point!  The journal.
+**Who finds the codechecker?** Good point!  The journal.  Community
+however think this is an important issue, and are volunteering.
 
@@
 - How to embed this in juornals.
-- ???
+- When is CODECHECK redundant?  will it become part of regular peer review?
+or, if ayuthors become more skillful, then everything becomes automated, and no need for codecheck.
+CODECHECK disappearing is then a measure of success.
+
+Or will it always be adjunct to peer review, and needed?
+
+working within the trad system of publishers to help push the conversation along.
 

### 2020-09-16 codecheck-paper.Rmd
Get the .tex output

@@
 #csl: plos.csl
-output: pdf_document
+output:
+  pdf_document:
+    keep_tex: true  
 ---
@@
 
-# Open questions
+# TODO
 
-* Prefer one sentence per line / short / one para per line.
 * Nust* and Eglen*. contributed equally.
-* latex.
-* Which journal to target -- maybe worry abotu that once we have a
-  preprint.  F1000R my current favourite.
 
-* Check with Giuliano re: "code works".  Write to Naomi to apologise / placeholder.
+* Check with Giuliano re: "code works".  Write to Naomi.
 
@@
   \centering
+	  \includegraphics{/home/stephen/papers/talks/talkfigs/openscience/codecheck_overview.pdf}
   \caption{The detective. Could we number the arrows as steps?
@@
 
-## Core principles
+# Core principles
 

### 2020-09-16 codecheck-paper.tex
Initial commit of .tex

\documentclass[12pt]{article}
\usepackage{f1000_styles}
\usepackage[numbers]{natbib}
\usepackage{hyperref,url,setspace}

\onehalfspacing

\begin{document}
\title{CODECHECK: An open-science initiative to facilitate sharing of computer programs and results presented in scientific publications.}
\author[1,$\ast$]{Daniel N\"{u}st}
\author[2,$\ast$]{Stephen J Eglen}
\affil[1]{Address of author1}
\affil[2]{Address of author2}
\affil[$\ast$]{Contributed equally}
\maketitle
\begin{abstract}
\ldots{}
\end{abstract}

\section*{TODO}\label{todo}

\begin{itemize}
\item Check with Giuliano re: ``code works''. Write to Naomi.
\end{itemize}

\section*{Introduction}\label{introduction}

Many areas of scientific research now use computation to either simulate
or analyse their data. These computations are increasingly complex, and
difficult to explain coherently in a paper. (MARWICK). To complement the
traditional route of sharing research by writing papers, there is a
growing trend/demand to share the underlying artifacts, notably code and
datasets, so that others can inspect, reproduce or expand that work
(Figure 1). Some of the earliest proponents were Buckheit and Donoho
(1995) who commented ``advert is scholarship for the paper''. (get
correct quote).

So, assuming that researchers begin to share more artifacts, how might
they be examined to test that they do what they claim? For example, most
scientific journals now require a data sharing statement that outlines
what data the authors have (or will) share. The implementation status of
this varies considerably according to the journal. At one of the
spectrum, there are specialist journals that have been created to accept
``data papers'' (e.g.~\emph{Scientific Data}, \ldots{} NAME COUPLE OF
OTHERS). These journals have established rigorous procedures by which
the data is validated according to standards in each field. At the other
end of the spectrum, we still witness today the infamous statement
``Data available upon reasonable request'' whereby authors,
well-intentioned, often cannot provide the data when readers ask for it.
(ANDREW H: any references to sum up poor/evolving state of data sharing;
Vines et al 2014).

Given that there are no clear standards yet for sharing data, what hope
might there be for sharing computer programs? Our anectodal experiences
on this matter suggest that often a variety of reasons are given for why
their code cannot be shared, e.g.~``there is no documentation / I do not
want to maintain it / I do not want to give away my code to my
competitors''. Our view, rooted in open science princples, is that
wherever possible sharing code is good for the community, as outlined
ten years ago \cite{Barnes2010-iv}. Having the code freely available, and
suitably archived, provides a valuable resource for others to learn
from, even when it doesn't run or if there is no documentation. However,
with a little effort, we believe that if someone independent can run/use
the code, this is worth documenting as early as possible. With this in
mind, we have developed a set of PRINCIPLES and an example workflow that
provides what he hope is a pragmatic way of checking that code works. We
call this system CODECHECK as outlined next.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/rr.pdf}
  \caption{The inverse problem in reproducible research.  The left
  half of the diagram shows the diverse range of materials used
  internally within a laboratory.  These materials are often then
  condensed into a paper for sharing with the outside world via the
  research paper, a static PDF document.  Working backwards from the
  PDF to the underlying materials is impossible.  By sharing the
  materials on the left, others outside the lab are able to reproduce
  or build on this work.  (Should we draw a backward arrow showing the
  impossibility of this backward flow?)}
  \label{fig:inverse}
\end{figure}

\section*{What is a CODECHECK?}\label{what-is-a-codecheck}

CODECHECK is best demonstrated by way of our example workflow, and later
we expand on the underlying principles. The workflow involves three
groups of people (1) the AUTHOR providing the code to be checked, (2)
the PUBLISHER of a journal intereseted in publishing the paper by the
AUTHOR and (3) the CODECHECKER who checks that the AUTHOR's code works.
The workflow that we have refined is documented in Figure 2.

Step 1: the author submit their manuscript along with code and data to
the publishers. The code and data need not be openly available at this
point.

Step 2: the publisher finds a codechecker to check the code. This is
analgous to the publisher finding one or more reviewers to evaluating
the code. However a cruical difference is that we suggest the codechcker
and the author can talk freely to each directly (see next).

Step 3: the codechecker runs the code, based on instructions provided by
the author, and checks if some/all of the results from the paper can be
reproduced. (See later for exact nature of ``reproduced''). If there are
any problems running the code, the codechecker asks the author for help
to resolve the problems. The codechecker then tries to run the code
again. This process iterates until either the codechecker is successful,
or the codechecker concludes the code does not (\emph{what do in this
case}). As part of this process, the codechecker could work entirely
locally on their own compute resource, or in the cloud, e.g.~using
mybinder infrastructure. Using such cloud-based infrastructure allows
for the codechecker and author to collaboratively run the code together.

Step 4: The codechecker writes a certificate stating how the code was
run and includes a copy of outputs (figures, tables) that were
independently generated.

Step 5: the certificate and code/data get deposited on an open archive
(currently Zenodo), and the certificate is given to the PUBLISHER.

Step 6: the Publisher can then use the certificate as part of their
supplementary records for a paper, and give credit to the codechecker
for their work by depositing appropriate records (\emph{bit hazy here}).

\begin{figure}
  \centering
      \includegraphics{/home/stephen/papers/talks/talkfigs/openscience/codecheck_overview.pdf}
  \caption{The detective. Could we number the arrows as steps?
(Implementing the CODECHECK process).  DANIEL: best source for this figure?}
  \label{fig:worfklow}
\end{figure}

\subsection*{Variations}\label{variations}

This example workflow leaves sufficient room for different
implementations of this process. There are several aspects that can be
considered:

\subsubsection*{When to do a codecheck}\label{when-to-do-a-codecheck}

preprint (no publisher available), pre-condition for peer review, during
peer review, after acceptance. Each of these have their own pros/cons.

\subsubsection*{Who does the codecheck}\label{who-does-the-codecheck}

in house staff, pool of ECRs, volunteeres, for inhouse staff, credit not
an issue.

\subsubsection*{Who knows who?}\label{who-knows-who}

single blind, double blind, private/public communication between author
and codechecker, collaborative. (This might need finessing with
principle 2 below).

\section*{Core principles}\label{core-principles}

The workflow and variations just outlined below reflect our current
views on how a codecheck should be performed. They are not set in
stone, but we do believe the following core principles underpinour
CODECHECK:

\textbf{1. Codecheckers record but don't investigate or fix.} The
codechecker follows the instructions provided by the author in running
the code. If any steps are unclear, or if code does not run correctly,
the codechecker records this fact, and then communicates this
information to the author. We believe that the job of the codechecker is
not to fix all these problems, but simply to report them, and await a
fix from the author.

\textbf{2. Communication between humans is key.} Some code may just work
without any interaction (e.g.~Manuel's certificate); however, more often
there are hidden dependencies that need to be found, or revised software
installations. Allowing the codechecker and human to communicate
directly and openly with each other should make this process as
constructive as possible. We feel that routing this conversation
(possibly anonymously) through a publisher would slow things down and
reduce the chances for community bulding.

\textbf{3. Credit is given to codecheckers.} (TODO Explain in a few
sentences what each means, e.g.~see 1,2) Ideally publons, or similar,
but can just be a citation to the publicly available DOI. (Community
DOIs use these.)

\textbf{4. Workflows must be auditable.} (TODO Explain in a few
sentences what each means, e.g.~see 1,2)

\textbf{5. Open by default.}  By default, unless there are strong
reasons to the contrary (e.g. clinically-sensitive data), all code and
data, both from author and codechecker, will be made freely available
when the certificate is published, following community principles.

\section*{Register}\label{register}

We have a curated list of 15 certificates available at
\url{https://codecheck.org.uk/register}. These are a mixture of
reproductions of papers for a variety of reasons (do we name them all?)
from interesting historical papers, to papers codechecked as part of
peer review, or because of public interest (covid). several covid models
have been done during the time of peer review of these papers, under our
initative rather than requested from journal, but certificate have then
been acknowledged/cited in the paper (LANCET / LSHTM).

We would like to briefly highlight three examples here.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Gigascience -- our first paper. Only visualisation, not machine
  learning (which would have taken sseveral days of compute time).
\item
  report 9 -- covid model of UK lockdown model.
\item
  Geoscience (add 6 for Agile).
\end{enumerate}

\subsection*{Annotated certificate}\label{annotated-certificate}

What can you expect to find a certificate? Draw an annotated example.

Should we annotate a certificate, at least the cover page, to
demonstrate what we mean by a certificate? We could print a four page
certificate 4-up so that it fits onto one page. Text might be too small
though to read? A 2-up, printed sideways would work well though.

\subsubsection*{Tools}\label{tools}

To run the system, we rely on github to host our work, and use a custom
R pacakge, codecheck, to help with the generation of certificates and
depositing work on zenodo. (Cite zen4R).

You can do all with open infrastructure (e.g.~automated rendering of
register). but you can use your own tools if you prefer.

\section*{Related work}\label{related-work}

TO BE CONTINUED.

All the related projectsfrom A-Z\ldots{} e.g.~CASCAID, Rescience,
modeldb, o2r. Community iniatives like reprohack along similar lines of
checking code. \url{https://reprohack.github.io/reprohack-hq/}

DELIBERATEY NOT COMPREHENSIVE HERE.

\section*{Limitations}\label{limitations}

\textbf{Handling failure}: we have no established process for this, as
so far all our codechecks succeeded. Open question remains as what to do
in this case -- should we publicly report that the code would not run.

\textbf{Compute time}: for those papers that take significant compute
time (think days, not minutes), who will pay for the compute time?

\textbf{Software}: authors can currently provide code that requires
propieratiry software. Given the prevalence of software like MATLAB,
this seems like a pragmatic choice, but then (a) we can't rely on open
infrastructure like mybinder, (b) codechecker must have that software.

\textbf{Can't someone cheat?} yes. we don't check it is correct code,
just simply that it runs. Very low bar. But by having the code open,
interested parties can then examine the code, and knowing that the code
will be open we hope will be an incentive for authors to share.

\textbf{Who's got time for more peer review?} Agree, but ECRS might be
attracted to working with groups (recall it is not anonymous)

\textbf{Who finds the codechecker?} Good point! The journal. Community
however think this is an important issue, and are volunteering.

As a principle, we do not require results to be exactly the same for a
codecheck to pass, simply that the code runs and generates the output
files that the author claims. Stochastic simulations mean that often we
cannot get exactly the same, or even different versions of libraries.
(Cite freesurfer example that generates different outputs on different
operating systems).

\begin{itemize}
\item
  modelDB principle: minimum of one figure should be reproducibe.
  Exampler certificate 2020-016 (I think) which had only a small subset
  of the possible figures. Still useful and worth sharing.
\end{itemize}

\section*{Future work and conclusions}\label{future-work-and-conclusions}

\begin{itemize}
\item
  possibility to allow CI for open papers such that for ``nearly free''
  we can find out when softwave breaks, and if provenance tools are
  used, when.
\item
  How to embed this in juornals.
\item
  When is CODECHECK redundant? will it become part of regular peer
  review? or, if ayuthors become more skillful, then everything becomes
  automated, and no need for codecheck. CODECHECK disappearing is then a
  measure of success.
\end{itemize}

Or will it always be adjunct to peer review, and needed?

working within the trad system of publishers to help push the
conversation along.

\subsection*{Competing interests}

SJE is on the senior editorial board at the journal \emph{Scientific
Data}.

\subsection*{Grant information}

This work was financially supported by the UK Software
Sustainability Institute and a Mozilla Science mini grant.

\subsection*{Acknowledgements}\label{acknowledgements}

Collaborators and potential collaborators:

We are grateful to the following individuals for discussions regarding
the work presented here: Andy Collings, Melissa Harrison, Giuliano
Maciocci (eLife), Rebecca Kirk (PLOS Computational Biology), Scott
Edmunds (Gigascience) Andrew Hufton (Scientific Data). Iain Davies and
Yuhao (Sebastian) Wang developed code and example certificates.

\subsection*{Author contributions}

DN and SJE contributed equally to all aspects of this project.

{\small\bibliographystyle{unsrtnat}
\bibliography{bibliography}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage
\begin{itemize}
\item
  role of data and code, data science
\item
  reproducible resarch
\item
  data enclaves
\item
  CASCAID
\item
  ROpenSci/PyOpenSci (and how their software review differs, though they
  also collaborate with journals,
  \url{https://devguide.ropensci.org/softwarereviewintro.html})
\item
  time capsules \textgreater{}
  \url{https://twitter.com/DougBlank/status/1135904909663068165?s=09}
\item
  science is changing:
  \url{https://theconversation.com/what-you-need-to-know-about-how-coronavirus-is-changing-science-137641}
\item
  reproducibility is important, also

  \begin{itemize}
  \item
    effects on well-being/mental health,
    \url{https://www.nature.com/articles/d41586-020-01642-9}
  \end{itemize}
\end{itemize}

\textbf{main contributions}

\begin{itemize}
\item
  a concept for integrating a minimal code review into common scholarly
  communication processes around peer review
\item
  set of shared principles (recognition value) that allow researchers to
  quickly understand the level of evaluation across
  journals/publishers/conferences, and allows helps these stakeholders
  to establish workflow checking in their domain/solution/product
\item
  a common language and rubric for classifying different levels of code
  and data evaluation as part of a peer-review
\end{itemize}

\section*{What is CODE CHECK?}\label{what-is-code-check}

\begin{itemize}
\item
  \textbf{Principles}: \url{https://codecheck.org.uk/}
\item
  relation to ``proper'' citation of data and software, and depositing
  of data and software in suitable repos \emph{besides} the bundle for
  CODE CHECK: do it, only provide the concrete analysis script for the
  check
\item
  \textbf{Why is it useful?}

  \begin{itemize}
  \item
    reduce the barrier to evaluating non-text parts of research
  \item
    acknowledges different skill sets and existing high load on
    reviewers
  \item
    CC breaks problem that publishers staff expertise today is not
    likely to suffice
  \item
    engaging ECRs is a perfet fit (they are up-to-date in methods, are
    introduced to peer review)
  \item
    relation to research compendia? point it out for authors unsure
    about how to structure their workflow
  \item
    CrossRef and Publons and ORCID have features around reviews
  \item
    shifting burden from reviewer to author (but not too much)
  \item
    What issues with (open) peer review does CODECHECK address?
  \end{itemize}
\item
  keynote \url{https://doi.org/10.7557/5.4963}: in discussion, the
  speaker mentioned value of peer review for ``baseline reassurance'',
  i.e.~a paper has at least been checked by someone with an
  understanding, to increase trust especially if looking at papers from
  other disciplines
\item
  \textbf{report} must fulfil some requirements to make clear what is
  really checked and not mislead - readers will have very different
  understandings of a green checkmark!
\item
  reproducibility is hard
  \url{https://twitter.com/annakrystalli/status/1144176149859377152?ref_src=twsrc\%5Etfw}
  \textgreater{} code check gives at least some benefits
\item
  what are the requirements / level or reproducibility required or
  checked?

  \begin{itemize}
  \item
    \href{https://stodden.net/icerm_report.pdf}{Stodden et al., 2013}:
    five-level hierarchy of research - reviewable, replicable,
    confirmable, auditable, and open or reproducible
  \item
    \href{https://figshare.com/articles/Showing_your_working_A_guide_to_reproducible_neuroimaging_analyses/4244996/2}{Whitaker
    2017}: reproducible, generalisable, replicability, robustnes
  \end{itemize}
\item
  can authors retract their code/data once a CODECHECK has started? In
  the eLife prepreint review, there seems no ``way back'', see
  \url{https://twitter.com/ceptional/status/1271528976763121664?s=09}
\end{itemize}

\section*{Related work}\label{related-work-1}

\begin{itemize}
\item
  CODECHECK vs.~Verification Reports \textgreater{} \textgreater{}
  \url{https://www.sciencedirect.com/science/article/pii/S0010945220301738\#fn3}
\item
  CODECHECK vs.~the ``Artifacts Evaluated - Functional'' badge of the
  ACM, see
  \url{https://www.acm.org/publications/policies/artifact-review-badging}
\item
  Does CODECHECK do internal replication, as defined by
  \url{https://gatesopenresearch.org/articles/4-17/v2} ?
\item
  \url{https://paperpile.com/shared/rVNwBS}
\item
  Are ``Replicated Computations Results (RCR) Report'' a thing?
  apparently so (using ACM terminology)

  \begin{itemize}
  \item
    \url{https://dl.acm.org/doi/10.1145/3185337}
  \item
    \url{http://users.iems.northwestern.edu/~nelsonb/Publications/Nelson_RCRgreenSim.pdf}
  \item
    \url{https://dl.acm.org/doi/epdf/10.1145/3341094}

    \begin{itemize}
    \item
      citation for replicated workflow lists it as ``to appear''
    \end{itemize}
  \item
    getting proper papers for reproductions != CODECHECK approach where
    we want credit as reviewers
  \item
    see also
    \href{https://www.google.com/search?q=\%22Replicated+Computations+Results+\%28RCR\%29+Report+for\%22}{Google
    Search}
  \end{itemize}
\item
  \url{https://medium.com/bits-and-behavior/a-modern-vision-for-peer-review-d5f73f0fae07}
  - work by Amy J. Ko:
  \url{https://faculty.washington.edu/ajko/publications}

  \begin{itemize}
  \item
    is CODE CHECK's process too small, too incremental a change?
  \end{itemize}
\item
  overview of peer review and bibliometrics (if not a reference itself,
  it gives plenty of references):
  \url{http://juser.fz-juelich.de/record/865096/files/Peer\%20Review\%20and\%20Bibliometrics.pdf}
\item
  Open Peer Review \textgreater{}
  \url{https://researchintegrityjournal.biomedcentral.com/articles/10.1186/s41073-019-0063-9}

  \begin{itemize}
  \item
    Are OPR lessons transferable to CODECHECK? Does CODECHECK work
    better with OPR than with (single/double) blind peer review?
  \item
    similarities to their advice: open reports, give credit
  \end{itemize}
\item
  \emph{Successes and struggles with computational reproducibility:
  Lessons from the Fragile Families Challenge} \textgreater{}
  \url{https://osf.io/preprints/socarxiv/g3pdb/}
\item
  community standards for software, cf.~slide 6 in
  \url{https://figshare.com/articles/_/5675104}

  \begin{itemize}
  \item
    ``good enough practice'' is ``get a colleague to try using it'',
    slide 19 in \url{https://figshare.com/articles/_/5675104}
  \end{itemize}
\item
  journals doing code reviews

  \begin{itemize}
  \item
    \emph{``We know of 16 journals that take this valuable step (ie TOP
    Level 3): AEA as already mentioned, plus others in polisci,
    psychology, biostats, and org chem: see Data Rep Policies here
    \url{https://t.co/plOF8j6ADU}''}
    (\url{https://twitter.com/EvoMellor/status/1202692360456589339?s=09});
    see also whole thread!
  \item
    \url{https://medium.com/@NeurIPSConf/call-for-papers-689294418f43}
  \item
    \url{https://www.journals.elsevier.com/information-systems/editorial-board/}
  \item
    SIGMOD Reproducibility Review?
  \item
    Biostatistics ``AER''
    (\url{https://academic.oup.com/biostatistics/article/10/3/405/293660})
  \end{itemize}
\item
  open science platforms/cloud services \textgreater{}
  \url{https://www.nature.com/articles/d41586-019-03366-x}
\item
  live code in articles \textgreater{} o2r, eLife RDS
  \url{https://www.nature.com/articles/d41586-019-00724-7}, \ldots{}
\item
  Artefact Evaluation - \url{https://www.artifact-eval.org/about.html}
\item
  continuous analysis - \url{https://www.nature.com/articles/nbt.3780}
\item
  ``A Universal Identifier for Computational Results'' \textgreater{}
  \url{https://www.sciencedirect.com/science/article/pii/S1877050911001256}
\item
  Discussion about using Docker containers during JOSS reviews
  \textgreater{}
  \url{https://github.com/openjournals/joss/issues/498\#issuecomment-462046912}
\item
  \url{https://scigen.report}, an independent site for registering
  reproductions, via
  \url{https://annakrystalli.me/talks/ro-reprohack.html\#32}

  \begin{itemize}
  \item
    example
    \url{https://scigen.report/reviews/get?doi=10.1063/1.1823034}
  \item
    seems to be closed infrastructure
  \end{itemize}
\item
  new journal \emph{ReScienceX} (about experiments) also seems to have
  an ECR angle, \url{https://www.nature.com/articles/d41586-020-01328-2}
\item
  really interesting work on reproductions in hydrology, with a neat
  survey that reproducers filled out which might be interesting to
  capture from codecheckers, too:
  \url{https://www.nature.com/articles/sdata201930\#Sec9}
\end{itemize}

\section*{CODE CHECK implementation concepts}\label{code-check-implementation-concepts}

\begin{itemize}
\item
  need to get wording straight: process/implementation/workflow (DN:
  suggest to use ``workflow'' for research-analysis workflow)
\item
  \url{https://codecheck.org.uk/process/}
\item
  \emph{dimensions} along which a check can vary
\item
  independent/pre-submission (Peer\_Coding) by author + colleagues
\item
  AGILE
\item
  premier (OA) publisher
\item
  community OA / scholarly society with public peer review
\item
  invited reproducibility paper track,
  cf.~\url{https://www.elsevier.com/journals/information-systems/0306-4379/guide-for-authors}
\item
  on giving credit: public databases

  \begin{itemize}
  \item
    \url{https://www.reviewercredits.com/}
  \item
    Publons
  \item
    ORCID
  \end{itemize}
\item
  compare with existing processes that are publicly documented

  \begin{itemize}
  \item
    AJPS Verification Policy:
    \url{https://ajps.org/ajps-verification-policy/}
  \item
    \ldots{}
  \end{itemize}
\item
  flexible level of detail? a check can be enhanced to (potentially
  semi-automatically) check for good practices

  \begin{itemize}
  \item
    ``A group of us in the ({\textbf{???}}) Software Citation
    Implementation Working Group produced a''Software Citation Checklist
    for Authors" \url{https://t.co/4Noe66FsoX} providing guidelines on
    how to cite software and describing why it is important and what
    software should be cited" via
    \url{https://twitter.com/alegonbel/status/1207664922932453376?s=09})
  \end{itemize}
\item
  focus on ECR as codecheckers

  \begin{itemize}
  \item
    ECR are the one's breaking new ground in schol comm, cf.~the
    percentages of career stages submitting registered reports on slide
    20 \textgreater{} \url{https://osf.io/2vjus/}
  \end{itemize}
\end{itemize}

\section*{Annotated Examples}\label{annotated-examples}

\begin{itemize}
\item
  \textbf{the default implementation}
\item
  critically discuss our/the default implementation vs.~the principles
\item
  HPC example
\item
  if we see the corpus of examles as a scientific dataset, it should be
  published in a citable way (maybe even submit to a data journal)
\end{itemize}

\section*{Open problems}\label{open-problems}

\begin{itemize}
\item
  Do publishers need to build up data science experts who can conduct
  checks?
\item
  Is an integration of software/data citation checks
  benefitial/realistic?
\item
  Bot-support and tools (linter, report authoring with R Markdown)
\item
  How can open platforms be used to enable a review in private
  (cf.~\url{https://www.cambridge.org/core/blog/2019/08/19/how-to-make-the-data-and-code-for-your-manuscript-available-to-peer-reviewers-before-making-it-public/}),
  if that is desired
\item
  Does nudging towards better practices work (highlight the successes)
  or is any check with opt-in doomed to fail?
\item
  Do we need more than one codechecker? Is there a similar diversity as
  in a peer review where more opinions are needed?
\item
  can CODECHECK work for preprints? how? is a re-check required?
\item
  Risk: is the hurdle too low? are ``levels'' of checks are solution to
  increase transparency of check's level of detail
\item
  CODECHECK in the context of innovations and disruptions in scientific
  publishing

  \begin{itemize}
  \item
    \url{https://doi.org/10.1042/ETLS20180172}
  \item
    \url{https://doi.org/10.3390/publications7020034}
  \end{itemize}
\item
  Should every code be checkd, or should the labour be capped by only
  checking ``important'' works or at important stages; if yes, isn't
  peer review such a stage?
\item
  see \url{https://github.com/codecheckers/discussion}
\item
  Can we establish code checks as a type of review within the schol comm
  metadata (or should we?) \textgreater{} we should not, put them on the
  same level as reviews!
\item
  How can we turn a CODECHECK into a positive experience?

  \begin{itemize}
  \item
    cf.~kindness in peer review in
    \url{https://www.mdpi.com/2304-6775/8/2/26}
  \end{itemize}
\item
  How can we allow reactions to checks? (commenting on reports)
\item
  Will bitwise reproducibility become easier than what a CODECHECK does?
  triggered by
  \url{https://twitter.com/khinsen/status/1242842759733665799?s=09})
\item
  If a report is not published for failures: Isn't there a bias for the
  codechecker to have a ``successful'' reproduction because then a
  report is published?
\end{itemize}

\end{document}

### 2020-09-17 codecheck-paper.tex
ship out ...

@@
 \usepackage{hyperref,url,setspace}
+
+\usepackage{booktabs}
 
@@
 
-What can you expect to find a certificate? Draw an annotated example.
+What can you expect to find in  a CODECHECK certificate? Draw an annotated example.
 
-Should we annotate a certificate, at least the cover page, to
-demonstrate what we mean by a certificate? We could print a four page
-certificate 4-up so that it fits onto one page. Text might be too small
-though to read? A 2-up, printed sideways would work well though.
+\begin{table}
+  \centering
+
+  \begin{tabular}{lllp{9cm}}  \toprule
+    Certificate & Ref & Research area & Description \\ \midrule
+    2020-001 & \cite{cert-2020-001} & Machine learning & CODECHECK undertaken post acceptance of manuscript and before its publication in \textit{Gigascience} \cite{Piccolo2020-lo}.
+    \\ 
+    2020-002 & \cite{cert-2020-002} & neuroscience &
+    Paper from 1997 showing unsupervised learning on images \cite{Hancock1992-mp} \\
+    2020-003 & \cite{cert-2020-003} & neuroscience &
+    Classic paper from 1982 on  models of associative memory \cite{Hopfield1982-mz} \\
+    2020-004 & \cite{cert-2020-004} & area & barto \\
+    2020-005 & \cite{cert-2020-005} & area & Larisch \\
+    2020-006 & \cite{cert-2020-006} & area & Detorakis \\
+    2020-007 & \cite{cert-2020-007} & area & in progress \\
+    2020-008 & \cite{cert-2020-008} & Covid & covid-uk Davies \\
+    2020-009 & \cite{cert-2020-009} & Covid &  tracing \\
+    2020-010 & \cite{cert-2020-010} & Covid &  report 9 \\
+    2020-011 & \cite{cert-2020-011} & Covid &  nature europe \\
+    \\ \bottomrule
+  \end{tabular}
+  \caption{Register of completed certificates.  An interactive version
+  of this is available at \url{http://codecheck.org.uk/register}.}
+  \label{tab:register}
+\end{table}
+
+As an example, we have taken the certificate 2020-012
+\cite{cert-2020-012}, and shown the first four pages (the remaining N
+pages contain other example outputs), see Figure~\ref{fig:cert}.  This
+certificate corresponded to a preprint on the modelling of COVID-19 in
+the USA.i
+%% TODO - has this been published yet?.
+Figure~\ref{fig:cert} shows an annotation of the first four (out of
+ten) pages for the certificate for this codecheck.
+Figure~\ref{fig:cert}A shows the certificate number, and its DOI,
+which points to the certificate as archived on Zenodo.
+Figure~\ref{fig:cert}B provides the key metadata about the check --
+which paper was checked, and who/when it was checked.  Materials
+relating to the codecheck are freely available via the Repository
+link.  Figure~\ref{fig:cert}C is a summary of how the codecheck was
+performed, and any interesting findings.
+
+Turning to page 2 of the certificate, in Figure~\ref{fig:cert}D we
+have a table that is a list of the outputs that were generated from
+the codecheck (this is termed the MANIFEST.)  For each output, there
+is a comment stating which figure/table it should be compared to in
+the original paper, along with the size of the file.  On Page 3,
+Figure~\ref{fig:cert}E shows some more detailed notes from the
+codechecker, in this case indicating what steps were needed to
+initialize the environment and run the code.  This particular example
+was quite long, taking about 17 hours to run.  Finally, on page 4 of
+the certificate, we see the first output that was generated by the
+codecheck (Figure~\ref{fig:cert}F).  In this case, the figure matches
+figure 4 of REF.  Subsequent pages of the certificate show other
+outputs (not shown here).
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{figs/annotate-cert-crop.pdf}
+  \caption{Annotated certificate 2020-012 (first four pages only).}
+  \label{fig:cert}
+\end{figure}
 

### 2020-09-17 codecheck-paper.tex
add reference article for 2020-012

--- 2020-09-16 codecheck-paper.tex
@@
 
+\usepackage{booktabs}
+
 \onehalfspacing
@@
 \begin{document}
-\title{CODECHECK: An open-science initiative to facilitate sharing of computer programs and results presented in scientific publications.}
+\title{CODECHECK: An open-science initiative to facilitate sharing of
+  computer programs and results presented in scientific publications.}
 \author[1,$\ast$]{Daniel N\"{u}st}
 \author[2,$\ast$]{Stephen J Eglen}
-\affil[1]{Address of author1}
-\affil[2]{Address of author2}
+\affil[1]{\url{https://orcid.org/0000-0002-0024-5046}}
+\affil[2]{\url{https://orcid.org/0000-0001-8607-8025}}
 \affil[$\ast$]{Contributed equally}
@@
 So, assuming that researchers begin to share more artifacts, how might
-they be examined to test that they do what they claim? For example, most
-scientific journals now require a data sharing statement that outlines
-what data the authors have (or will) share. The implementation status of
-this varies considerably according to the journal. At one of the
-spectrum, there are specialist journals that have been created to accept
-``data papers'' (e.g.~\emph{Scientific Data}, \ldots{} NAME COUPLE OF
+they be examined to test that they do what they claim? For example,
+most scientific journals now require a data sharing statement that
+outlines what data the authors have (or will) share. The
+implementation status of this varies considerably according to the
+journal. At one of the spectrum, there are specialist journals that
+have been created to accept ``data papers'' (e.g.~\emph{Scientific
+  Data}, \textit{Geoscience Data Journal}\ldots{} NAME COUPLE OF
 OTHERS). These journals have established rigorous procedures by which
-the data is validated according to standards in each field. At the other
-end of the spectrum, we still witness today the infamous statement
-``Data available upon reasonable request'' whereby authors,
-well-intentioned, often cannot provide the data when readers ask for it.
-(ANDREW H: any references to sum up poor/evolving state of data sharing;
-Vines et al 2014).
+the data is validated according to standards in each field. At the
+other end of the spectrum, we still witness today the infamous
+statement ``Data available upon reasonable request'' whereby authors,
+well-intentioned, often cannot provide the data when readers ask for
+it.  (ANDREW H: any references to sum up poor/evolving state of data
+sharing; Vines et al 2014).
 
@@
 
-Step 1: the author submit their manuscript along with code and data to
+\textbf{Step 1:} the author submit their manuscript along with code and data to
 the publishers. The code and data need not be openly available at this
@@
 
-Step 2: the publisher finds a codechecker to check the code. This is
+\textbf{Step 2:} the publisher finds a codechecker to check the code. This is
 analgous to the publisher finding one or more reviewers to evaluating
@@
 
-Step 3: the codechecker runs the code, based on instructions provided by
+\textbf{Step 3:} the codechecker runs the code, based on instructions provided by
 the author, and checks if some/all of the results from the paper can be
@@
 
-Step 4: The codechecker writes a certificate stating how the code was
+\textbf{Step 4:} the codechecker writes a certificate stating how the code was
 run and includes a copy of outputs (figures, tables) that were
@@
 
-Step 5: the certificate and code/data get deposited on an open archive
+\textbf{Step 5:} the certificate and code/data get deposited on an open archive
 (currently Zenodo), and the certificate is given to the PUBLISHER.
 
-Step 6: the Publisher can then use the certificate as part of their
+\textbf{Step 6:} the PUBLISHER can then use the certificate as part of their
 supplementary records for a paper, and give credit to the codechecker
@@
   \centering
-      \includegraphics{/home/stephen/papers/talks/talkfigs/openscience/codecheck_overview.pdf}
+      \includegraphics{figs/codecheck_overview.pdf}
   \caption{The detective. Could we number the arrows as steps?
-(Implementing the CODECHECK process).  DANIEL: best source for this figure?}
+(Implementing the CODECHECK process).  DANIEL: could you check whether
+the icons are reusable.}
   \label{fig:worfklow}
@@
 
+DANIEL: COULD YOU DO THIS?
 This example workflow leaves sufficient room for different
@@
 been acknowledged/cited in the paper (LANCET / LSHTM).
+Table~\ref{tab:register} lists the certificates completed to date,
+along with citations to the original papers.
 
@@
 
+\begin{table}
+  \centering
+
+  \begin{tabular}{lllp{9cm}}  \toprule
+    Certificate & Ref & Research area & Description \\ \midrule
+    2020-001 & \cite{cert-2020-001} & Machine learning & CODECHECK undertaken post acceptance of manuscript and before its publication in \textit{Gigascience} \cite{Piccolo2020-lo}.
+    \\ 
+    2020-002 & \cite{cert-2020-002} & neuroscience &
+    Paper from 1997 showing unsupervised learning from natural images \cite{Hancock1992-mp}.
+    \\
+    2020-003 & \cite{cert-2020-003} & neuroscience &
+    Classic paper from 1982 on  models of associative memory \cite{Hopfield1982-mz}. \\
+    2020-004 & \cite{cert-2020-004} & area & barto \\
+    2020-005 & \cite{cert-2020-005} & area & Larisch \\
+    2020-006 & \cite{cert-2020-006} & area & Detorakis \\
+    2020-007 & \cite{cert-2020-007} & area & in progress \\
+    2020-008 & \cite{cert-2020-008} & Covid & covid-uk Davies \\
+    2020-009 & \cite{cert-2020-009} & Covid &  tracing \\
+    2020-010 & \cite{cert-2020-010} & Covid &  report 9 \\
+    2020-011 & \cite{cert-2020-011} & Covid &  nature europe \\
+    2020-012 & \cite{cert-2020-012} & Covid &  USA Covid modelling \cite{Unwin2020}. \\
+    \\ \bottomrule
+  \end{tabular}
+  \caption{Register of completed certificates.  An interactive version
+  of this is available at \url{http://codecheck.org.uk/register}.}
+  \label{tab:register}
+\end{table}
+
 \subsection*{Annotated certificate}\label{annotated-certificate}
 
-What can you expect to find a certificate? Draw an annotated example.
+At the completion of each CODECHECK, a certificate is written that
+states which figures and tables from the original article could be
+reproduced by the CODECHECKER.  This certificate is made freely
+available so that all readers of the article can see which elements
+have been reproduced.  The certificate also includes links to the code
+and data used by the codechecker, allowing others to build on the
+work.  The format of the certificates has evolved slightly during the
+project, as we have learnt to automate different aspects of the
+certification.
 
-Should we annotate a certificate, at least the cover page, to
-demonstrate what we mean by a certificate? We could print a four page
-certificate 4-up so that it fits onto one page. Text might be too small
-though to read? A 2-up, printed sideways would work well though.
+To demonstrate the typical contents of a certificate, in
+Figure~\ref{fig:cert} we have annotated the first four (out of ten)
+pages of the certificate 2020-012 \cite{cert-2020-012}.  The reference
+article \cite{Unwin2020} describes the modelling COVID-19 across the
+USA.
+%% TODO - has this been published yet?.  No, not yet... in review...
+Figure~\ref{fig:cert}A shows the certificate number, and its DOI,
+which points to the certificate as archived on Zenodo.  Our logo is
+attached to the certificate to denote successful reproduction.
+Figure~\ref{fig:cert}B provides the key metadata about the check --
+which paper was checked, and who/when it was checked.  Materials
+relating to the codecheck are freely available via the Repository
+link.  Figure~\ref{fig:cert}C is a summary of how the codecheck was
+performed, and any interesting findings.
 
-\subsubsection*{Tools}\label{tools}
+Figure~\ref{fig:cert}D (page 2 of the certificate) shows a table that
+lists the outputs that were generated from the codecheck (this is
+termed the MANIFEST.)  For each output, there is a comment stating
+which figure/table it should be compared to in the original paper,
+along with the size of the file.  On page 3 of the certificate,
+Figure~\ref{fig:cert}E shows some more detailed notes from the
+codechecker, in this case indicating what steps were needed to
+initialize the environment and run the code.  This particular example
+was quite long, taking about 17 hours to run.  Finally, on page 4 of
+the certificate, we see the first output that was generated by the
+codecheck (Figure~\ref{fig:cert}F).  In this case, the figure matches
+figure 4 of REF.  Subsequent pages of the certificate show other
+outputs (not shown here).
 
-To run the system, we rely on github to host our work, and use a custom
-R pacakge, codecheck, to help with the generation of certificates and
-depositing work on zenodo. (Cite zen4R).
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{figs/annotate-cert-crop.pdf}
+  \caption{Annotated certificate 2020-012 (first four pages only).}
+  \label{fig:cert}
+\end{figure}
 
-You can do all with open infrastructure (e.g.~automated rendering of
+\subsubsection*{Tools and resources}\label{tools}
+
+To run the system, we rely on github to host our work, and use a
+custom R pacakge, codecheck, to help with the generation of
+certificates and depositing work on zenodo. (Cite zen4R).  You can do
+everything with open infrastructure (e.g.~automated rendering of
 register). but you can use your own tools if you prefer.
@@
 
+\textbf{Checking multiple times?}  If you check a preprint before peer
+review, it might need checking again as it is modified during peer
+review.  Unavoidable, and as happened to us (2020-012 has been
+revised).  But checking the manuscript the second time is likely to be
+much less work than the first time.
+
 As a principle, we do not require results to be exactly the same for a
 codecheck to pass, simply that the code runs and generates the output
-files that the author claims. Stochastic simulations mean that often we
-cannot get exactly the same, or even different versions of libraries.
-(Cite freesurfer example that generates different outputs on different
-operating systems).
+files that the author claims. Stochastic simulations mean that often
+we cannot get exactly the same, or even different versions of
+libraries.  (Cite freesurfer example \cite{Gronenschild2012-pp} that
+generates different outputs on different operating systems).
 
 \begin{itemize}
-\item
-  modelDB principle: minimum of one figure should be reproducibe.
-  Exampler certificate 2020-016 (I think) which had only a small subset
+\item modelDB principle: minimum of one figure should be reproducibe.
+  Example certificate 2020-016 (I think) which had only a small subset
   of the possible figures. Still useful and worth sharing.

### 2020-09-17 codecheck-paper.tex
working through the register, adding bibliographies...

@@
 
-DANIEL: COULD YOU DO THIS?
+DANIEL: COULD YOU DO THIS SECTION ON DIMENSIONS?
 This example workflow leaves sufficient room for different
@@
 (possibly anonymously) through a publisher would slow things down and
-reduce the chances for community bulding.
+reduce the chances for community building.
 
@@
 
-We would like to briefly highlight three examples here.
+The bulk of the certificates fall into three themes: (1) `classic'
+papers from computational neuroscience (2) Covid modelling preprints
+(3) AGILE.  We would like to briefly highlight three examples here.
 
@@
 \item
-  report 9 -- covid model of UK lockdown model.
+  report 9 -- covid model of UK lockdown model -- unlike many claims
+  in the popular media at the time, the model was reproducible
+  [NOORDEN news piece]. 
 \item
-  Geoscience (add 6 for Agile).
+  Manuel's certificate -- approached us as he requested a certificate
+  as paper was entering peer review.
 \end{enumerate}
+
+(DANIEL: say something about the Agile bundle, and the origins?)
 
@@
 
-  \begin{tabular}{lllp{9cm}}  \toprule
-    Certificate & Ref & Research area & Description \\ \midrule
-    2020-001 & \cite{cert-2020-001} & Machine learning & CODECHECK undertaken post acceptance of manuscript and before its publication in \textit{Gigascience} \cite{Piccolo2020-lo}.
+  \begin{tabular}{lllp{9cm}} \toprule Certificate & Ref & Research
+    area & Description \\ \midrule 2020-001 & \cite{cert-2020-001} &
+    Machine learning & CODECHECK undertaken post acceptance of
+    manuscript and before its publication in \textit{Gigascience}
+    \cite{Piccolo2020-lo}.
     \\
-    2020-002 & \cite{cert-2020-002} & neuroscience &
-    Paper from 1997 showing unsupervised learning from natural images \cite{Hancock1992-mp}.
+    2020-002 & \cite{cert-2020-002} & neuroscience & Paper from 1997
+                                                     showing
+                                                     unsupervised
+                                                     learning from
+                                                     natural images
+                                                     \cite{Hancock1992-mp}.
     \\
     2020-003 & \cite{cert-2020-003} & neuroscience &
-    Classic paper from 1982 on  models of associative memory \cite{Hopfield1982-mz}. \\
-    2020-004 & \cite{cert-2020-004} & area & barto \\
+                                                     Classic paper
+                                                     from 1982 on
+                                                     models of
+                                                     associative
+                                                     memory
+                                                     \cite{Hopfield1982-mz}.
+    (Simulation code was written specifically for this project.)\\
+    2020-004 & \cite{cert-2020-004} & area & Barto \\
     2020-005 & \cite{cert-2020-005} & area & Larisch \\
@@
     2020-007 & \cite{cert-2020-007} & area & in progress \\
-    2020-008 & \cite{cert-2020-008} & Covid & covid-uk Davies \\
+    2020-008 & \cite{cert-2020-008} & Covid & Modelling of
+    interventions on Covid  cases in the UK; checked as preprint (NOW REMOVED XXX)
+    and now published \cite{Davies2020-vj} \\
     2020-009 & \cite{cert-2020-009} & Covid &  tracing \\
     2020-010 & \cite{cert-2020-010} & Covid &  report 9 \\
-    2020-011 & \cite{cert-2020-011} & Covid &  nature europe \\
-    2020-012 & \cite{cert-2020-012} & Covid &  USA Covid modelling \cite{Unwin2020}. \\
+    2020-011 & \cite{cert-2020-011} & Covid & Modelling of Covid
+    spread across Europe;  paper was checked when in press (now  published) \cite{Flaxman2020-yb}. \\
+    2020-012 & \cite{cert-2020-012} & Covid &   Modelling of Covid spread across the USA; preprint currently review \cite{Unwin2020}. \\
     \\ \bottomrule
@@
   \caption{Register of completed certificates.  An interactive version
-  of this is available at \url{http://codecheck.org.uk/register}.}
+  of this is available at \url{http://codecheck.org.uk/register}.  SJE
+to complete...}
   \label{tab:register}
@@
 \item
-  How to embed this in juornals.
+  How to embed this in journals.
 \item

### 2020-09-17 codecheck-paper.tex
okay, no more citation hunting tonight...

@@
 
-  \begin{tabular}{lllp{9cm}} \toprule Certificate & Ref & Research
-    area & Description \\ \midrule 2020-001 & \cite{cert-2020-001} &
+  \begin{tabular}{llp{11cm}}
+    \toprule
+    Certificate & Research area & Description \\ \midrule
+    2020-001 \cite{cert-2020-001} &
     Machine learning & CODECHECK undertaken post acceptance of
@@
     \\
-    2020-002 & \cite{cert-2020-002} & neuroscience & Paper from 1997
-                                                     showing
-                                                     unsupervised
-                                                     learning from
-                                                     natural images
-                                                     \cite{Hancock1992-mp}.
+    2020-002  \cite{cert-2020-002} & neuroscience & Paper from 1997
+ showing unsupervised learning from natural images \cite{Hancock1992-mp}.
+ Code written for this project.
     \\
-    2020-003 & \cite{cert-2020-003} & neuroscience &
+    2020-003  \cite{cert-2020-003} & neuroscience &
                                                      Classic paper
@@
     (Simulation code was written specifically for this project.)\\
-    2020-004 & \cite{cert-2020-004} & area & Barto \\
-    2020-005 & \cite{cert-2020-005} & area & Larisch \\
-    2020-006 & \cite{cert-2020-006} & area & Detorakis \\
-    2020-007 & \cite{cert-2020-007} & area & in progress \\
-    2020-008 & \cite{cert-2020-008} & Covid & Modelling of
+    2020-004  \cite{cert-2020-004} & area & Cart-pole balancing problem solved using reinforcement learning \cite{Barto1983-rg}.  Code written for this reimplementation.\\
+    2020-005  \cite{cert-2020-005} & area & Larisch \\
+    2020-006  \cite{cert-2020-006} & area & Detorakis \\
+    2020-007  \cite{cert-2020-007} & area & in progress \\
+    2020-008  \cite{cert-2020-008} & Covid & Modelling of
     interventions on Covid  cases in the UK; checked as preprint (NOW REMOVED XXX)
     and now published \cite{Davies2020-vj} \\
-    2020-009 & \cite{cert-2020-009} & Covid &  tracing \\
-    2020-010 & \cite{cert-2020-010} & Covid &  report 9 \\
-    2020-011 & \cite{cert-2020-011} & Covid & Modelling of Covid
+    2020-009  \cite{cert-2020-009} & Covid &  tracing \\
+    2020-010  \cite{cert-2020-010} & Covid &  report 9 \\
+    2020-011  \cite{cert-2020-011} & Covid & Modelling of Covid
     spread across Europe;  paper was checked when in press (now  published) \cite{Flaxman2020-yb}. \\
-    2020-012 & \cite{cert-2020-012} & Covid &   Modelling of Covid spread across the USA; preprint currently review \cite{Unwin2020}. \\
+    2020-012  \cite{cert-2020-012} & Covid &   Modelling of Covid spread across the USA; preprint currently review \cite{Unwin2020}. \\
     \\ \bottomrule

### 2020-09-28 codecheck-paper.tex
go through introduction section

@@
 \item Check with Giuliano re: ``code works''. Write to Naomi.
+\item Use "execution" insteand of "sharing" in title?
+\item Can we introduce "research compendium" as a term?
 \end{itemize}
@@
 or analyse their data. These computations are increasingly complex, and
-difficult to explain coherently in a paper. (MARWICK). To complement the
+difficult to explain coherently in a paper \citep{marwick_how_2015}. To complement the
 traditional route of sharing research by writing papers, there is a
@@
 datasets, so that others can inspect, reproduce or expand that work
-(Figure 1). Some of the earliest proponents were Buckheit and Donoho
-(1995) who commented ``advert is scholarship for the paper''. (get
-correct quote).
+(Figure 1). Some of the earliest proponents were Buckheit and Donoho \cite{buckheit_wavelab_1995} who coined what has been called 
+\emph{Claerbout's claim} (extending on \citet{claerbout_electronic_1992}):
+\emph{"An article about computational science in a scientific publication 
+is \text{not} the scholarship itself, it is merely \textbf{advertising} of
+the scholarship. The actual scholarship is the complete software development
+environment and the complete set of instructions which generated the 
+figures."}
+%  Donoho (2010) has another paraphrasing of the same quote, which might work better:
+% "an article about computational result is advertising, not scholarship. The actual scholarship is the full software environment, code and data, that produced the result."
 
@@
 journal. At one of the spectrum, there are specialist journals that
-have been created to accept ``data papers'' (e.g.~\emph{Scientific
-  Data}, \textit{Geoscience Data Journal}\ldots{} NAME COUPLE OF
-OTHERS). These journals have established rigorous procedures by which
+have been created to accept ``data papers'' (e.g.~\href{}{\emph{Scientific
+  Data}}, 
+\href{https://essd.copernicus.org/}{\emph{Earth System Science Data}}, 
+\href{https://rmets.onlinelibrary.wiley.com/journal/20496060}{\emph{Geoscience Data Journal}},
+\href{https://bdj.pensoft.net/}{\emph{Biodiversity Data Journal}},
+\href{https://openpsychologydata.metajnl.com/}{\emph{Journal of Open Psychology Data}},
+\href{https://odjar.org/}{\emph{Open Data Journal for Agricultural Research}},
+\href{https://openhealthdata.metajnl.com}{\emph{Journal of Open Health Data}}).
+These journals have established rigorous procedures by which
 the data is validated according to standards in each field. At the
@@
 statement ``Data available upon reasonable request'' whereby authors,
-well-intentioned, often cannot provide the data when readers ask for
+while possible well-intentioned at the time of writing the article,
+often cannot provide the data when readers ask for
 it.  (ANDREW H: any references to sum up poor/evolving state of data
@@
 from, even when it doesn't run or if there is no documentation. However,
-with a little effort, we believe that if someone independent can run/use
+with a little effort, we believe that if someone independent can run/execute
 the code, this is worth documenting as early as possible. With this in
@@
   research paper, a static PDF document.  Working backwards from the
-  PDF to the underlying materials is impossible.  By sharing the
+  PDF to the underlying materials is impossible. This probihibits reuse
+  and is not only non-transparent for a specific paper, but also 
+  ineffective for science as a whole. By sharing the
   materials on the left, others outside the lab are able to reproduce
-  or build on this work.  (Should we draw a backward arrow showing the
-  impossibility of this backward flow?)}
+  or build on this work.}
   \label{fig:inverse}

### 2020-09-28 codecheck-paper.tex
add remaining checks to the table

@@
     2020-012  \cite{cert-2020-012} & Covid &   Modelling of Covid spread across the USA; preprint currently review \cite{Unwin2020}. \\
+    2020-013  \cite{cert-2020-013} & area & biorXiv \cite{Spitschan2020.06.02.129502}. \\
+    2020-014  \cite{cert-2020-014} & area & eLife \cite{Sadeh2020}. \\
+    2020-015  \cite{cert-2020-015} & area & eLife \cite{Liou2020}. \\
+    2020-016  \cite{cert-2020-016} & GIScience & 
+    R models demonstrating the Modifiable Aral Unit Problem (MAUP) in spatial data science \cite{Brunsdon2020}. \\
+    2020-017  \cite{cert-2020-017} & GIScience & 
+    Spatial data handling, analysis, and visualisation using a variety of R packages \cite{Bivand2020}. \\
+    2020-018  \cite{cert-2020-018} & GIScience & 
+    Cellular automaton in R for modeling dynamic phenomena within a Discrete Global Grid System (DGGS) data model; using a subset for demonstration \cite{Hojati2020}. \\
+    2020-019  \cite{cert-2020-019} & GIScience & Reachability analysis of suburban transportation using analysis of privately owned shared cars; workflow uses Python, and reproduction with a considerably subsampled dataset \cite{Illium2020}.
+\\
+    2020-020  \cite{cert-2020-020} & GIScience & In-database windows operators for processing spatio-temporal data streams based on Python and PostgreSQL/PostGIS \cite{Werner2020}; reproduction using container.
+\\
+    2020-021  \cite{cert-2020-021} & GIScience & Python code for comparing supervised machine learning models for spatial nominal entity recognition. \cite{Medad2020}. \\
+    2020-022  \cite{cert-2020-022} & GIScience & Python code for visualising text analysis on intents and concepts from geo-analytic questions \cite{Xu2020}. \\
+    2020-023  \cite{cert-2020-023} & GIScience & Analysis of spatial footprints of geo-tagged extreme weather events from social media using R \cite{Owuor2020}. \\
     \\ \bottomrule
   \end{tabular}
-  \caption{Register of completed certificates.  An interactive version
+  \caption{Register of completed certificates as of October 2020.  An interactive version
   of this is available at \url{http://codecheck.org.uk/register}.  SJE

### 2020-09-28 codecheck-paper.tex
edit description of steps, update the detective figure

@@
 we expand on the underlying principles. The workflow involves three
-groups of people (1) the AUTHOR providing the code to be checked, (2)
+groups of people: (1) the AUTHOR of a paper providing the code to be checked, (2)
 the PUBLISHER of a journal intereseted in publishing the paper by the
@@
 
-\textbf{Step 1:} the author submit their manuscript along with code and data to
-the publishers. The code and data need not be openly available at this
-point.
+\textbf{Step 1:} the AUTHOR submit their manuscript along with code and data to
+the PUBLISHER. The code and data need not be openly available at this point.
+However, in many cases the code and data may be published on a code hosting platform,
+such as GitHub/Lab.
 
-\textbf{Step 2:} the publisher finds a codechecker to check the code. This is
-analgous to the publisher finding one or more reviewers to evaluating
-the code. However a cruical difference is that we suggest the codechcker
-and the author can talk freely to each directly (see next).
+\textbf{Step 2:} the publisher finds a CODECHECKER to check the code. This is
+analogous to the publisher finding one or more reviewers to evaluating
+the paper. However, a cruical difference is that we suggest the CODECHECKER
+and the author can talk freely to each other directly (see next).
 
-\textbf{Step 3:} the codechecker runs the code, based on instructions provided by
+\textbf{Step 3:} the CODECHECKER runs the code, based on instructions provided by
 the author, and checks if some/all of the results from the paper can be
 reproduced. (See later for exact nature of ``reproduced''). If there are
-any problems running the code, the codechecker asks the author for help
-to resolve the problems. The codechecker then tries to run the code
-again. This process iterates until either the codechecker is successful,
-or the codechecker concludes the code does not (\emph{what do in this
-case}). As part of this process, the codechecker could work entirely
-locally on their own compute resource, or in the cloud, e.g.~using
-mybinder infrastructure. Using such cloud-based infrastructure allows
-for the codechecker and author to collaboratively run the code together.
+any problems running the code, the CODECHECKER asks the author for help
+to resolve the problems. The CODECHECKER then tries to run the code
+again. This process iterates until either the CODECHECKER is successful,
+or the CODECHECKER concludes the code does not work (\emph{what do in this
+case}). As part of this process, the CODECHECKER could work entirely
+locally on their own compute resource, or in the cloud, e.g.~using the open
+MyBinder infrastructure \cite{jupyter_binder_2018}. Using such cloud-based 
+infrastructure allows for the CODECHECKER and author to collaboratively improve
+the code and enforces a much more complete definition of the computing environment,
+but, unless infrastructure is provided by, e.g., by the publisher, requires code
+and data to be published openly online.
 
-\textbf{Step 4:} the codechecker writes a certificate stating how the code was
+\textbf{Step 4:} the CODECHECKER writes a certificate stating how the code was
 run and includes a copy of outputs (figures, tables) that were
-independently generated.
+independently generated. The certificate may include further recommendations to
+the author how to improve the material.
 
-\textbf{Step 5:} the certificate and code/data get deposited on an open archive
-(currently Zenodo), and the certificate is given to the PUBLISHER.
+\textbf{Step 5:} the certificate as well as auxiliary files created during the check,
+e.g., a specifiction of a computing environment, helper scripts, and the code and data,
+if available under open licenses, get deposited in an open archive (currently Zenodo);
+the certificate is given to the PUBLISHER.
 
 \textbf{Step 6:} the PUBLISHER can then use the certificate as part of their
-supplementary records for a paper, and give credit to the codechecker
-for their work by depositing appropriate records (\emph{bit hazy here}).
+supplementary records for a paper or provide it to scientific reviewers; the
+PUBLISHER also gives credit to the CODECHECKER for their work by depositing the
+activity appropriately in scholarly publication databases, such as ORCID.
+The PULISHER also ensures proper publication metadata, e.g., links from the 
+certificate repository to the published paper or the original code repository.
 
 \begin{figure}
+% original figure source: https://docs.google.com/drawings/d/1tGOZFNZle-oE1Ynw_tLZFmNcv3wgxD0HPs5bXzDOrng/edit
+% emojis: https://github.com/googlefonts/noto-emoji and https://github.com/twitter/twemoji
+% emoji codes:
+% book stack: 1f4da
+% package: 1f4e6
+% file drawer: 1f5c4
+% disk: 1f4be
+% document: 1f4c4
+% phd: 1f393
+% detective: u1f575
+% laptop: 1f4bb
+% folder: 1f4c1 1f4c2
+% recept: 1f9fe
   \centering
       \includegraphics{figs/codecheck_overview.pdf}
-  \caption{The detective. Could we number the arrows as steps?
-(Implementing the CODECHECK process).  DANIEL: could you check whether
-the icons are reusable.}
+  \caption{The CODECHECK example process implementation. Codechecker act as a detectives:
+  The investigate and record, but do not fix issues.}
   \label{fig:worfklow}

### 2020-09-28 codecheck-paper.tex
edit text on principles

@@
 views on how a codecheck should be performed. They are not set in
-stone, but we do believe the following core principles underpinour
+stone, but we do believe the following core principles underpin our
 CODECHECK:
@@
 \textbf{2. Communication between humans is key.} Some code may just work
-without any interaction (e.g.~Manuel's certificate); however, more often
+without any interaction (e.g.~Manuel's certificate TODO add cert number); 
+however, more often
 there are hidden dependencies that need to be found, or revised software
@@
 (possibly anonymously) through a publisher would slow things down and
-reduce the chances for community building.
+reduce the chances for community building as well as the lessons learned
+for both author and codecheckers.
 
-\textbf{3. Credit is given to codecheckers.} (TODO Explain in a few
-sentences what each means, e.g.~see 1,2) Ideally publons, or similar,
-but can just be a citation to the publicly available DOI. (Community
-DOIs use these.)
+\textbf{3. Credit is given to codecheckers.} The value of contributing a
+CODECHECK is comparable to providing a scientific peer review. Indiviual
+variations aside, it may also require a similar amount of time. Therefore,
+the codechecker's activity should be publicly recorded and mentioned in the
+published paper.
+The public record can be realised by publishing the certificate in a 
+citable form (i.e., with a DOI; this is done within the community codecheck
+process), by listing codecheckers on the journal's website or, ideally, by
+publishing the checks alongside peer review activities in public databases,
+such as ORCID \footnote{\href{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}}
+or Publons.
 
-\textbf{4. Workflows must be auditable.} (TODO Explain in a few
-sentences what each means, e.g.~see 1,2)
+\textbf{4. Workflows must be auditable.} This requires that the codechecker 
+has at least enough material to validate the workflow outputs submitted by 
+the authors. Tode could be executed at least once without critical errors or
+warnings using the provided instructions, and, therefore, all code and data 
+was given and could be investigated more deeply or extended in the future.
+Ideally, this is a "one~click" step, but achieving this requires particular 
+skills. It is also hard to create a sufficient level of documentation for 
+third parties. That is why the possibility to communicate is more important
+than perfected workflows, and why a CODECHECK is not automated.
+A codechecker does \emph{not} conduct a code review.
+% more content from the website, may fit better elsewhere:
+% The CODECHECK is not automated on purpose: automation may (a) lead to people gaming the system, (b) hide details that eventually decrease level of certainty that a codechecker has in their assessment, and (c) reduce the understandability of instructions in the long term, which is more important than short term ease of use (see Principle 2).
 
 \textbf{5. Open by default.}  By default, unless there are strong
-reasons to the contrary (e.g. clinically-sensitive data), all code and
-data, both from author and codechecker, will be made freely available
-when the certificate is published, following community principles.
+reasons to the contrary (e.g., clinically-sensitive data), all code and
+data, both from author and codechecker, will be made freely available, i.e., 
+under open licenses, when the certificate is published. The code and data 
+publication should follow community good practices.
 

### 2020-09-28 codecheck-paper.tex
update grant info for Daniel

@@
 Sustainability Institute and a Mozilla Science mini grant.
+DN is supported by grant \href{https://gepris.dfg.de/gepris/projekt/415851837}{PE~1632/17-1} from the German Research Foundation (DFG).
 
@@
 Maciocci (eLife), Rebecca Kirk (PLOS Computational Biology), Scott
-Edmunds (Gigascience) Andrew Hufton (Scientific Data). Iain Davies and
+Edmunds (GigaScience), and Andrew Hufton (Scientific Data). Iain Davies and
 Yuhao (Sebastian) Wang developed code and example certificates.

### 2020-09-28 codecheck-paper.tex
describe the "themes" of checks

@@
 The bulk of the certificates fall into three themes: (1) `classic'
-papers from computational neuroscience (2) Covid modelling preprints
-(3) AGILE.  We would like to briefly highlight three examples here.
+papers from computational neuroscience (2) Covid modelling preprints, and
+(3) GIScience.  
+The first theme was the initial set of papers used to explore the concept
+of CODECHECK. The idea was to take very well known articles from a specific
+domain, in this case the work area of co-author S.~Eglen, possibly as as a
+conversation starter for potential collaborating journals.
+The second theme is rather opportunistic. As the COVID-19 pandemic swept
+over the whole Earth, we hoped to contribute to science's answer to the
+unknown virus. The checks were solicited through community interaction.
+The third theme represents co-author D.~N\"ust's work area.
+Independent of CODECHECK, the \emph{Reproducible AGILE} Initiative \cite{reproducible_agile}
+established a process for reproducing workflows at the AGILE conference
+series. % TODO add reference to MUNIN talk
+While using slightly different terms and infrastructue, e.g., `reproducibility
+reports' are published on OSF, instead of certificates on Zenodo, the 
+reproducibility reviews adhere to the codecheck principles.
+Furthermore, two checks were completed as part of a peer review for a
+GIScience journal.
+
+Out of these areas, we would like to briefly highlight three examples here.
 
@@
 \end{enumerate}
-
-(DANIEL: say something about the Agile bundle, and the origins?)
 

### 2020-09-30 codecheck-paper.tex
add tools and resources description

@@
 
-To run the system, we rely on github to host our work, and use a
-custom R pacakge, codecheck, to help with the generation of
-certificates and depositing work on zenodo. (Cite zen4R).  You can do
-everything with open infrastructure (e.g.~automated rendering of
-register). but you can use your own tools if you prefer.
+To run the system, we rely on freely available infrastructure, more 
+specifically GitHub and Zenodo.
+
+The \texttt{codecheckers} GitHub organisation\footnote{
+\url{https://github.com/codecheckers}} contains projects managing the
+project website, the community of codecheckers, community discussions,
+forks of codechecked code repositories, and the main register for
+managing the collection of checks. Both the website\footnote{
+\url{https://codecheck.org.uk/}} and the register
+\footnote{\url{https://codecheck.org.uk/register}} are hosted as GitHub
+pages. The register is effectively a single table in CSV format that
+connects the certificate identifier with the repository associated with
+a specific conducted codecheck. Each of these repositories, which 
+currently can be hosted on GitHub or Open Science Framework (OSF), 
+contains the CODECHECK metadata file \texttt{codecheck.yml}. The register
+further contains a column for the type of check, e.g., community, journal,
+or conference, and the respective GitHub issue where communications and 
+assignments around a specific check are organised. No information is 
+duplicated between the register and the metadata files. When a new
+certificate is entered into the register, an automated process using 
+continuous integration infrastructure of GitHub, GitHub Actions, retrieves
+the metadata files and builds different representations of the register
+for viewing (HTML) or integration with other services (CSV, JSON).
+
+Zenodo is an open repository for scientific data and documents. It
+mints DOIs for deposits and ensures long-term availability of all digital
+artefacts related to the project. These include the community 
+certificates as well as the register, which is regularly archived to 
+Zenodo \cite{codecheck_register} (including all file formats).
+
+A custom R package, \texttt{codecheck}\footnote{
+\url{https://github.com/codecheckers/codecheck}} helps with the generation
+of certificates and depositing certificates and reated files on Zenodo 
+using the R package \texttt{zen4R} \cite{zen4r}.
+Codecheckers may choose to not use the package and rely on their own tools
+for creating the certificate. This is a deliberate flexibility to
+accomodate both different skill sets and unforseen technical advances or
+challenges.
+
+These tools and resources demonstrate that a CODECHECK process can be
+managed without infrastructure costs on freely available platforms.
+The scripted and automated parts should also ensure that more checks can
+be conducted in the future.
+The remaining costs are the project's domain, which is negligible, and
+the human resources for managing the project and processes, which are
+considerable and currently soley based on, partly grant-based, public funding.
+Naturally, similar set-ups are possible, such as using GitLab/GitLab.com 
+and the GitLab CI instead of GitHub and GitHub Actions, or using a different 
+data repository for depositing certificates and data.
 

### 2020-10-01 codecheck-paper.tex
use only DOI of references, make DOIs clickable in PDF
update description of register and certificate

@@
 \usepackage{booktabs}
+\usepackage{doi} % turn DOIs into links
 
@@
 \section*{Register}\label{register}
+% count certificates: 
+% curl https://raw.githubusercontent.com/codecheckers/register/master/register.csv | wc -l
+% minus 1 for Hathway-Goodman (no empty last line, so no -1 for the headline needed)
 
-We have a curated list of 15 certificates available at
+We have a curated list of 22 certificates available at
 \url{https://codecheck.org.uk/register}. These are a mixture of
-reproductions of papers for a variety of reasons (do we name them all?)
-from interesting historical papers, to papers codechecked as part of
-peer review, or because of public interest (covid). several covid models
-have been done during the time of peer review of these papers, under our
-initative rather than requested from journal, but certificate have then
-been acknowledged/cited in the paper (LANCET / LSHTM).
-Table~\ref{tab:register} lists the certificates completed to date,
-along with citations to the original papers.
+reproductions of papers for a variety of reasons and fall into three themes:
+(1) `classic' papers from computational neuroscience
+(2) COVID-19 modelling preprints, and
+(3) GIScience.  
+%from interesting historical papers, to papers codechecked as part of
+%peer review, or because of public interest (COVID-19).
 
-The bulk of the certificates fall into three themes: (1) `classic'
-papers from computational neuroscience (2) Covid modelling preprints, and
-(3) GIScience.  
 The first theme was the initial set of papers used to explore the concept
-of CODECHECK. The idea was to take very well known articles from a specific
-domain, in this case the work area of co-author S.~Eglen, possibly as as a
-conversation starter for potential collaborating journals.
+of CODECHECK. The idea was to take very well known historical articles from a 
+specific domain, in this case the work area of co-author S.~Eglen, possibly 
+as as a conversation starter for potential collaborating journals.
 The second theme is rather opportunistic. As the COVID-19 pandemic swept
-over the whole Earth, we hoped to contribute to science's answer to the
-unknown virus. The checks were solicited through community interaction.
+over the whole Earth, we hoped to contribute a small part to science's answer
+to the unknown virus. The checks were solicited through community interaction
+%Several COVID-19 models have been done during the time of peer review of these papers,
+and under our initative rather than requested from journal, but the certificate
+has then been acknowledged/cited in the paper (LANCET/LSHTM). % TODO SJE: clarify references in brackets
 The third theme represents co-author D.~N\"ust's work area.
@@
 
+Table~\ref{tab:register} lists the certificates completed to date,
+along with citations to the original papers.
 Out of these areas, we would like to briefly highlight three examples here.
@@
     manuscript and before its publication in \textit{Gigascience}
-    \cite{Piccolo2020-lo}.
-    \\
+    \cite{Piccolo2020-lo}. \\
     2020-002  \cite{cert-2020-002} & neuroscience & Paper from 1997
  showing unsupervised learning from natural images \cite{Hancock1992-mp}.
- Code written for this project.
-    \\
-    2020-003  \cite{cert-2020-003} & neuroscience &
-                                                     Classic paper
+ Code written for this project. \\
+    2020-003  \cite{cert-2020-003} & neuroscience &  Classic paper
                                                      from 1982 on
@@
     2020-007  \cite{cert-2020-007} & area & in progress \\
-    2020-008  \cite{cert-2020-008} & Covid & Modelling of
-    interventions on Covid  cases in the UK; checked as preprint (NOW REMOVED XXX)
+    2020-008  \cite{cert-2020-008} & COVID-19 & Modelling of
+    interventions on COVID-19 cases in the UK; checked as preprint (NOW REMOVED XXX)
     and now published \cite{Davies2020-vj} \\
-    2020-009  \cite{cert-2020-009} & Covid &  tracing \\
-    2020-010  \cite{cert-2020-010} & Covid &  report 9 \\
-    2020-011  \cite{cert-2020-011} & Covid & Modelling of Covid
+    2020-009  \cite{cert-2020-009} & COVID-19 & tracing \\
+    2020-010  \cite{cert-2020-010} & COVID-19 & report 9 \\
+    2020-011  \cite{cert-2020-011} & COVID-19 & Modelling of Covid
     spread across Europe;  paper was checked when in press (now  published) \cite{Flaxman2020-yb}. \\
-    2020-012  \cite{cert-2020-012} & Covid &   Modelling of Covid spread across the USA; preprint currently review \cite{Unwin2020}. \\
+    2020-012  \cite{cert-2020-012} & COVID-19 & Modelling of COVID-19 spread across the USA; preprint currently review \cite{Unwin2020}. \\
     2020-013  \cite{cert-2020-013} & area & biorXiv \cite{Spitschan2020.06.02.129502}. \\
@@
   \caption{Register of completed certificates as of October 2020.  An interactive version
-  of this is available at \url{http://codecheck.org.uk/register}.  SJE
-to complete...}
+  of this is available at \url{http://codecheck.org.uk/register}. % TODO: SJE to complete...
+  }
   \label{tab:register}
@@
 At the completion of each CODECHECK, a certificate is written that
-states which figures and tables from the original article could be
-reproduced by the CODECHECKER.  This certificate is made freely
+states which outputs, i.e., numbers, figures and tables, from the 
+original article could be reproduced by the CODECHECKER.
+This certificate is made openly
 available so that all readers of the article can see which elements
@@
 and data used by the codechecker, allowing others to build on the
-work.  The format of the certificates has evolved slightly during the
+work.
+The format of the certificates has evolved slightly during the
 project, as we have learnt to automate different aspects of the
@@
 To demonstrate the typical contents of a certificate, in
-Figure~\ref{fig:cert} we have annotated the first four (out of ten)
-pages of the certificate 2020-012 \cite{cert-2020-012}.  The reference
-article \cite{Unwin2020} describes the modelling COVID-19 across the
-USA.
+Figure~\ref{fig:cert} we have annotated the first four out of ten
+pages of the certificate \texttt{2020-012} \cite{cert-2020-012}.
+The referenced article \cite{Unwin2020} describes the modelling COVID-19
+across the USA.
 %% TODO - has this been published yet?.  No, not yet... in review...
 Figure~\ref{fig:cert}A shows the certificate number, and its DOI,
-which points to the certificate as archived on Zenodo.  Our logo is
+which points to the certificate as archived on Zenodo.
+The CODECHECK logo is
 attached to the certificate to denote successful reproduction.
 Figure~\ref{fig:cert}B provides the key metadata about the check --
-which paper was checked, and who/when it was checked.  Materials
+which paper was checked, who the authors and the codechecker were,
+and when it was checked. Materials
 relating to the codecheck are freely available via the Repository
-link.  Figure~\ref{fig:cert}C is a summary of how the codecheck was
+link. This table is generated from machine readable text file, the
+CODECHECK configuration file \texttt{codecheck.yml}.
+The configuration file collects the core pieces of metadata surrounding
+a check and enables current and future automation of workflows and 
+meta-analyses.
+Figure~\ref{fig:cert}C is a textual summary of how the codecheck was
 performed, and any interesting findings.

### 2020-10-01 codecheck-paper.tex
add link to repo and add overall TODO
minor updates in introduction, scale Figure 1 to use less than full page

@@
 \affil[$\ast$]{Contributed equally}
+\affil[$$]{Public project for this article is at \url{https://github.com/codecheckers/paper}}
 \maketitle
@@
 \item Can we introduce "research compendium" as a term?
+\item Should we make code and data sharing a prerequisite or leave that door open?
 \end{itemize}
@@
 or analyse their data. These computations are increasingly complex, and
-difficult to explain coherently in a paper \citep{marwick_how_2015}. To complement the
-traditional route of sharing research by writing papers, there is a
-growing trend/demand to share the underlying artifacts, notably code and
-datasets, so that others can inspect, reproduce or expand that work
-(Figure 1). Some of the earliest proponents were Buckheit and Donoho \cite{buckheit_wavelab_1995} who coined what has been called 
+difficult to explain coherently in a paper \citep{marwick_how_2015}.
+To complement the traditional route of sharing research by writing papers,
+there is a growing trend/demand to share the underlying artifacts, notably 
+code and datasets, so that others can inspect, reproduce or expand that work
+(see Figure~\ref{fig:inverse}).
+Some of the earliest proponents were Buckheit and Donoho
+\cite{buckheit_wavelab_1995} who coined what has been called 
 \emph{Claerbout's claim} (extending on \citet{claerbout_electronic_1992}):
@@
 most scientific journals now require a data sharing statement that
-outlines what data the authors have (or will) share. The
-implementation status of this varies considerably according to the
+outlines what data the authors have (or will) share.
+The implementation status of this varies considerably according to the
 journal. At one of the spectrum, there are specialist journals that
@@
   \centering
-  \includegraphics[width=\textwidth]{figs/rr.pdf}
+  \includegraphics[width=0.7\textwidth]{figs/rr.pdf}
   \caption{The inverse problem in reproducible research.  The left

### 2020-10-01 codecheck-paper.tex
move overview figure up and add dimensions figure
add text for variations/dimensions

@@
 AUTHOR and (3) the CODECHECKER who checks that the AUTHOR's code works.
-The workflow that we have refined is documented in Figure 2.
-
-\textbf{Step 1:} the AUTHOR submit their manuscript along with code and data to
-the PUBLISHER. The code and data need not be openly available at this point.
-However, in many cases the code and data may be published on a code hosting platform,
-such as GitHub/Lab.
-
-\textbf{Step 2:} the publisher finds a CODECHECKER to check the code. This is
-analogous to the publisher finding one or more reviewers to evaluating
-the paper. However, a cruical difference is that we suggest the CODECHECKER
-and the author can talk freely to each other directly (see next).
-
-\textbf{Step 3:} the CODECHECKER runs the code, based on instructions provided by
-the author, and checks if some/all of the results from the paper can be
-reproduced. (See later for exact nature of ``reproduced''). If there are
-any problems running the code, the CODECHECKER asks the author for help
-to resolve the problems. The CODECHECKER then tries to run the code
-again. This process iterates until either the CODECHECKER is successful,
-or the CODECHECKER concludes the code does not work (\emph{what do in this
-case}). As part of this process, the CODECHECKER could work entirely
-locally on their own compute resource, or in the cloud, e.g.~using the open
-MyBinder infrastructure \cite{jupyter_binder_2018}. Using such cloud-based 
-infrastructure allows for the CODECHECKER and author to collaboratively improve
-the code and enforces a much more complete definition of the computing environment,
-but, unless infrastructure is provided by, e.g., by the publisher, requires code
-and data to be published openly online.
-
-\textbf{Step 4:} the CODECHECKER writes a certificate stating how the code was
-run and includes a copy of outputs (figures, tables) that were
-independently generated. The certificate may include further recommendations to
-the author how to improve the material.
-
-\textbf{Step 5:} the certificate as well as auxiliary files created during the check,
-e.g., a specifiction of a computing environment, helper scripts, and the code and data,
-if available under open licenses, get deposited in an open archive (currently Zenodo);
-the certificate is given to the PUBLISHER.
-
-\textbf{Step 6:} the PUBLISHER can then use the certificate as part of their
-supplementary records for a paper or provide it to scientific reviewers; the
-PUBLISHER also gives credit to the CODECHECKER for their work by depositing the
-activity appropriately in scholarly publication databases, such as ORCID.
-The PULISHER also ensures proper publication metadata, e.g., links from the 
-certificate repository to the published paper or the original code repository.
+The workflow that we have refined is documented in Figure~\ref{fig:workflow}.
+
+% TODO: how is a CODECHECK different from a "specialist reviewer" ? Is it because it is not special, but regular?
 
@@
   \centering
-      \includegraphics{figs/codecheck_overview.pdf}
+      \includegraphics[width=\textwidth]{figs/codecheck_overview.pdf}
   \caption{The CODECHECK example process implementation. Codechecker act as a detectives:
@@
 
+\textbf{Step 1:} the AUTHOR submit their manuscript along with code and data to
+the PUBLISHER. The code and data need not be openly available at this point.
+However, in many cases the code and data may be published on a code hosting platform,
+such as GitHub/Lab.
+
+\textbf{Step 2:} the publisher finds a CODECHECKER to check the code. This is
+analogous to the publisher finding one or more reviewers to evaluating
+the paper. However, a cruical difference is that we suggest the CODECHECKER
+and the author can talk freely to each other directly (see next).
+
+\textbf{Step 3:} the CODECHECKER runs the code, based on instructions provided by
+the author, and checks if some/all of the results from the paper can be
+reproduced. (See later for exact nature of ``reproduced''). If there are
+any problems running the code, the CODECHECKER asks the author for help
+to resolve the problems. The CODECHECKER then tries to run the code
+again. This process iterates until either the CODECHECKER is successful,
+or the CODECHECKER concludes the code does not work (\emph{what do in this
+case}). As part of this process, the CODECHECKER could work entirely
+locally on their own compute resource, or in the cloud, e.g.~using the open
+MyBinder infrastructure \cite{jupyter_binder_2018}. Using such cloud-based 
+infrastructure allows for the CODECHECKER and author to collaboratively improve
+the code and enforces a much more complete definition of the computing environment,
+but, unless infrastructure is provided by, e.g., by the publisher, requires code
+and data to be published openly online.
+
+\textbf{Step 4:} the CODECHECKER writes a certificate stating how the code was
+run and includes a copy of outputs (figures, tables) that were
+independently generated. The certificate may include further recommendations to
+the author how to improve the material.
+
+\textbf{Step 5:} the certificate as well as auxiliary files created during the check,
+e.g., a specifiction of a computing environment, helper scripts, and the code and data,
+if available under open licenses, get deposited in an open archive (currently Zenodo);
+the certificate is given to the PUBLISHER.
+
+\textbf{Step 6:} the PUBLISHER can then use the certificate as part of their
+supplementary records for a paper or provide it to scientific reviewers; the
+PUBLISHER also gives credit to the CODECHECKER for their work by depositing the
+activity appropriately in scholarly publication databases, such as ORCID.
+The PULISHER also ensures proper publication metadata, e.g., links from the 
+certificate repository to the published paper or the original code repository.
+
 \subsection*{Variations}\label{variations}
 
-DANIEL: COULD YOU DO THIS SECTION ON DIMENSIONS?
-This example workflow leaves sufficient room for different
-implementations of this process. There are several aspects that can be
-considered:
-
-\subsubsection*{When to do a codecheck}\label{when-to-do-a-codecheck}
-
-preprint (no publisher available), pre-condition for peer review, during
-peer review, after acceptance. Each of these have their own pros/cons.
+\subsubsection*{Dimensions of CODECHECK processes}\label{dimensions-of-processes}
+
+The example workflow leaves sufficient room for different
+implementations of this process. To elaborate on these options, we consider
+the following dimensions in a space of possible CODECHECKs, as shown in 
+Figure~\ref{fig:dimensions}, to represent relevant aspects for the involved
+stakeholders. These aspects touch on timing, responsibilities, and 
+transparency and will be discussed in the following sections.
+Each of these have their own pros and cons and can for the most part be
+freely combined.
+
+\begin{figure}
+% original figure source: https://docs.google.com/drawings/d/1q2EdW3Ad-IpcD-CeoR1eIFMJW2fF6OVmDhCtCUmLO1k/edit
+  \centering
+      \includegraphics[width=0.8\textwidth]{figs/codecheck_dimensions.pdf}
+  \caption{The dimensions of implementing a CODECHECK process.}
+  \label{fig:dimensions}
+\end{figure}
+
+\subsubsection*{When to do a codecheck and with what power}
+\label{when-to-do-a-codecheck}
+
+The time for a CODECHECK greatly depends on the intentions and the ascribed
+importance. The earlier a CODECHECK happens in the process, the more the
+results can play a role in the most important decision in scholarly 
+communication: is a paper published, sent back for revisions, or rejected.
+The later in a review process the check happens, the less challenging is it
+to allow a bidirectional communication between author and codechecker, e.g.,
+because the author might already be notified of the acceptance and may more
+freely share materials online.
+Even earlier checks, i.e., a CODECHECK of a preprint, may even help to 
+improve the workflow itself even before a publisher is involved. The 
+possibility for codechecking papers could be part of a preprint server's
+policy, or simply be initiated by interested readers and community members.
+
+A pre-review CODECHECK allows editors to use it as a filter to only send 
+a submission out to peer review if it passes the check. A less strict
+pre-review check would simply be included in the submission package provided
+to the reviewers, so they can include the CODECHECK's results in their 
+evaluation of the work.
+A CODECHECK may also be conducted in parallel to the scientific review. This
+puts less burden on the turnaround time for the check yet only makes the 
+outcomes available for the final consideration of the handling editor.
+The check could also be tasked on demand, e.g., only if one of the reviewers
+suggests/requests one because previous screenings might not have caught the 
+fact that a computational workflow is part of the submission. However, 
+soliciting such a "specialist review" is much more undesirable than having
+a regular CODECHECK, without any special treatment for some submissions.
+A CODECHECK could then be awared equal weight as the other reviews.
+A post acceptance CODECHECK puts the least requirements on the author and may
+simply lead to an award of excellence on top of the acceptance of the
+submission. This is the least strict solution while still properly 
+acknowleding the check results, because the CODECHECK is to be completed 
+before the publication of the final paper.
+The GIScience group of checks (see \nameref{register}) falls into this
+category: the AGILE conference highlights articles where the reproducibility
+was successfully reviewed with a badge on the volume and article landing
+pages.
+
+Finally, a CODECHECK may also be conducted after publication, though it
+would required to update the article later with a note to reference the check,
+otherwise readers would presumably not be aware of it. While this has the
+least impact on the current publishing practices, it also has the smallest
+effect and does not acknowledge the importance of reproducible workflows for
+ensuring good scientific practice.
 
@@
 
-in house staff, pool of ECRs, volunteeres, for inhouse staff, credit not
-an issue.
+The options for the person who can conduct a CODECHECK require a not quite
+simple matching of a combination of skills and availability. Ideally, the
+codechecker has a matching code \emph{and} domain expertise, for example to
+check a workflow based on Python code analysing MRT brain imagery. However,
+a well-documented workflow should be executable by any person with a basic
+understanding of computers. Naturally, the more prerequisite knowledge the
+codechecker has, the less time is spend to understand the goals and 
+the mechanics of an analysis. From our experiences, the priority should be
+given to matching technical expertise first, as the a lack of knowledge in
+setting  up a computing environment with a particular language or tool is 
+much more discouraging and the assessment of the outcome, e.g., comparing
+created figures with the original, is largely possible without 
+understanding.
+The time and expertise alloted to the check
+are the main drivers for the depth of checking, though in general, we 
+expect a CODECHECK not to evaluate performance, e.g., does a racing car
+have the best apex speed, but to assess operation free of failure, i.e., 
+is it a car with 4 wheels and at least one door, and does it move forwards
+when accelerating. This depth can be seen independent from the power of
+the check (see above), as long as it is communicated clearly to all
+stakeholders, most importantly authors and readers.
+
+The concrete people to provide these expertises can range from researchers,
+possibly drawn from the regular pool of volunteer reviewers or from a 
+special group, to editorial staff of a publisher. If the regular reviewers
+are relied upon, the editor's task to find suitable ones becomes more
+tedious; extra information such as software familiarity would have to be
+modelled in review platforms.
+We see a great opportunity to especially involve early career 
+researchers as codecheckers. They arguably have a high interest in learning
+about the latest tools and technologies as they are building up their own
+expertise and specialisation. They are also introduced to peer review and 
+may transition into the role of scientific reviewer over time.
+Not to solve the general problems of peer review, which is largely an
+unsupervised process, i.e., rarely are reviewers taught how to do it, but
+a junior codechecker may be supported by a senior codechecker, too.
+We see a high potential in setting up the new processes for CODECHECKING with
+a clear committment on opennes and transparency, independent of the
+respective current peer review process (see \nameref{who-knows-who}).
+
+Publisher staff represents the most controlled yet also not quickly scalable
+option. Hiring the required technical and or domain expertise puts a financial
+burden on the publisher, of course, but such a committment also shows the 
+value given to publishing reproducibility and would positively impact the 
+submission experience, because of more controllability, and external 
+perception of the journal or conference. A staff codechecker reduces the 
+challenges when it comes to open communication and awarding credit, as
+in-house personnel may be more readily be given access to non-anonymised
+submission material and receives a proper salary instead of public credit
+as part of the scientific community.
+
+In contrast, for researchers it can be very important to be publicly credited
+with their activity as a reviewer (see \nameref{who-knows-who}). A regular
+review may be listed in public databases (e.g., ORCID
+\footnote{\href{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}} or commercial offerings such as Publons
+\footnote{\href{https://publons.com/}{https://publons.com/}} or 
+ReviewerCredits\footnote{\href{https://www.reviewercredits.com/}{https://www.reviewercredits.com/}}).
+A codechecker should be listed just the same. The number of scientific reviews
+if of course a very coarse indicator for the community contribution by an 
+individual researcher, so it's unclear if a regular reviewer, who in addition
+conducts a CODECHECK to the scientific review, should be credited twice.
+
+The codecheckers community\footnote{\href{https://github.com/codecheckers/codecheckers/}{https://github.com/codecheckers/codecheckers/}}
+currently has over 20 individuals who signed up as volunteer codecheckers
+following talks given on the project in the last 12 months.
+That is even though no checks are actively elicited.
+We see a high potential in an open shared list of potential codecheckers 
+compared to a private, in-house group, as it may better match expertise
+and share the workload.
+% TODO tell more about the motivations and commitments that we know about?
 
@@
 
-single blind, double blind, private/public communication between author
-and codechecker, collaborative. (This might need finessing with
-principle 2 below).
+The topic of blinding is broadly discussed, not the least in the push towards
+open peer review as part of the Open Science movement [REF]. Without taking
+a strong stand on this discussion, the general motivation for more 
+transparency and reproducibility does indeed favour a more open review 
+process. However, anonymity can also serve to protect individuals 
+\ref{tennant_limitations_2020}, e.g.,  juniour scientists, whereas it might 
+be acceptable to require revealing of the names of tenured researchers.
+The possible negative effects of a non-anonymous review are greatly reduced
+if a CODECHECK is not relevant for the decision to accept or reject, but that
+is of course not desirable. Instead, we argue that a CODECHECK is technical
+process that should in any case be fixable and not a question of opinion or
+faulty approach. It is possible for a nonsense workflow to receive a 
+CODECHECK certificate.
+Furthermore, the level of documentation that is needed for third parties
+to reproduce a workflow is extremely hard to get right. Too often, this 
+uncertainty leads to researchers giving up and not documenting at all.
+
+The technical nature of the check and the challenge of good enough 
+documentation is why we see great benefits in a effective bidirectional
+means of communication between author and codechecker. Instead of trying to
+fix problems or guess the next step, the codechecker can ask the author to 
+rework the docs  or the code.
+Instead of lenghtily describing a small change, the codechecker may simply
+provide a fix within the code.
+Instead of hiding useful and  instructive files created during the codecheck
+(e.g., a machine-readable computing environment specification), the author 
+and readers can profit from the software-related expertise of the
+codechecker.
+While such a communication may be facilitated in an anonymous way by the 
+publisher, it most likely only helps to protect the identity of the 
+codechecker, because, especially when of a certain size, not trivial, or 
+if very specialised, it is close to impossible to fully anonymise a 
+software analysis stack.
+Therefore, the most effective means with the most desirable result for
+the stakeholders is a non-anonymous and collaborative CODECHECK.
+The possible contributions by the codechecker may even be integrated into
+the code of the workflow and be acknowledged as code commits. This way, 
+proper credit is given within the research software development community.
+Nevertheless, depending on other needs represented by other dimensions,
+the aspect of `Who knows who?' may be the most flexible one.
+
+\subsubsection*{How is the CODECHECK integrated in other processes?}\label{turnaround-time}
+
+As a prerequisite, we see a CODECHECK as part of a typical scholarly
+communication process. While more extreme ideas to change the paradigms
+how researchers share their work exist [REF Octopus], there is a 
+value in an evolutionary approach. Thereby, existing communities can 
+transition towards more open practices at their own pace and ensure
+no one is left behind unfairly.
+When integrating a CODECHECK in existing review and publication processes,
+the \emph{turnaround time} is crucial. Depending on when and who conducts
+the check, it may be done faster, or add considerably to the overall
+time from submission to publication.
+% TODO write about our experiences how long a CODECHECK takes?
 

### 2020-10-01 codecheck-paper.tex
small update to principles, editorial changes

@@
 
-The workflow and variations just outlined below reflect our current
+The example workflow and variations just outlined above reflect our current
 views on how a codecheck should be performed. They are not set in
@@
 process), by listing codecheckers on the journal's website or, ideally, by
-publishing the checks alongside peer review activities in public databases,
-such as ORCID \footnote{\href{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}}
-or Publons.
+publishing the checks alongside peer review activities in public databases.
 
@@
 has at least enough material to validate the workflow outputs submitted by 
-the authors. Tode could be executed at least once without critical errors or
+the authors. Stark~\cite{stark_before_2018} calls this `preproducibility'.
+Code could be executed at least once without critical errors or
 warnings using the provided instructions, and, therefore, all code and data 
@@
 % count certificates: 
-% curl https://raw.githubusercontent.com/codecheckers/register/master/register.csv | wc -l
-% minus 1 for Hathway-Goodman (no empty last line, so no -1 for the headline needed)
+% curl https://codecheck.org.uk/register/stats.json
+% -1 for Hathway-Goodman (no empty last line, so no -1 for the column headers needed)
 
@@
 states which outputs, i.e., numbers, figures and tables, from the 
-original article could be reproduced by the CODECHECKER.
+original article could be reproduced by the codechecker.
 This certificate is made openly
@@
 lists the outputs that were generated from the codecheck (this is
-termed the MANIFEST.)  For each output, there is a comment stating
+termed the MANIFEST).  For each output, there is a comment stating
 which figure/table it should be compared to in the original paper,
@@
 codecheck (Figure~\ref{fig:cert}F).  In this case, the figure matches
-figure 4 of REF.  Subsequent pages of the certificate show other
+figure~4 of \cite{Unwin2020}.  Subsequent pages of the certificate show other
 outputs (not shown here).
@@
   \includegraphics[width=\textwidth]{figs/annotate-cert-crop.pdf}
-  \caption{Annotated certificate 2020-012 (first four pages only).}
+  \caption{Annotated certificate 2020-012 \cite{cert-2020-012} (first four pages only).}
   \label{fig:cert}

### 2020-10-01 codecheck-paper.tex
update text on R package, notes for related work

@@
 A custom R package, \texttt{codecheck}\footnote{
-\url{https://github.com/codecheckers/codecheck}} helps with the generation
-of certificates and depositing certificates and reated files on Zenodo 
-using the R package \texttt{zen4R} \cite{zen4r}.
+\url{https://github.com/codecheckers/codecheck}}, helps with the authoring
+of certificates, and with the deposition of certificates and reated files to
+Zenodo using the R package \texttt{zen4R} \cite{zen4r}.
 Codecheckers may choose to not use the package and rely on their own tools
@@
 
+\url{https://twitter.com/Code_Copilot}
+
+% first glimpse into "Code exeuction processes in scholarly publishing" ?
+% `https://docs.google.com/document/d/1fMWrFvBTwAYXg5J_bYIjgb-gY7DfQ0DOjx9Vlg_Rbmg/edit# ?
+
 DELIBERATEY NOT COMPREHENSIVE HERE.

### 2020-10-01 codecheck-paper.tex
extend text in limitations, add notes for future work

@@
 
+\textbf{Just peer review}: Wait, isn't that what peer review should be 
+doing already? That is correct.
+The general observation is that the increasing
+number of researchers, submissions, and reviews to conduct under the
+pressure to publish, puts the scholarly publication system under a lot
+of strain [REF].  We also see parallels to the terms 
+`reproducible research' and `open science'. Establishing a CODECHECK 
+process is largely an acknowledgement that peer reviewing practices 
+have faltered to adopt to the challenges of digitisation and computer-based
+research or data science. The concept of a CODECHECK, just as the 
+reproducibility of research and the openness of science, may be of 
+transitional nature. If the activites described here as being part of a 
+CODECHECK become the norm for scientific reviews and publications, then
+the initiative will have succeeded.
+However, the rapidly changing technology used in  research leads us to 
+believe that a seperate role of codechecker is benefitial in general.
+
 \textbf{Handling failure}: we have no established process for this, as
@@
 in this case -- should we publicly report that the code would not run.
+Be careful as to the impact of acceptance or volunteers this decision
+has.
 
@@
 time (think days, not minutes), who will pay for the compute time?
+The openness of CODECHECK certificates could help to establish community
+peer pressure, as submitting authors can be evaluated on their engagement
+as codecheckers.
 
@@
 
-\textbf{Can't someone cheat?} yes. we don't check it is correct code,
-just simply that it runs. Very low bar. But by having the code open,
-interested parties can then examine the code, and knowing that the code
-will be open we hope will be an incentive for authors to share.
+\textbf{Can't someone cheat?} Yes. we don't check it is correct code,
+just simply that it runs. This `mechanical' test is indeed a very low bar.
+Yet, by having the code and data openly deposited, interested parties can
+then examine the code, and knowing that the code will be open we hope will
+be an incentive for authors to share. It will also give the opportunity 
+to future researchers, potentially with new methods, to look for errors.
+This is a much more effective means to avoid cheating than an arms race
+trying to catch the cheater with closed datasets and code.
+% give credit to Everything Hertz podcast for the idea? they compared it
+% to the blood samples of drug tests in sports (stored for n years)
 
@@
 \item
+  CODECHECK works! we have created 23 certificates to demonstrate it.
+\item 
+  CODECHECK and can increase trust in results, underlies the same
+  limitations as peer review (needs people with time), but also the same
+  power (community, reasoning of human)
+\item
   possibility to allow CI for open papers such that for ``nearly free''
@@
   How to embed this in journals.
+\item
+  When some journals have it - what are crucial decisions/pain points?
+  How open do they make the check itself and its outputs, and how does it
+  relate/influence the openness of regular peer review
 \item

### 2020-10-01 codecheck-paper.tex
reflow grant info, update acknowledgements
remove incorporated notes from the end

@@
 Sustainability Institute and a Mozilla Science mini grant.
-DN is supported by grant \href{https://gepris.dfg.de/gepris/projekt/415851837}{PE~1632/17-1} from the German Research Foundation (DFG).
+DN is supported by grant
+\href{https://gepris.dfg.de/gepris/projekt/415851837}{PE~1632/17-1}
+from the German Research Foundation (DFG).
 
 \subsection*{Acknowledgements}\label{acknowledgements}
-
-Collaborators and potential collaborators:
+%Collaborators and potential collaborators:
 
@@
   \item
-    CrossRef and Publons and ORCID have features around reviews
-  \item
     shifting burden from reviewer to author (but not too much)
@@
 \begin{itemize}
+\item The limitations to our understanding of peer review: https://researchintegrityjournal.biomedcentral.com/articles/10.1186/s41073-020-00092-1
 \item
@@
 \item
-  on giving credit: public databases
-
-  \begin{itemize}
-  \item
-    \url{https://www.reviewercredits.com/}
-  \item
-    Publons
-  \item
-    ORCID
-  \end{itemize}
-\item
   compare with existing processes that are publicly documented

### 2020-10-01 codecheck-paper.tex
add idea about independence from financial models of journals
codecheck for independent journals

@@
 % TODO tell more about the motivations and commitments that we know about?
+The overarching community is also a mean to ensure CODECHECK is a viable
+option not just for large publishers, independent of their financial model,
+but also for independent no-cost open access journals.
 
@@
   measure of success.
+\item
+  codecheck allows innovation independent of finance model of the journal
+  and current peer review practices; evolutionary now, the "norm" later
 \end{itemize}

### 2020-10-01 codecheck-paper.tex
add link for Octopus and Hypergraph

@@
 As a prerequisite, we see a CODECHECK as part of a typical scholarly
-communication process. While more extreme ideas to change the paradigms
-how researchers share their work exist [REF Octopus], there is a 
+communication process. While more radical ideas to change the paradigms
+how researchers share their work exist\foonote{For example Octopus, 
+\url{https://science-octopus.org/about}, or Hypergraph
+\url{https://www.libscie.org/hypergraph}.}, there is a 
 value in an evolutionary approach. Thereby, existing communities can 

### 2020-12-04 codecheck-paper.tex
add todo
try to fix namerefs
add "main contribution" statement

@@
 \usepackage[numbers]{natbib}
-\usepackage{hyperref,url,setspace}
+\usepackage{hyperref}
+\usepackage{url}
+\usepackage{setspace}
 
@@
 \usepackage{doi} % turn DOIs into links
+\usepackage{nameref}
 
@@
 Given that there are no clear standards yet for sharing data, what hope
-might there be for sharing computer programs? Our anectodal experiences
+might there be for sharing computer programs? While both data and software
+together are required to validate outcomes of a computational analysis,
+the do differ in the respect that the former can be seen as static/inert,
+while the latter requires an environment and interaction. This makes 
+software harder to share.
+Our anectodal experiences
 on this matter suggest that often a variety of reasons are given for why
-their code cannot be shared, e.g.~``there is no documentation / I do not
+code cannot be shared, e.g.~``there is no documentation / I do not
 want to maintain it / I do not want to give away my code to my
-competitors''. Our view, rooted in open science princples, is that
+competitors''. Our view, rooted in Open Science princples, is that
 wherever possible sharing code is good for the community, as outlined
@@
 mind, we have developed a set of PRINCIPLES and an example workflow that
-provides what he hope is a pragmatic way of checking that code works. We
-call this system CODECHECK as outlined next.
+provides what he hope is a pragmatic way of checking that code works.
+
+The main contribution of this work is a thorough description of a process
+and its variations to integrate a much needed evaluation of the 
+computational reproducibility \cite{barba_terminologies_2018} into peer
+review, and a demonstration of its feasibility by means of over 20
+reproductions across different scientific disciplines.
+We call this system CODECHECK as outlined next.
 
@@
 policy, or simply be initiated by interested readers and community members.
+% TODO add a sentence about the possible effect of just asking for
+% code/data early on in the submission process, see https://twitter.com/tsuyomiyakawa/status/1230672758163402752?s=09)
 

### 2020-12-04 codecheck-paper.tex
smaller editorial changes and fixes
remove some comment

@@
 % folder: 1f4c1 1f4c2
-% recept: 1f9fe
+% receipt: 1f9fe
   \centering
@@
 However, in many cases the code and data may be published on a code hosting platform,
-such as GitHub/Lab.
+such as GitHub or GitLab.
 
@@
 
-The example workflow leaves sufficient room for different
+The example workflow leaves room for different
 implementations of this process. To elaborate on these options, we consider
-the following dimensions in a space of possible CODECHECKs, as shown in 
+the following dimensions in a space of possible CODECHECK workflows, as shown in 
 Figure~\ref{fig:dimensions}, to represent relevant aspects for the involved
@@
 
-\subsubsection*{When to do a codecheck and with what power}
+\subsubsection*{When to do a CODECHECK and with what power}
 \label{when-to-do-a-codecheck}
@@
 
-\subsubsection*{Who does the codecheck}\label{who-does-the-codecheck}
+\subsubsection*{Who does the CODECHECK?}\label{who-does-the-codecheck}
 
@@
 communication process. While more radical ideas to change the paradigms
-how researchers share their work exist\foonote{For example Octopus, 
+how researchers share their work exist\footnote{For example Octopus, 
 \url{https://science-octopus.org/about}, or Hypergraph
@@
 (3) GIScience.  
-%from interesting historical papers, to papers codechecked as part of
-%peer review, or because of public interest (COVID-19).
-
 The first theme was the initial set of papers used to explore the concept
@@
   \caption{Register of completed certificates as of October 2020.  An interactive version
-  of this is available at \url{http://codecheck.org.uk/register}. % TODO: SJE to complete...
+  is available at \url{http://codecheck.org.uk/register}. % TODO: SJE to complete...
   }
@@
 Data}.
+DN is reproducibility editor at the Association of Geographic Information 
+Laboratories in Europe's (AGILE) annual conference.
 

### 2020-12-04 codecheck-paper.tex
change heading levels
clarify types of checks

@@
 
+\subsection*{Workflow and people}\label{workflow-people}
+
 CODECHECK is best demonstrated by way of our example workflow, and later
@@
 
-\subsubsection*{Dimensions of CODECHECK processes}\label{dimensions-of-processes}
+\subsubsection*{Dimensions of CODECHECK workflows}\label{dimensions-of-workflows}
 
@@
 data, both from author and codechecker, will be made freely available, i.e., 
-under open licenses, when the certificate is published. The code and data 
-publication should follow community good practices.
-
-\section*{Register}\label{register}
+under open licenses, before or when the certificate is published.
+The code and data  publication should follow community good practices.
+
+\section*{Implementation}\label{register}
+
+\subsection*{Register}\label{register}
 % count certificates: 
@@
 and under our initative rather than requested from journal, but the certificate
-has then been acknowledged/cited in the paper (LANCET/LSHTM). % TODO SJE: clarify references in brackets
-The third theme represents co-author D.~N\"ust's work area.
+¸has then been acknowledged/cited in the paper (LANCET/LSHTM). % TODO SJE: clarify references in brackets
+The third theme represents co-author D.~N\"ust's service as Reproducibility
+Reviewer at the AGILE conference series.
 Independent of CODECHECK, the \emph{Reproducible AGILE} Initiative \cite{reproducible_agile}
 established a process for reproducing workflows at the AGILE conference
-series. % TODO add reference to MUNIN talk
-While using slightly different terms and infrastructue, e.g., `reproducibility
-reports' are published on OSF, instead of certificates on Zenodo, the 
+series \cite{nust_improving_2020}.
+While using slightly different terms and infrastructue, e.g., ``reproducibility
+reports'' are published on OSF instead of certificates on Zenodo, the 
 reproducibility reviews adhere to the codecheck principles.
-Furthermore, two checks were completed as part of a peer review for a
-GIScience journal.
+Furthermore, a small number of checks was completed as part of a peer reviews
+for a GIScience journals.
 
@@
 
-\subsubsection*{Tools and resources}\label{tools}
+\subsection*{Tools and resources}\label{tools}
 

### 2020-12-04 codecheck-paper.tex
extend limitations

@@
 
-\textbf{Just peer review}: Wait, isn't that what peer review should be 
+\textbf{Just peer review}: wait, isn't CODECHECK what peer review should be 
 doing already? That is correct.
-The general observation is that the increasing
+However, the general observation is that the increasing
 number of researchers, submissions, and reviews to conduct under the
@@
 process is largely an acknowledgement that peer reviewing practices 
-have faltered to adopt to the challenges of digitisation and computer-based
+have faltered to adapt to the challenges of digitisation and computer-based
 research or data science. The concept of a CODECHECK, just as the 
@@
 However, the rapidly changing technology used in research leads us to 
-believe that a seperate role of codechecker is benefitial in general.
-
-\textbf{Handling failure}: we have no established process for this, as
-so far all our codechecks succeeded. Open question remains as what to do
-in this case -- should we publicly report that the code would not run.
-Be careful as to the impact of acceptance or volunteers this decision
-has.
+believe that a seperate role of codechecker and the expertise it brings
+will be benefitial in the long term.
+
+\textbf{Not going deep enough}: 
+It should be noted that a codechecker is not taking the same roles as the
+\emph{statistical reviewer}, as it is practices by some journals in the 
+biomedical domain (cf.~\cite{petrovecki_role_2009,greenwood_how_2015}).
+The statistical reviewer actually evaluates the appropriateness of
+statistical methods \cite{greenwood_how_2015} and can support topical
+reviewers if, e.g., complex methods or sophisticated variants of statistical
+tests are applied \cite{petrovecki_role_2009}.
+The codechecker may go equally deep into the review, if they have, by 
+intention of by chance, the required expertise and the required time, 
+but they are by no means expected to be able to provide
+such a level of scrutiny. We can imagine a tiered workflow where the 
+codechecker, just as the conventional reviewer, could recommend a detailed
+statistical review to the editor as they might come across upon different
+hints.
+%The CODECHECK provides a certain level of safety with regard to the inherent
+%problems that a computational workflow might have. We want to capture the
+%obvious errors, but ... https://www.quora.com/What-is-the-difference-between-safety-and-security
+
+\textbf{Handling failure}: we have no established process for the case
+that a reproduction fails. This may be the sign of a pre-selection bias
+of the examples, as so far all our codechecks succeeded.
+In case of a journal adopting CODECHECK for all submissions, the open 
+question remains as what to do in this case will have to be answered.
+At the start, we doubt publicly reporting failures (i.e., the code 
+would not run) will increase overall reproducibility and recommend not
+to share the outcome publicly but to share the latest draft of the 
+certific with the authors.
+One must be careful as to the impact of acceptance by authors and the
+concerns of volunteers this approach, or a contrary one, has.
 
@@
 
-\textbf{Software}: authors can currently provide code that requires
-propieratiry software. Given the prevalence of software like MATLAB,
-this seems like a pragmatic choice, but then (a) we can't rely on open
-infrastructure like mybinder, (b) codechecker must have that software.
-
-\textbf{Can't someone cheat?} Yes. we don't check it is correct code,
+\textbf{Proprietary software}: authors can currently provide code that requires
+proprietary software. Given the prevalence of proprietary software in
+some disciplines, such as MATLAB or eCognition,
+this seems like a pragmatic choice.
+However, it prohibits to benefit from open infrastructure for reproducibility
+(cf.~\cite{konkol_publishing_2020}) and requires the codechecker to have access 
+to that software.
+Non-open software also considerably hampers reuse, especially by researchers
+from the global south. Therefore, allowing proprietary software is a compromise
+that each implementation of CODECHECK should critically reconsider.
+
+\textbf{Can't someone cheat?} Yes. We don't check it is correct code,
 just simply that it runs. This `mechanical' test is indeed a very low bar.
-Yet, by having the code and data openly deposited, interested parties can
+Yet, by having the code and (raw) data openly deposited, interested parties can
 then examine the code, and knowing that the code will be open we hope will
@@
 This is a much more effective means to avoid cheating than an arms race
-trying to catch the cheater with closed datasets and code.
-% give credit to Everything Hertz podcast for the idea? they compared it
-% to the blood samples of drug tests in sports (stored for n years)
-
-\textbf{Who's got time for more peer review?} Agree, but ECRS might be
-attracted to working with groups (recall it is not anonymous)
-
-\textbf{Who finds the codechecker?} Good point! The journal. Community
+trying to catch the cheater with closed datasets and code. This approach
+is comparable to the storing of blood samples in sports to detect
+doping in the future (cf. \cite{everythinghertz97}).
+Another picture that helped us to define the scope of a CODECHECK is:
+the codechecker is the forensic photographer, making sure all the details
+are captured so that an investigator may, now or in the future, scrutinize
+them.
+
+\textbf{Who's got time for more peer review?} Agree, but the technical
+nature and specific skill set might allow to introduce different groups
+into the peer review process. ECRs might be attracted to working with
+groups to learn more about recent methods and reproducibility practices.
+Research software experts (Research Software Engineers, RSE) who might
+not regularly be involved in writing or reviewing papers could be 
+interested to increase their connection with scholarly practices.
+An extra codechecker may simplify the matchmaking an editor has to do
+when identifying suitable reviewers for a submission, as technical and
+topical expertise can be provided by different people, who must not even
+be traditional peer reviewers. In fact, the codechecker could even be
+a software professional employed by a publisher.
+Also, recall that we expect CODECHECKSs to be not anonymous and always
+publicly deposited, which is common practice for some journals, but not
+broadly practiced. We also found that the focus on the mechanics of a
+workflow and the ability to interact with the authors can make
+reproductions very educational. It's a different role and as such could
+be a welcome alternation/diversion for reviewers.
+
+\textbf{Who finds the codechecker?} Good point! The journal, using either
+their own team of codecheckers or the public database. Community
 however think this is an important issue, and are volunteering.
@@
 \textbf{Checking multiple times?} If you check a preprint before peer
-review, it might need checking again as it is modified during peer
-review.  Unavoidable, and as happened to us (2020-012 has been
-revised).  But checking the manuscript the second time is likely to be
-much less work than the first time.
+review or a submission to a journal with non-public review,
+it might need checking again as it is modified during peer
+review.  This is in part unavoidable, and as happened to us (e.g., 
+`2020-012` has been revised). This can also be desired, if the
+interaction between author, reviewer, and codechecker lead to
+improvements.
+Since checking the manuscript the second time is likely to be
+much less work than the first time, this is largely mitigated by 
+open and regular communication.
 

### 2020-12-04 codecheck-paper.tex
first version of conclusions

@@
 
-\begin{itemize}
-\item
-  CODECHECK works! we have created 23 certificates to demonstrate it.
-\item 
-  CODECHECK and can increase trust in results, underlies the same
-  limitations as peer review (needs people with time), but also the same
-  power (community, reasoning of human)
+CODECHECK works!
+We have created a considerable number of certificates to demonstrate it.
+The creation of the certificates and the interactions with authors and
+editors not only improved the principles and the workflow, but also
+confirmed the approach taken. CODECHECKs and can increase trust in
+scientific results. Yet, they underly the same limitations as peer review
+in general and are closely connected to larger disruptions and challenges
+in scholarly communication \cite{eglen_recent_2018,tennant_ten_2019}, 
+including the tensions between commercial publishing and reviewer's often
+free labour.
+Even more, it is clear that establishing CODECHECK-like systems can be 
+impeded by and must be seen in connection to much larger issues in 
+science, such as broken metrics or malpractices triggered by publication
+pressure \cite{piwowar_altmetrics:_2013,nosek_promoting_2015}.
+While developed for the current "paper"-centric publication process and
+being compatible with different styles of peer review, the CODECHECK 
+principles would also work well with novel publication paradigms that enable
+much more iterative and granular communication of scientific outputs,
+e.g., Octopus \cite{}. An explicit segmentation of research steps could
+even make the focus of a CODECHECK, e.g., on the analysis sub-publication,
+easier.
+
+If adopted today, a CODECHECK system, even if temporal in a sustainable 
+transition towards more open publication and review practices, can contribute
+to increased trust in the results of research.
+More reproducible practices initiated by CODECHECKs could lead
+communities/disciplines to reach a state where authors do provide good
+enough material and reviewers have acquired broady enough skills that
+the latter will generally conduct a CODECHECK-level of checking, and
+only in especially sophisticated cases will a specialised codechecker
+be included.
+The main challenge for us remains to get journals to embrace the
+idea behind CODECHECK, whether they use the name for it or not.
+This \emph{cultural change}, however, is needed for the valuation of the
+efforts that go into proper evaluation of papers.
+With the implementation of CODECHECK workflows, the question of training
+codecheckers will become relevant. We expect both a mentoring scheme within
+the CODECHECK community (experienced codecheckers provide on the job training
+or serve as fallback advisors) as well as collaboration with reproducible
+research initiatives such as
+ReproHack\footnote{https://reprohack.github.io/reprohack-hq/},
+ReproducibiliTea\footnote{https://reproducibilitea.org/}
+\cite{fitzgibbon_brewing_2020},
+and Repro4Everyone\footnote{https://repro4everyone.org/}
+\cite{auer_reproducibility_2020},
+to be able to grow the pool of codecheckers as needed.
+The initial reaction of the scientific community, where up to today over 
+twenty people spontaneously signed up as volunteer codecheckers after talks
+on the initiative, makes us feel optimistic that the time is right for 
+putting scholarly peer review on the path to facilitate sharing and execution
+of computer programs and documenting reproducibility of results.
+
+\begin{itemize}
+\item
+
+\item
+
 \item
@@
 \item
-  role of data and code, data science
-\item
-  reproducible resarch
+  The Lancet's statistical review process: areas for improvement by authors
+  \url{https://www.thelancet.com/journals/lancet/article/PII0140-6736(92)90409-V/fulltext}
 \item

### 2020-12-10 codecheck-paper.tex
turn notes into content

@@
 \item Should we make code and data sharing a prerequisite or leave that door open?
+\item Take a look at the Lancet's statistical review process (areas for improvement by authors) \url{https://www.thelancet.com/journals/lancet/article/PII0140-6736(92)90409-V/fulltext}
+\item How to relate to "reproducibility does not matter but trust does", see \url{https://gmbecker.github.io/MayInstituteKeynote2019/outline.html}
+\item Should we ask the people currently listed in the acknowledgements?
+\item revisit \url{https://github.com/codecheckers/discussion}
 \end{itemize}
@@
 journal. At one of the spectrum, there are specialist journals that
-have been created to accept ``data papers'' (e.g.~\href{}{\emph{Scientific
-  Data}}, 
+have been created to accept ``data papers''
+(e.g.~\href{https://www.nature.com/sdata/}{\emph{Scientific~Data}}, 
 \href{https://essd.copernicus.org/}{\emph{Earth System Science Data}}, 
@@
 with a little effort, we believe that if someone independent can run/execute
-the code, this is worth documenting as early as possible. With this in
-mind, we have developed a set of PRINCIPLES and an example workflow that
-provides what he hope is a pragmatic way of checking that code works.
+the code, this is worth documenting as early as possible and that the barrier
+to evaluating non-text parts of research can be greatly reduced with a 
+focused approach.
+Just like peer review can provide a ``baseline reasurance'', i.e., a paper
+has at least ben checked by someone with an understanding of the topic
+\cite{fyfe_mission_2019}, we think the same baseline should be provided
+regarding the workflow underlying a paper.
+With this in mind, we have developed a set of PRINCIPLES and an example 
+workflow that provides what he hope is a pragmatic way of checking that 
+code works.
 
@@
 However, in many cases the code and data may be published on a code hosting platform,
-such as GitHub or GitLab.
+such as GitHub or GitLab. Ideally, the AUTHOR is expecting the CODECHECK and
+prepares for it, e.g., by asking a colleague to briefly attempt a reproduction
+of the submitted workflow.
 
@@
 any problems running the code, the CODECHECKER asks the author for help
-to resolve the problems. The CODECHECKER then tries to run the code
-again. This process iterates until either the CODECHECKER is successful,
+to resolve the problems. The burden to provide reproducible material lies with
+the author.
+The CODECHECKER then tries to run the code again.
+This process iterates until either the CODECHECKER is successful,
 or the CODECHECKER concludes the code does not work (\emph{what do in this
@@
 locally on their own compute resource, or in the cloud, e.g.~using the open
-MyBinder infrastructure \cite{jupyter_binder_2018}. Using such cloud-based 
+MyBinder infrastructure \cite{jupyter_binder_2018} or alternatives, some of which
+are more tailored for scientific publications while others offer commercial
+options to, e.g., publishers \cite{}.
+Using such cloud-based 
 infrastructure allows for the CODECHECKER and author to collaboratively improve
@@
 and data to be published openly online.
+Note that the taks of the CODECHECKER is to check the ``mechanics'' of the
+workflow, not to evaluate its correctness. 
+Stodden et~al. \cite{stodden_setting_2013} destinguish between verification and
+validation in the context of Mathematics, and in the same vein a CODECHECK
+ensures a verification\footnote{Note this is a different use as the \emph{Verification Reports} article type in the journal \emph{Cortex}
+\cite{chambers_verification_2020}},
+i.e., checking that the code correctly produces the 
+output it claims to create, but not a validation, i.e., does the code implement
+the right algorithm to solve the problem that is researched, 
+of the computational results.
 
@@
 the author how to improve the material.
+The plain text in the certificate is the most flexible and feasible way to 
+make clear what was really checked. Readers will have very different 
+understandings of what a ``green'' checkmark (cf. Step~5) means.
 
@@
 certificate repository to the published paper or the original code repository.
+A badge or other visual aid may be applied to highlight the fact of a completed
+CODECHECK and provide a incentive and reward for the, today often considered
+``additional'' efforts of providing readily computationally reproducible workflows,
+despite the simplification into a binary value and the possible confusion of
+the meaning, especially across disciplines and methods.
 
@@
 tedious; extra information such as software familiarity would have to be
-modelled in review platforms.
-We see a great opportunity to especially involve early career 
-researchers as codecheckers. They arguably have a high interest in learning
+modelled in review platforms. Keeping diverse specialists as staff to
+realise a codechking process is unlikely to be feasible, especially for 
+small or independent publishers.
+
+We consider a single codechecker to be enough, unlike peer review where
+multiple referees are common. Because of the technical focus, the 
+determination if the concrete outputs of author and codechecker, who can
+collaborate to solve any issues, should be a very factual process and not
+require much debate or opinion. Depending on the power of the CODECHECK
+(see \ref{when-to-do-a-codecheck}), a mediator, e.g., the editor, or a
+second opinion might be warranted in rare cases.
+
+We see a great opportunity to especially involve early-career researchers
+(ECRs) as codecheckers. They arguably have a high interest in learning
 about the latest tools and technologies as they are building up their own
-expertise and specialisation. They are also introduced to peer review and 
+expertise and specialisation. The possibility to gain insights into latest
+science for ECRs and at the same time increase the appreciation of
+reproduction efforts is a sweet spot that CODECHECK shares
+with \emph{ReScience X}, a journal devoted to reproduction and replication 
+experiments \cite{roesch_new_2020}.
+In fact, ECRs are often much more familiar with latest technology and are 
+likely to be early adoptors as creators, i.e., be authors of CODECHECK-ready manuscripts, too\footnote{Cf. the 77\% of 141 registered reports who were 
+submitted to leading Neuroscience journals by ECRs
+\cite{chambers_registered_2019}.}
+They are also introduced to peer review and 
 may transition into the role of scientific reviewer over time.
@@
 We see a high potential in setting up the new processes for CODECHECKING with
-a clear committment on opennes and transparency, independent of the
+a clear committment on openness and transparency, independent of the
 respective current peer review process (see \nameref{who-knows-who}).
@@
 A codechecker should be listed just the same. The number of scientific reviews
-if of course a very coarse indicator for the community contribution by an 
-individual researcher, so it's unclear if a regular reviewer, who in addition
+is of course a very coarse indicator for the community contribution by an 
+individual, yet is is unclear if a regular reviewer, who in addition
 conducts a CODECHECK to the scientific review, should be credited twice.
@@
 time from submission to publication.
-% TODO write about our experiences how long a CODECHECK takes?
+In our first checks, we found that a codecheck may generally take from 2-5
+hours, with some outliers on the higher end. This time includes writing and
+publishing the certificate, but excludes the time for actual computations.
+The efforts include considerable time to communicate about the process, 
+especially around who publishes what and when, so that proper 
+cross-referencing between paper and certificate is ensured.
+When integrated into peer review systems, this handling of the steps should
+become much more effective.
 
@@
 
-\textbf{4. Workflows must be auditable.} This requires that the codechecker 
+\textbf{4. Workflows must be auditable.}
+This requires that the codechecker 
 has at least enough material to validate the workflow outputs submitted by 
-the authors. Stark~\cite{stark_before_2018} calls this `preproducibility'.
-Code could be executed at least once without critical errors or
+the authors. Stark~\cite{stark_before_2018} calls this `preproducibility'
+and the ICERM report \cite{} defines the level ``Auditable Research'' 
+similarly. Note that openness is not required here, as to accommodate 
+communities of practice and review workflows that (currently) do not want
+to or are not able to work openly.
+A completed check eans that code could be executed at least once without
+critical errors or
 warnings using the provided instructions, and, therefore, all code and data 
@@
 skills. It is also hard to create a sufficient level of documentation for 
-third parties. That is why the possibility to communicate is more important
+third parties. Furthermore automation may lead to people gaming the system
+or reliance on technology hiding important details. All such aspects can
+reduce the understandability of the material, which we estimate to be the
+best mean to ensure long-term transparency and usefulness.
+That is why the possibility to communicate is more important
 than perfected workflows, and why a CODECHECK is not automated.
-A codechecker does \emph{not} conduct a code review.
-% more content from the website, may fit better elsewhere:
-% The CODECHECK is not automated on purpose: automation may (a) lead to people gaming the system, (b) hide details that eventually decrease level of certainty that a codechecker has in their assessment, and (c) reduce the understandability of instructions in the long term, which is more important than short term ease of use (see Principle 2).
-
-\textbf{5. Open by default.}  By default, unless there are strong
+We acknowledge that others have argued in favour of bitwise reproducibility
+because in the long run it can be automated\footnote{Konrad Hinsen on 
+Twitter: \url{https://twitter.com/khinsen/status/1242842759733665799}};
+but until then we need CODECHECK's approach.
+
+\textbf{5. Open by default and transitional by disposition.} % or "design"?
+By default, unless there are strong
 reasons to the contrary (e.g., clinically-sensitive data), all code and
@@
 The code and data  publication should follow community good practices.
+By disposition, a CODECHECK workflow should be designed as a recognition of
+shortcomings in asserting reproducibility of submissions and understand the
+complex situation that lead to this. However, establishing a CODECHECK workflow
+should go hand in hand with initiatives and a perspective to improve education
+around reproducibility of methods and communication of comprehensive 
+computational workflows, with the long term goal for a specific community or
+journal's target audience to reach a state where the level of verification 
+and trust provided by codecheckers becomes part of peer review.
 
@@
 and under our initative rather than requested from journal, but the certificate
-¸has then been acknowledged/cited in the paper (LANCET/LSHTM). % TODO SJE: clarify references in brackets
+has then been acknowledged/cited in the paper (LANCET/LSHTM). % TODO SJE: clarify references in brackets
 The third theme represents co-author D.~N\"ust's service as Reproducibility
@@
 The scripted and automated parts should also ensure that more checks can
-be conducted in the future.
+be conducted in the future, though the management by a paid CODECHECK
+editor or further automation, e.g., with a bot, could improve turnaround time.
 The remaining costs are the project's domain, which is negligible, and
@@
 
-TO BE CONTINUED.
-
-All the related projectsfrom A-Z\ldots{} e.g.~CASCAID, Rescience,
-modeldb, o2r. Community iniatives like reprohack along similar lines of
-checking code. \url{https://reprohack.github.io/reprohack-hq/}
-
-\url{https://twitter.com/Code_Copilot}
-
+The journal \emph{ACM Transactions on Mathematical Software (TOMS)} 
+established a "Replicated Computational Results" (RCR) review proces
+\cite{heroux_editorial_2015} (using "replicable" with the same meaning 
+as CODECHECK uses "reproducible").
+The CODECHECK principles were independently developed as we only later 
+learned about TOMS RCR, which we judge as fulfilling CODECHECK principles
+1 to 4.
+This work provides more extensive discussion of the possible variants, 
+while TOMS RCR realises one particular workflow for reproducibility 
+reviews for one particular journal. \cite{heroux_editorial_2015}
+include similar concerns about
+selection of reviewers as above. Unlike existing CODECHECK certificates, 
+the RCR reports undergo editorial review, but this could very well become
+part of CODECHECK workflows at journals.
+The publication of the RCR report is seen as a reward for the efforts of
+the reproducing person, while the possible conflict of interest of this 
+motivation is acknowledged. TOMS also reconises reviewer activity in a 
+partnership with Publons
+\footnote{\url{https://authors.acm.org/author-services/publons}},
+though it is unclear if the author of an RCR report receives both the report
+authorship and the reviewer recognition. In any case, this is different
+from CODECHECK, which specifically aims at recognition as reviewers.
+In our view, this removes the possible conflict of interest while keeping
+the public acknowledgement.
+
+Specific to the field of Mathematics, the RCR was also  expected to apply a
+review of the software itself if the system it runs on cannot be evaluated by
+an independent party.
+The TOMS RCR also notes the importance of communication and expects
+collaboration between author and RCR reviewers.
+The TOMS creators share thoughts around reviewer selection and also put trust
+in reviewer judgement over numerical bit-wise perfection.
+A key difference is that authors opt-in with an \emph{RCR Review Request} and
+the RCR reports, which are very similar to CODECHECK certificates, are
+published in the TOMS journal next to the actual papers.
+A search on Crossref\footnote{\url{https://search.crossref.org/?q=\%22Replicated+Computations+Results+\%28RCR\%29+Report\%22&from_ui=yes&page=2}; search executed
+on 2020-12-10} shows 15 RCR Reports to be published and the workflow to be 
+extended to the ACM journal \emph{Transactions on Modeling and Computer 
+Simulation (TOMACS)}\footnote{\url{https://dl.acm.org/journal/tomacs/author-guidelines#rcrini}}.
+
+% 8: https://scholar.google.de/scholar?start=10&q=Replicated+Computations+Results+%22(RCR)+Report%22&hl=de&as_sdt=0,5
+% 15: https://search.crossref.org/?q=Replicated+Computations+Results+%22%28RCR%29+Report%22&from_ui=yes&publication=ACM+Transactions+on+Mathematical+Software
+% 15: https://academic.microsoft.com/search?q=Replicated%20Computations%20Results%20%22(RCR)%20Report%22&f=&orderBy=0&skip=10&take=10
+
+A number of journals or conferences/workshops provide guidelines for ensuring reproducible publications and their verification.
+These include, for example, \ldots{}
 % first glimpse into "Code exeuction processes in scholarly publishing" ?
@@
 
-DELIBERATEY NOT COMPREHENSIVE HERE.
+\emph{Information Systems}, with an invited reproducibility paper track
+\footnote{\url{https://www.elsevier.com/journals/information-systems/0306-4379/guide-for-authors}}
+
+However, no approach besides CODECHECK spans across journals and is
+intentionally designed to be picked up and implemented by many different 
+journals or events, and to build a community of codecheckers. CODECHECK shares
+this interdisciplinary nature with other community iniatives concered with 
+reproducibility awareness, education, and supports, such as ReproHack
+\footnote{\url{https://reprohack.github.io/reprohack-hq/}}, Code Copilot
+\footnote{\url{https://twitter.com/Code_Copilot}}, or Papers with Code
+\footenote{\url{https://paperswithcode.com/about}}.
+The latter recently announced a collaborations with the preprint server
+\emph{arXiv} on providing data and code supplements for machine learning
+manuscripts\footnote{\url{https://medium.com/paperswithcode/papers-with-code-partners-with-arxiv-ecc362883167}} and runs a reproduciblity challenge
+\footnote{\url{https://paperswithcode.com/rc2020}}.
+
+The journal \emph{ReScience C} publishes replications,
+\url{https://rescience.github.io/faq/};
+they do not accept proprietary software and require open source code;
+they also required a third party to replicate; they also use Zenodo
+to issue DOIs.
+The journal \emph{Cortex} has a special article type 
+\emph{Verification Reports}, which  actually are about replication of results 
+but are very well designed/reasoned \cite{chambers_verification_2020}.
+In a similar vein, the CODECHECK certificates
+could also be published as a special article type within journals.
+
+For research with very high stakes, where reproduction would be too weak and
+post publication replication possibly to late because of policiy impact,
+\cite{benjamin-chung_internal_2020} propose \emph{internal replication}.
+A workflow that underwent internal
+replication would very likely be of high quality and relatively easy to codecheck.
+In a similar vein, internal codechecks may be used, with the same limitations such
+as group think \cite{benjamin-chung_internal_2020},
+to ensure reproducibility before submission. The guidelines
+for CODECHECK can be used by fellow/not independent lab members just as well.
+
+Gavis and Donoho \cite{gavish_universal_2011} propose a new discipline and 
+infrastructure for reproducible computational research with its own tools and
+API. Their specific packaging format, provenance record, and cryptographic
+\emph{Verifiable Result Identifier} would indeed provide excellent
+reproduciblity. However, the system is also complex and was not taken up by
+any publisher as far as we know; the system is indeed not open source itself.
+In comparison, CODECHECK is less powerful but also much more flexible and
+less dependent on specific tools or infrastructure. If data and code are 
+deposited properly, i.e., very unlike to disappear, then the certificate's 
+DOI is practically quite close to the cryptographic identifier.
+
+Local reproduction initiatives/services:
+\href{https://ciser.cornell.edu/research/results-reproduction-r-squared-service/}{CISER R-squared}
+(established/permanent),
+\href{https://go.wwu.de/r2s2}{WWU R2S2} (pilot only, focused on reproductions),
+\href{https://github.com/OxfordCodeReviewNet/forum}{Oxford code review network (OxCRN)}
+(community, broader scope).
+For these, the CODECHECK scope, tools, and practical experiences may be
+helpful resources.
+
+\emph{SciGen.Report} is a community-run platform to foster communication on
+reproducibility\footnote{\url{https://scigen.report/}}. People can report on
+fully, particially or un-successful reproductions post publication.
 
@@
 
-\textbf{Not going deep enough}: 
+\textbf{Not going deep enough/too vague}:
+CODECHECK by design does not require authors to provide (and sustain) an
+eternally functional worklow and neither suggests a specific software stack
+or pracitcal approaches, such as research compendia
+\footnote{\url{https://research-compendium.science/}},
+even if they are very valuable.
+Creating something that anyone can reproduce has been called a 
+fool's errand
+\footnote{\url{https://twitter.com/DougBlank/status/1135904909663068165?s=09}}
+and we tend to agree.
+However, the package of data, code, and documentation
+collaboratively created by authors and codecheckers is a snapshot of a 
+working analysis that greatly increases the likelyhood of a successful 
+reproduction and possibility to extensions by third parties, with access
+to suitable resources and matching skill set, in the future.
+Concrete implementations of CODECHECK workflows, especially for specific
+disciplines, may reify much more helpful guidelines for authors how to
+create reproducibility packages.
+The author-friendly ``low bar'' should not stay low forever, but cultural
+change takes time and the encouragement and guidance that CODECHECK,
+as part of the widely accepted peer review concept, can provide may
+eventually allow to raise the bar as high as possible, e.g., with
+executable research compendia \cite{nust_opening_2017}
+or continuous analysis \cite{beaulieu-jones_reproducibility_2017-1}.
+However, considering that missing artefacts and lack of documentation
+have repeatedly been identified as main barriers to reproducibility
+(e.g., \cite{stagge_assessing_2019,nust_improving_2020}),
+we would not underestimate the power of a simple check.
+
 It should be noted that a codechecker is not taking the same roles as the
@@
 
+An important distinction is that a codechecker does not conduct a 
+\emph{code review}. Code reviews are without doubt extremely valuable in 
+terms of improving reproducibility and re-usability, and their proponents 
+even believe they can improve the science \cite{petre_code_2014}.
+Code reviews, however, have quite different structural challenges and 
+require even more resources. That said, a well-reviewed codebase is likely
+to be easier to codecheck, and the awareness for high-quality code raised
+through CODECHECK may lead to more support for code reviewing.
+Initiatives and journals that conduct software reviews independently of 
+a specific publication or workfloware, e.g.,
+ROpenSci\footnote{\url{https://ropensci.org/}},
+PyOpenSci\footnote{\url{https://www.pyopensci.org/}},
+and JOSS\footnote{\url{https://joss.theoj.org/}}.
+
+Furthermore, the codechecker's task list is intentionally not overloaded
+with related concerns, even though they are very important, such as ensuring
+proper citation of data and software, or depositing data and software in 
+suitablerepositories.
+We do expect codecheckers to be aware of these though and possibly point 
+them out.
+
 \textbf{Handling failure}: we have no established process for the case
@@
 concerns of volunteers this approach, or a contrary one, has.
+Private feedback for failures and public acknowledgement/encouragement
+for successes seems to be the way to go.
+\cite{Rosenthal2016b} discuss incentives for different actors around the
+implementation of reproducibility. We see CODECHECK as one mean for
+organisations to invest in reproducibility by creating incentives
+until reproducible computations become the norm.
 
@@
 as codecheckers.
+The challegnes of computational resources can also quickly be reduced when
+authors provide a small sample or a synthetic dataset together with the
+expected outputs for said subset, which can equally wel demonstrate the 
+usefulness and validity of their workflow.
 
@@
 However, it prohibits to benefit from open infrastructure for reproducibility
-(cf.~\cite{konkol_publishing_2020}) and requires the codechecker to have access 
-to that software.
+(cf.~\cite{konkol_publishing_2020,perkel_make_2019})
+and requires the codechecker to have access to that software.
 Non-open software also considerably hampers reuse, especially by researchers
@@
 that each implementation of CODECHECK should critically reconsider.
+
+\textbf{Not possible for my data}: 
+Solutions for proprietary and sensitive data exists.
+Authors can provide synthetic data (cf. \cite{shannon_opening_2018}), some
+data can effectively be redacted \cite{oloughlin_data_2015}, and publishers or
+independent entities can provide infrastructure for sharing data and workflows
+confidentially \cite{perignon_certify_2019} or with access to derived results
+but not raw data \cite{shannon_opening_2018},
+i.e., data enclaves \cite{foster_research_2018},
+or domains of reproducibility \cite{harris_more_2017})
 
@@
 
-\textbf{Who's got time for more peer review?} Agree, but the technical
-nature and specific skill set might allow to introduce different groups
+\textbf{Who's got time for more peer review?} Agree.
+However, there are differences to peer review.
+First, the technical nature of a CODECHECK can provide a more 
+clearly defined scope and expectations compared to conventional 
+peer review. The authors are told what to provide and the codechecker
+can be told when to stop, for example.
+Second, the specific skill set allows to introduce different groups
 into the peer review process. ECRs might be attracted to working with
@@
 a software professional employed by a publisher.
-Also, recall that we expect CODECHECKSs to be not anonymous and always
+Third, recall that we expect CODECHECKSs to be not anonymous and always
 publicly deposited, which is common practice for some journals, but not
-broadly practiced. We also found that the focus on the mechanics of a
+broadly practiced. The possibility to ask for clarification or more 
+material/information can greatly reduce the digging required and thereby
+the time spend.
+We also found that the focus on the mechanics of a
 workflow and the ability to interact with the authors can make
 reproductions very educational. It's a different role and as such could
-be a welcome alternation/diversion for reviewers.
+be a welcome alternation/diversion for researchers to give back their 
+time to the community.
 
@@
 however think this is an important issue, and are volunteering.
+With increasing demand
 
@@
 open and regular communication.
+The higher the reproducibility and automation, the more aspects of 
+a CODECHECK may eventually be supported by automation, though we believe
+the human codecheckers should remain the decision maker.
+Automatically detecting if results are ``the same'' is a question
+intrinsically linked with each unique process and quite hard to realise
+for common outputs, such as plots or maps.
+Nevertheless, continuous integration (CI) infrastructure may help 
+codecheckers, and, without much further work, use scientific workflows as
+detectors for when a piece of software breaks if a workflow is set up to
+use the latest versions of tools.
 
@@
 
-\section*{Future work and conclusions}\label{future-work-and-conclusions}
+\textbf{Not revolutionary enough?}
+CODECHECK's approach is to acknowledge the shortcomings around 
+computational reproducibility and iteratively improve within the current
+system of scholarly communication through journals.
+It remains to be proved if this approach is welcomed broadly and if 
+involving the traditional system of publishers helps to push the 
+conversation along or not.
+We have discussed stronger more drastic rules at length, such as only 
+considering diamond Open Access journals, community-led no-fee journals,
+or journals with collaborative/open peer review, but eventually decided
+against it.
+Note that the focused nature of the CODECHECK principles allows journals
+and publishers to innovate on the finance model and peer review practices
+at their own pace.
+
+\section*{Conclusions and future work}\label{future-work-and-conclusions}
 
@@
 editors not only improved the principles and the workflow, but also
-confirmed the approach taken. CODECHECKs and can increase trust in
-scientific results. Yet, they underly the same limitations as peer review
-in general and are closely connected to larger disruptions and challenges
-in scholarly communication \cite{eglen_recent_2018,tennant_ten_2019}, 
+confirmed the approach taken. This result corroborates findings from
+similar evaluations of reproducible computational research in journals and
+at conferences.
+CODECHECKs and can increase trust in scientific results and the set of 
+shared principles and name can through recognition value allow researchers
+across publication venues to judge the level of scrutiny that results have
+faced. CODECHECK requires direct acknowledgement of the contribution to the
+scientific community in the role, not indirectly via citations of 
+reproductions, and can be a worthwhile educational activity, especially for
+ECRs. A technically simple yet far reaching change would be a special type of 
+reviewer activity for codechecking in public researcher profiles next to
+conventional peer reviews.
+
+Yet, CODECHECK underlies the same limitations as peer review
+in general and is closely connected to larger disruptions and challenges
+in scholarly communication
+\cite{eglen_recent_2018,tennant_ten_2019,fyfe_mission_2019}, 
 including the tensions between commercial publishing and reviewer's often
-free labour.
+free labour, and a global pandemic jumbling up scientific publishing
+and exposing a broader general audience to science, even preliminary 
+results \cite{munafo_what_2020}.
 Even more, it is clear that establishing CODECHECK-like systems can be 
@@
 pressure \cite{piwowar_altmetrics:_2013,nosek_promoting_2015}.
+We certainly do not want the binary attribute of "CODECHECK completed"
+to become a factor in bibliometric approaches for performance assesments.
+% Mittermaier, Peer Review and Bibliometrics: http://hdl.handle.net/2128/22745
 While developed for the current "paper"-centric publication process and
 being compatible with different styles of peer review, the CODECHECK 
-principles would also work well with novel publication paradigms that enable
-much more iterative and granular communication of scientific outputs,
-e.g., Octopus \cite{}. An explicit segmentation of research steps could
-even make the focus of a CODECHECK, e.g., on the analysis sub-publication,
-easier.
+principles would also work well with novel publication paradigms, e.g.,
+peer-reviewed computational notebooks \cite{earthcube_new_2020},
+iterative and granular communication of scientific outputs,
+such as Octopus\footnote{\url{https://science-octopus.org/}}, 
+articles with live-code \cite{perkel_pioneering_2019-1}
+\footnote{For example, eLife's ERA, see 
+\url{https://elifesciences.org/labs/dc5acbde/welcome-to-a-new-era-of-reproducible-publishing}.},
+and completely new visions for scholarly communication and peer review
+\footnote{Cf., for example, \emph{A modern vision for peer review} by 
+Amy J. Ko: \url{https://medium.com/bits-and-behavior/a-modern-vision-for-peer-review-d5f73f0fae07}}.
+An explicit segmentation of research steps could even make the focus 
+of a CODECHECK easier by only checking the ``analysis'' sub-publication.
+The discovery of CODECHECKs could be increased by depositing the reports
+into public databases of reproductions, such as \emph{SciGen.Report}.
+Public researcher profiles, such as ORCID, may consider different
+types of reviewer activity to capture the contributions of independent 
+code execution to science.
+Notably, the discussed limitations are largely self-imposed for easier
+acceptance and evolutionary integration, as to not break the current 
+system and increase demands over time without leaving practitioners behind.
 
@@
 transition towards more open publication and review practices, can contribute
-to increased trust in the results of research.
+to increased trust in the results of research. Introducing CODECHECK
+should be informed by lessons learned from (introducing) open peer review
+(OPR) \cite{ross-hellauer_guidelines_2019}.
+All our conversations with publishers and editors at some point
+touched on the procedural and technical implications of openness, be it 
+that OPR is neither applied nor well-known by stakeholders, or that process
+management software would not support such open communication.
+We would be keen to use the flexibility of the principles and cooperate
+on non-public CODECHECKs to learn more about the advantages and yet unclear
+specific challenges. Does CODECHECK work better with OPR?
+Each rollout of CODECHECK should be accompanied by studies to ensure 
+codechecking does not fall into the same gaps as peer review did \cite{tennant_limitations_2020}
+and to ensure positive change within the respective peer review system.
 More reproducible practices initiated by CODECHECKs could lead
@@
 The main challenge for us remains to get journals to embrace the
-idea behind CODECHECK, whether they use the name for it or not.
+idea behind CODECHECK and realise processes conform to the principles,
+whether they use the actual name for it or not.
+The CODECHECK badge is, intentionally, not branded and simply states
+``code works''.
 This \emph{cultural change}, however, is needed for the valuation of the
 efforts that go into proper evaluation of papers.
+Multiple adopting journals will also help to answer relevant questions on
+the establishing process (What are crucial decisions or pain points? Can
+authors retract code/data once a CODECHECK has started?),
+what variants of CODECHECKs will be most common, and how open CODECHECKs
+may influence the scope and anonymity of conventional review over time.
+
 With the implementation of CODECHECK workflows, the question of training
@@
 
-\begin{itemize}
-\item
-
-\item
-
-\item
-  possibility to allow CI for open papers such that for ``nearly free''
-  we can find out when softwave breaks, and if provenance tools are
-  used, when.
-\item
-  How to embed this in journals.
-\item
-  When some journals have it - what are crucial decisions/pain points?
-  How open do they make the check itself and its outputs, and how does it
-  relate/influence the openness of regular peer review
-\item
-  When is CODECHECK redundant? will it become part of regular peer
-  review? or, if ayuthors become more skillful, then everything becomes
-  automated, and no need for codecheck. CODECHECK disappearing is then a
-  measure of success.
-\item
-  codecheck allows innovation independent of finance model of the journal
-  and current peer review practices; evolutionary now, the "norm" later
-\end{itemize}
-
-Or will it always be adjunct to peer review, and needed?
-
-working within the trad system of publishers to help push the
-conversation along.
-
 \subsection*{Competing interests}
@@
 %Collaborators and potential collaborators:
+% someone from SpringerNature
 
@@
 Yuhao (Sebastian) Wang developed code and example certificates.
+We thank Antonio P\'{a}ez (Journal of Geographical Systems) for enabling
+CODECHECKs, Carlos Granell and Frank Ostermann for contributing
+certificates as reproducibility reviewers at the AGILE conference, and 
+all authors of auditable workflows for their participation.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\clearpage
-\begin{itemize}
-\item
-  The Lancet's statistical review process: areas for improvement by authors
-  \url{https://www.thelancet.com/journals/lancet/article/PII0140-6736(92)90409-V/fulltext}
-\item
-  data enclaves
-\item
-  CASCAID
-\item
-  ROpenSci/PyOpenSci (and how their software review differs, though they
-  also collaborate with journals,
-  \url{https://devguide.ropensci.org/softwarereviewintro.html})
-\item
-  time capsules \textgreater{}
-  \url{https://twitter.com/DougBlank/status/1135904909663068165?s=09}
-\item
-  science is changing:
-  \url{https://theconversation.com/what-you-need-to-know-about-how-coronavirus-is-changing-science-137641}
-\item
-  reproducibility is important, also
-
-  \begin{itemize}
-  \item
-    effects on well-being/mental health,
-    \url{https://www.nature.com/articles/d41586-020-01642-9}
-  \end{itemize}
-\end{itemize}
-
-\textbf{main contributions}
-
-\begin{itemize}
-\item
-  a concept for integrating a minimal code review into common scholarly
-  communication processes around peer review
-\item
-  set of shared principles (recognition value) that allow researchers to
-  quickly understand the level of evaluation across
-  journals/publishers/conferences, and allows helps these stakeholders
-  to establish workflow checking in their domain/solution/product
-\item
-  a common language and rubric for classifying different levels of code
-  and data evaluation as part of a peer-review
-\end{itemize}
-
-\section*{What is CODE CHECK?}\label{what-is-code-check}
-
-\begin{itemize}
-\item
-  \textbf{Principles}: \url{https://codecheck.org.uk/}
-\item
-  relation to ``proper'' citation of data and software, and depositing
-  of data and software in suitable repos \emph{besides} the bundle for
-  CODE CHECK: do it, only provide the concrete analysis script for the
-  check
-\item
-  \textbf{Why is it useful?}
-
-  \begin{itemize}
-  \item
-    reduce the barrier to evaluating non-text parts of research
-  \item
-    acknowledges different skill sets and existing high load on
-    reviewers
-  \item
-    CC breaks problem that publishers staff expertise today is not
-    likely to suffice
-  \item
-    engaging ECRs is a perfet fit (they are up-to-date in methods, are
-    introduced to peer review)
-  \item
-    relation to research compendia? point it out for authors unsure
-    about how to structure their workflow
-  \item
-    shifting burden from reviewer to author (but not too much)
-  \item
-    What issues with (open) peer review does CODECHECK address?
-  \end{itemize}
-\item
-  keynote \url{https://doi.org/10.7557/5.4963}: in discussion, the
-  speaker mentioned value of peer review for ``baseline reassurance'',
-  i.e.~a paper has at least been checked by someone with an
-  understanding, to increase trust especially if looking at papers from
-  other disciplines
-\item
-  \textbf{report} must fulfil some requirements to make clear what is
-  really checked and not mislead - readers will have very different
-  understandings of a green checkmark!
-\item
-  reproducibility is hard
-  \url{https://twitter.com/annakrystalli/status/1144176149859377152?ref_src=twsrc\%5Etfw}
-  \textgreater{} code check gives at least some benefits
-\item
-  what are the requirements / level or reproducibility required or
-  checked?
-
-  \begin{itemize}
-  \item
-    \href{https://stodden.net/icerm_report.pdf}{Stodden et al., 2013}:
-    five-level hierarchy of research - reviewable, replicable,
-    confirmable, auditable, and open or reproducible
-  \item
-    \href{https://figshare.com/articles/Showing_your_working_A_guide_to_reproducible_neuroimaging_analyses/4244996/2}{Whitaker
-    2017}: reproducible, generalisable, replicability, robustnes
-  \end{itemize}
-\item
-  can authors retract their code/data once a CODECHECK has started? In
-  the eLife prepreint review, there seems no ``way back'', see
-  \url{https://twitter.com/ceptional/status/1271528976763121664?s=09}
-\end{itemize}
-
-\section*{Related work}\label{related-work-1}
-
-\begin{itemize}
-\item The limitations to our understanding of peer review: https://researchintegrityjournal.biomedcentral.com/articles/10.1186/s41073-020-00092-1
-\item
-  CODECHECK vs.~Verification Reports \textgreater{} \textgreater{}
-  \url{https://www.sciencedirect.com/science/article/pii/S0010945220301738\#fn3}
-\item
-  CODECHECK vs.~the ``Artifacts Evaluated - Functional'' badge of the
-  ACM, see
-  \url{https://www.acm.org/publications/policies/artifact-review-badging}
-\item
-  Does CODECHECK do internal replication, as defined by
-  \url{https://gatesopenresearch.org/articles/4-17/v2} ?
-\item
-  \url{https://paperpile.com/shared/rVNwBS}
-\item
-  Are ``Replicated Computations Results (RCR) Report'' a thing?
-  apparently so (using ACM terminology)
-
-  \begin{itemize}
-  \item
-    \url{https://dl.acm.org/doi/10.1145/3185337}
-  \item
-    \url{http://users.iems.northwestern.edu/~nelsonb/Publications/Nelson_RCRgreenSim.pdf}
-  \item
-    \url{https://dl.acm.org/doi/epdf/10.1145/3341094}
-
-    \begin{itemize}
-    \item
-      citation for replicated workflow lists it as ``to appear''
-    \end{itemize}
-  \item
-    getting proper papers for reproductions != CODECHECK approach where
-    we want credit as reviewers
-  \item
-    see also
-    \href{https://www.google.com/search?q=\%22Replicated+Computations+Results+\%28RCR\%29+Report+for\%22}{Google
-    Search}
-  \end{itemize}
-\item
-  \url{https://medium.com/bits-and-behavior/a-modern-vision-for-peer-review-d5f73f0fae07}
-  - work by Amy J. Ko:
-  \url{https://faculty.washington.edu/ajko/publications}
-
-  \begin{itemize}
-  \item
-    is CODE CHECK's process too small, too incremental a change?
-  \end{itemize}
-\item
-  overview of peer review and bibliometrics (if not a reference itself,
-  it gives plenty of references):
-  \url{http://juser.fz-juelich.de/record/865096/files/Peer\%20Review\%20and\%20Bibliometrics.pdf}
-\item
-  Open Peer Review \textgreater{}
-  \url{https://researchintegrityjournal.biomedcentral.com/articles/10.1186/s41073-019-0063-9}
-
-  \begin{itemize}
-  \item
-    Are OPR lessons transferable to CODECHECK? Does CODECHECK work
-    better with OPR than with (single/double) blind peer review?
-  \item
-    similarities to their advice: open reports, give credit
-  \end{itemize}
-\item
-  \emph{Successes and struggles with computational reproducibility:
-  Lessons from the Fragile Families Challenge} \textgreater{}
-  \url{https://osf.io/preprints/socarxiv/g3pdb/}
-\item
-  community standards for software, cf.~slide 6 in
-  \url{https://figshare.com/articles/_/5675104}
-
-  \begin{itemize}
-  \item
-    ``good enough practice'' is ``get a colleague to try using it'',
-    slide 19 in \url{https://figshare.com/articles/_/5675104}
-  \end{itemize}
-\item
-  journals doing code reviews
-
-  \begin{itemize}
-  \item
-    \emph{``We know of 16 journals that take this valuable step (ie TOP
-    Level 3): AEA as already mentioned, plus others in polisci,
-    psychology, biostats, and org chem: see Data Rep Policies here
-    \url{https://t.co/plOF8j6ADU}''}
-    (\url{https://twitter.com/EvoMellor/status/1202692360456589339?s=09});
-    see also whole thread!
-  \item
-    \url{https://medium.com/@NeurIPSConf/call-for-papers-689294418f43}
-  \item
-    \url{https://www.journals.elsevier.com/information-systems/editorial-board/}
-  \item
-    SIGMOD Reproducibility Review?
-  \item
-    Biostatistics ``AER''
-    (\url{https://academic.oup.com/biostatistics/article/10/3/405/293660})
-  \end{itemize}
-\item
-  open science platforms/cloud services \textgreater{}
-  \url{https://www.nature.com/articles/d41586-019-03366-x}
-\item
-  live code in articles \textgreater{} o2r, eLife RDS
-  \url{https://www.nature.com/articles/d41586-019-00724-7}, \ldots{}
-\item
-  Artefact Evaluation - \url{https://www.artifact-eval.org/about.html}
-\item
-  continuous analysis - \url{https://www.nature.com/articles/nbt.3780}
-\item
-  ``A Universal Identifier for Computational Results'' \textgreater{}
-  \url{https://www.sciencedirect.com/science/article/pii/S1877050911001256}
-\item
-  Discussion about using Docker containers during JOSS reviews
-  \textgreater{}
-  \url{https://github.com/openjournals/joss/issues/498\#issuecomment-462046912}
-\item
-  \url{https://scigen.report}, an independent site for registering
-  reproductions, via
-  \url{https://annakrystalli.me/talks/ro-reprohack.html\#32}
-
-  \begin{itemize}
-  \item
-    example
-    \url{https://scigen.report/reviews/get?doi=10.1063/1.1823034}
-  \item
-    seems to be closed infrastructure
-  \end{itemize}
-\item
-  new journal \emph{ReScienceX} (about experiments) also seems to have
-  an ECR angle, \url{https://www.nature.com/articles/d41586-020-01328-2}
-\item
-  really interesting work on reproductions in hydrology, with a neat
-  survey that reproducers filled out which might be interesting to
-  capture from codecheckers, too:
-  \url{https://www.nature.com/articles/sdata201930\#Sec9}
-\end{itemize}
-
-\section*{CODE CHECK implementation concepts}\label{code-check-implementation-concepts}
-
-\begin{itemize}
-\item
-  need to get wording straight: process/implementation/workflow (DN:
-  suggest to use ``workflow'' for research-analysis workflow)
-\item
-  \url{https://codecheck.org.uk/process/}
-\item
-  \emph{dimensions} along which a check can vary
-\item
-  independent/pre-submission (Peer\_Coding) by author + colleagues
-\item
-  AGILE
-\item
-  premier (OA) publisher
-\item
-  community OA / scholarly society with public peer review
-\item
-  invited reproducibility paper track,
-  cf.~\url{https://www.elsevier.com/journals/information-systems/0306-4379/guide-for-authors}
-\item
-  compare with existing processes that are publicly documented
-
-  \begin{itemize}
-  \item
-    AJPS Verification Policy:
-    \url{https://ajps.org/ajps-verification-policy/}
-  \item
-    \ldots{}
-  \end{itemize}
-\item
-  flexible level of detail? a check can be enhanced to (potentially
-  semi-automatically) check for good practices
-
-  \begin{itemize}
-  \item
-    ``A group of us in the ({\textbf{???}}) Software Citation
-    Implementation Working Group produced a''Software Citation Checklist
-    for Authors" \url{https://t.co/4Noe66FsoX} providing guidelines on
-    how to cite software and describing why it is important and what
-    software should be cited" via
-    \url{https://twitter.com/alegonbel/status/1207664922932453376?s=09})
-  \end{itemize}
-\item
-  focus on ECR as codecheckers
-
-  \begin{itemize}
-  \item
-    ECR are the one's breaking new ground in schol comm, cf.~the
-    percentages of career stages submitting registered reports on slide
-    20 \textgreater{} \url{https://osf.io/2vjus/}
-  \end{itemize}
-\end{itemize}
-
-\section*{Annotated Examples}\label{annotated-examples}
-
-\begin{itemize}
-\item
-  \textbf{the default implementation}
-\item
-  critically discuss our/the default implementation vs.~the principles
-\item
-  HPC example
-\item
-  if we see the corpus of examles as a scientific dataset, it should be
-  published in a citable way (maybe even submit to a data journal)
-\end{itemize}
-
-\section*{Open problems}\label{open-problems}
-
-\begin{itemize}
-\item
-  Do publishers need to build up data science experts who can conduct
-  checks?
-\item
-  Is an integration of software/data citation checks
-  benefitial/realistic?
-\item
-  Bot-support and tools (linter, report authoring with R Markdown)
-\item
-  How can open platforms be used to enable a review in private
-  (cf.~\url{https://www.cambridge.org/core/blog/2019/08/19/how-to-make-the-data-and-code-for-your-manuscript-available-to-peer-reviewers-before-making-it-public/}),
-  if that is desired
-\item
-  Does nudging towards better practices work (highlight the successes)
-  or is any check with opt-in doomed to fail?
-\item
-  Do we need more than one codechecker? Is there a similar diversity as
-  in a peer review where more opinions are needed?
-\item
-  can CODECHECK work for preprints? how? is a re-check required?
-\item
-  Risk: is the hurdle too low? are ``levels'' of checks are solution to
-  increase transparency of check's level of detail
-\item
-  CODECHECK in the context of innovations and disruptions in scientific
-  publishing
-
-  \begin{itemize}
-  \item
-    \url{https://doi.org/10.1042/ETLS20180172}
-  \item
-    \url{https://doi.org/10.3390/publications7020034}
-  \end{itemize}
-\item
-  Should every code be checkd, or should the labour be capped by only
-  checking ``important'' works or at important stages; if yes, isn't
-  peer review such a stage?
-\item
-  see \url{https://github.com/codecheckers/discussion}
-\item
-  Can we establish code checks as a type of review within the schol comm
-  metadata (or should we?) \textgreater{} we should not, put them on the
-  same level as reviews!
-\item
-  How can we turn a CODECHECK into a positive experience?
-
-  \begin{itemize}
-  \item
-    cf.~kindness in peer review in
-    \url{https://www.mdpi.com/2304-6775/8/2/26}
-  \end{itemize}
-\item
-  How can we allow reactions to checks? (commenting on reports)
-\item
-  Will bitwise reproducibility become easier than what a CODECHECK does?
-  triggered by
-  \url{https://twitter.com/khinsen/status/1242842759733665799?s=09})
-\item
-  If a report is not published for failures: Isn't there a bias for the
-  codechecker to have a ``successful'' reproduction because then a
-  report is published?
-\end{itemize}
-
 \end{document}

### 2020-12-14 codecheck-paper.tex
add further reference, try to fix long URL in footnote, fix URLs

@@
 \usepackage{hyperref}
-\usepackage{url}
+\usepackage[hyphenbreaks]{breakurl}
+\usepackage[hyphens]{url}
 \usepackage{setspace}
@@
 \item revisit \url{https://github.com/codecheckers/discussion}
+\item 
 \end{itemize}
@@
 \url{https://elifesciences.org/labs/dc5acbde/welcome-to-a-new-era-of-reproducible-publishing}.},
+decentralized infrastructure and public reviewer reputation systems
+\cite{tenorio-fornes_towards_2019},
 and completely new visions for scholarly communication and peer review
@@
 research initiatives such as
-ReproHack\footnote{https://reprohack.github.io/reprohack-hq/},
-ReproducibiliTea\footnote{https://reproducibilitea.org/}
+ReproHack\footnote{\url{https://reprohack.github.io/reprohack-hq/}},
+ReproducibiliTea\footnote{\url{https://reproducibilitea.org/}}
 \cite{fitzgibbon_brewing_2020},
-and Repro4Everyone\footnote{https://repro4everyone.org/}
+and Repro4Everyone\footnote{\url{https://repro4everyone.org/}}
 \cite{auer_reproducibility_2020},

### 2020-12-14 codecheck-paper.tex
fix rendering
add two more references and fix some citations

@@
 \usepackage{hyperref}
-\usepackage[hyphenbreaks]{breakurl}
-\usepackage[hyphens]{url}
+\usepackage{url}
 \usepackage{setspace}
@@
 \item revisit \url{https://github.com/codecheckers/discussion}
-\item 
 \end{itemize}
@@
 are more tailored for scientific publications while others offer commercial
-options to, e.g., publishers \cite{}.
+options to, e.g., publishers (cf. \cite{konkol_publishing_2020}).
 Using such cloud-based 
@@
 The topic of blinding is broadly discussed, not the least in the push towards
-open peer review as part of the Open Science movement [REF]. Without taking
-a strong stand on this discussion, the general motivation for more 
-transparency and reproducibility does indeed favour a more open review 
-process. However, anonymity can also serve to protect individuals 
+open peer review as part of the Open Science movement 
+(cf. \cite{ross-hellauer_guidelines_2019}).
+Without taking a strong stand on this discussion, the general motivation 
+for more  transparency and reproducibility does indeed favour a more open 
+review process.
+However, anonymity can also serve to protect individuals 
 \ref{tennant_limitations_2020}, e.g.,  juniour scientists, whereas it might 
@@
 the authors. Stark~\cite{stark_before_2018} calls this `preproducibility'
-and the ICERM report \cite{} defines the level ``Auditable Research'' 
-similarly. Note that openness is not required here, as to accommodate 
+and the ICERM report \cite{stodden_setting_2013} defines the level
+``Auditable Research''  similarly.
+Note that openness is not required here, as to accommodate 
 communities of practice and review workflows that (currently) do not want
@@
 % `https://docs.google.com/document/d/1fMWrFvBTwAYXg5J_bYIjgb-gY7DfQ0DOjx9Vlg_Rbmg/edit# ?
-
 \emph{Information Systems}, with an invited reproducibility paper track
 \footnote{\url{https://www.elsevier.com/journals/information-systems/0306-4379/guide-for-authors}}
+
+Another journal initiative is more indirectly connected with reproductions.
+\emph{Nature Machine Intelligence} recently introduced a new type of article,
+the reusability report \cite{noauthor_research_2020}.
+Inspired by the detailed and nuanced submissions to a reproducibility 
+challenge, the reusability report focuses on the exploration of robustness
+and generalizability of the original paper's claims
+\cite{noauthor_research_2020}. This answers the specific community's 
+challenges around computational reproducibility and also values these kind
+of contributions as independent publications, which goes beyond the goals 
+of CODECHECK.
 
@@
 \footnote{\url{https://paperswithcode.com/rc2020}}.
+% Papers with Code now also does more disciplines than just ML.
+Likewise, different disciplines and journals provide reproducibility
+checklists, e.g., \cite{rosenberg_next_2020}, which naturally share some
+aspects while addressing particularities as well as addressing researchers
+from different fields. Regarding the education and guidance for authors, 
+we see CODECHECK's role in referencing and linking educational efforts
+and helpful material, not as a creator and maintainer of content.
 
@@
 pressure to publish, puts the scholarly publication system under a lot
-of strain [REF].  We also see parallels to the terms 
+of strain.  We also see parallels to the terms 
 `reproducible research' and `open science'. Establishing a CODECHECK 

### 2020-12-14 codecheck-paper.tex
fix label and URLs as footnotes

@@
 
-\section*{Implementation}\label{register}
+\section*{Implementation}\label{implementation}
 
@@
 \footnote{\url{https://twitter.com/Code_Copilot}}, or Papers with Code
-\footenote{\url{https://paperswithcode.com/about}}.
+\footnote{\url{https://paperswithcode.com/about}}.
 The latter recently announced a collaborations with the preprint server
@@
 
-The journal \emph{ReScience C} publishes replications,
-\url{https://rescience.github.io/faq/};
+The journal \emph{ReScience C} publishes replications
+\footnote{\url{https://rescience.github.io/faq/}};
 they do not accept proprietary software and require open source code;

### 2020-12-14 codecheck-paper.tex
add Yale YARD

@@
 \href{https://github.com/OxfordCodeReviewNet/forum}{Oxford code review network (OxCRN)}
-(community, broader scope).
+(community, broader scope), \href{https://isps.yale.edu/research/data/approach}{YARD 
+at Yale} with dedicated staff for replications.
 For these, the CODECHECK scope, tools, and practical experiences may be

### 2020-12-18 codecheck-paper.tex
add HDSR to todo list
codecheck can help enforce journal policies

@@
 \item revisit \url{https://github.com/codecheckers/discussion}
+\item \url{https://hdsr.mitpress.mit.edu/pub/f0obb31j/release/1} and other articles form the HDSR special issue
 \end{itemize}
@@
 of the computational results.
+Nevertheless, the simple attempt to reproduce may highlight shortcomings in
+a submission with respect to transparency as it may be required by a journal's
+policy and help to resolve misunderstandings and discrepancies
+(cf.~\cite{christian_journal_2020}).
 

### 2021-01-05 codecheck-paper.tex
add possible source
adding a couple of old notes (forgot to commit)

@@
 \item \url{https://hdsr.mitpress.mit.edu/pub/f0obb31j/release/1} and other articles form the HDSR special issue
+\item \url{https://av.tib.eu/media/42484} > "Code will make systematic mistakes, not random mistakes", "code is more reliable (and auditable) than a human"
+\item \url{https://doi.org/10.1016/j.isci.2020.101698} - "Rigor and Transparency Index may serve as a proxy for methodological quality" - CODECHECK doesn't evaluate that, could be used for distinction
+\item Make code Feedback and leave better than found a principle, not an option?
+Just checking the code does not give enough credit (?), but improving it certainly does, and good peer review improves (watch out to not do full software review though)
 \end{itemize}
@@
 
-The time for a CODECHECK greatly depends on the intentions and the ascribed
+The point in time when a CODECHECK happens greatly depends on the 
+intentions and the ascribed
 importance. The earlier a CODECHECK happens in the process, the more the
@@
 However, anonymity can also serve to protect individuals 
-\ref{tennant_limitations_2020}, e.g.,  juniour scientists, whereas it might 
+\cite{tennant_limitations_2020}, e.g.,  juniour scientists, whereas it might 
 be acceptable to require revealing of the names of tenured researchers.
@@
 process that should in any case be fixable and not a question of opinion or
-faulty approach. It is possible for a nonsense workflow to receive a 
-CODECHECK certificate.
+faulty approach. It is technically possible for a nonsense workflow to receive
+a CODECHECK certificate.
+If passing a CODECHECK becomes mandatory, the full transparency may have to be 
+revisited as the relations between authors and codecheckers fall under the
+same social and acommunity challenges as non-anonymous peer review
+(cf.~\cite{everythinghertz123}).
+% previous sentenc inspired by https://osf.io/9cftx/
 Furthermore, the level of documentation that is needed for third parties
@@
 \url{https://science-octopus.org/about}, or Hypergraph
-\url{https://www.libscie.org/hypergraph}.}, there is a 
+\url{https://www.libscie.org/hypergraph}.},
+% mimosa is another platform idea, but no real prototype AFAICT:
+% https://projects.invisionapp.com/share/E9Z3RKE7W3P#/screens/436978924
+% https://docs.google.com/presentation/d/1V_K8hghgnvGfEtW7TdwtowTveZyy4Qs8rMemNyHx_Dg/edit#slide=id.p
+there is a 
 value in an evolutionary approach. Thereby, existing communities can 
@@
 
-\textbf{Handling failure}: we have no established process for the case
+\textbf{Handling failure}: We have no established process for the case
 that a reproduction fails. This may be the sign of a pre-selection bias
@@
 to share the outcome publicly but to share the latest draft of the 
-certific with the authors.
-One must be careful as to the impact of acceptance by authors and the
-concerns of volunteers this approach, or a contrary one, has.
+certificate with the authors.
+A negative comment in a CODECHECK certificate or a failed check does
+not necessarily mean the paper or the science is bad
+(cf. discussion on negative comments in \cite{everythinghertz123}).
+Therefore, one must be careful as to the impact of acceptance by authors 
+and the concerns of volunteers this approach, or a contrary one, has.
 Private feedback for failures and public acknowledgement/encouragement
 for successes seems to be the way to go.
+% eLife's new model (see https://osf.io/9cftx/) will publish reports evantually, e.g., after preprint has been accepted elsewhere
 \cite{Rosenthal2016b} discuss incentives for different actors around the
@@
 \textbf{Who's got time for more peer review?} Agree.
+Ignoring the complex community contract behind volunteered peer review,
+the time spent with CODECHECKs likely reduces the individual's time to work
+as a reviewer.
 However, there are differences to peer review.

### 2021-01-07 codecheck-paper.tex
about to make edits after talking to Daniel

--- 2020-12-18 codecheck-paper.tex
@@
 
-\begin{itemize}
+\begin{enumerate}
 \item Check with Giuliano re: ``code works''. Write to Naomi.
@@
 \item \url{https://hdsr.mitpress.mit.edu/pub/f0obb31j/release/1} and other articles form the HDSR special issue
-\end{itemize}
+\item \url{https://av.tib.eu/media/42484} > "Code will make systematic mistakes, not random mistakes", "code is more reliable (and auditable) than a human"
+\end{enumerate}
 
@@
 extended to the ACM journal \emph{Transactions on Modeling and Computer 
-Simulation (TOMACS)}\footnote{\url{https://dl.acm.org/journal/tomacs/author-guidelines#rcrini}}.
+  Simulation (TOMACS)}
+%% SJE TODO - fix footnote
+%% \footnote{\url{https://dl.acm.org/journal/tomacs/author-guidelines-%35-rcrini}}.
 

### 2021-01-07 codecheck-paper.tex
enough to send this out

@@
 \begin{document}
-\title{CODECHECK: An open-science initiative to facilitate sharing of
-  computer programs and results presented in scientific publications.}
+\title{CODECHECK: An open-science initiative for the independent execution of computations underlying research articles.}
 \author[1,$\ast$]{Daniel N\"{u}st}
@@
 \begin{abstract}
+  This is a draft paper that we hope to submit to F1000R (and perhaps
+  as a preprint before) in the coming weeks.
   \ldots{}
@@
 \item Check with Giuliano re: ``code works''. Write to Naomi.
-\item Use "execution" insteand of "sharing" in title?
-\item Can we introduce "research compendium" as a term?
-\item Should we make code and data sharing a prerequisite or leave that door open?
+\item Can we introduce "research compendium" as a term?  (yes).
+\item Should we make code and data sharing a prerequisite or leave
+  that door open? (make it the norm)
 \item Take a look at the Lancet's statistical review process (areas for improvement by authors) \url{https://www.thelancet.com/journals/lancet/article/PII0140-6736(92)90409-V/fulltext}
 \item How to relate to "reproducibility does not matter but trust does", see \url{https://gmbecker.github.io/MayInstituteKeynote2019/outline.html}
-\item Should we ask the people currently listed in the acknowledgements?
-\item revisit \url{https://github.com/codecheckers/discussion}
-\item \url{https://hdsr.mitpress.mit.edu/pub/f0obb31j/release/1} and other articles form the HDSR special issue
-\item \url{https://av.tib.eu/media/42484} > "Code will make systematic mistakes, not random mistakes", "code is more reliable (and auditable) than a human"
+\item Check with people  in the acknowledgements
+%\item revisit \url{https://github.com/codecheckers/discussion}
+% \item \url{https://hdsr.mitpress.mit.edu/pub/f0obb31j/release/1} and other articles form the HDSR special issue
+% \item \url{https://av.tib.eu/media/42484} > "Code will make systematic mistakes, not random mistakes", "code is more reliable (and auditable) than a human"
 \end{enumerate}
@@
 To complement the traditional route of sharing research by writing papers,
-there is a growing trend/demand to share the underlying artifacts, notably 
+there is a growing trend/demand to share the underlying artefacts, notably 
 code and datasets, so that others can inspect, reproduce or expand that work
@@
 
-So, assuming that researchers begin to share more artifacts, how might
+So, assuming that researchers begin to share more artefacts, how might
 they be examined to test that they do what they claim? For example,
@@
 other end of the spectrum, we still witness today the infamous
-statement ``Data available upon reasonable request'' whereby authors,
+statement ``Data available upon reasonable request''. Authors,
 while possible well-intentioned at the time of writing the article,
 often cannot provide the data when readers ask for
-it.  (ANDREW H: any references to sum up poor/evolving state of data
-sharing; Vines et al 2014).
+it as data disappears over time \cite{Vines2014-hf}.
 
@@
 together are required to validate outcomes of a computational analysis,
-the do differ in the respect that the former can be seen as static/inert,
-while the latter requires an environment and interaction. This makes 
+they do differ in  that data can be seen as static/inert,
+while code requires an environment and interaction. This makes 
 software harder to share.
@@
 want to maintain it / I do not want to give away my code to my
-competitors''. Our view, rooted in Open Science princples, is that
+competitors''. Our view, rooted in Open Science principles, is that
 wherever possible sharing code is good for the community, as outlined
-ten years ago \cite{Barnes2010-iv}. Having the code freely available, and
+in 2010 \cite{Barnes2010-iv}. Having the code freely available, and
 suitably archived, provides a valuable resource for others to learn
@@
 focused approach.
-Just like peer review can provide a ``baseline reasurance'', i.e., a paper
-has at least ben checked by someone with an understanding of the topic
+Just like peer review can provide a ``baseline reassurance'', i.e., a paper
+has at least been checked by someone with an understanding of the topic
 \cite{fyfe_mission_2019}, we think the same baseline should be provided
@@
   half of the diagram shows the diverse range of materials used
-  internally within a laboratory.  These materials are often then
-  condensed into a paper for sharing with the outside world via the
+  within a laboratory.  These materials are often then
+  condensed for sharing with the outside world via the
   research paper, a static PDF document.  Working backwards from the
-  PDF to the underlying materials is impossible. This probihibits reuse
+  PDF to the underlying materials is impossible. This prohibits reuse
   and is not only non-transparent for a specific paper, but also 
   ineffective for science as a whole. By sharing the
-  materials on the left, others outside the lab are able to reproduce
+  materials on the left, others outside the lab can reproduce
   or build on this work.}
@@
 groups of people: (1) the AUTHOR of a paper providing the code to be checked, (2)
-the PUBLISHER of a journal intereseted in publishing the paper by the
+the PUBLISHER of a journal interested in publishing the paper by the
 AUTHOR and (3) the CODECHECKER who checks that the AUTHOR's code works.
@@
   \caption{The CODECHECK example process implementation. Codechecker act as a detectives:
-  The investigate and record, but do not fix issues.}
-  \label{fig:worfklow}
+  They investigate and record, but do not fix issues.}
+  \label{fig:workflow}
 \end{figure}
@@
 analogous to the publisher finding one or more reviewers to evaluating
-the paper. However, a cruical difference is that we suggest the CODECHECKER
+the paper. However, a crucial difference is that we suggest the CODECHECKER
 and the author can talk freely to each other directly (see next).
@@
 reproduced. (See later for exact nature of ``reproduced''). If there are
-any problems running the code, the CODECHECKER asks the author for help
+any problems running the code, the CODECHECKER asks the AUTHOR for help
 to resolve the problems. The burden to provide reproducible material lies with
@@
 This process iterates until either the CODECHECKER is successful,
-or the CODECHECKER concludes the code does not work (\emph{what do in this
-case}). As part of this process, the CODECHECKER could work entirely
+or the CODECHECKER concludes the code does not work (\emph{TODO NOT
+  YET HANDLED}). As part of this process, the CODECHECKER could work entirely
 locally on their own compute resource, or in the cloud, e.g.~using the open
@@
 the code and enforces a much more complete definition of the computing environment,
-but, unless infrastructure is provided by, e.g., by the publisher, requires code
+but, unless secure infrastructure is provided by, e.g., by the publisher, requires code
 and data to be published openly online.
-Note that the taks of the CODECHECKER is to check the ``mechanics'' of the
+Note that the task of the CODECHECKER is to check the ``mechanics'' of the
 workflow, not to evaluate its correctness. 
-Stodden et~al. \cite{stodden_setting_2013} destinguish between verification and
+Stodden et~al. \cite{stodden_setting_2013} distinguish between verification and
 validation in the context of Mathematics, and in the same vein a CODECHECK
@@
 
-\textbf{Step 5:} the certificate as well as auxiliary files created during the check,
-e.g., a specifiction of a computing environment, helper scripts, and the code and data,
-if available under open licenses, get deposited in an open archive (currently Zenodo);
-the certificate is given to the PUBLISHER.
+\textbf{Step 5:} the certificate as well as auxiliary files created
+during the check, e.g., a specification of a computing environment,
+helper scripts, and the code and data, if available under open
+licenses, get deposited in an open archive (currently Zenodo); the
+certificate is given to the PUBLISHER.
 
@@
 activity appropriately in scholarly publication databases, such as ORCID.
-The PULISHER also ensures proper publication metadata, e.g., links from the 
+The PUBLISHER also ensures proper publication metadata, e.g., links from the 
 certificate repository to the published paper or the original code repository.
@@
 \subsubsection*{Dimensions of CODECHECK workflows}\label{dimensions-of-workflows}
+
+\textbf{Daniel to fix by merging turnaround time into other
+  dimensions, and labelling these as the four dimensions: When?
+  Openness? Who? Importance?}
 
@@
 a regular CODECHECK, without any special treatment for some submissions.
-A CODECHECK could then be awared equal weight as the other reviews.
+A CODECHECK could then be awarded equal weight as the other reviews.
 A post acceptance CODECHECK puts the least requirements on the author and may
@@
 submission. This is the least strict solution while still properly 
-acknowleding the check results, because the CODECHECK is to be completed 
+acknowledging the check results, because the CODECHECK is to be completed 
 before the publication of the final paper.
@@
 codechecker has a matching code \emph{and} domain expertise, for example to
-check a workflow based on Python code analysing MRT brain imagery. However,
+check a workflow based on Python code analysing neuroimaging datasets. However,
 a well-documented workflow should be executable by any person with a basic
 understanding of computers. Naturally, the more prerequisite knowledge the
-codechecker has, the less time is spend to understand the goals and 
+codechecker has, the less time is needed to understand the goals and 
 the mechanics of an analysis. From our experiences, the priority should be
@@
 understanding.
-The time and expertise alloted to the check
+The time and expertise allotted to the check
 are the main drivers for the depth of checking, though in general, we 
@@
 modelled in review platforms. Keeping diverse specialists as staff to
-realise a codechking process is unlikely to be feasible, especially for 
+realise a codechecking process is unlikely to be feasible, especially for 
 small or independent publishers.
@@
 require much debate or opinion. Depending on the power of the CODECHECK
-(see \ref{when-to-do-a-codecheck}), a mediator, e.g., the editor, or a
+(see TODO: check ref \ref{when-to-do-a-codecheck}), a mediator, e.g., the editor, or a
 second opinion might be warranted in rare cases.
@@
 In fact, ECRs are often much more familiar with latest technology and are 
-likely to be early adoptors as creators, i.e., be authors of CODECHECK-ready manuscripts, too\footnote{Cf. the 77\% of 141 registered reports who were 
+likely to be early adopters as creators, i.e., be authors of CODECHECK-ready manuscripts, too\footnote{Cf. the 77\% of 141 registered reports who were 
 submitted to leading Neuroscience journals by ECRs
@@
 We see a high potential in setting up the new processes for CODECHECKING with
-a clear committment on openness and transparency, independent of the
+a clear commitment on openness and transparency, independent of the
 respective current peer review process (see \nameref{who-knows-who}).
 
-Publisher staff represents the most controlled yet also not quickly scalable
+Publisher staff represents the most controlled but also resource-intensive
 option. Hiring the required technical and or domain expertise puts a financial
-burden on the publisher, of course, but such a committment also shows the 
+burden on the publisher, of course, but such a commitment also shows the 
 value given to publishing reproducibility and would positively impact the 
@@
 In contrast, for researchers it can be very important to be publicly credited
-with their activity as a reviewer (see \nameref{who-knows-who}). A regular
+with their activity as a reviewer (see TODO CHECK REF \nameref{who-knows-who}). A regular
 review may be listed in public databases (e.g., ORCID
@@
 is of course a very coarse indicator for the community contribution by an 
-individual, yet is is unclear if a regular reviewer, who in addition
+individual, yet it is unclear if a regular reviewer, who in addition
 conducts a CODECHECK to the scientific review, should be credited twice.
@@
 (cf. \cite{ross-hellauer_guidelines_2019}).
-Without taking a strong stand on this discussion, the general motivation 
+Without taking a strong stance on this discussion, the general motivation 
 for more  transparency and reproducibility does indeed favour a more open 
@@
 However, anonymity can also serve to protect individuals 
-\ref{tennant_limitations_2020}, e.g.,  juniour scientists, whereas it might 
+\cite{tennant_limitations_2020}, e.g.,  junior scientists, whereas it might 
 be acceptable to require revealing of the names of tenured researchers.
@@
 rework the docs  or the code.
-Instead of lenghtily describing a small change, the codechecker may simply
+Instead of lengthily describing a small change, the codechecker may simply
 provide a fix within the code.
@@
 
-As a prerequisite, we see a CODECHECK as part of a typical scholarly
-communication process. While more radical ideas to change the paradigms
-how researchers share their work exist\footnote{For example Octopus, 
+We see a CODECHECK as part of a typical scholarly communication
+process. While more radical ideas to change the paradigms how
+researchers share their work exist\footnote{For example Octopus,
   \url{https://science-octopus.org/about}, or Hypergraph
-\url{https://www.libscie.org/hypergraph}.}, there is a 
-value in an evolutionary approach. Thereby, existing communities can 
-transition towards more open practices at their own pace and ensure
-no one is left behind unfairly.
-When integrating a CODECHECK in existing review and publication processes,
-the \emph{turnaround time} is crucial. Depending on when and who conducts
-the check, it may be done faster, or add considerably to the overall
-time from submission to publication.
-In our first checks, we found that a codecheck may generally take from 2-5
-hours, with some outliers on the higher end. This time includes writing and
-publishing the certificate, but excludes the time for actual computations.
-The efforts include considerable time to communicate about the process, 
-especially around who publishes what and when, so that proper 
-cross-referencing between paper and certificate is ensured.
-When integrated into peer review systems, this handling of the steps should
-become much more effective.
+  \url{https://www.libscie.org/hypergraph}.}, there is a value in an
+evolutionary approach. Thereby, existing communities can transition
+towards more open practices at their own pace and ensure no one is
+left behind unfairly.  When integrating a CODECHECK in existing review
+and publication processes, the \emph{turnaround time} is
+crucial. Depending on when and who conducts the check, it may be done
+faster, or add considerably to the overall time from submission to
+publication.  In our first checks, we found that a codecheck may
+generally take from 2-5 hours, with some outliers on the higher
+end. This time includes writing and publishing the certificate, but
+excludes the time for actual computations.  The efforts include
+considerable time to communicate about the process, especially around
+who publishes what and when, so that proper cross-referencing between
+paper and certificate is ensured.  When integrated into peer review
+systems, this handling of the steps should become much more effective.
 
@@
 
-\textbf{2. Communication between humans is key.} Some code may just work
-without any interaction (e.g.~Manuel's certificate TODO add cert number); 
-however, more often
-there are hidden dependencies that need to be found, or revised software
+\textbf{2. Communication between humans is key.} Some code may just
+work without any interaction (e.g. CODECHECK certificate 2020-013
+\cite{cert-2020-013}); however, more often there are hidden
+dependencies that need to be found, or revised software
 installations. Allowing the codechecker and human to communicate
@@
 (possibly anonymously) through a publisher would slow things down and
-reduce the chances for community building as well as the lessons learned
-for both author and codecheckers.
+reduce the chances for community building as well as the lessons
+learned for both author and codecheckers.
 
 \textbf{3. Credit is given to codecheckers.} The value of contributing a
-CODECHECK is comparable to providing a scientific peer review. Indiviual
+CODECHECK is comparable to providing a scientific peer review. Individual
 variations aside, it may also require a similar amount of time. Therefore,
@@
 to or are not able to work openly.
-A completed check eans that code could be executed at least once without
+A completed check means that code could be executed at least once without
 critical errors or
@@
 of CODECHECK. The idea was to take very well known historical articles from a 
-specific domain, in this case the work area of co-author S.~Eglen, possibly 
+specific domain, in this case the work area of co-author SJE, possibly 
 as as a conversation starter for potential collaborating journals.
@@
 %Several COVID-19 models have been done during the time of peer review of these papers,
-and under our initative rather than requested from journal, but the certificate
-has then been acknowledged/cited in the paper (LANCET/LSHTM). % TODO SJE: clarify references in brackets
-The third theme represents co-author D.~N\"ust's service as Reproducibility
+and under our initiative rather than requested from journal, but the certificate
+has then been acknowledged/cited in the paper \cite{Davies2020-vj}.
+The third theme represents co-author DN's service as Reproducibility
 Reviewer at the AGILE conference series.
@@
 series \cite{nust_improving_2020}.
-While using slightly different terms and infrastructue, e.g., ``reproducibility
+While using slightly different terms and infrastructure, e.g., ``reproducibility
 reports'' are published on OSF instead of certificates on Zenodo, the 
@@
 
+\textbf{TODO: SJE, tidy this paragraph.}
 \begin{enumerate}
@@
   Gigascience -- our first paper.  Only visualisation, not machine
-  learning (which would have taken sseveral days of compute time).
+  learning (which would have taken several days of compute time).
 \item
@@
     (Simulation code was written specifically for this project.)\\
-    2020-004  \cite{cert-2020-004} & area & Cart-pole balancing problem solved using reinforcement learning \cite{Barto1983-rg}.  Code written for this reimplementation.\\
-    2020-005  \cite{cert-2020-005} & area & Larisch \\
-    2020-006  \cite{cert-2020-006} & area & Detorakis \\
-    2020-007  \cite{cert-2020-007} & area & in progress \\
+    2020-004  \cite{cert-2020-004} & neuroscience & Cart-pole balancing problem solved using reinforcement learning \cite{Barto1983-rg}.  Code written for this reimplementation.\\
+    2020-005  \cite{cert-2020-005} & neuroscience & Larisch \\
+    2020-006  \cite{cert-2020-006} & neuroscience & Detorakis \\
+    2020-007  \cite{cert-2020-007} & neuroscience & \textbf{in progress} \\
     2020-008  \cite{cert-2020-008} & COVID-19 & Modelling of
@@
     2020-012  \cite{cert-2020-012} & COVID-19 & Modelling of COVID-19 spread across the USA; preprint currently review \cite{Unwin2020}. \\
-    2020-013  \cite{cert-2020-013} & area & biorXiv \cite{Spitschan2020.06.02.129502}. \\
-    2020-014  \cite{cert-2020-014} & area & eLife \cite{Sadeh2020}. \\
-    2020-015  \cite{cert-2020-015} & area & eLife \cite{Liou2020}. \\
+    2020-013  \cite{cert-2020-013} & neuroscience & biorXiv \cite{Spitschan2020.06.02.129502}. \\
+    2020-014  \cite{cert-2020-014} & neuroscience & eLife \cite{Sadeh2020}. \\
+    2020-015  \cite{cert-2020-015} & neuroscience & eLife \cite{Liou2020}. \\
     2020-016  \cite{cert-2020-016} & GIScience & 
@@
 the human resources for managing the project and processes, which are
-considerable and currently soley based on, partly grant-based, public funding.
+considerable and currently solely based on, partly grant-based, public funding.
 Naturally, similar set-ups are possible, such as using GitLab/GitLab.com 
@@
 The journal \emph{ACM Transactions on Mathematical Software (TOMS)} 
-established a "Replicated Computational Results" (RCR) review proces
+established a "Replicated Computational Results" (RCR) review process
 \cite{heroux_editorial_2015} (using "replicable" with the same meaning 
@@
 1 to 4.
-This work provides more extensive discussion of the possible variants, 
+Our CODECHECK provides more extensive discussion of the possible variants, 
 while TOMS RCR realises one particular workflow for reproducibility 
@@
 the reproducing person, while the possible conflict of interest of this 
-motivation is acknowledged. TOMS also reconises reviewer activity in a 
+motivation is acknowledged. TOMS also recognises reviewer activity in a 
 partnership with Publons
@@
 journals or events, and to build a community of codecheckers. CODECHECK shares
-this interdisciplinary nature with other community iniatives concered with 
+this interdisciplinary nature with other community initiatives concerned with 
 reproducibility awareness, education, and supports, such as ReproHack
@@
 For research with very high stakes, where reproduction would be too weak and
-post publication replication possibly to late because of policiy impact,
+post publication replication possibly too late because of policy impact,
 \cite{benjamin-chung_internal_2020} propose \emph{internal replication}.
@@
 
-\textbf{Just peer review}: wait, isn't CODECHECK what peer review should be 
-doing already? That is correct.
-However, the general observation is that the increasing
-number of researchers, submissions, and reviews to conduct under the
-pressure to publish, puts the scholarly publication system under a lot
-of strain.  We also see parallels to the terms 
+\textbf{Isn't CODECHECK what peer review should be doing already?}
+That is correct.  However, the general observation is that the
+increasing number of researchers, submissions, and reviews to conduct
+under the pressure to publish, puts the scholarly publication system
+under a lot of strain.  We also see parallels to the terms
 `reproducible research' and `open science'. Establishing a CODECHECK
 process is largely an acknowledgement that peer reviewing practices
-have faltered to adapt to the challenges of digitisation and computer-based
-research or data science. The concept of a CODECHECK, just as the 
-reproducibility of research and the openness of science, may be of 
-transitional nature. If the activites described here as being part of a 
-CODECHECK become the norm for scientific reviews and publications, then
-the initiative will have succeeded.
-However, the rapidly changing technology used in research leads us to 
-believe that a seperate role of codechecker and the expertise it brings
-will be benefitial in the long term.
-
-\textbf{Not going deep enough/too vague}:
+have faltered to adapt to the challenges of digitisation and
+computer-based research or data science. The concept of a CODECHECK,
+just as the reproducibility of research and the openness of science,
+may be of transitional nature. If the activities described here as
+being part of a CODECHECK become the norm for scientific reviews and
+publications, then the initiative will have succeeded.  However, the
+rapidly changing technology used in research leads us to believe that
+a separate role of codechecker and the expertise it brings will be
+beneficial in the long term.
+
+\textbf{CODECHECK requirements are not demanding enough}:
 CODECHECK by design does not require authors to provide (and sustain) an
-eternally functional worklow and neither suggests a specific software stack
-or pracitcal approaches, such as research compendia
+eternally functional workflow and neither suggests a specific software stack
+or practical approaches, such as research compendia
 \footnote{\url{https://research-compendium.science/}},
@@
 collaboratively created by authors and codecheckers is a snapshot of a 
-working analysis that greatly increases the likelyhood of a successful 
+working analysis that greatly increases the likelihood of a successful 
 reproduction and possibility to extensions by third parties, with access
@@
 Initiatives and journals that conduct software reviews independently of 
-a specific publication or workfloware, e.g.,
+a specific publication or workflow include
 ROpenSci\footnote{\url{https://ropensci.org/}},
@@
 
-Furthermore, the codechecker's task list is intentionally not overloaded
-with related concerns, even though they are very important, such as ensuring
-proper citation of data and software, or depositing data and software in 
-suitablerepositories.
-We do expect codecheckers to be aware of these though and possibly point 
-them out.
+Furthermore, the codechecker's task list is intentionally not
+overloaded with related concerns, even though they are very important,
+such as ensuring proper citation of data and software, or depositing
+data and software in suitable repositories. We do expect codecheckers
+to be aware of these though and possibly highlight them.
 
@@
 to share the outcome publicly but to share the latest draft of the 
-certific with the authors.
+certificate with the authors.
 One must be careful as to the impact of acceptance by authors and the
@@
 as codecheckers.
-The challegnes of computational resources can also quickly be reduced when
+The challenges of computational resources can also quickly be reduced when
 authors provide a small sample or a synthetic dataset together with the
-expected outputs for said subset, which can equally wel demonstrate the 
+expected outputs for this small example, which can equally well demonstrate the 
 usefulness and validity of their workflow.
@@
 proprietary software. Given the prevalence of proprietary software in
-some disciplines, such as MATLAB or eCognition,
-this seems like a pragmatic choice.
+some disciplines, e.g MATLAB, this seems like a pragmatic choice.
 However, it prohibits to benefit from open infrastructure for reproducibility
@@
 
-\textbf{Who's got time for more peer review?} Agree.
+\textbf{Who's got time for more peer review?} We agree.
 However, there are differences to peer review.
@@
 a software professional employed by a publisher.
-Third, recall that we expect CODECHECKSs to be not anonymous and always
+Third, recall that we expect CODECHECKs to be not anonymous and always
 publicly deposited, which is common practice for some journals, but not
@@
 \textbf{Who finds the codechecker?} Good point! The journal, using either
-their own team of codecheckers or the public database. Community
+their own team of codecheckers or the public database. Researchers
 however think this is an important issue, and are volunteering.
-With increasing demand
 
 \textbf{Checking multiple times?} If you check a preprint before peer
-review or a submission to a journal with non-public review,
-it might need checking again as it is modified during peer
-review.  This is in part unavoidable, and as happened to us (e.g., 
-`2020-012` has been revised). This can also be desired, if the
-interaction between author, reviewer, and codechecker lead to
-improvements.
-Since checking the manuscript the second time is likely to be
-much less work than the first time, this is largely mitigated by 
-open and regular communication.
-The higher the reproducibility and automation, the more aspects of 
-a CODECHECK may eventually be supported by automation, though we believe
-the human codecheckers should remain the decision maker.
-Automatically detecting if results are ``the same'' is a question
-intrinsically linked with each unique process and quite hard to realise
-for common outputs, such as plots or maps.
-Nevertheless, continuous integration (CI) infrastructure may help 
-codecheckers, and, without much further work, use scientific workflows as
-detectors for when a piece of software breaks if a workflow is set up to
-use the latest versions of tools.
+review or a submission to a journal with non-public review, it might
+need checking again as it is modified during peer review.  This is in
+part unavoidable, and as happened to us (e.g., certificate 2020-012
+\cite{cert-2020-012} has been revised). This can also be desired, if
+the interaction between author, reviewer, and codechecker lead to
+improvements.  Since checking the manuscript the second time is likely
+to be much less work than the first time, this is largely mitigated by
+open and regular communication.  The higher the reproducibility and
+automation, the more aspects of a CODECHECK may eventually be
+supported by automation, though we believe the human codecheckers
+should remain the decision maker.  Automatically detecting if results
+are ``the same'' is a question intrinsically linked with each unique
+process and quite hard to realise for common outputs, such as plots or
+maps.  Nevertheless, continuous integration (CI) infrastructure may
+help codecheckers, and, without much further work, use scientific
+workflows as detectors for when a piece of software breaks if a
+workflow is set up to use the latest versions of tools.
 
@@
 files that the author claims. Stochastic simulations mean that often
-we cannot get exactly the same, or even different versions of
-libraries.  (Cite freesurfer example \cite{Gronenschild2012-pp} that
-generates different outputs on different operating systems).
+we cannot get exactly the same, or even the same versions of libraries
+can generate outputs that differ by operating system
+\cite{Gronenschild2012-pp}.
 
@@
 
-SJE is on the senior editorial board at the journal \emph{Scientific
-Data}.
-DN is reproducibility editor at the Association of Geographic Information 
-Laboratories in Europe's (AGILE) annual conference.
+SJE is on the editorial board at the journal \emph{Scientific Data}.
+DN is reproducibility editor at the Association of Geographic
+Information Laboratories in Europe's (AGILE) annual conference.
 
@@
 \end{document}
+
+% LocalWords:  CODECHECK CODECHECKER Zenodo ORCID ECRs codecheck

### 2021-01-07 codecheck-paper.tex
Naomi

@@
 the work presented here: Andy Collings, Melissa Harrison, Giuliano
-Maciocci (eLife), Rebecca Kirk (PLOS Computational Biology), Scott
+Maciocci, Naomi Penfold (eLife), Rebecca Kirk (PLOS Computational Biology), Scott
 Edmunds (GigaScience), and Andrew Hufton (Scientific Data). Iain Davies and

### 2021-01-07 codecheck-paper.tex
this version is good enough to send out for comment/checking acks

@@
   that door open? (make it the norm)
-\item Take a look at the Lancet's statistical review process (areas for improvement by authors) \url{https://www.thelancet.com/journals/lancet/article/PII0140-6736(92)90409-V/fulltext}
-\item How to relate to "reproducibility does not matter but trust does", see \url{https://gmbecker.github.io/MayInstituteKeynote2019/outline.html}
-\item Check with people  in the acknowledgements
+% \item How to relate to "reproducibility does not matter but trust does", see \url{https://gmbecker.github.io/MayInstituteKeynote2019/outline.html}
+\item Check people  are okay being listed in acknowledgements
 %\item revisit \url{https://github.com/codecheckers/discussion}
 % \item \url{https://hdsr.mitpress.mit.edu/pub/f0obb31j/release/1} and other articles form the HDSR special issue
-% \item \url{https://av.tib.eu/media/42484} > "Code will make systematic mistakes, not random mistakes", "code is more reliable (and auditable) than a human"
+% \item \url{https://av.tib.eu/media/42484} > "Code will make
+% systematic mistakes, not random mistakes", "code is more reliable
+% (and auditable) than a human"
+% \item Take a look at the Lancet's statistical review process (areas for improvement by authors) \url{https://www.thelancet.com/journals/lancet/article/PII0140-6736(92)90409-V/fulltext}
+
 \end{enumerate}

### 2021-01-07 codecheck-paper.tex

--- 2021-01-07 codecheck-paper.tex
@@
 \begin{enumerate}
-\item Check with Giuliano re: ``code works''. Write to Naomi.
-\item Can we introduce "research compendium" as a term?  (yes).
-\item Should we make code and data sharing a prerequisite or leave
-  that door open? (make it the norm)
+\item Check with Giuliano re: ``code works''.
+\item Can we introduce "research compendium" as a term?
+\item Should we make code and data sharing a prerequisite or leave that door open?
 \item Take a look at the Lancet's statistical review process (areas for improvement by authors) \url{https://www.thelancet.com/journals/lancet/article/PII0140-6736(92)90409-V/fulltext}
 \item How to relate to "reproducibility does not matter but trust does", see \url{https://gmbecker.github.io/MayInstituteKeynote2019/outline.html}
-\item Check with people  in the acknowledgements
-%\item revisit \url{https://github.com/codecheckers/discussion}
-% \item \url{https://hdsr.mitpress.mit.edu/pub/f0obb31j/release/1} and other articles form the HDSR special issue
-% \item \url{https://av.tib.eu/media/42484} > "Code will make systematic mistakes, not random mistakes", "code is more reliable (and auditable) than a human"
-\end{enumerate}
+\item Should we ask the people currently listed in the acknowledgements?
+\item revisit \url{https://github.com/codecheckers/discussion}
+\item \url{https://hdsr.mitpress.mit.edu/pub/f0obb31j/release/1} and other articles form the HDSR special issue
+\item \url{https://av.tib.eu/media/42484} > "Code will make systematic mistakes, not random mistakes", "code is more reliable (and auditable) than a human"
+\item \url{https://doi.org/10.1016/j.isci.2020.101698} - "Rigor and Transparency Index may serve as a proxy for methodological quality" - CODECHECK doesn't evaluate that, could be used for distinction
+\item Make code Feedback and leave better than found a principle, not an option?
+Just checking the code does not give enough credit (?), but improving it certainly does, and good peer review improves (watch out to not do full software review though)
+\end{itemize}
 
@@
 
-We see a CODECHECK as part of a typical scholarly communication
-process. While more radical ideas to change the paradigms how
-researchers share their work exist\footnote{For example Octopus,
+As a prerequisite, we see a CODECHECK as part of a typical scholarly
+communication process. While more radical ideas to change the paradigms
+how researchers share their work exist\footnote{For example Octopus, 
 \url{https://science-octopus.org/about}, or Hypergraph
-  \url{https://www.libscie.org/hypergraph}.}, there is a value in an
-evolutionary approach. Thereby, existing communities can transition
-towards more open practices at their own pace and ensure no one is
-left behind unfairly.  When integrating a CODECHECK in existing review
-and publication processes, the \emph{turnaround time} is
-crucial. Depending on when and who conducts the check, it may be done
-faster, or add considerably to the overall time from submission to
-publication.  In our first checks, we found that a codecheck may
-generally take from 2-5 hours, with some outliers on the higher
-end. This time includes writing and publishing the certificate, but
-excludes the time for actual computations.  The efforts include
-considerable time to communicate about the process, especially around
-who publishes what and when, so that proper cross-referencing between
-paper and certificate is ensured.  When integrated into peer review
-systems, this handling of the steps should become much more effective.
+\url{https://www.libscie.org/hypergraph}.},
+% mimosa is another platform idea, but no real prototype AFAICT:
+% https://projects.invisionapp.com/share/E9Z3RKE7W3P#/screens/436978924
+% https://docs.google.com/presentation/d/1V_K8hghgnvGfEtW7TdwtowTveZyy4Qs8rMemNyHx_Dg/edit#slide=id.p
+there is a 
+value in an evolutionary approach. Thereby, existing communities can 
+transition towards more open practices at their own pace and ensure
+no one is left behind unfairly.
+When integrating a CODECHECK in existing review and publication processes,
+the \emph{turnaround time} is crucial. Depending on when and who conducts
+the check, it may be done faster, or add considerably to the overall
+time from submission to publication.
+In our first checks, we found that a codecheck may generally take from 2-5
+hours, with some outliers on the higher end. This time includes writing and
+publishing the certificate, but excludes the time for actual computations.
+The efforts include considerable time to communicate about the process, 
+especially around who publishes what and when, so that proper 
+cross-referencing between paper and certificate is ensured.
+When integrated into peer review systems, this handling of the steps should
+become much more effective.
 
@@
 certificate with the authors.
-One must be careful as to the impact of acceptance by authors and the
-concerns of volunteers this approach, or a contrary one, has.
+A negative comment in a CODECHECK certificate or a failed check does
+not necessarily mean the paper or the science is bad
+(cf. discussion on negative comments in \cite{everythinghertz123}).
+Therefore, one must be careful as to the impact of acceptance by authors 
+and the concerns of volunteers this approach, or a contrary one, has.
 Private feedback for failures and public acknowledgement/encouragement
 for successes seems to be the way to go.
+% eLife's new model (see https://osf.io/9cftx/) will publish reports evantually, e.g., after preprint has been accepted elsewhere
 \cite{Rosenthal2016b} discuss incentives for different actors around the
@@
 
-\textbf{Who's got time for more peer review?} We agree.
+\textbf{Who's got time for more peer review?} Agree.
+Ignoring the complex community contract behind volunteered peer review,
+the time spent with CODECHECKs likely reduces the individual's time to work
+as a reviewer.
 However, there are differences to peer review.

### 2021-01-08 codecheck-paper.tex
fix latex error
questions marks denote waiting to hear from person

@@
 Just checking the code does not give enough credit (?), but improving it certainly does, and good peer review improves (watch out to not do full software review though)
-\end{itemize}
+\end{enumerate}
 
@@
 
+% question mark indicates we are waiting for person to respond.
 We are grateful to the following individuals for discussions regarding
-the work presented here: Andy Collings, Melissa Harrison, Giuliano
-Maciocci, Naomi Penfold (eLife), Rebecca Kirk (PLOS Computational Biology), Scott
-Edmunds (GigaScience), and Andrew Hufton (Scientific Data). Iain Davies and
-Yuhao (Sebastian) Wang developed code and example certificates.
-We thank Antonio P\'{a}ez (Journal of Geographical Systems) for enabling
-CODECHECKs, Carlos Granell and Frank Ostermann for contributing
+the work presented here: Andy Collings?, Melissa Harrison?, Giuliano
+Maciocci, Naomi Penfold? (eLife), Rebecca Kirk? (PLOS Computational Biology), Scott
+Edmunds (GigaScience), and Andrew Hufton? (Scientific Data). Iain Davies? and
+Yuhao (Sebastian) Wang? developed code and example certificates.
+We thank Antonio P\'{a}ez? (Journal of Geographical Systems) for enabling
+CODECHECKs, Carlos Granell? and Frank Ostermann? for contributing
 certificates as reproducibility reviewers at the AGILE conference, and 

### 2021-01-08 codecheck-paper.tex
few more acks confirmed

@@
 We are grateful to the following individuals for discussions regarding
-the work presented here: Andy Collings?, Melissa Harrison?, Giuliano
-Maciocci, Naomi Penfold? (eLife), Rebecca Kirk? (PLOS Computational Biology), Scott
-Edmunds (GigaScience), and Andrew Hufton? (Scientific Data). Iain Davies? and
-Yuhao (Sebastian) Wang? developed code and example certificates.
+the work presented here: Andy Collings, Melissa Harrison?, Giuliano
+Maciocci, Naomi Penfold? (eLife), Rebecca Kirk (PLOS Computational Biology), Scott
+Edmunds (GigaScience), and Andrew Hufton? (Scientific Data). Iain Davies and
+Yuhao (Sebastian) Wang developed code and example certificates.
 We thank Antonio P\'{a}ez? (Journal of Geographical Systems) for enabling
-CODECHECKs, Carlos Granell? and Frank Ostermann? for contributing
+CODECHECKs, Carlos Granell and Frank Ostermann? for contributing
 certificates as reproducibility reviewers at the AGILE conference, and 

### 2021-01-10 codecheck-paper.tex
Naomi okay -- possibly add Emmy?

@@
     2020-008  \cite{cert-2020-008} & COVID-19 & Modelling of
-    interventions on COVID-19 cases in the UK; checked as preprint (NOW REMOVED XXX)
+    interventions on COVID-19 cases in the UK; checked as preprint
     and now published \cite{Davies2020-vj} \\
@@
 the work presented here: Andy Collings, Melissa Harrison?, Giuliano
-Maciocci, Naomi Penfold? (eLife), Rebecca Kirk (PLOS Computational Biology), Scott
-Edmunds (GigaScience), and Andrew Hufton? (Scientific Data). Iain Davies and
-Yuhao (Sebastian) Wang developed code and example certificates.
-We thank Antonio P\'{a}ez? (Journal of Geographical Systems) for enabling
-CODECHECKs, Carlos Granell and Frank Ostermann? for contributing
-certificates as reproducibility reviewers at the AGILE conference, and 
-all authors of auditable workflows for their participation.
+Maciocci, Naomi Penfold, Emmy Tsang? (eLife), Rebecca Kirk (PLOS
+Computational Biology), Scott Edmunds (GigaScience), and Andrew
+Hufton? (Scientific Data). Iain Davies and Yuhao (Sebastian) Wang
+developed code and example certificates.  We thank Antonio P\'{a}ez?
+(Journal of Geographical Systems) for enabling CODECHECKs, Carlos
+Granell and Frank Ostermann? for contributing certificates as
+reproducibility reviewers at the AGILE conference, and all authors of
+auditable workflows for their participation.
 

### 2021-01-10 codecheck-paper.tex
emmy ok

@@
 the work presented here: Andy Collings, Melissa Harrison?, Giuliano
-Maciocci, Naomi Penfold, Emmy Tsang? (eLife), Rebecca Kirk (PLOS
+Maciocci, Naomi Penfold, Emmy Tsang (eLife), Rebecca Kirk (PLOS
 Computational Biology), Scott Edmunds (GigaScience), and Andrew

### 2021-01-11 codecheck-paper.tex
more acks confirmed

--- 2021-01-08 codecheck-paper.tex
@@
 whether they use the actual name for it or not.
-The CODECHECK badge is, intentionally, not branded and simply states
-``code works''.
+The CODECHECK badge is, intentionally, not branded beyond the checkmark
+and green colour and simply states ``code works''.
 This \emph{cultural change}, however, is needed for the valuation of the
@@
 the work presented here: Andy Collings, Melissa Harrison?, Giuliano
-Maciocci, Naomi Penfold? (eLife), Rebecca Kirk (PLOS Computational Biology), Scott
+Maciocci, Naomi Penfold (eLife), Rebecca Kirk (PLOS Computational Biology), Scott
 Edmunds (GigaScience), and Andrew Hufton? (Scientific Data). Iain Davies and
 Yuhao (Sebastian) Wang developed code and example certificates.
-We thank Antonio P\'{a}ez? (Journal of Geographical Systems) for enabling
-CODECHECKs, Carlos Granell and Frank Ostermann? for contributing
+We thank Antonio P\'{a}ez (Journal of Geographical Systems) for enabling
+CODECHECKs, Carlos Granell and Frank Ostermann for contributing
 certificates as reproducibility reviewers at the AGILE conference, and 

### 2021-01-11 codecheck-paper.tex
update table
add 024 and 025, update all description texts, update references

@@
 Furthermore, a small number of checks was completed as part of a peer reviews
-for a GIScience journals.
+for GIScience journals.
 
@@
 \begin{table}
+  \footnotesize
+
   \centering
 
-  \begin{tabular}{llp{11cm}}
+  \begin{tabular}{llp{12cm}}
     \toprule
-    Certificate & Research area & Description \\ \midrule
-    2020-001 \cite{cert-2020-001} &
-    Machine learning & CODECHECK undertaken post acceptance of
-    manuscript and before its publication in \textit{Gigascience}
-    \cite{Piccolo2020-lo}. \\
-    2020-002  \cite{cert-2020-002} & neuroscience & Paper from 1997
- showing unsupervised learning from natural images \cite{Hancock1992-mp}.
- Code written for this project. \\
-    2020-003  \cite{cert-2020-003} & neuroscience &  Classic paper
-                                                     from 1982 on
-                                                     models of
-                                                     associative
-                                                     memory
-                                                     \cite{Hopfield1982-mz}.
-    (Simulation code was written specifically for this project.)\\
-    2020-004  \cite{cert-2020-004} & neuroscience & Cart-pole balancing problem solved using reinforcement learning \cite{Barto1983-rg}.  Code written for this reimplementation.\\
-    2020-005  \cite{cert-2020-005} & neuroscience & Larisch \\
-    2020-006  \cite{cert-2020-006} & neuroscience & Detorakis \\
-    2020-007  \cite{cert-2020-007} & neuroscience & \textbf{in progress} \\
-    2020-008  \cite{cert-2020-008} & COVID-19 & Modelling of
-    interventions on COVID-19 cases in the UK; checked as preprint (NOW REMOVED XXX)
-    and now published \cite{Davies2020-vj} \\
-    2020-009  \cite{cert-2020-009} & COVID-19 & tracing \\
-    2020-010  \cite{cert-2020-010} & COVID-19 & report 9 \\
-    2020-011  \cite{cert-2020-011} & COVID-19 & Modelling of Covid
-    spread across Europe;  paper was checked when in press (now  published) \cite{Flaxman2020-yb}. \\
-    2020-012  \cite{cert-2020-012} & COVID-19 & Modelling of COVID-19 spread across the USA; preprint currently review \cite{Unwin2020}. \\
-    2020-013  \cite{cert-2020-013} & neuroscience & biorXiv \cite{Spitschan2020.06.02.129502}. \\
-    2020-014  \cite{cert-2020-014} & neuroscience & eLife \cite{Sadeh2020}. \\
-    2020-015  \cite{cert-2020-015} & neuroscience & eLife \cite{Liou2020}. \\
-    2020-016  \cite{cert-2020-016} & GIScience & 
-    R models demonstrating the Modifiable Aral Unit Problem (MAUP) in spatial data science \cite{Brunsdon2020}. \\
-    2020-017  \cite{cert-2020-017} & GIScience & 
-    Spatial data handling, analysis, and visualisation using a variety of R packages \cite{Bivand2020}. \\
-    2020-018  \cite{cert-2020-018} & GIScience & 
-    Cellular automaton in R for modeling dynamic phenomena within a Discrete Global Grid System (DGGS) data model; using a subset for demonstration \cite{Hojati2020}. \\
-    2020-019  \cite{cert-2020-019} & GIScience & Reachability analysis of suburban transportation using analysis of privately owned shared cars; workflow uses Python, and reproduction with a considerably subsampled dataset \cite{Illium2020}.
+    \textbf{Certificate} & \textbf{Research area} & \textbf{Description} \\ \midrule
+    2020-001 \cite{cert-2020-001} & Machine learning & Code for benchmarking ML classification tool checked post acceptance of manuscript and before its publication in \textit{Gigascience} \cite{Piccolo2020-lo}.
 \\
-    2020-020  \cite{cert-2020-020} & GIScience & In-database windows operators for processing spatio-temporal data streams based on Python and PostgreSQL/PostGIS \cite{Werner2020}; reproduction using container.
+    2020-002  \cite{cert-2020-002} & Neuroscience & Code written for this project checked by second project member as demonstration using paper from 1997 showing unsupervised learning from natural images \cite{Hancock1992-mp}.
 \\
-    2020-021  \cite{cert-2020-021} & GIScience & Python code for comparing supervised machine learning models for spatial nominal entity recognition. \cite{Medad2020}. \\
-    2020-022  \cite{cert-2020-022} & GIScience & Python code for visualising text analysis on intents and concepts from geo-analytic questions \cite{Xu2020}. \\
-    2020-023  \cite{cert-2020-023} & GIScience & Analysis of spatial footprints of geo-tagged extreme weather events from social media using R \cite{Owuor2020}. \\
+    2020-003  \cite{cert-2020-003} & Neuroscience &  Code written for this project checked by second project member as demonstration using classic paper on models of associative memory \cite{Hopfield1982-mz}.
+\\
+    2020-004  \cite{cert-2020-004} & Neuroscience & Code written for this project checked by second project member as demonstration using classic paper on cart-pole balancing problem \cite{Barto1983-rg}.
+\\
+    2020-005  \cite{cert-2020-005} & Neuroscience & Check of independent reimplementation of spike-timing-dependent plasticity (STDP) model \cite{larisch_re_2019} conducted as demonstration for this paper.
+\\ % Python ANNarchy
+    2020-006  \cite{cert-2020-006} & Neuroscience & Check of independent reimplementation of a generalized linear integrate-and-fire neural model \cite{detorakis_re_2017} conducted as demonstration for this paper.
+\\
+    2020-007  \cite{cert-2020-007} & Neuroscience & Check of independent reimplementation of analysing spike patterns of neurons \cite{hathway_re_2018} conducted as demonstration for this paper. \\
+    2020-008  \cite{cert-2020-008} & COVID-19 & Code for modelling of interventions on COVID-19 cases in the UK checked at preprint stage \cite{davies-preprint-2020} and later published \cite{Davies2020-vj}.
+\\
+    2020-009  \cite{cert-2020-009} & COVID-19 & Code for analysis of effectiveness of measures to reduce transmission of SARS-CoV-2 checked as preprint \cite{kucharski-preprint-2020} and later published \cite{kucharski_effectiveness_2020}.
+\\ % R, check mentiond in acks, but not linked
+    2020-010  \cite{cert-2020-010} & COVID-19 & Code for analysis of non-pharmaceutical interventions (Report 9) checked as a preprint \cite{ferguson_report_2020}. \\ % CovidSim
+    2020-011  \cite{cert-2020-011} & COVID-19 & Code for modelling of COVID-19 spread across Europe was provided by authors and checked while paper was in press \cite{flaxman_estimating_2020}. \\
+    2020-012  \cite{cert-2020-012} & COVID-19 & Code for modelling of COVID-19 spread across the USA was checked as preprint \cite{unwin_report_2020} and later published \cite{unwin_state-level_2020}.
+\\
+    2020-013  \cite{cert-2020-013} & Neuroscience & Code for analsis of rest-activity patterns in people wihout con-mediated vision was checked as a preprint \cite{Spitschan2020.06.02.129502} after direct contact with the authors.
+\\ % MATLAB
+    2020-014  \cite{cert-2020-014} & Neuroscience & Code for analysis of pertubation patterns of neural activity was checked after publication as part of publisher collaboration \cite{Sadeh2020}.
+\\ % MATLAB
+    2020-015  \cite{cert-2020-015} & Neuroscience & Code for a neural network model for human focal seizures was checked after publication as part of publisher collaboration \cite{Liou2020}.
+\\ % MATLAB
+    2020-016  \cite{cert-2020-016} & GIScience & Code for models demonstrating the Modifiable Aral Unit Problem (MAUP) in spatial data science \cite{Brunsdon2020} was checked during peer review.
+\\ % R
+    2020-017  \cite{cert-2020-017} & GIScience & Code for spatial data handling, analysis, and visualisation using a variety of R packages \cite{Bivand2020} was checked after peer review before publication.
+\\ % R
+    2020-018  \cite{cert-2020-018} & GIScience & AGILE conference reproducibility report using a demonstration data subset with cellular automaton for modeling dynamic phenomena \cite{Hojati2020}.
+\\ % R
+    2020-019  \cite{cert-2020-019} & GIScience & AGILE conference reproducibility report with subsampled dataset for reachability analysis of suburban transportation using shared cars \cite{Illium2020}.
+\\
+    2020-020  \cite{cert-2020-020} & GIScience & AGILE conference reproducibility report using a container for checking in-database windows operators for processing spatio-temporal data \cite{Werner2020}.
+\\ % Python + PostGIS
+    2020-021  \cite{cert-2020-021} & GIScience & AGILE conference reproducibility report checking code for comparing supervised machine learning models for spatial nominal entity recognition \cite{Medad2020}.
+\\ % Python
+    2020-022  \cite{cert-2020-022} & GIScience & AGILE conference reproducibility report checking code for visualising text analysis on intents and concepts from geo-analytic questions \cite{Xu2020}.
+\\ % Python
+    2020-023  \cite{cert-2020-023} & GIScience & AGILE conference reproducibility report on analysis of spatial footprints of geo-tagged extreme weather events from social media \cite{Owuor2020}. \\
+    2020-024  \cite{cert-2020-024} & Neuroscience & Code for multi-agent system for concept drift detection in electromyography \cite{vieira_driftage_2020} was checked during peer review. \\ % R
+    2020-025  \cite{cert-2020-025} & GIScience & Adaptation and application of Local Indicators for Categorical Data (LICD) to archaeological data \cite{carrer_application_2021} was checked after peer review before publication.
+\\ % R
     \\ \bottomrule
@@
 pages of the certificate \texttt{2020-012} \cite{cert-2020-012}.
-The referenced article \cite{Unwin2020} describes the modelling COVID-19
+The referenced article \cite{unwin_report_2020} describes the modelling COVID-19
 across the USA.
@@
 codecheck (Figure~\ref{fig:cert}F).  In this case, the figure matches
-figure~4 of \cite{Unwin2020}.  Subsequent pages of the certificate show other
-outputs (not shown here).
+figure~4 of \cite{unwin_report_2020}.
+Subsequent pages of the certificate show other outputs (not shown here).
 

### 2021-01-11 codecheck-paper.tex
update dimensions figure and corresponding text

@@
 computational reproducibility \cite{barba_terminologies_2018} into peer
-review, and a demonstration of its feasibility by means of over 20
+review, and a demonstration of its feasibility by means of 25
 reproductions across different scientific disciplines.
@@
 
-\textbf{Daniel to fix by merging turnaround time into other
-  dimensions, and labelling these as the four dimensions: When?
-  Openness? Who? Importance?}
-
 The example workflow leaves room for different
@@
 
-\subsubsection*{When to do a CODECHECK and with what power}
+\subsubsection*{When to do a CODECHECK and with what importance?}
 \label{when-to-do-a-codecheck}
@@
 The point in time when a CODECHECK happens greatly depends on the 
-intentions and the ascribed
-importance. The earlier a CODECHECK happens in the process, the more the
-results can play a role in the most important decision in scholarly 
-communication: is a paper published, sent back for revisions, or rejected.
+intentions and the ascribed importance.
+The earlier a CODECHECK happens in the process, the more the
+results can play a role in editorial decisions:
+is a paper published, sent back for revisions, or rejected.
+Such a \textbf{strict prerequisite} can reduce the workload of reviewers,
+and therefore seems out of place late in the review process.
 The later in a review process the check happens, the less challenging is it
@@
 
-A pre-review CODECHECK allows editors to use it as a filter to only send 
-a submission out to peer review if it passes the check. A less strict
-pre-review check would simply be included in the submission package provided
+We see a CODECHECK as part of a typical scholarly communication process.
+While more radical ideas to change the paradigms
+how researchers share their work exist\footnote{For example Octopus, 
+\url{https://science-octopus.org/about}, or Hypergraph
+\url{https://www.libscie.org/hypergraph}.},
+% mimosa is another platform idea, but no real prototype AFAICT:
+% https://projects.invisionapp.com/share/E9Z3RKE7W3P#/screens/436978924
+% https://docs.google.com/presentation/d/1V_K8hghgnvGfEtW7TdwtowTveZyy4Qs8rMemNyHx_Dg/edit#slide=id.p
+there is a value in an evolutionary approach.
+Thereby, existing communities can 
+transition towards more open practices at their own pace and ensure
+no one is left behind unfairly.
+When integrating a CODECHECK in existing review and publication processes,
+the \emph{turnaround time} is crucial. Depending on when and who conducts
+the check, it may be done faster, or add considerably to the overall
+time from submission to publication.
+In our first checks, we found that a codecheck may generally take from 2-5
+hours, with some outliers on the higher end. This time includes writing and
+publishing the certificate, but excludes the time for actual computations.
+These efforts include considerable time to communicate about the process, 
+especially around who publishes which document when, so that proper 
+cross-referencing between paper and certificate is ensured via persistent
+identifiers.
+When integrated into a peer review system, this handling of documents should
+become much more effective and reduce organisational efforts.
+
+A \textbf{pre-review} CODECHECK allows editors to use it as a filter to only
+send  a submission out to peer review if it passes the check.
+The certificate would  be included in the submission package provided
 to the reviewers, so they can include the CODECHECK's results in their 
 evaluation of the work.
-A CODECHECK may also be conducted in parallel to the scientific review. This
-puts less burden on the turnaround time for the check yet only makes the 
+Reviewers may then judge the the relevance of the computations for the results
+of the work, or follow journal guidelines as how strict to apply the check's
+outcomes.
+
+A CODECHECK may also be conducted in \textbf{parallel} to the scientific review.
+This puts less burden on the turnaround time for the check, yet only makes the 
 outcomes available for the final consideration of the handling editor.
@@
 a regular CODECHECK, without any special treatment for some submissions.
-A CODECHECK could then be awarded equal weight as the other reviews.
-A post acceptance CODECHECK puts the least requirements on the author and may
-simply lead to an award of excellence on top of the acceptance of the
-submission. This is the least strict solution while still properly 
-acknowledging the check results, because the CODECHECK is to be completed 
+A CODECHECK could then be awarded \textbf{equal weight} as the other reviews.
+
+A \textbf{post acceptance} CODECHECK puts the least requirements on the author 
+and may simply lead to an award of excellence on top of the acceptance of the
+submission. This \textbf{extra merit} is the least strict solution while still
+properly  acknowledging the check results,
+because the CODECHECK is to be completed 
 before the publication of the final paper.
@@
 was successfully reviewed with a badge on the volume and article landing
-pages.
-
-Finally, a CODECHECK may also be conducted after publication, though it
-would required to update the article later with a note to reference the check,
-otherwise readers would presumably not be aware of it. While this has the
-least impact on the current publishing practices, it also has the smallest
-effect and does not acknowledge the importance of reproducible workflows for
+pages. Similarly, the GIScience papers in collaborations with journals were 
+checked while the authors worked on revisions.
+
+Finally, a CODECHECK may also be conducted \textbf{post publication}, though it
+would required to update the article and article metadata later to reference 
+the check, otherwise readers would presumably not be aware of it. In general,
+publishers are hesitant to make such a revision of published articles.
+While this has the least impact on the current publishing practices and also
+enables open discourse,
+it also has the smallest effect on the original submission
+and does not acknowledge the importance of reproducible workflows for
 ensuring good scientific practice.
 
-\subsubsection*{Who does the CODECHECK?}\label{who-does-the-codecheck}
-
-The options for the person who can conduct a CODECHECK require a not quite
-simple matching of a combination of skills and availability. Ideally, the
-codechecker has a matching code \emph{and} domain expertise, for example to
-check a workflow based on Python code analysing neuroimaging datasets. However,
-a well-documented workflow should be executable by any person with a basic
-understanding of computers. Naturally, the more prerequisite knowledge the
-codechecker has, the less time is needed to understand the goals and 
-the mechanics of an analysis. From our experiences, the priority should be
-given to matching technical expertise first, as the a lack of knowledge in
-setting  up a computing environment with a particular language or tool is 
-much more discouraging and the assessment of the outcome, e.g., comparing
-created figures with the original, is largely possible without 
-understanding.
-The time and expertise allotted to the check
-are the main drivers for the depth of checking, though in general, we 
-expect a CODECHECK not to evaluate performance, e.g., does a racing car
-have the best apex speed, but to assess operation free of failure, i.e., 
-is it a car with 4 wheels and at least one door, and does it move forwards
-when accelerating. This depth can be seen independent from the power of
-the check (see above), as long as it is communicated clearly to all
-stakeholders, most importantly authors and readers.
-
-The concrete people to provide these expertises can range from researchers,
-possibly drawn from the regular pool of volunteer reviewers or from a 
-special group, to editorial staff of a publisher. If the regular reviewers
-are relied upon, the editor's task to find suitable ones becomes more
-tedious; extra information such as software familiarity would have to be
-modelled in review platforms. Keeping diverse specialists as staff to
-realise a codechecking process is unlikely to be feasible, especially for 
-small or independent publishers.
-
-We consider a single codechecker to be enough, unlike peer review where
-multiple referees are common. Because of the technical focus, the 
-determination if the concrete outputs of author and codechecker, who can
-collaborate to solve any issues, should be a very factual process and not
-require much debate or opinion. Depending on the power of the CODECHECK
-(see TODO: check ref \ref{when-to-do-a-codecheck}), a mediator, e.g., the editor, or a
-second opinion might be warranted in rare cases.
-
-We see a great opportunity to especially involve early-career researchers
-(ECRs) as codecheckers. They arguably have a high interest in learning
-about the latest tools and technologies as they are building up their own
-expertise and specialisation. The possibility to gain insights into latest
-science for ECRs and at the same time increase the appreciation of
-reproduction efforts is a sweet spot that CODECHECK shares
-with \emph{ReScience X}, a journal devoted to reproduction and replication 
-experiments \cite{roesch_new_2020}.
-In fact, ECRs are often much more familiar with latest technology and are 
-likely to be early adopters as creators, i.e., be authors of CODECHECK-ready manuscripts, too\footnote{Cf. the 77\% of 141 registered reports who were 
-submitted to leading Neuroscience journals by ECRs
-\cite{chambers_registered_2019}.}
-They are also introduced to peer review and 
-may transition into the role of scientific reviewer over time.
-Not to solve the general problems of peer review, which is largely an
-unsupervised process, i.e., rarely are reviewers taught how to do it, but
-a junior codechecker may be supported by a senior codechecker, too.
-We see a high potential in setting up the new processes for CODECHECKING with
-a clear commitment on openness and transparency, independent of the
-respective current peer review process (see \nameref{who-knows-who}).
-
-Publisher staff represents the most controlled but also resource-intensive
-option. Hiring the required technical and or domain expertise puts a financial
-burden on the publisher, of course, but such a commitment also shows the 
-value given to publishing reproducibility and would positively impact the 
-submission experience, because of more controllability, and external 
-perception of the journal or conference. A staff codechecker reduces the 
-challenges when it comes to open communication and awarding credit, as
-in-house personnel may be more readily be given access to non-anonymised
-submission material and receives a proper salary instead of public credit
-as part of the scientific community.
-
-In contrast, for researchers it can be very important to be publicly credited
-with their activity as a reviewer (see TODO CHECK REF \nameref{who-knows-who}). A regular
-review may be listed in public databases (e.g., ORCID
-\footnote{\href{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}} or commercial offerings such as Publons
-\footnote{\href{https://publons.com/}{https://publons.com/}} or 
-ReviewerCredits\footnote{\href{https://www.reviewercredits.com/}{https://www.reviewercredits.com/}}).
-A codechecker should be listed just the same. The number of scientific reviews
-is of course a very coarse indicator for the community contribution by an 
-individual, yet it is unclear if a regular reviewer, who in addition
-conducts a CODECHECK to the scientific review, should be credited twice.
-
-The codecheckers community\footnote{\href{https://github.com/codecheckers/codecheckers/}{https://github.com/codecheckers/codecheckers/}}
-currently has over 20 individuals who signed up as volunteer codecheckers
-following talks given on the project in the last 12 months.
-That is even though no checks are actively elicited.
-We see a high potential in an open shared list of potential codecheckers 
-compared to a private, in-house group, as it may better match expertise
-and share the workload.
-% TODO tell more about the motivations and commitments that we know about?
-The overarching community is also a mean to ensure CODECHECK is a viable
-option not just for large publishers, independent of their financial model,
-but also for independent no-cost open access journals.
-
-\subsubsection*{Who knows who?}\label{who-knows-who}
+\subsubsection*{Openness, or: Who knows who?}\label{who-knows-who}
 
@@
 
-\subsubsection*{How is the CODECHECK integrated in other processes?}\label{turnaround-time}
-
-As a prerequisite, we see a CODECHECK as part of a typical scholarly
-communication process. While more radical ideas to change the paradigms
-how researchers share their work exist\footnote{For example Octopus, 
-\url{https://science-octopus.org/about}, or Hypergraph
-\url{https://www.libscie.org/hypergraph}.},
-% mimosa is another platform idea, but no real prototype AFAICT:
-% https://projects.invisionapp.com/share/E9Z3RKE7W3P#/screens/436978924
-% https://docs.google.com/presentation/d/1V_K8hghgnvGfEtW7TdwtowTveZyy4Qs8rMemNyHx_Dg/edit#slide=id.p
-there is a 
-value in an evolutionary approach. Thereby, existing communities can 
-transition towards more open practices at their own pace and ensure
-no one is left behind unfairly.
-When integrating a CODECHECK in existing review and publication processes,
-the \emph{turnaround time} is crucial. Depending on when and who conducts
-the check, it may be done faster, or add considerably to the overall
-time from submission to publication.
-In our first checks, we found that a codecheck may generally take from 2-5
-hours, with some outliers on the higher end. This time includes writing and
-publishing the certificate, but excludes the time for actual computations.
-The efforts include considerable time to communicate about the process, 
-especially around who publishes what and when, so that proper 
-cross-referencing between paper and certificate is ensured.
-When integrated into peer review systems, this handling of the steps should
-become much more effective.
+\subsubsection*{Who does the CODECHECK?}\label{who-does-the-codecheck}
+
+The options for the person who can conduct a CODECHECK require a not quite
+simple matching of a combination of skills and availability. Ideally, the
+codechecker has a matching code \emph{and} domain expertise, for example to
+check a workflow based on Python code analysing neuroimaging datasets. However,
+a well-documented workflow should be executable by any person with a basic
+understanding of computers. Naturally, the more prerequisite knowledge the
+codechecker has, the less time is needed to understand the goals and 
+the mechanics of an analysis. From our experiences, the priority should be
+given to matching technical expertise first, as the a lack of knowledge in
+setting  up a computing environment with a particular language or tool is 
+much more discouraging and the assessment of the outcome, e.g., comparing
+created figures with the original, is largely possible without 
+understanding the science.
+The time and expertise allotted to the check
+are the main drivers for the depth of checking, though in general, we 
+expect a CODECHECK not to evaluate performance, e.g., does a racing car
+have the best apex speed, but to assess operation free of failure, i.e., 
+is it a car with 4 wheels and at least one door, and does it move forwards
+when accelerating. This depth can be seen independent from the power of
+the check (see above), as long as it is communicated clearly to all
+stakeholders, most importantly authors and readers.
+
+The concrete people to provide these expertises can range from researchers,
+possibly drawn from the regular pool of volunteer
+\textbf{scientific reviewers} or from a special group of
+\textbf{reproducibility reviewers}, via specific roles
+such as \textbf{reproducibility editors}, to experts working as part of
+\textbf{editorial staff} with a publisher. If the regular reviewers
+are relied upon, the editor's task to find suitable ones becomes more
+tedious; extra information such as software familiarity would have to be
+modelled in review platforms.
+
+We consider a \emph{single codechecker} to be enough, unlike peer review where
+multiple referees are common. Because of the technical focus, the 
+determination if the concrete outputs of author and codechecker, who can
+collaborate to solve any issues, actually match should be a very factual 
+process and not
+(see TODO: check ref \ref{when-to-do-a-codecheck}), a mediator, e.g., the editor, or a
+second opinion might be warranted in rare cases.
+
+We see a great opportunity to especially involve
+\emph{early-career researchers}
+(ECRs) as codecheckers. They arguably have a high interest in learning
+about the latest tools and technologies as they are building up their own
+expertise and specialisation. The possibility to gain insights into latest
+science for ECRs and at the same time increase the appreciation of
+reproduction efforts is a sweet spot that CODECHECK shares
+with \emph{ReScience X}, a journal devoted to reproduction and replication 
+experiments \cite{roesch_new_2020}.
+In fact, ECRs are often much more familiar with latest technology and are 
+likely to be early adopters as creators, i.e., be authors of CODECHECK-ready manuscripts, too\footnote{Cf. the 77\% of 141 registered reports who were 
+submitted to leading Neuroscience journals by ECRs
+\cite{chambers_registered_2019}.}
+They are also introduced to peer review and 
+may transition into the role of scientific reviewer over time.
+Not to solve the general problems of peer review, which is largely an
+unsupervised process, i.e., rarely are reviewers taught how to do it, but
+a junior codechecker may be supported by a senior codechecker, too.
+We see a high potential in setting up the new processes for CODECHECKING with
+a clear commitment on openness and transparency, independent of the
+respective current peer review process (see \nameref{who-knows-who}).
+
+Publisher staff represents the most controlled but also resource-intensive
+option. Hiring the required technical and or domain expertise puts a financial
+burden on the publisher, of course, but such a commitment also shows the 
+value given to publishing reproducibility and would positively impact the 
+submission experience, because of more controllability, and external 
+perception of the journal or conference. A staff codechecker reduces the 
+challenges when it comes to open communication and awarding credit, as
+in-house personnel may be more readily be given access to non-anonymised
+submission material and receives a proper salary instead of public credit
+as part of the scientific community.
+
+In contrast, for researchers it can be very important to be publicly credited
+with their activity as a reviewer (see TODO CHECK REF \nameref{who-knows-who}). A regular
+review may be listed in public databases (e.g., ORCID
+\footnote{\href{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}} or commercial offerings such as Publons
+\footnote{\href{https://publons.com/}{https://publons.com/}} or 
+ReviewerCredits\footnote{\href{https://www.reviewercredits.com/}{https://www.reviewercredits.com/}}).
+A codechecker should be listed just the same. The number of scientific reviews
+is of course a very coarse indicator for the community contribution by an 
+individual, yet it is unclear if a regular reviewer, who in addition
+conducts a CODECHECK to the scientific review, should be credited twice.
+
+The codecheckers community\footnote{\href{https://github.com/codecheckers/codecheckers/}{https://github.com/codecheckers/codecheckers/}}
+currently has over 20 individuals who signed up as volunteer codecheckers
+following talks given on the project in the last 12 months.
+That is even though no checks are actively elicited.
+We see a high potential in an open shared list of potential codecheckers 
+compared to a private, in-house group, as it may better match expertise
+and share the workload.
+% TODO tell more about the motivations and commitments that we know about?
+The overarching community is also a mean to ensure CODECHECK is a viable
+option not just for large publishers, independent of their financial model,
+but also for independent no-cost open access journals.
 

### 2021-01-11 codecheck-paper.tex
resolve TODOs

@@
 \end{abstract}
-
-\section*{TODO}\label{todo}
-
-\begin{enumerate}
-\item Check with Giuliano re: ``code works''.
-\item Can we introduce "research compendium" as a term?
-\item Should we make code and data sharing a prerequisite or leave that door open?
-\item Take a look at the Lancet's statistical review process (areas for improvement by authors) \url{https://www.thelancet.com/journals/lancet/article/PII0140-6736(92)90409-V/fulltext}
-\item How to relate to "reproducibility does not matter but trust does", see \url{https://gmbecker.github.io/MayInstituteKeynote2019/outline.html}
-\item Should we ask the people currently listed in the acknowledgements?
-\item revisit \url{https://github.com/codecheckers/discussion}
-\item \url{https://hdsr.mitpress.mit.edu/pub/f0obb31j/release/1} and other articles form the HDSR special issue
-\item \url{https://av.tib.eu/media/42484} > "Code will make systematic mistakes, not random mistakes", "code is more reliable (and auditable) than a human"
-\item \url{https://doi.org/10.1016/j.isci.2020.101698} - "Rigor and Transparency Index may serve as a proxy for methodological quality" - CODECHECK doesn't evaluate that, could be used for distinction
-\item Make code Feedback and leave better than found a principle, not an option?
-Just checking the code does not give enough credit (?), but improving it certainly does, and good peer review improves (watch out to not do full software review though)
-\end{enumerate}
 
@@
 process and not
-(see TODO: check ref \ref{when-to-do-a-codecheck}), a mediator, e.g., the editor, or a
+require much debate or opinion. Code usually makes systematic and repeatable
+mistakes is is thereby more reliable and auditable than processes controlled
+by humans \cite{tibav:42484}, e.g., in a laboratory.
+
+Depending on the importance of the CODECHECK
+(see \ref{when-to-do-a-codecheck}), a mediator, e.g., the editor, or a
 second opinion might be warranted in rare cases.
@@
 
-Publisher staff represents the most controlled but also resource-intensive
+A publisher's \textbf{editorial staff} represents the most controlled but also
+resource-intensive
 option. Hiring the required technical and or domain expertise puts a financial
@@
 as part of the scientific community.
+Keeping diverse specialists as staff to realise a codechecking process is 
+unlikely to be feasible, especially for small or independent publishers.
 
 In contrast, for researchers it can be very important to be publicly credited
-with their activity as a reviewer (see TODO CHECK REF \nameref{who-knows-who}). A regular
+with their activity as a reviewer (see TODO CHECK REF \nameref{who-knows-who}).
+A regular
 review may be listed in public databases (e.g., ORCID
@@
 ``Auditable Research''  similarly.
-Note that openness is not required here, as to accommodate 
-communities of practice and review workflows that (currently) do not want
-to or are not able to work openly.
+Every community may establish their own good practices or adapt generic
+concepts and practical tools, such as publishing all building blocks
+of science in a research compendium
+\footnote{\url{https://research-compendium.science/}}.
 A completed check means that code could be executed at least once without
@@
 under open licenses, before or when the certificate is published.
+Note that openness is not required for the paper itself as to accommodate 
+journals to transition to sustainable open access models.
 The code and data publication should follow community good practices.
-By disposition, a CODECHECK workflow should be designed as a recognition of
-shortcomings in asserting reproducibility of submissions and understand the
+By disposition, a CODECHECK workflow should be designed (a) as a recognition of
+shortcomings in asserting reproducibility of submissions and (b)
+with an understanding of the
 complex situation that lead to this. However, establishing a CODECHECK workflow
@@
 journal's target audience to reach a state where the level of verification 
-and trust provided by codecheckers becomes part of peer review.
+and trust provided by codecheckers becomes part of peer review and regular
+merit of all accepted papers.
 
@@
 \subsection*{Register}\label{register}
-% count certificates: 
-% curl https://codecheck.org.uk/register/stats.json
-% -1 for Hathway-Goodman (no empty last line, so no -1 for the column headers needed)
-
-We have a curated list of 22 certificates available at
+
+We have a curated list of 25 certificates available at
 \url{https://codecheck.org.uk/register}. These are a mixture of
 reproductions of papers for a variety of reasons and fall into three themes:
-(1) `classic' papers from computational neuroscience
+(1) classic and current papers from computational neuroscience,
 (2) COVID-19 modelling preprints, and
@@
 
-\textbf{Can't someone cheat?} Yes. We don't check it is correct code,
+\textbf{Can't someone cheat?} Yes. We don't check it is correct code or
+sound science,
 just simply that it runs. This `mechanical' test is indeed a very low bar.
@@
 time to the community.
-
-\textbf{Who finds the codechecker?} Good point! The journal, using either
-their own team of codecheckers or the public database. Researchers
-however think this is an important issue, and are volunteering.
+With code and workflows, there also is a good chance that a codechecker's
+feedback can have direct impact and lead to improvements of the author's 
+work that comes with a rewarding experience or even tangible contributions.
+While such impact is also part of idealistic peer review, it is
+rather rare beyond paraphrased anonymous acknowledgement.
 
@@
 or journals with collaborative/open peer review, but eventually decided
-against it.
+against them.
+We also have considered to require latest technologial solutions which can
+support reproducibility (cf.~\cite{konkol_publishing_2020}), but have 
+decided to focus on the human interface and the judgement of an experienced
+researchers/developer as a more sustainable approach.
+The structurelessness of the required data and software gives the needed
+freedom to enable all types of research. It could be complemented by
+automated scoring (e.g.,~\cite{menke_rigor_2020}, yet automation and 
+metrics bears risks that lead us to put our trust in the codechecker.
 Note that the focused nature of the CODECHECK principles allows journals
-and publishers to innovate on the finance model and peer review practices
+and publishers to innovate on the financial model and peer review practices
 at their own pace.
@@
 SJE is on the editorial board at the journal \emph{Scientific Data}.
-DN is reproducibility editor at the Association of Geographic
+DN is reproducibility chair at the Association of Geographic
 Information Laboratories in Europe's (AGILE) annual conference.

### 2021-01-12 codecheck-paper.tex
final acks confirmation

@@
 \usepackage{doi} % turn DOIs into links
-\usepackage{nameref}
 
@@
 \begin{document}
-\title{CODECHECK: An open-science initiative for the independent execution of computations underlying research articles.}
+\title{CODECHECK: An open-science initiative for the independent execution of computations underlying research articles}
 \author[1,$\ast$]{Daniel N\"{u}st}
@@
 Maciocci, Naomi Penfold (eLife), Rebecca Kirk (PLOS Computational Biology), Scott
-Edmunds (GigaScience), and Andrew Hufton? (Scientific Data). Iain Davies and
+Edmunds (GigaScience), and Andrew Hufton (Scientific Data). Iain Davies and
 Yuhao (Sebastian) Wang developed code and example certificates.

### 2021-01-12 codecheck-paper.tex
adjust figure style to rest, add terms
restructure small bits and slightly shorten conclusions

@@
 \begin{figure}
+% emojis: https://github.com/googlefonts/noto-emoji
+% cross mark: u274c
+% check mark: u2714
   \centering
@@
 at conferences.
-CODECHECKs and can increase trust in scientific results and the set of 
-shared principles and name can through recognition value allow researchers
+CODECHECKs can increase trust in scientific results.
+The set of 
+shared principles and common name, through recognition value,
+allow researchers
 across publication venues to judge the level of scrutiny that results have
 faced. CODECHECK requires direct acknowledgement of the contribution to the
-scientific community in the role, not indirectly via citations of 
-reproductions, and can be a worthwhile educational activity, especially for
-ECRs. A technically simple yet far reaching change would be a special type of 
-reviewer activity for codechecking in public researcher profiles next to
-conventional peer reviews.
+scientific community in the role of codechecker, not indirectly via citations 
+of reproductions or informal crediation and closed inner circles.
+Codechecking can be a worthwhile educational activity, especially for
+ECRs.
 
@@
 results \cite{munafo_what_2020}.
-Even more, it is clear that establishing CODECHECK-like systems can be 
+It is clear that establishing CODECHECK-like systems can be 
 impeded by and must be seen in connection to much larger issues in 
@@
 pressure \cite{piwowar_altmetrics:_2013,nosek_promoting_2015}.
-We certainly do not want the binary attribute of "CODECHECK completed"
+We certainly do not want the binary attribute of ``code works''
 to become a factor in bibliometric approaches for performance assesments.
 % Mittermaier, Peer Review and Bibliometrics: http://hdl.handle.net/2128/22745
+
 While developed for the current "paper"-centric publication process and
@@
 iterative and granular communication of scientific outputs, 
-such as Octopus\footnote{\url{https://science-octopus.org/}}, 
 articles with live-code \cite{perkel_pioneering_2019-1}
@@
 \cite{tenorio-fornes_towards_2019},
-and completely new visions for scholarly communication and peer review
-\footnote{Cf., for example, \emph{A modern vision for peer review} by 
+and completely new visions for scholarly communication and peer 
+review\footnote{Cf., for example, \emph{A modern vision for peer review} by 
 Amy J. Ko: \url{https://medium.com/bits-and-behavior/a-modern-vision-for-peer-review-d5f73f0fae07}}.
@@
 management software would not support such open communication.
-We would be keen to use the flexibility of the principles and cooperate
-on non-public CODECHECKs to learn more about the advantages and yet unclear
-specific challenges. Does CODECHECK work better with OPR?
-Each rollout of CODECHECK should be accompanied by studies to ensure 
-codechecking does not fall into the same gaps as peer review did \cite{tennant_limitations_2020}
-and to ensure positive change within the respective peer review system.
+
 More reproducible practices initiated by CODECHECKs could lead
@@
 whether they use the actual name for it or not.
+We would be keen to use the flexibility of the principles and cooperate
+with journals to learn more about the advantages and yet unclear
+specific challenges -- Do CODECHECKs really work better with OPR?
+Each rollout of CODECHECK should be accompanied by studies to ensure 
+codechecking does not fall into the same gaps as peer review did
+\cite{tennant_limitations_2020}
+and to ensure positive change within the respective peer review system.
 The CODECHECK badge is, intentionally, not branded beyond the checkmark
@@
 to be able to grow the pool of codecheckers as needed.
-The initial reaction of the scientific community, where up to today over 
-twenty people spontaneously signed up as volunteer codecheckers after talks
-on the initiative, makes us feel optimistic that the time is right for 
+The initial reaction of the scientific community and the spontaneously signing
+up volunteer codecheckers make us feel optimistic that the time is right for 
 putting scholarly peer review on the path to facilitate sharing and execution
-of computer programs and documenting reproducibility of results.
+of computational workflows.
 

### 2021-01-12 codecheck-paper.tex
update introduction (clarify, ..)

@@
 
-Many areas of scientific research now use computation to either simulate
+Many areas of scientific research use computations to either simulate
 or analyse their data. These computations are increasingly complex, and
@@
 To complement the traditional route of sharing research by writing papers,
-there is a growing trend/demand to share the underlying artefacts, notably 
+there is a growing demand to share the underlying artefacts, notably 
 code and datasets, so that others can inspect, reproduce or expand that work
@@
 The implementation status of this varies considerably according to the
-journal. At one of the spectrum, there are specialist journals that
+journal. At one end of the spectrum, there are specialist journals that
 have been created to accept ``data papers''
@@
 together are required to validate outcomes of a computational analysis,
-they do differ in  that data can be seen as static/inert,
-while code requires an environment and interaction. This makes 
-software harder to share.
+they do differ in  that data can be seen as static and inert,
+while code requires an environment to be run in and interacted with.
+This makes software harder to share.
 Our anectodal experiences
 on this matter suggest that often a variety of reasons are given for why
-code cannot be shared, e.g.~``there is no documentation / I do not
-want to maintain it / I do not want to give away my code to my
-competitors''. Our view, rooted in Open Science principles, is that
-wherever possible sharing code is good for the community, as outlined
-in 2010 \cite{Barnes2010-iv}. Having the code freely available, and
-suitably archived, provides a valuable resource for others to learn
-from, even when it doesn't run or if there is no documentation. However,
-with a little effort, we believe that if someone independent can run/execute
-the code, this is worth documenting as early as possible and that the barrier
-to evaluating non-text parts of research can be greatly reduced with a 
+code cannot be shared, e.g.,~``there is no documentation'', ``I cannot 
+maintain it'', or ``I do not want to give away my code to competitors''.
+Our view, rooted in Open~Science principles
+\footnote{\url{https://en.wikipedia.org/wiki/Open_science}}, is that
+wherever possible sharing code is good for the community and the 
+individual \cite{Barnes2010-iv,markowetz_five_2015}.
+Having code and data openly available and archived, 
+provides a valuable resource for others to learn
+from, even when it doesn't run anymore or if there is no documentation.
+However,
+with a little effort, we believe that if someone independent can execute
+an analysis, this is worth documenting as early as possible and that the barrier
+to evaluating non-text materials of research can be greatly reduced with a 
 focused approach.
-Just like peer review can provide a ``baseline reassurance'', i.e., a paper
-has at least been checked by someone with an understanding of the topic
+Just like data journals' validations of data and all journals' peer review
+can provide a ``baseline reassurance'', i.e.,~a paper
+has been checked by someone with an understanding of the topic
 \cite{fyfe_mission_2019}, we think the same baseline should be provided
 regarding the workflow underlying a paper.
-With this in mind, we have developed a set of PRINCIPLES and an example 
-workflow that provides what he hope is a pragmatic way of checking that 
-code works.
+With this in mind, we have developed a set of principles and an example 
+workflow that provides what he hope is a pragmatic way of checking that a
+paper's code works.
 
@@
 and its variations to integrate a much needed evaluation of the 
-computational reproducibility \cite{barba_terminologies_2018} into peer
+computational reproducibility\footnote{Following the 
+Claerbout/Donoho/Peng terminilogy, see 
+\cite{barba_terminologies_2018}} into peer
 review, and a demonstration of its feasibility by means of 25

### 2021-01-12 codecheck-paper.tex
update workflow description (clarifycations)

@@
 we expand on the underlying principles. The workflow involves three
-groups of people: (1) the AUTHOR of a paper providing the code to be checked, (2)
-the PUBLISHER of a journal interested in publishing the paper by the
-AUTHOR and (3) the CODECHECKER who checks that the AUTHOR's code works.
-The workflow that we have refined is documented in Figure~\ref{fig:workflow}.
-
-% TODO: how is a CODECHECK different from a "specialist reviewer" ? Is it because it is not special, but regular?
+groups of people:
+(1) the AUTHOR of a paper providing the code to be checked,
+(2) the PUBLISHER of a journal interested in publishing the paper 
+by the author, and
+(3) the CODECHECKER who checks that the authors's code works.
+The workflow that we have refined is documented in Figure~\ref{fig:workflow}
+and consists of six steps.
 
@@
 
-\textbf{Step 1:} the AUTHOR submit their manuscript along with code and data to
+\textbf{Step 1:} the author submits their manuscript along with code and data to
 the PUBLISHER. The code and data need not be openly available at this point.
@@
 such as GitHub or GitLab. Ideally, the AUTHOR is expecting the CODECHECK and
-prepares for it, e.g., by asking a colleague to briefly attempt a reproduction
-of the submitted workflow.
+prepares for it, e.g.,~by asking a colleague to attempt a reproduction or by 
+following the PUBLISHER's guidelines for sharing computational workflows.
 
 \textbf{Step 2:} the publisher finds a CODECHECKER to check the code. This is
-analogous to the publisher finding one or more reviewers to evaluating
+analogous to the publisher finding one or more reviewers to evaluate
 the paper. However, a crucial difference is that we suggest the CODECHECKER
-and the author can talk freely to each other directly (see next).
+and the author can talk freely to each other directly (see next), independent
+of the openness of the scientific peer review process.
 
 \textbf{Step 3:} the CODECHECKER runs the code, based on instructions provided by
-the author, and checks if some/all of the results from the paper can be
-reproduced. (See later for exact nature of ``reproduced''). If there are
-any problems running the code, the CODECHECKER asks the AUTHOR for help
-to resolve the problems. The burden to provide reproducible material lies with
-the author.
+the author. They check if some or all of the results from the paper can be
+reproduced. If there are
+any problems running the code, the CODECHECKER asks the AUTHOR for help,
+updates, or further documentation.
+The burden to provide reproducible material lies with the author.
 The CODECHECKER then tries to run the code again.
 This process iterates until either the CODECHECKER is successful,
-or the CODECHECKER concludes the code does not work (\emph{TODO NOT
-  YET HANDLED}). As part of this process, the CODECHECKER could work entirely
-locally on their own compute resource, or in the cloud, e.g.~using the open
+or the CODECHECKER concludes the code does not work.
+As part of this process, the CODECHECKER could work entirely
+locally on their own compute resource, or in the cloud, e.g.,~using the open
 MyBinder infrastructure \cite{jupyter_binder_2018} or alternatives, some of which
 are more tailored for scientific publications while others offer commercial
-options to, e.g., publishers (cf. \cite{konkol_publishing_2020}).
-Using such cloud-based 
+options to, e.g.,~publishers (cf. \cite{konkol_publishing_2020}).
+Cloud-based 
 infrastructure allows for the CODECHECKER and author to collaboratively improve
-the code and enforces a much more complete definition of the computing environment,
-but, unless secure infrastructure is provided by, e.g., by the publisher, requires code
-and data to be published openly online.
+the code and enforces a complete definition of the computing environment,
+but, unless secure infrastructure is provided, e.g.,~by the PUBLISHER,
+requires code and data to be published openly online.
+
 Note that the task of the CODECHECKER is to check the ``mechanics'' of the
 workflow, not to evaluate its correctness. 
-Stodden et~al. \cite{stodden_setting_2013} distinguish between verification and
-validation in the context of Mathematics, and in the same vein a CODECHECK
-ensures a verification\footnote{Note this is a different use as the \emph{Verification Reports} article type in the journal \emph{Cortex}
-\cite{chambers_verification_2020}},
-i.e., checking that the code correctly produces the 
-output it claims to create, but not a validation, i.e., does the code implement
-the right algorithm to solve the problem that is researched, 
-of the computational results.
+Stodden et~al. \cite{stodden_setting_2013} distinguish between \emph{verification}
+and \emph{validation}
+in the context of Mathematics. Following their definition, a CODECHECK
+ensures a verification\footnote{Note this is a different use as the
+\emph{Verification Reports} 
+article type in the journal \emph{Cortex} \cite{chambers_verification_2020}},
+of computational results, i.e.,~checking that the code correctly produces the 
+output it claims to create, but not a validation, i.e.,~does the code implement
+the right algorithm to solve the problem that is researched.
 Nevertheless, the simple attempt to reproduce may highlight shortcomings in
@@
 the author how to improve the material.
-The plain text in the certificate is the most flexible and feasible way to 
-make clear what was really checked. Readers will have very different 
-understandings of what a ``green'' checkmark (cf. Step~5) means.
+The free text in the certificate is the most flexible and feasible way to 
+make clear what was really checked, because each workflow is unique.
+Since no specific tool or platform should be required so that no authors are
+excluded, automation or fixed checklists are futile.
 
 \textbf{Step 5:} the certificate as well as auxiliary files created
-during the check, e.g., a specification of a computing environment,
-helper scripts, and the code and data, if available under open
-licenses, get deposited in an open archive (currently Zenodo); the
-certificate is given to the PUBLISHER.
-
-\textbf{Step 6:} the PUBLISHER can then use the certificate as part of their
-supplementary records for a paper or provide it to scientific reviewers; the
-PUBLISHER also gives credit to the CODECHECKER for their work by depositing the
-activity appropriately in scholarly publication databases, such as ORCID.
-The PUBLISHER also ensures proper publication metadata, e.g., links from the 
+during the check, e.g.,~a specification of a computing environment, data 
+subsets or helper scripts, and the original code and data get deposited in
+an open archive unless restrictions regarding data size, license or 
+sensitivity apply.
+Currently, the CODECHECKERs deposit the material on Zenodo themselves, 
+while processes managed by the PUBLISHER may have them complete this step.
+A badge or other visual aid may be added to the deposit and the paper and
+link to the certificate.
+Despite the simplification into a binary value and the risk of confusion of
+the extent of the check, especially across disciplines and methods,
+a badge can provide recognition value and creditation for the completed
+CODECHECK. The badge and the actual check are incentives for the efforts
+of providing a readily computationally reproducible workflows.
+
+\textbf{Step 6:} the PUBLISHER can, depending on the timing, provide the
+certificate to scientific reviewers or handling editors, or publish it
+and link between the certificate, the paper, and any repositories 
+through metadata. Currently, the CODECHECKER creates these on Zenodo\footnote{
+See ``related identifiers'' and ``alternate identifiers'' in the right-hand
+sidebar of certificate \texttt{2020-025} \cite{cert-2020-025}.}.
+The PUBLISHER also gives credit to the CODECHECKER for their work by depositing
+the activity appropriately in scholarly profiles, such as ORCID\footnote{
+See peer review contributions in ORCID records at
+\url{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}.},
+The PUBLISHER also ensures proper publication metadata, e.g.,~links from the 
 certificate repository to the published paper or the original code repository.
-A badge or other visual aid may be applied to highlight the fact of a completed
-CODECHECK and provide a incentive and reward for the, today often considered
-``additional'' efforts of providing readily computationally reproducible workflows,
-despite the simplification into a binary value and the possible confusion of
-the meaning, especially across disciplines and methods.
 

### 2021-01-12 codecheck-paper.tex
restructure dimensions for better readability

@@
 implementations of this process. To elaborate on these options, we consider
-the following dimensions in a space of possible CODECHECK workflows, as shown in 
+the several dimensions in a space of possible CODECHECK workflows, as shown in 
 Figure~\ref{fig:dimensions}, to represent relevant aspects for the involved
@@
 transparency and will be discussed in the following sections.
-Each of these have their own pros and cons and can for the most part be
-freely combined.
+The options on each dimension have their own pros and cons and can,
+for the most part, be freely combined.
 
@@
 
-The point in time when a CODECHECK happens greatly depends on the 
-intentions and the ascribed importance.
+The point in time for a CODECHECK and the ascribed importance are closely
+connected, so the dimensions \emph{When} and \emph{Imporance} are described
+together.
 The earlier a CODECHECK happens in the process, the more the
@@
 is a paper published, sent back for revisions, or rejected.
-Such a \textbf{strict prerequisite} can reduce the workload of reviewers,
-and therefore seems out of place late in the review process.
+Even earlier checks, i.e.,~a CODECHECK of a preprint, may even help to 
+improve the workflow itself even before a publisher is involved. The 
+possibility for codechecking papers could be part of a preprint server's
+policy, or simply be initiated by interested readers and community members.
+
+A \textbf{strict prerequisite} for passing a check can reduce the workload
+of reviewers, and therefore seems out of place late in the review process.
 The later in a review process the check happens, the less challenging is it
@@
 because the author might already be notified of the acceptance and may more
-freely share materials online.
-Even earlier checks, i.e., a CODECHECK of a preprint, may even help to 
-improve the workflow itself even before a publisher is involved. The 
-possibility for codechecking papers could be part of a preprint server's
-policy, or simply be initiated by interested readers and community members.
-% TODO add a sentence about the possible effect of just asking for
-% code/data early on in the submission process, see https://twitter.com/tsuyomiyakawa/status/1230672758163402752?s=09)
-
-We see a CODECHECK as part of a typical scholarly communication process.
+freely share materials online closer to the paper publication date.
+
+A \textbf{pre-review} CODECHECK allows editors to use it as a filter to only
+send  a submission out to peer review if it passes the check.
+The certificate would  be included in the submission package provided
+to the reviewers, so they can include the CODECHECK's results in their 
+evaluation of the work.
+Reviewers may then judge the the relevance of the computations for the results
+of the work, or follow journal guidelines as how strict to apply the check's
+outcomes.
+
+A CODECHECK may also be conducted in \textbf{parallel} to the scientific review.
+This puts less burden on the turnaround time for the check, yet only makes the 
+outcomes available for the final consideration of the handling editor.
+The check could also be tasked after suggestion by a reviewer and remove the
+need for a pre-review screening for computational workflows. However, 
+soliciting such a ``specialist review'' is much less desirable than having
+a regular CODECHECK, without any special treatment for some submissions.
+In both cases, a CODECHECK could then be awarded \textbf{equal weight} as 
+the other reviews for the editor's decision.
+
+A \textbf{post acceptance} CODECHECK has the smallest impact on editorial 
+decisions and may simply provide an \textbf{extra merit} on top of the 
+acceptance of the submission. 
+This  is the least impactful solution while still evaluating all material
+and properly acknowledging the check results from the paper, because the 
+check can be completed  before the publication of the paper.
+The GIScience group of checks (see below) falls into this
+category: the AGILE conference highlights articles where the reproducibility
+was successfully reviewed with a badge on the volume and article landing
+pages. Similarly, the GIScience papers in collaborations with journals were 
+checked while the authors worked on revisions.
+
+A CODECHECK may also be conducted \textbf{post publication}, though it
+would required to update the article and article metadata later to reference 
+the check, otherwise readers would presumably not be aware of it. In general,
+publishers are hesitant to make such a revision of published articles.
+While this has the least impact on the current publishing practices and also
+enables open discourse,
+it also has the smallest effect on the original submission
+and does not acknowledge the importance of reproducible workflows for
+ensuring good scientific practice.
+
+In general, the timing of a CODECHECK needs to be connected with 
+established or typical scholarly communication process.
 While more radical ideas to change the paradigms
-how researchers share their work exist\footnote{For example Octopus, 
+how researchers can share their work much more granually exist\footnote{
+For example Octopus, 
 \url{https://science-octopus.org/about}, or Hypergraph
@@
 % https://docs.google.com/presentation/d/1V_K8hghgnvGfEtW7TdwtowTveZyy4Qs8rMemNyHx_Dg/edit#slide=id.p
-there is a value in an evolutionary approach.
-Thereby, existing communities can 
+there is a value in an evolutionary approach and we do not suggest
+to completely remove the paper as a point in time to share work just yet.
+Enhancing existing processes with CODECHECKs allows communities to 
 transition towards more open practices at their own pace and ensure
@@
 publishing the certificate, but excludes the time for actual computations.
-These efforts include considerable time to communicate about the process, 
-especially around who publishes which document when, so that proper 
+These efforts are comparable to the time needed to peer review a submission,
+which concurs with the efforts some volunteer codecheckers are willing to
+spend.
+However, the time includes a considerable share of communicating about the
+process,  especially around who publishes which document when, so that proper
 cross-referencing between paper and certificate is ensured via persistent
@@
 When integrated into a peer review system, this handling of documents should
-become much more effective and reduce organisational efforts.
-
-A \textbf{pre-review} CODECHECK allows editors to use it as a filter to only
-send  a submission out to peer review if it passes the check.
-The certificate would  be included in the submission package provided
-to the reviewers, so they can include the CODECHECK's results in their 
-evaluation of the work.
-Reviewers may then judge the the relevance of the computations for the results
-of the work, or follow journal guidelines as how strict to apply the check's
-outcomes.
-
-A CODECHECK may also be conducted in \textbf{parallel} to the scientific review.
-This puts less burden on the turnaround time for the check, yet only makes the 
-outcomes available for the final consideration of the handling editor.
-The check could also be tasked on demand, e.g., only if one of the reviewers
-suggests/requests one because previous screenings might not have caught the 
-fact that a computational workflow is part of the submission. However, 
-soliciting such a "specialist review" is much more undesirable than having
-a regular CODECHECK, without any special treatment for some submissions.
-A CODECHECK could then be awarded \textbf{equal weight} as the other reviews.
-
-A \textbf{post acceptance} CODECHECK puts the least requirements on the author 
-and may simply lead to an award of excellence on top of the acceptance of the
-submission. This \textbf{extra merit} is the least strict solution while still
-properly  acknowledging the check results,
-because the CODECHECK is to be completed 
-before the publication of the final paper.
-The GIScience group of checks (see \nameref{register}) falls into this
-category: the AGILE conference highlights articles where the reproducibility
-was successfully reviewed with a badge on the volume and article landing
-pages. Similarly, the GIScience papers in collaborations with journals were 
-checked while the authors worked on revisions.
-
-Finally, a CODECHECK may also be conducted \textbf{post publication}, though it
-would required to update the article and article metadata later to reference 
-the check, otherwise readers would presumably not be aware of it. In general,
-publishers are hesitant to make such a revision of published articles.
-While this has the least impact on the current publishing practices and also
-enables open discourse,
-it also has the smallest effect on the original submission
-and does not acknowledge the importance of reproducible workflows for
-ensuring good scientific practice.
+become much more effective and reduce organisational overhead.
 
@@
 
-The topic of blinding is broadly discussed, not the least in the push towards
+The topic of anonymity is broadly discussed, not the least in the push towards
 open peer review as part of the Open Science movement 
@@
 Without taking a strong stance on this discussion, the general motivation 
-for more  transparency and reproducibility does indeed favour a more open 
+behind CODECHECK
+for higher transparency and reproducibility does indeed favour a more open 
 review process.
 However, anonymity can also serve to protect individuals 
-\cite{tennant_limitations_2020}, e.g.,  junior scientists, whereas it might 
-be acceptable to require revealing of the names of tenured researchers.
-The possible negative effects of a non-anonymous review are greatly reduced
+\cite{tennant_limitations_2020}, e.g.~junior scientists, whereas it might 
+be acceptable to require revealing of names from tenured researchers.
+The possible negative effects of a non-anonymous review are reduced
 if a CODECHECK is not relevant for the decision to accept or reject, but that
@@
 process that should in any case be fixable and not a question of opinion or
-faulty approach. It is technically possible for a nonsense workflow to receive
+a faulty approach.
+It is technically possible for a nonsense workflow to receive
 a CODECHECK certificate.
@@
 % previous sentenc inspired by https://osf.io/9cftx/
-Furthermore, the level of documentation that is needed for third parties
-to reproduce a workflow is extremely hard to get right. Too often, this 
-uncertainty leads to researchers giving up and not documenting at all.
 
@@
 fix problems or guess the next step, the codechecker can ask the author to 
-rework the docs  or the code.
-Instead of lengthily describing a small change, the codechecker may simply
-provide a fix within the code.
-Instead of hiding useful and  instructive files created during the codecheck
-(e.g., a machine-readable computing environment specification), the author 
-and readers can profit from the software-related expertise of the
-codechecker.
-While such a communication may be facilitated in an anonymous way by the 
+rework the docs  or the update code.
+Of course, the codechecker may also provide a fix within the code instead
+of lengthily describing a small change if it is quickly done.
+Authors and readers can profit from codecheckers' experience and approach,
+as the latter may create useful and instructive files, e.g.,~a machine-readable
+computing environment specification, during the check.
+While communication between author and codechecker may be facilitated in an 
+anonymous way by the 
 publisher, it most likely only helps to protect the identity of the 
 codechecker, because, especially when of a certain size, not trivial, or 
-if very specialised, it is close to impossible to fully anonymise a 
-software analysis stack.
-Therefore, the most effective means with the most desirable result for
+if very specialised, it is close to impossible to fully anonymise a codebase.
+Therefore, the most effective means with the most desirable variant for
 the stakeholders is a non-anonymous and collaborative CODECHECK.
-The possible contributions by the codechecker may even be integrated into
+The contributions by the codechecker may even be integrated into
 the code of the workflow and be acknowledged as code commits. This way, 
 proper credit is given within the research software development community.
-Nevertheless, depending on other needs represented by other dimensions,
+Nevertheless, depending on needs represented by other dimensions,
 the aspect of `Who knows who?' may be the most flexible one.
@@
 setting  up a computing environment with a particular language or tool is 
-much more discouraging and the assessment of the outcome, e.g., comparing
+much more discouraging and the assessment of the outcome, e.g.,~comparing
 created figures with the original, is largely possible without 
@@
 are the main drivers for the depth of checking, though in general, we 
-expect a CODECHECK not to evaluate performance, e.g., does a racing car
-have the best apex speed, but to assess operation free of failure, i.e., 
-is it a car with 4 wheels and at least one door, and does it move forwards
+expect a CODECHECK not to evaluate performance, e.g.,~does a racing car
+have the best apex speed, but to assess operation free of failure, i.e.,~is
+it a car with 4 wheels and at least one door, and does it move forwards
 when accelerating. This depth can be seen independent from the power of
@@
 mistakes is is thereby more reliable and auditable than processes controlled
-by humans \cite{tibav:42484}, e.g., in a laboratory.
-
+by humans \cite{tibav:42484}, e.g.,~in a laboratory.
 Depending on the importance of the CODECHECK
-(see \ref{when-to-do-a-codecheck}), a mediator, e.g., the editor, or a
+(see above), a mediator, e.g.,~the editor, or a
 second opinion might be warranted in rare cases.
@@
 In fact, ECRs are often much more familiar with latest technology and are 
-likely to be early adopters as creators, i.e., be authors of CODECHECK-ready manuscripts, too\footnote{Cf. the 77\% of 141 registered reports who were 
+likely to be early adopters as creators, i.e.,~be authors of CODECHECK-ready manuscripts, too\footnote{Cf. the 77\% of 141 registered reports who were 
 submitted to leading Neuroscience journals by ECRs
@@
 Not to solve the general problems of peer review, which is largely an
-unsupervised process, i.e., rarely are reviewers taught how to do it, but
+unsupervised process, i.e.,~rarely are reviewers taught how to do it, but
 a junior codechecker may be supported by a senior codechecker, too.
@@
 a clear commitment on openness and transparency, independent of the
-respective current peer review process (see \nameref{who-knows-who}).
+respective current peer review process (see \emph{Openness} dimension).
 
@@
 In contrast, for researchers it can be very important to be publicly credited
-with their activity as a reviewer (see TODO CHECK REF \nameref{who-knows-who}).
+with their activity as a reviewer.
 A regular
-review may be listed in public databases (e.g., ORCID
+review may be listed in public databases (e.g.,~ORCID
 \footnote{\href{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}} or commercial offerings such as Publons
@@
 following talks given on the project in the last 12 months.
-That is even though no checks are actively elicited.
-We see a high potential in an open shared list of potential codecheckers 
-compared to a private, in-house group, as it may better match expertise
+That is even though no checks are actively elicited in collaboration
+with journals.
+Their motivations, mentioned in the sign-up issues\footnote{
+\url{https://github.com/codecheckers/codecheckers/labels/registration}},
+are:
+supporting reproducible research and open science,
+establishing a culture of open and reproducible science and being part
+of making it the norm,
+improving coding skills,
+giving back to the community,
+gaining experience in helping scientists with their code,
+reducing the amount of bad scientific code and encouraging a culture that
+reduces friction in passing on academic codebases,
+learning from other people's mistakes,
+and curiosity.
+We see a benefits in an open shared list of codecheckers
+across journals
+compared to a private in-house group, as it may better match expertise
 and share the workload.
-% TODO tell more about the motivations and commitments that we know about?
 The overarching community is also a mean to ensure CODECHECK is a viable
 option not just for large publishers, independent of their financial model,
-but also for independent no-cost open access journals.
+but also for independent no-cost Open Access journals.
 

### 2021-01-12 codecheck-paper.tex
update principles and register (consistent styling, understandability)

@@
 fix from the author.
+The level of documentation that is needed for third parties
+to reproduce a workflow is extremely hard to get right. Too often this 
+uncertainty leads to researchers giving up and not documenting at all.
+The conversation with a codechecker fixes this problem.
 
 \textbf{2. Communication between humans is key.} Some code may just
-work without any interaction (e.g. CODECHECK certificate 2020-013
+work without any interaction (e.g.,~CODECHECK certificate \texttt{2020-013}
 \cite{cert-2020-013}); however, more often there are hidden
@@
 The public record can be realised by publishing the certificate in a 
-citable form (i.e., with a DOI; this is done within the community codecheck
+citable form (i.e.,~with a DOI; this is done within the community codecheck
 process), by listing codecheckers on the journal's website or, ideally, by
@@
 By default, unless there are strong
-reasons to the contrary (e.g., clinically-sensitive data), all code and
-data, both from author and codechecker, will be made freely available, i.e., 
-under open licenses, before or when the certificate is published.
+reasons to the contrary (e.g.,~clinically-sensitive data), all code and
+data, both from author and codechecker, will be made freely available, 
+i.e.,~under open licenses, before or when the certificate is published.
 Note that openness is not required for the paper itself as to accommodate 
-journals to transition to sustainable open access models.
+journals to transition to sustainable Open Access models.
 The code and data publication should follow community good practices.
@@
 We have a curated list of 25 certificates available at
-\url{https://codecheck.org.uk/register}. These are a mixture of
-reproductions of papers for a variety of reasons and fall into three themes:
+\url{https://codecheck.org.uk/register}.
+Table~\ref{tab:register} lists the certificates completed to date,
+along with citations to the original papers.
+These CODECHECKs fall into three themes:
 (1) classic and current papers from computational neuroscience,
@@
 (3) GIScience.
+
 The first theme was the initial set of papers used to explore the concept
-of CODECHECK. The idea was to take very well known historical articles from a 
-specific domain, in this case the work area of co-author SJE, possibly 
+of CODECHECK. 
+The idea was to take very well known historical articles from a 
+specific domain, in this case the work area of SJE, possibly 
 as as a conversation starter for potential collaborating journals.
+
 The second theme is rather opportunistic. As the COVID-19 pandemic swept
@@
 %Several COVID-19 models have been done during the time of peer review of these papers,
-and under our initiative rather than requested from journal, but the certificate
-has then been acknowledged/cited in the paper \cite{Davies2020-vj}.
-The third theme represents co-author DN's service as Reproducibility
+and under our initiative rather than requested from journal, but some certificates
+have since then been acknowledged in reviewed papers
+\cite{Davies2020-vj,kucharski_effectiveness_2020}.
+
+The third theme represents DN's service as Reproducibility
 Reviewer at the AGILE conference series.
@@
 series \cite{nust_improving_2020}.
-While using slightly different terms and infrastructure, e.g., ``reproducibility
-reports'' are published on OSF instead of certificates on Zenodo, the 
-reproducibility reviews adhere to the codecheck principles.
+While using slightly different terms and infrastructure (``reproducibility
+reports'' are published on OSF instead of certificates on Zenodo) the 
+AGILE reproducibility reviews adhere to the codecheck principles.
 Furthermore, a small number of checks was completed as part of a peer reviews
@@
 
-Table~\ref{tab:register} lists the certificates completed to date,
-along with citations to the original papers.
 Out of these areas, we would like to briefly highlight three examples here.
-
-\textbf{TODO: SJE, tidy this paragraph.}
+\textbf{TODO SJE: tidy this paragraph.}
 \begin{enumerate}
@@
     \textbf{Certificate} & \textbf{Research area} & \textbf{Description} \\ \midrule
-    2020-001 \cite{cert-2020-001} & Machine learning & Code for benchmarking ML classification tool checked post acceptance of manuscript and before its publication in \textit{Gigascience} \cite{Piccolo2020-lo}.
+    \texttt{2020-001} \cite{cert-2020-001} & Machine learning & Code for benchmarking ML classification tool checked post acceptance of manuscript and before its publication in \textit{Gigascience} \cite{Piccolo2020-lo}.
 \\
-    2020-002  \cite{cert-2020-002} & Neuroscience & Code written for this project checked by second project member as demonstration using paper from 1997 showing unsupervised learning from natural images \cite{Hancock1992-mp}.
+    \texttt{2020-002}  \cite{cert-2020-002} & Neuroscience & Code written for this project checked by second project member as demonstration using paper from 1997 showing unsupervised learning from natural images \cite{Hancock1992-mp}.
 \\
-    2020-003  \cite{cert-2020-003} & Neuroscience &  Code written for this project checked by second project member as demonstration using classic paper on models of associative memory \cite{Hopfield1982-mz}.
+    \texttt{2020-003}  \cite{cert-2020-003} & Neuroscience &  Code written for this project checked by second project member as demonstration using classic paper on models of associative memory \cite{Hopfield1982-mz}.
 \\
-    2020-004  \cite{cert-2020-004} & Neuroscience & Code written for this project checked by second project member as demonstration using classic paper on cart-pole balancing problem \cite{Barto1983-rg}.
+    \texttt{2020-004}  \cite{cert-2020-004} & Neuroscience & Code written for this project checked by second project member as demonstration using classic paper on cart-pole balancing problem \cite{Barto1983-rg}.
 \\
-    2020-005  \cite{cert-2020-005} & Neuroscience & Check of independent reimplementation of spike-timing-dependent plasticity (STDP) model \cite{larisch_re_2019} conducted as demonstration for this paper.
+    \texttt{2020-005}  \cite{cert-2020-005} & Neuroscience & Check of independent reimplementation of spike-timing-dependent plasticity (STDP) model \cite{larisch_re_2019} conducted as demonstration for this paper.
 \\ % Python ANNarchy
-    2020-006  \cite{cert-2020-006} & Neuroscience & Check of independent reimplementation of a generalized linear integrate-and-fire neural model \cite{detorakis_re_2017} conducted as demonstration for this paper.
+    \texttt{2020-006}  \cite{cert-2020-006} & Neuroscience & Check of independent reimplementation of a generalized linear integrate-and-fire neural model \cite{detorakis_re_2017} conducted as demonstration for this paper.
 \\
-    2020-007  \cite{cert-2020-007} & Neuroscience & Check of independent reimplementation of analysing spike patterns of neurons \cite{hathway_re_2018} conducted as demonstration for this paper. \\
-    2020-008  \cite{cert-2020-008} & COVID-19 & Code for modelling of interventions on COVID-19 cases in the UK checked at preprint stage \cite{davies-preprint-2020} and later published \cite{Davies2020-vj}.
+    \texttt{2020-007}  \cite{cert-2020-007} & Neuroscience & Check of independent reimplementation of analysing spike patterns of neurons \cite{hathway_re_2018} conducted as demonstration for this paper.
 \\
-    2020-009  \cite{cert-2020-009} & COVID-19 & Code for analysis of effectiveness of measures to reduce transmission of SARS-CoV-2 checked as preprint \cite{kucharski-preprint-2020} and later published \cite{kucharski_effectiveness_2020}.
+    \texttt{2020-008}  \cite{cert-2020-008} & COVID-19 & Code for modelling of interventions on COVID-19 cases in the UK checked at preprint stage \cite{davies-preprint-2020} and later published \cite{Davies2020-vj}.
+\\
+    \texttt{2020-009}  \cite{cert-2020-009} & COVID-19 & Code for analysis of effectiveness of measures to reduce transmission of SARS-CoV-2 checked as preprint \cite{kucharski-preprint-2020} and later published \cite{kucharski_effectiveness_2020}.
 \\ % R, check mentiond in acks, but not linked
-    2020-010  \cite{cert-2020-010} & COVID-19 & Code for analysis of non-pharmaceutical interventions (Report 9) checked as a preprint \cite{ferguson_report_2020}. \\ % CovidSim
-    2020-011  \cite{cert-2020-011} & COVID-19 & Code for modelling of COVID-19 spread across Europe was provided by authors and checked while paper was in press \cite{flaxman_estimating_2020}. \\
-    2020-012  \cite{cert-2020-012} & COVID-19 & Code for modelling of COVID-19 spread across the USA was checked as preprint \cite{unwin_report_2020} and later published \cite{unwin_state-level_2020}.
+    \texttt{2020-010}  \cite{cert-2020-010} & COVID-19 & Code for analysis of non-pharmaceutical interventions (Report 9) checked as a preprint \cite{ferguson_report_2020}. \\ % CovidSim
+    \texttt{2020-011}  \cite{cert-2020-011} & COVID-19 & Code for modelling of COVID-19 spread across Europe was provided by authors and checked while paper was in press \cite{flaxman_estimating_2020}.
 \\
-    2020-013  \cite{cert-2020-013} & Neuroscience & Code for analsis of rest-activity patterns in people wihout con-mediated vision was checked as a preprint \cite{Spitschan2020.06.02.129502} after direct contact with the authors.
+    \texttt{2020-012}  \cite{cert-2020-012} & COVID-19 & Code for modelling of COVID-19 spread across the USA was checked as preprint \cite{unwin_report_2020} and later published \cite{unwin_state-level_2020}.
+\\
+    \texttt{2020-013}  \cite{cert-2020-013} & Neuroscience & Code for analsis of rest-activity patterns in people wihout con-mediated vision was checked as a preprint \cite{Spitschan2020.06.02.129502} after direct contact with the authors.
 \\ % MATLAB
-    2020-014  \cite{cert-2020-014} & Neuroscience & Code for analysis of pertubation patterns of neural activity was checked after publication as part of publisher collaboration \cite{Sadeh2020}.
+    \texttt{2020-014}  \cite{cert-2020-014} & Neuroscience & Code for analysis of pertubation patterns of neural activity was checked after publication as part of publisher collaboration \cite{Sadeh2020}.
 \\ % MATLAB
-    2020-015  \cite{cert-2020-015} & Neuroscience & Code for a neural network model for human focal seizures was checked after publication as part of publisher collaboration \cite{Liou2020}.
+    \texttt{2020-015}  \cite{cert-2020-015} & Neuroscience & Code for a neural network model for human focal seizures was checked after publication as part of publisher collaboration \cite{Liou2020}.
 \\ % MATLAB
-    2020-016  \cite{cert-2020-016} & GIScience & Code for models demonstrating the Modifiable Aral Unit Problem (MAUP) in spatial data science \cite{Brunsdon2020} was checked during peer review.
+    \texttt{2020-016}  \cite{cert-2020-016} & GIScience & Code for models demonstrating the Modifiable Aral Unit Problem (MAUP) in spatial data science \cite{Brunsdon2020} was checked during peer review.
 \\ % R
-    2020-017  \cite{cert-2020-017} & GIScience & Code for spatial data handling, analysis, and visualisation using a variety of R packages \cite{Bivand2020} was checked after peer review before publication.
+    \texttt{2020-017}  \cite{cert-2020-017} & GIScience & Code for spatial data handling, analysis, and visualisation using a variety of R packages \cite{Bivand2020} was checked after peer review before publication.
 \\ % R
-    2020-018  \cite{cert-2020-018} & GIScience & AGILE conference reproducibility report using a demonstration data subset with cellular automaton for modeling dynamic phenomena \cite{Hojati2020}.
+    \texttt{2020-018}  \cite{cert-2020-018} & GIScience & AGILE conference reproducibility report using a demonstration data subset with cellular automaton for modeling dynamic phenomena \cite{Hojati2020}.
 \\ % R
-    2020-019  \cite{cert-2020-019} & GIScience & AGILE conference reproducibility report with subsampled dataset for reachability analysis of suburban transportation using shared cars \cite{Illium2020}.
+    \texttt{2020-019}  \cite{cert-2020-019} & GIScience & AGILE conference reproducibility report with subsampled dataset for reachability analysis of suburban transportation using shared cars \cite{Illium2020}.
 \\
-    2020-020  \cite{cert-2020-020} & GIScience & AGILE conference reproducibility report using a container for checking in-database windows operators for processing spatio-temporal data \cite{Werner2020}.
+    \texttt{2020-020}  \cite{cert-2020-020} & GIScience & AGILE conference reproducibility report using a container for checking in-database windows operators for processing spatio-temporal data \cite{Werner2020}.
 \\ % Python + PostGIS
-    2020-021  \cite{cert-2020-021} & GIScience & AGILE conference reproducibility report checking code for comparing supervised machine learning models for spatial nominal entity recognition \cite{Medad2020}.
+    \texttt{2020-021}  \cite{cert-2020-021} & GIScience & AGILE conference reproducibility report checking code for comparing supervised machine learning models for spatial nominal entity recognition \cite{Medad2020}.
 \\ % Python
-    2020-022  \cite{cert-2020-022} & GIScience & AGILE conference reproducibility report checking code for visualising text analysis on intents and concepts from geo-analytic questions \cite{Xu2020}.
+    \texttt{2020-022}  \cite{cert-2020-022} & GIScience & AGILE conference reproducibility report checking code for visualising text analysis on intents and concepts from geo-analytic questions \cite{Xu2020}.
 \\ % Python
-    2020-023  \cite{cert-2020-023} & GIScience & AGILE conference reproducibility report on analysis of spatial footprints of geo-tagged extreme weather events from social media \cite{Owuor2020}. \\
-    2020-024  \cite{cert-2020-024} & Neuroscience & Code for multi-agent system for concept drift detection in electromyography \cite{vieira_driftage_2020} was checked during peer review. \\ % R
-    2020-025  \cite{cert-2020-025} & GIScience & Adaptation and application of Local Indicators for Categorical Data (LICD) to archaeological data \cite{carrer_application_2021} was checked after peer review before publication.
+    \texttt{2020-023}  \cite{cert-2020-023} & GIScience & AGILE conference reproducibility report on analysis of spatial footprints of geo-tagged extreme weather events from social media \cite{Owuor2020}.
+\\
+    \texttt{2020-024}  \cite{cert-2020-024} & Neuroscience & Code for multi-agent system for concept drift detection in electromyography \cite{vieira_driftage_2020} was checked during peer review.
+\\ % R
+    \texttt{2020-025}  \cite{cert-2020-025} & GIScience & Adaptation and application of Local Indicators for Categorical Data (LICD) to archaeological data \cite{carrer_application_2021} was checked after peer review before publication.
 \\ % R
@@
   \end{tabular}
-  \caption{Register of completed certificates as of October 2020.  An interactive version
-  is available at \url{http://codecheck.org.uk/register}. % TODO: SJE to complete...
+  \caption{Register of completed certificates as of December 2020.  An interactive version
+  is available at \url{http://codecheck.org.uk/register}.
   }

### 2021-01-12 codecheck-paper.tex
update annotated certificate AND METADATA

@@
 
-\subsection*{Annotated certificate}\label{annotated-certificate}
-
-At the completion of each CODECHECK, a certificate is written that
-states which outputs, i.e., numbers, figures and tables, from the 
-original article could be reproduced by the codechecker.
+\subsection*{Annotated certificate and check metadata}\label{annotated-certificate}
+
+At the completion of each CODECHECK, the codechecker writes a certificate
+stating which outputs, i.e.,~numbers, figures or tables, from the 
+original article, could be reproduced by the codechecker.
 This certificate is made openly
 available so that all readers of the article can see which elements
-have been reproduced.  The certificate also includes links to the code
+have been reproduced and what limitations or issues were faced.
+The certificate also includes links to the code
 and data used by the codechecker, allowing others to build on the
@@
 project, as we have learnt to automate different aspects of the
-certification.
-
-To demonstrate the typical contents of a certificate, in
-Figure~\ref{fig:cert} we have annotated the first four out of ten
-pages of the certificate \texttt{2020-012} \cite{cert-2020-012}.
+certification, such as inserting figures into certificate documents
+and managing core metadata surrounding a check.
+The metadata is stored in a machine-readable structured file in YAML,
+the CODECHECK configuration file \texttt{codecheck.yml}\footnote{
+See technical specification of the CODECHECK configuration file at
+\url{https://codecheck.org.uk/spec/config/latest/}}.
+The configuration file enables current and future automation of workflows and
+meta-analyses.
+
+To demonstrate the typical contents of a certificate the the underlying
+metadata, 
+Figure~\ref{fig:cert} shows the first four out of ten
+pages of the certificate \texttt{2020-012} \cite{cert-2020-012} with
+annotations.
 The referenced article \cite{unwin_report_2020} describes the modelling COVID-19
 across the USA.
-%% TODO - has this been published yet?.  No, not yet... in review...
-Figure~\ref{fig:cert}A shows the certificate number, and its DOI,
-which points to the certificate as archived on Zenodo.
+Figure~\ref{fig:cert}A shows the certificate number and its DOI,
+which points to the certificate and any supplemental files on Zenodo.
 The CODECHECK logo is
-attached to the certificate to denote successful reproduction.
-Figure~\ref{fig:cert}B provides the key metadata about the check --
-which paper was checked, who the authors and the codechecker were,
-and when it was checked. Materials
-relating to the codecheck are freely available via the Repository
-link. This table is generated from machine readable text file, the
-CODECHECK configuration file \texttt{codecheck.yml}.
-The configuration file collects the core pieces of metadata surrounding
-a check and enables current and future automation of workflows and 
-meta-analyses.
-Figure~\ref{fig:cert}C is a textual summary of how the codecheck was
+added for recognition and to denote successful reproduction.
+Figure~\ref{fig:cert}B provides the key metadata as a table generated
+from the \texttt{codecheck.yml} --
+which paper was checked (title, DOI), who the authors and the codechecker 
+were, when it was checked, and in which repository related material is
+accessible.
+Figure~\ref{fig:cert}C is a textual summary of how the CODECHECK was
 performed, and any interesting findings.
@@
 Figure~\ref{fig:cert}D (page 2 of the certificate) shows a table that
-lists the outputs that were generated from the codecheck (this is
-termed the MANIFEST).  For each output, there is a comment stating
-which figure/table it should be compared to in the original paper,
-along with the size of the file.  On page 3 of the certificate,
-Figure~\ref{fig:cert}E shows some more detailed notes from the
-codechecker, in this case indicating what steps were needed to
-initialize the environment and run the code.  This particular example
-was quite long, taking about 17 hours to run.  Finally, on page 4 of
+lists the outputs that were generated from the check.
+This table is generated from the manifest of output files in the CODECHECK
+configuration file. The manifest lists output files with a short description.
+The certificate's file table shows the file name, the description stating to
+which figure/table each file should be compared to in the original paper, and
+the size of the file.
+Page 3 of the certificate,
+as Figure~\ref{fig:cert}E shows, provides detailed notes from the
+codechecker. In this particular example, the notes document what steps were 
+needed to initialize the environment and run the code, and that the execution
+took quite long to run -- about 17 hours.  Finally, on page 4 of
 the certificate, we see the first output that was generated by the
-codecheck (Figure~\ref{fig:cert}F).  In this case, the figure matches
+CODECHECK (Figure~\ref{fig:cert}F).  In this case, the figure matches
 figure~4 of \cite{unwin_report_2020}.
-Subsequent pages of the certificate show other outputs (not shown here).
+Subsequent pages of the certificate show other outputs and the computing
+environment in which the certificate itself was created (not shown here).
 
@@
   \includegraphics[width=\textwidth]{figs/annotate-cert-crop.pdf}
-  \caption{Annotated certificate 2020-012 \cite{cert-2020-012} (first four pages only).}
+  \caption{Annotated certificate \texttt{2020-012} \cite{cert-2020-012} (first four pages only).}
   \label{fig:cert}

### 2021-01-12 codecheck-paper.tex
update tools

@@
 
-The \texttt{codecheckers} GitHub organisation\footnote{
-\url{https://github.com/codecheckers}} contains projects managing the
+The \texttt{codecheckers} \textbf{GitHub} organisation\footnote{
+\url{https://github.com/codecheckers}} contains projects for managing the
 project website, the community of codecheckers, community discussions,
-forks of codechecked code repositories, and the main register for
-managing the collection of checks. Both the website\footnote{
-\url{https://codecheck.org.uk/}} and the register
-\footnote{\url{https://codecheck.org.uk/register}} are hosted as GitHub
-pages. The register is effectively a single table in CSV format that
+forks of checked code repositories, and the main register of 
+CODECHECKs. Both the website\footnote{
+\url{https://codecheck.org.uk/}} and the register\footnote{
+\url{https://codecheck.org.uk/register}} are hosted as GitHub
+pages. The register is a single table in CSV format that
 connects the certificate identifier with the repository associated with
-a specific conducted codecheck. Each of these repositories, which 
+a specific conducted CODECHECK. Each of these repositories, which 
 currently can be hosted on GitHub or Open Science Framework (OSF), 
 contains the CODECHECK metadata file \texttt{codecheck.yml}. The register
-further contains a column for the type of check, e.g., community, journal,
+further contains a column for the type of check, e.g.,~community, journal,
 or conference, and the respective GitHub issue where communications and 
@@
 duplicated between the register and the metadata files. When a new
-certificate is entered into the register, an automated process using 
+certificate is entered into the register, an automated process using the
 continuous integration infrastructure of GitHub, GitHub Actions, retrieves
-the metadata files and builds different representations of the register
-for viewing (HTML) or integration with other services (CSV, JSON).
-
-Zenodo is an open repository for scientific data and documents. It
+all metadata files and builds different representations of the register
+for viewing (HTML) and for integration with other services (CSV, JSON).
+
+\textbf{Zenodo} is an open repository for scientific data and documents. It
 mints DOIs for deposits and ensures long-term availability of all digital
 artefacts related to the project. These include the community 
-certificates as well as the register, which is regularly archived to 
-Zenodo \cite{codecheck_register} (including all file formats).
-
-A custom R package, \texttt{codecheck}\footnote{
+certificates as well as the regularly archived register \cite{codecheck_register}.
+
+A \textbf{custom R package}, \texttt{codecheck}\footnote{
 \url{https://github.com/codecheckers/codecheck}}, helps with the authoring
-of certificates, and with the deposition of certificates and reated files to
+of certificates, and with the deposition of certificates and realed files to
 Zenodo using the R package \texttt{zen4R} \cite{zen4r}.
 Codecheckers may choose to not use the package and rely on their own tools
-for creating the certificate. This is a deliberate flexibility to
-accomodate both different skill sets and unforseen technical advances or
+for creating the certificate, even a simple word processor.
+This is a deliberate flexibility to
+accomodate different skill sets and unforseen technical advances or
 challenges.
@@
 managed without infrastructure costs on freely available platforms.
-The scripted and automated parts should also ensure that more checks can
+The automation should ensure that more checks can
 be conducted in the future, though the management by a paid CODECHECK
-editor or further automation, e.g., with a bot, could improve turnaround time.
-The remaining costs are the project's domain, which is negligible, and
-the human resources for managing the project and processes, which are
-considerable and currently solely based on, partly grant-based, public funding.
-Naturally, similar set-ups are possible, such as using GitLab/GitLab.com 
-and the GitLab CI instead of GitHub and GitHub Actions, or using a different 
-data repository for depositing certificates and data.
+editor or further automation, e.g.,~with a bot, could improve turnaround time.
+The only costs are the project's domain, which is negligible.
+This leaves the initiative's main resource -- the \textbf{humans}
+for managing the project and processes and the codecheckers.
+All contributions are currently rely on (partly grant-based) public funding
+and volunteering.
+Naturally, similar set-ups are conceivable, such as using GitLab.com or
+a hosted GitLab instance and the GitLab continuous integration.
 

### 2021-01-12 codecheck-paper.tex
restructure related work and resolve open issues

@@
 as CODECHECK uses "reproducible").
+A search on Crossref\footnote{\url{https://search.crossref.org/?q=\%22Replicated+Computations+Results+\%28RCR\%29+Report\%22&from_ui=yes&page=2}; search executed
+on 2020-12-10.} shows 15 RCR Reports to be published and the 
+% Number of found TOMAS RCR reports:
+% 8: https://scholar.google.de/scholar?start=10&q=Replicated+Computations+Results+%22(RCR)+Report%22&hl=de&as_sdt=0,5
+% 15: https://search.crossref.org/?q=Replicated+Computations+Results+%22%28RCR%29+Report%22&from_ui=yes&publication=ACM+Transactions+on+Mathematical+Software
+% 15: https://academic.microsoft.com/search?q=Replicated%20Computations%20Results%20%22(RCR)%20Report%22&f=&orderBy=0&skip=10&take=10
+workflow to be 
+extended to the ACM journal \emph{Transactions on Modeling and Computer 
+Simulation (TOMACS)}\footnote{\url{https://dl.acm.org/journal/tomacs/author-guidelines}}.
+
 The CODECHECK principles were independently developed as we only later 
@@
 1 to 4.
-Our CODECHECK provides more extensive discussion of the possible variants, 
+This work provides an extensive discussion of the possible variants, 
 while TOMS RCR realises one particular workflow for reproducibility 
-reviews for one particular journal. \cite{heroux_editorial_2015}
-include similar concerns about
+reviews for one particular journal. The TOMS editorial
+\cite{heroux_editorial_2015} shares similar concerns about
 selection of reviewers as above. Unlike existing CODECHECK certificates, 
 the RCR reports undergo editorial review, but this could very well become
-part of CODECHECK workflows at journals.
+part of CODECHECK processes at journals.
 The publication of the RCR report is seen as a reward for the efforts of
@@
 motivation is acknowledged. TOMS also recognises reviewer activity in a 
-partnership with Publons
-\footnote{\url{https://authors.acm.org/author-services/publons}},
-though it is unclear if the author of an RCR report receives both the report
+partnership with Publons\footnote{
+\url{https://authors.acm.org/author-services/publons}},
+though it is unclear if the author of an RCR report receives both the report's
 authorship and the reviewer recognition. In any case, this is different
 from CODECHECK, which specifically aims at recognition as reviewers.
-In our view, this removes the possible conflict of interest while keeping
-the public acknowledgement.
-
-Specific to the field of Mathematics, the RCR was also  expected to apply a
+In our view, the reviewer role removes the possible conflict of interest while
+keeping the public acknowledgement.
+Specific to the field of Mathematics, the RCR is also expected to apply a
 review of the software itself if the system it runs on cannot be evaluated by
 an independent party.
-The TOMS RCR also notes the importance of communication and expects
-collaboration between author and RCR reviewers.
-The TOMS creators share thoughts around reviewer selection and also put trust
+The TOMS RCR creators concur with the importance of communication, expect
+collaboration between author and RCR reviewers,
+share the considerations around reviewer selection, and also put trust
 in reviewer judgement over numerical bit-wise perfection.
@@
 published in the TOMS journal next to the actual papers.
-A search on Crossref\footnote{\url{https://search.crossref.org/?q=\%22Replicated+Computations+Results+\%28RCR\%29+Report\%22&from_ui=yes&page=2}; search executed
-on 2020-12-10} shows 15 RCR Reports to be published and the workflow to be 
-extended to the ACM journal \emph{Transactions on Modeling and Computer 
-  Simulation (TOMACS)}
-%% SJE TODO - fix footnote
-%% \footnote{\url{https://dl.acm.org/journal/tomacs/author-guidelines-%35-rcrini}}.
-
-% 8: https://scholar.google.de/scholar?start=10&q=Replicated+Computations+Results+%22(RCR)+Report%22&hl=de&as_sdt=0,5
-% 15: https://search.crossref.org/?q=Replicated+Computations+Results+%22%28RCR%29+Report%22&from_ui=yes&publication=ACM+Transactions+on+Mathematical+Software
-% 15: https://academic.microsoft.com/search?q=Replicated%20Computations%20Results%20%22(RCR)%20Report%22&f=&orderBy=0&skip=10&take=10
-
-A number of journals or conferences/workshops provide guidelines for ensuring reproducible publications and their verification.
-These include, for example, \ldots{}
+
+A number of journals provide guidelines for 
+ensuring reproducible publications and their verification,
+e.g.,~\emph{Information Systems} with its invited 
+reproducibility paper track\footnote{
+\url{https://www.elsevier.com/journals/information-systems/0306-4379/guide-for-authors}.}.
 % first glimpse into "Code exeuction processes in scholarly publishing" ?
 % `https://docs.google.com/document/d/1fMWrFvBTwAYXg5J_bYIjgb-gY7DfQ0DOjx9Vlg_Rbmg/edit# ?
-\emph{Information Systems}, with an invited reproducibility paper track
-\footnote{\url{https://www.elsevier.com/journals/information-systems/0306-4379/guide-for-authors}}
-
-Another journal initiative is more indirectly connected with reproductions.
 \emph{Nature Machine Intelligence} recently introduced a new type of article,
@@
 of CODECHECK.
-
-However, no approach besides CODECHECK spans across journals and is
+The journal \emph{Cortex} has a special article type 
+\emph{Verification Reports}, which  actually are about replication of results 
+but are very well designed/reasoned \cite{chambers_verification_2020}.
+In a similar vein, the CODECHECK certificates
+could also be published as a special article type within journals.
+
+Going beyond article types, the journal \emph{ReScience C} publishes only
+replications\footnote{\url{https://rescience.github.io/faq/}} requiring
+open code and replication by a third party to. ReScience also relies on
+free infrastructure (GitHub and Zenodo).
+For research with very high stakes, where reproduction would be too weak and
+post publication replication possibly too late because of policy impact,
+Benjamin-Chung et~al. \cite{benjamin-chung_internal_2020} propose
+\emph{internal replication}.
+A workflow that underwent internal
+replication would very likely be of high quality and relatively easy to check.
+Similarly, internal CODECHECKs may be used, with the same limitations such
+as group think \cite{benjamin-chung_internal_2020},
+to ensure reproducibility before submission.
+Such internal checks are professionalised in local reproduction services,
+such as CISER R-squared\footnote{
+\url{https://ciser.cornell.edu/research/results-reproduction-r-squared-service/}},
+} (established/permanent),
+%Uni Münster's Reproducible Research Support Service\footnote{
+%\url{https://go.wwu.de/r2s2}} (pilot),
+Oxford's code review network (OxCRN)\footnote{
+\url{https://github.com/OxfordCodeReviewNet/forum}} (community, broader scope), or
+Yale's YARD\footnote{
+\url{https://isps.yale.edu/research/data/approach}} (staff for replications).
+
+Gavis and Donoho \cite{gavish_universal_2011} propose a new discipline and 
+infrastructure for reproducible computational research with its own tools and
+API. Their specific packaging format, provenance record, and cryptographic
+\emph{Verifiable Result Identifier} would indeed provide excellent
+reproduciblity. However, the system is also complex and was not taken up by
+any publisher as far as we know; the system is indeed not open source itself.
+In comparison, CODECHECK is less powerful but also much more flexible and
+less dependent on specific tools or infrastructure. If data and code are 
+deposited properly, i.e.,~very unlikely to disappear, then the certificate's 
+DOI is practically close to the cryptographic identifier.
+
+\emph{SciGen.Report} is a community-run platform to foster communication on
+reproducibility\footnote{\url{https://scigen.report/}}. People can report on
+fully, particially or un-successful reproductions post publication.
+
+No approach besides CODECHECK spans across journals and is
 intentionally designed to be picked up and implemented by many different 
 journals or events, and to build a community of codecheckers. CODECHECK shares
-this interdisciplinary nature with other community initiatives concerned with 
+its interdisciplinary nature with other community initiatives concerned with 
 reproducibility awareness, education, and supports, such as ReproHack
@@
 Likewise, different disciplines and journals provide reproducibility
-checklists, e.g., \cite{rosenberg_next_2020}, which naturally share some
+checklists, e.g.,~\cite{rosenberg_next_2020}, which naturally share some
 aspects while addressing particularities as well as addressing researchers
@@
 we see CODECHECK's role in referencing and linking educational efforts
-and helpful material, not as a creator and maintainer of content.
-
-The journal \emph{ReScience C} publishes replications
-\footnote{\url{https://rescience.github.io/faq/}};
-they do not accept proprietary software and require open source code;
-they also required a third party to replicate; they also use Zenodo
-to issue DOIs.
-The journal \emph{Cortex} has a special article type 
-\emph{Verification Reports}, which  actually are about replication of results 
-but are very well designed/reasoned \cite{chambers_verification_2020}.
-In a similar vein, the CODECHECK certificates
-could also be published as a special article type within journals.
-
-For research with very high stakes, where reproduction would be too weak and
-post publication replication possibly too late because of policy impact,
-\cite{benjamin-chung_internal_2020} propose \emph{internal replication}.
-A workflow that underwent internal
-replication would very likely be of high quality and relatively easy to codecheck.
-In a similar vein, internal codechecks may be used, with the same limitations such
-as group think \cite{benjamin-chung_internal_2020},
-to ensure reproducibility before submission. The guidelines
-for CODECHECK can be used by fellow/not independent lab members just as well.
-
-Gavis and Donoho \cite{gavish_universal_2011} propose a new discipline and 
-infrastructure for reproducible computational research with its own tools and
-API. Their specific packaging format, provenance record, and cryptographic
-\emph{Verifiable Result Identifier} would indeed provide excellent
-reproduciblity. However, the system is also complex and was not taken up by
-any publisher as far as we know; the system is indeed not open source itself.
-In comparison, CODECHECK is less powerful but also much more flexible and
-less dependent on specific tools or infrastructure. If data and code are 
-deposited properly, i.e., very unlike to disappear, then the certificate's 
-DOI is practically quite close to the cryptographic identifier.
-
-Local reproduction initiatives/services:
-\href{https://ciser.cornell.edu/research/results-reproduction-r-squared-service/}{CISER R-squared}
-(established/permanent),
-\href{https://go.wwu.de/r2s2}{WWU R2S2} (pilot only, focused on reproductions),
-\href{https://github.com/OxfordCodeReviewNet/forum}{Oxford code review network (OxCRN)}
-(community, broader scope), \href{https://isps.yale.edu/research/data/approach}{YARD 
-at Yale} with dedicated staff for replications.
-For these, the CODECHECK scope, tools, and practical experiences may be
-helpful resources.
-
-\emph{SciGen.Report} is a community-run platform to foster communication on
-reproducibility\footnote{\url{https://scigen.report/}}. People can report on
-fully, particially or un-successful reproductions post publication.
+and helpful material, not as a creator and maintainer of such content.
 

### 2021-01-12 codecheck-paper.tex
minor formatting changes

@@
 as part of the widely accepted peer review concept, can provide may
-eventually allow to raise the bar as high as possible, e.g., with
+eventually allow to raise the bar as high as possible, e.g.,~with
 executable research compendia \cite{nust_opening_2017}
@@
 have repeatedly been identified as main barriers to reproducibility
-(e.g., \cite{stagge_assessing_2019,nust_improving_2020}),
+(e.g.,~\cite{stagge_assessing_2019,nust_improving_2020}),
 we would not underestimate the power of a simple check.
@@
 statistical methods \cite{greenwood_how_2015} and can support topical
-reviewers if, e.g., complex methods or sophisticated variants of statistical
+reviewers if, e.g.,~complex methods or sophisticated variants of statistical
 tests are applied \cite{petrovecki_role_2009}.
@@
 that a reproduction fails. This may be the sign of a pre-selection bias
-of the examples, as so far all our codechecks succeeded.
+of the examples, as so far all our CODECHECKs succeeded.
 In case of a journal adopting CODECHECK for all submissions, the open 
 question remains as what to do in this case will have to be answered.
-At the start, we doubt publicly reporting failures (i.e., the code 
+At the start, we doubt publicly reporting failures (i.e.,~the code 
 would not run) will increase overall reproducibility and recommend not
@@
 for successes seems to be the way to go.
-% eLife's new model (see https://osf.io/9cftx/) will publish reports evantually, e.g., after preprint has been accepted elsewhere
+% eLife's new model (see https://osf.io/9cftx/) will publish reports evantually, e.g.,~after preprint has been accepted elsewhere
 \cite{Rosenthal2016b} discuss incentives for different actors around the
@@
 but not raw data \cite{shannon_opening_2018},
-i.e., data enclaves \cite{foster_research_2018},
+i.e.,~data enclaves \cite{foster_research_2018},
 or domains of reproducibility \cite{harris_more_2017})
@@
 need checking again as it is modified during peer review.  This is in
-part unavoidable, and as happened to us (e.g., certificate 2020-012
+part unavoidable, and as happened to us (e.g.,~certificate \texttt{2020-012}
 \cite{cert-2020-012} has been revised). This can also be desired, if
@@
 As a principle, we do not require results to be exactly the same for a
-codecheck to pass, simply that the code runs and generates the output
+CODECHECK to pass, simply that the code runs and generates the output
 files that the author claims. Stochastic simulations mean that often
@@
 \begin{itemize}
-\item modelDB principle: minimum of one figure should be reproducibe.
-  Example certificate 2020-016 (I think) which had only a small subset
+\item TODO SJE: modelDB principle: minimum of one figure should be reproducibe.
+  Example certificate \texttt{2020-016} (I think) which had only a small subset
   of the possible figures. Still useful and worth sharing.

### 2021-01-12 codecheck-paper.tex
update limitations

@@
 \textbf{Isn't CODECHECK what peer review should be doing already?}
-That is correct.  However, the general observation is that the
+On the surface, that is correct.
+However, the general observation is that the
 increasing number of researchers, submissions, and reviews to conduct
 under the pressure to publish, puts the scholarly publication system
-under a lot of strain.  We also see parallels to the terms
-`reproducible research' and `open science'. Establishing a CODECHECK
+under a lot of strain.  
+Miyakawa described how simply requesting raw data from authors as an 
+editor lead to withdrawal or rejection of nearly all submissions where the
+request was necessary \cite{miyakawa_no_2020}. Such basic checks that
+sufficient data supports the results requires large efforts from editors,
+yet they can be based on availability of data files and the content of the
+paper. These checks can be enhanced by more complex CODECHECKs asking for
+the code, and, because code is not inert but needs to be run in an specific
+environment, executing the code.
+This might fall within the idealistic expectations for peer review,
+but is far from common in reality.
+Establishing a CODECHECK
 process is largely an acknowledgement that peer reviewing practices
-have faltered to adapt to the challenges of digitisation and
-computer-based research or data science. The concept of a CODECHECK,
+have faltered to adapt to the challenges of digitisation data science.
+The concept of a CODECHECK,
 just as the reproducibility of research and the openness of science,
@@
 
-% LocalWords:  CODECHECK CODECHECKER Zenodo ORCID ECRs codecheck
+% LocalWords:  CODECHECK CODECHECKER Zenodo ORCID ECRs codecheck codechecker

### 2021-01-14 codecheck-paper.tex
trust vs. transparency

@@
 Nevertheless, the simple attempt to reproduce may highlight shortcomings in
-a submission with respect to transparency as it may be required by a journal's
-policy and help to resolve misunderstandings and discrepancies
-(cf.~\cite{christian_journal_2020}).
+a submission to meet journal policy requirements
+(cf.~\cite{christian_journal_2020})
+and can effectively increase transparency and thereby improve practices
+(cf.~\cite{nosek_scientific_2012})
+even if the check goes not into all details.
 
@@
 journal's target audience to reach a state where the level of verification 
-and trust provided by codecheckers becomes part of peer review and regular
+ensured by codecheckers becomes part of peer review and regular
 merit of all accepted papers.
@@
 at conferences.
-CODECHECKs can increase trust in scientific results.
+CODECHECKs increase transparency of the checked papers and can contribute
+to building trust in scientific results.
+% "science earns trust through transparency" > https://twitter.com/briannosek/status/1290240244101521408
 The set of 

### 2021-01-20 codecheck-paper.tex
last missing acknowledgement confirmed

@@
 We are grateful to the following individuals for discussions regarding
-the work presented here: Andy Collings, Melissa Harrison?, Giuliano
+the work presented here: Andy Collings, Melissa Harrison, Giuliano
 Maciocci, Naomi Penfold (eLife), Rebecca Kirk (PLOS Computational Biology), Scott

### 2021-01-25 codecheck-paper.tex
fix some warnings in bibliography, fix latex error
add icons to authors

@@
 
+\usepackage{academicons} % for ORCID icon, via https://tex.stackexchange.com/a/534476/212961
+\definecolor{idcolor}{HTML}{A6CE39}
+\newcommand{\orcid}[1]{\href{https://orcid.org/#1}{\color{idcolor}\aiOrcid}}
+
+\newcommand{\email}[1]{\href{mailto:#1}{#1}}
+
+\usepackage{fontawesome}
+\newcommand{\github}[1]{\href{https://github.com/#1}{\faGithub}}
+\newcommand{\twitter}[1]{\href{https://twitter.com/#1}{\faTwitter}}
+
 \onehalfspacing
@@
 \title{CODECHECK: An open-science initiative for the independent execution of computations underlying research articles}
-\author[1,$\ast$]{Daniel N\"{u}st}
-\author[2,$\ast$]{Stephen J Eglen}
-\affil[1]{\url{https://orcid.org/0000-0002-0024-5046}}
-\affil[2]{\url{https://orcid.org/0000-0001-8607-8025}}
+\author[1,$\ast$]{Daniel~N\"{u}st \orcid{0000-0002-0024-5046} \github{nuest} \twitter{nordholmen}}
+\author[2,$\ast$]{Stephen~J~Eglen \orcid{0000-0001-8607-8025} \github{sje30} \twitter{StephenEglen}}
+\affil[1]{Institute for Geoinformatics, University of M\"unster, Germany; \email{daniel.nuest@uni-muenster.de}}
+\affil[2]{Department of Applied Mathematics and Theoretical Physics, University of Cambridge, GB; \email{sje30@cam.ac.uk}}
 \affil[$\ast$]{Contributed equally}
@@
 such as CISER R-squared\footnote{
-\url{https://ciser.cornell.edu/research/results-reproduction-r-squared-service/}},
-} (established/permanent),
+\url{https://ciser.cornell.edu/research/results-reproduction-r-squared-service/}} (established/permanent),
 %Uni Münster's Reproducible Research Support Service\footnote{

### 2021-01-25 codecheck-paper.tex
fix last few TODOs

@@
 as as a conversation starter for potential collaborating journals.
+This intention worked out with the first CODECHECK (certificate number
+\texttt{2020-001}) before publication of an article for the journal
+\emph{GigaScience}, which successfully reproduced the visualisations
+though not the full machine learning process.
 
@@
 have since then been acknowledged in reviewed papers
-\cite{Davies2020-vj,kucharski_effectiveness_2020}.
+\cite{Davies2020-vj,kucharski_effectiveness_2020}. Unlike many claims in
+popular media at the time,
+% reference missing here TODO SJE, you had a note about some Noorden news piece
+we found the results of Report 9, a model of the UK lockdown procedures,
+to be reproducible (certificate number \texttt{2020-10}).
 
@@
 for GIScience journals.
-
-Out of these areas, we would like to briefly highlight three examples here.
-\textbf{TODO SJE: tidy this paragraph.}
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  Gigascience -- our first paper.  Only visualisation, not machine
-  learning (which would have taken several days of compute time).
-\item
-  report 9 -- covid model of UK lockdown model -- unlike many claims
-  in the popular media at the time, the model was reproducible
-  [NOORDEN news piece]. 
-\item
-  Manuel's certificate -- approached us as he requested a certificate
-  as paper was entering peer review.
-\end{enumerate}
 
@@
 established a "Replicated Computational Results" (RCR) review process
-\cite{heroux_editorial_2015} (using "replicable" with the same meaning 
-as CODECHECK uses "reproducible").
+\cite{heroux_editorial_2015}, using "replicable" with the meaning 
+as this work calls "reproducible".
 A search on Crossref\footnote{\url{https://search.crossref.org/?q=\%22Replicated+Computations+Results+\%28RCR\%29+Report\%22&from_ui=yes&page=2}; search executed
@@
 Simulation (TOMACS)}\footnote{\url{https://dl.acm.org/journal/tomacs/author-guidelines}}.
-
 The CODECHECK principles were independently developed as we only later 
@@
 we would not underestimate the power of a simple check.
+ModelDB curation policies ensure manually to reproduce one figure only 
+\cite{mcdougal_reproducibility_2016}, but that has not limited the usefulness
+nor success of the platform.
 
@@
 
-\begin{itemize}
-\item TODO SJE: modelDB principle: minimum of one figure should be reproducibe.
-  Example certificate \texttt{2020-016} (I think) which had only a small subset
-  of the possible figures. Still useful and worth sharing.
-\end{itemize}
-
 \textbf{Not revolutionary enough?}

### 2021-01-25 codecheck-paper.tex
add abstract, change link colours, update title

@@
 \usepackage{hyperref}
+\hypersetup{colorlinks=false, pdfborderstyle={/S/U/W 1}, pdfborder=0 0 1}
 \usepackage{url}
@@
 \begin{document}
-\title{CODECHECK: An open-science initiative for the independent execution of computations underlying research articles}
+%\title{CODECHECK: An open-science initiative for the independent execution of computations underlying research articles}
+\title{CODECHECK: An open-science initiative for the independent execution of computations underlying research articles during peer review improves reproducibility}
 \author[1,$\ast$]{Daniel~N\"{u}st \orcid{0000-0002-0024-5046} \github{nuest} \twitter{nordholmen}}
@@
 \affil[$\ast$]{Contributed equally}
-\affil[$$]{Public project for this article is at \url{https://github.com/codecheckers/paper}}
+\affil[$$]{Public project for this article is at \url{https://github.com/codecheckers/paper}.}
 \maketitle
 \begin{abstract}
-  This is a draft paper that we hope to submit to F1000R (and perhaps
-  as a preprint before) in the coming weeks.
-  \ldots{}
+  The scientific paper falls short of effectively communicating computational research.
+  To break out of the stalemate between multiple lacks of incentives, guidance, benefits, recognitions, and reproducibility, we propose to introduce execution of workflows underlying research articles as part of peer review.
+  The CODECHECK system relies on five principles and can be integrated into review and publications processes in multiple ways, which we describe along multiple dimensions (importance, who, openness, when).
+  We demonstrate CODECHECK with 25 reproductions of diverse scientific publications in collaboration with academic publishers and conferences.
+  These CODECHECKs show that asking for reproducible workflows in a collaborative review can effectively improve executability.
+  CODECHECK relies on free infrastructure and open tools.
+  While it has clear limitations, it can be a building block in the open-science and publishing ecosystems to improve the reproducibility, appreciation, and potentially quality of non-textual research artefacts.
 \end{abstract}

### 2021-02-01 codecheck-paper.tex
small edits by DN before revision

@@
 by the author, and
-(3) the CODECHECKER who checks that the authors's code works.
-The workflow that we have refined is documented in Figure~\ref{fig:workflow}
+The workflow we have refined is shown in Figure~\ref{fig:workflow}
 and consists of six steps.
@@
 and link between the certificate, the paper, and any repositories 
-through metadata. Currently, the CODECHECKER creates these on Zenodo\footnote{
+through metadata. Currently, the CODECHECKER creates these connections on
+Zenodo \footnote{
 See ``related identifiers'' and ``alternate identifiers'' in the right-hand
@@
 See peer review contributions in ORCID records at
-\url{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}.},
+\url{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}.}.
 The PUBLISHER also ensures proper publication metadata, e.g.,~links from the 
@@
 is of course a very coarse indicator for the community contribution by an 
-individual, yet it is unclear if a regular reviewer, who in addition
+In this case, the published CODECHECK certificate and possibly the published peer review can clearly describe the possibly large efforts put into both.
 conducts a CODECHECK to the scientific review, should be credited twice.
@@
 The public record can be realised by publishing the certificate in a 
-citable form (i.e.,~with a DOI; this is done within the community codecheck
+citable form (i.e.,~with a DOI, as conducted in the community CODECHECK
 process), by listing codecheckers on the journal's website or, ideally, by
@@
 
-The third theme represents DN's service as Reproducibility
-Reviewer at the AGILE conference series.
-Independent of CODECHECK, the \emph{Reproducible AGILE} Initiative \cite{reproducible_agile}
-established a process for reproducing workflows at the AGILE conference
-series \cite{nust_improving_2020}.
+The third theme represents DN's service as a Reproducibility Reviewer at the AGILE conference series, where the \emph{Reproducible AGILE} Initiative \cite{reproducible_agile} established a process for reproducing workflows at the AGILE conference series independent of CODECHECK \cite{nust_improving_2020}.
 While using slightly different terms and infrastructure (``reproducibility
-reports'' are published on OSF instead of certificates on Zenodo) the 
-AGILE reproducibility reviews adhere to the codecheck principles.
+reports'' are published on the Open Science Framework (OSF) instead of certificates on Zenodo) the 
+AGILE reproducibility reviews adhere to the CODECHECK principles.
 Furthermore, a small number of checks was completed as part of a peer reviews
@@
 specifically GitHub and Zenodo.
-
 The \texttt{codecheckers} \textbf{GitHub} organisation\footnote{
@@
 for viewing (HTML) and for integration with other services (CSV, JSON).
-
 \textbf{Zenodo} is an open repository for scientific data and documents. It
 mints DOIs for deposits and ensures long-term availability of all digital
-artefacts related to the project. These include the community 
-certificates as well as the regularly archived register \cite{codecheck_register}.
+artefacts related to the project. The CODECHECK community on Zenodo
+\footnote{\url{https://zenodo.org/communities/codecheck/}} includes 
+certificates as well as the regularly archived register
+\cite{codecheck_register_jan2021}.
 
 A \textbf{custom R package}, \texttt{codecheck}\footnote{
-\url{https://github.com/codecheckers/codecheck}}, helps with the authoring
-of certificates, and with the deposition of certificates and realed files to
-Zenodo using the R package \texttt{zen4R} \cite{zen4r}.
+\url{https://github.com/codecheckers/codecheck}}, is used to automate repetitive
+tasks around the authoring of certificates and the management of the register.
+For example, it includes scripts for the deposition of certificates and related 
+files to Zenodo using the R package \texttt{zen4R} \cite{zen4r} and for the
+register update process outlined above.
 Codecheckers may choose to not use the package and rely on their own tools
-for creating the certificate, even a simple word processor.
+for creating the certificate, even a regular word processor.
 This is a deliberate flexibility to
@@
 
-\textbf{CODECHECK requirements are not demanding enough}:
+\textbf{Aren't CODECHECK requirements not demanding enough?}
 CODECHECK by design does not require authors to provide (and sustain) an
@@
 and JOSS\footnote{\url{https://joss.theoj.org/}}.
-
 Furthermore, the codechecker's task list is intentionally not
-overloaded with related concerns, even though they are very important,
-such as ensuring proper citation of data and software, or depositing
-data and software in suitable repositories. We do expect codecheckers
-to be aware of these though and possibly highlight them.
-
-\textbf{Handling failure}: We have no established process for the case
+overloaded with related concerns, albeit potentially important ones,
+such as ensuring proper citation of data and software or depositing
+data and software in suitable repositories. Nevertheless,
+we expect codecheckers to often recognise these issues and highlight them.
+
+\textbf{How are failures during checks handled?}
+We have no established process for the case
 that a reproduction fails. This may be the sign of a pre-selection bias
@@
 
-\textbf{Compute time}: for those papers that take significant compute
-time (think days, not minutes), who will pay for the compute time?
+\textbf{Who will pay for the compute time?}
+For those papers that take significant compute
+time (days, not minutes), it is unclear who will pay for it.
 The openness of CODECHECK certificates could help to establish community
-peer pressure, as submitting authors can be evaluated on their engagement
-as codecheckers.
+peer pressure, or recognition, as submitting authors can be evaluated on their engagement
+as codecheckers. This may lead to a balance of the shared load to reproduce
+computations.
+However, one must carfully consider the sustainability of rerunning computations and the
+environmental impact large calculations, such as training machine learning models, have.
 The challenges of computational resources can also quickly be reduced when
@@
 
-\textbf{Proprietary software}: authors can currently provide code that requires
+\textbf{What about my proprietary software and sensitive data?}
+Authors can currently provide code that requires
 proprietary software. Given the prevalence of proprietary software in
@@
 (cf.~\cite{konkol_publishing_2020,perkel_make_2019})
-and requires the codechecker to have access to that software.
+and requires the codechecker to have access to that particular software.
 Non-open software also considerably hampers reuse, especially by researchers
@@
 
-\textbf{Not possible for my data}: 
 Solutions for proprietary and sensitive data exists.
@@
 the time spent with CODECHECKs likely reduces the individual's time to work
-as a reviewer.
+First, the \textbf{technical nature} of a CODECHECK can provide a more 
 However, there are differences to peer review.
 First, the technical nature of a CODECHECK can provide a more 
-clearly defined scope and expectations compared to conventional 
+clearly defined scope and expectations and thereby time budget compared to conventional 
 peer review. The authors are told what to provide and the codechecker
@@
 material/information can greatly reduce the digging required and thereby
-the time spend.
-We also found that the focus on the mechanics of a
+increase \textbf{efficiency}.
+Fourth, we also found that the focus on the mechanics of a
 workflow and the ability to interact with the authors can make
-reproductions very educational. It's a different role and as such could
+reproductions very \textbf{educational}. It's a different role and as such could
 be a welcome alternation/diversion for researchers to give back their 
@@
 \cite{Gronenschild2012-pp}.
+While reproducibility practices can mitigate some of these problems, e.g., by using a seed, the flexibility of the human judgement is still needed.
 
@@
 We have discussed stronger more drastic rules at length, such as only 
-considering diamond Open Access journals, community-led no-fee journals,
-or journals with collaborative/open peer review, but eventually decided
-against them.
+considering fully free and open source software, diamond Open Access journals or community-led no-fee journals,
+and journals with collaborative/open peer review, but we eventually decided
+against them on the level of the principles.
+For the CODECHECK community process\footnote{\url{https://codecheck.org.uk/guide/community-process}}
+and the volunteer codechecker community, these requirements are reconsidered.
+
 We also have considered to require latest technologial solutions which can
@@
 The structurelessness of the required data and software gives the needed
-freedom to enable all types of research. It could be complemented by
-automated scoring (e.g.,~\cite{menke_rigor_2020}, yet automation and 
-metrics bears risks that lead us to put our trust in the codechecker.
-Note that the focused nature of the CODECHECK principles allows journals
+CODECHECK could be complemented by automated scoring (e.g.,~\cite{menke_rigor_2020}), yet automation and metrics bear new risks.
+Note that the focus of the CODECHECK principles on code execution allows journals
 and publishers to innovate on the financial model and peer review practices

### 2021-02-01 codecheck-paper.tex
integrate editorial review by CB
also adjust wording in figure

@@
 %\title{CODECHECK: An open-science initiative for the independent execution of computations underlying research articles}
-\title{CODECHECK: An open-science initiative for the independent execution of computations underlying research articles during peer review improves reproducibility}
+\title{CODECHECK: An Open Science initiative for the independent execution of computations underlying research articles during peer review improves reproducibility}
 \author[1,$\ast$]{Daniel~N\"{u}st \orcid{0000-0002-0024-5046} \github{nuest} \twitter{nordholmen}}
@@
 \begin{abstract}
-  The scientific paper falls short of effectively communicating computational research.
-  To break out of the stalemate between multiple lacks of incentives, guidance, benefits, recognitions, and reproducibility, we propose to introduce execution of workflows underlying research articles as part of peer review.
-  The CODECHECK system relies on five principles and can be integrated into review and publications processes in multiple ways, which we describe along multiple dimensions (importance, who, openness, when).
-  We demonstrate CODECHECK with 25 reproductions of diverse scientific publications in collaboration with academic publishers and conferences.
-  These CODECHECKs show that asking for reproducible workflows in a collaborative review can effectively improve executability.
-  CODECHECK relies on free infrastructure and open tools.
-  While it has clear limitations, it can be a building block in the open-science and publishing ecosystems to improve the reproducibility, appreciation, and potentially quality of non-textual research artefacts.
+  In its traditional format, the scientific paper falls short of effectively communicating computational research.
+  Since this issue is difficult to address due to lacking incentives, guidance, benefits, recognitions, and reproducibility, we come at this problem from a different angle:
+  We propose to introduce, as part of peer review, a system by which the workflows underlying research articles are executed.
+  The CODECHECK system, which uses free intrastructure and open tools, relies on five principles and can be integrated into review and publication processes in multiple ways.
+  We describe these integrations along multiple dimensions (importance, who, openness, when).
+  In collaboration with academic publishers and conferences, we demonstrate CODECHECK with 25 reproductions of diverse scientific publications.
+  These CODECHECKs show that asking for reproducible workflows during a collaborative review can effectively improve executability.
+  While CODECHECK has clear limitations, it may represent a building block in the Open Science and publishing ecosystems for improving the reproducibility, appreciation, and, potentially, the quality of non-textual research artefacts.
 \end{abstract}
@@
 
-Many areas of scientific research use computations to either simulate
-or analyse their data. These computations are increasingly complex, and
-difficult to explain coherently in a paper \citep{marwick_how_2015}.
+Many areas of scientific research use computations to either simulate or analyse their data.
+These computations are becoming increasingly complex and are difficult to explain coherently in a paper \citep{marwick_how_2015}.
 To complement the traditional route of sharing research by writing papers,
@@
 (see Figure~\ref{fig:inverse}).
-Some of the earliest proponents were Buckheit and Donoho
-\cite{buckheit_wavelab_1995} who coined what has been called 
-\emph{Claerbout's claim} (extending on \citet{claerbout_electronic_1992}):
-\emph{"An article about computational science in a scientific publication 
+Some of the earliest proponents of this initiative were Buckheit and Donoho \cite{buckheit_wavelab_1995}, who coined what has been called \emph{Claerbout's claim} (extending on \citet{claerbout_electronic_1992}):
+\emph{``An article about computational science in a scientific publication 
 is \text{not} the scholarship itself, it is merely \textbf{advertising} of
@@
 environment and the complete set of instructions which generated the 
-figures."}
-%  Donoho (2010) has another paraphrasing of the same quote, which might work better:
-% "an article about computational result is advertising, not scholarship. The actual scholarship is the full software environment, code and data, that produced the result."
-
-So, assuming that researchers begin to share more artefacts, how might
-they be examined to test that they do what they claim? For example,
+figures.''}
+
+So, assuming that researchers begin to share more artefacts, how might these artefacts be examined to ensure that they do what they claim?
+For example,
 most scientific journals now require a data sharing statement that
 outlines what data the authors have (or will) share.
-The implementation status of this varies considerably according to the
-journal. At one end of the spectrum, there are specialist journals that
-have been created to accept ``data papers''
-(e.g.~\href{https://www.nature.com/sdata/}{\emph{Scientific~Data}}, 
+Yet, each journal can implement this requirement differently.
+At one end of the spectrum, journals have been created to accept ``data papers''
+(e.g.,~\href{https://www.nature.com/sdata/}{\emph{Scientific~Data}}, 
 \href{https://essd.copernicus.org/}{\emph{Earth System Science Data}}, 
@@
 \href{https://odjar.org/}{\emph{Open Data Journal for Agricultural Research}},
-\href{https://openhealthdata.metajnl.com}{\emph{Journal of Open Health Data}}).
-These journals have established rigorous procedures by which
-the data is validated according to standards in each field. At the
-other end of the spectrum, we still witness today the infamous
-statement ``Data available upon reasonable request''. Authors,
-while possible well-intentioned at the time of writing the article,
-often cannot provide the data when readers ask for
-it as data disappears over time \cite{Vines2014-hf}.
-
-Given that there are no clear standards yet for sharing data, what hope
-might there be for sharing computer programs? While both data and software
-together are required to validate outcomes of a computational analysis,
-they do differ in  that data can be seen as static and inert,
+\href{https://openhealthdata.metajnl.com}{\emph{Journal of Open Health Data}});
+these journals have established rigorous procedures by which
+the data is validated according to standards in each field.
+At the other end of the spectrum, many journals still simply allow authors to state ``Data available upon reasonable request''.
+Authors, while possibly well intentioned at the time of writing the article, often cannot provide the data when readers ask for it, as data disappears over time \cite{Vines2014-hf}.
+
+Given that no clear standards yet exist for sharing data, what hope might there be for sharing computer programs?
+Both data and software together are required to validate outcomes of a computational analysis, but they differ in that data can be seen as static and inert,
 while code requires an environment to be run in and interacted with.
 This makes software harder to share.
-Our anectodal experiences
-on this matter suggest that often a variety of reasons are given for why
-code cannot be shared, e.g.,~``there is no documentation'', ``I cannot 
-maintain it'', or ``I do not want to give away my code to competitors''.
-Our view, rooted in Open~Science principles
-\footnote{\url{https://en.wikipedia.org/wiki/Open_science}}, is that
-wherever possible sharing code is good for the community and the 
-individual \cite{Barnes2010-iv,markowetz_five_2015}.
-Having code and data openly available and archived, 
-provides a valuable resource for others to learn
-from, even when it doesn't run anymore or if there is no documentation.
-However,
-with a little effort, we believe that if someone independent can execute
-an analysis, this is worth documenting as early as possible and that the barrier
-to evaluating non-text materials of research can be greatly reduced with a 
-focused approach.
-Just like data journals' validations of data and all journals' peer review
-can provide a ``baseline reassurance'', i.e.,~a paper
-has been checked by someone with an understanding of the topic
-\cite{fyfe_mission_2019}, we think the same baseline should be provided
-regarding the workflow underlying a paper.
-With this in mind, we have developed a set of principles and an example 
-workflow that provides what he hope is a pragmatic way of checking that a
-paper's code works.
-
-The main contribution of this work is a thorough description of a process
-and its variations to integrate a much needed evaluation of the 
-computational reproducibility\footnote{Following the 
-Claerbout/Donoho/Peng terminilogy, see 
-\cite{barba_terminologies_2018}} into peer
-review, and a demonstration of its feasibility by means of 25
-reproductions across different scientific disciplines.
+Our anectodal experiences on this matter suggest that researchers often give a variety of reasons for why code cannot be shared, e.g.,~``there is no documentation'', ``I cannot maintain it'', or ``I do not want to give away my code to competitors''.
+Our view, rooted in Open~Science principles\footnote{\url{https://en.wikipedia.org/wiki/Open_science}}, is that sharing code, wherever possible, is good for the community and the individual \cite{Barnes2010-iv,markowetz_five_2015}.
+Having code and data openly available and archived provides a valuable resource for others to learn from, even if the code no longer runs or if there is no documentation.
+However, with a little effort, we believe that if an independent person can execute an analysis, this is worth documenting as early as possible and that such a focused approach can greatly reduce the barrier to evaluating non-text research materials.
+Just as data journals' validations of data and all journals' peer review can provide a ``baseline reassurance'', i.e.,~that a paper has been checked by someone with an understanding of the topic \cite{fyfe_mission_2019}, we think the same baseline should be provided for the workflow underlying a paper.
+With this in mind, we have developed a set of principles and an example workflow that provides what we hope is a pragmatic way of checking that a paper's code works.
+
+As this work's main contribution, we offer a thorough description of a process and its variations to integrate a much needed evaluation of computational reproducibility\footnote{Following the Claerbout/Donoho/Peng terminology, see \cite{barba_terminologies_2018}.} into peer review, and we demonstrate its feasibility by means of 25 reproductions across different scientific disciplines.
 We call this system CODECHECK as outlined next.
@@
   PDF to the underlying materials is impossible. This prohibits reuse
-  and is not only non-transparent for a specific paper, but also 
+  and is not only non-transparent for a specific paper but is also 
   ineffective for science as a whole. By sharing the
@@
 (1) the AUTHOR of a paper providing the code to be checked,
-(2) the PUBLISHER of a journal interested in publishing the paper 
-by the author, and
+(2) the PUBLISHER of a journal interested in publishing the author's paper, and
+(3) the CODECHECKER, who checks that the authors's code works.
 The workflow we have refined is shown in Figure~\ref{fig:workflow}
@@
 
-\textbf{Step 1:} the author submits their manuscript along with code and data to
-the PUBLISHER. The code and data need not be openly available at this point.
+\textbf{Step 1:} The author submits their manuscript along with the code and data to the PUBLISHER.
+The code and data need not be openly available at this point.
 However, in many cases the code and data may be published on a code hosting platform,
@@
 
-\textbf{Step 2:} the publisher finds a CODECHECKER to check the code. This is
+\textbf{Step 2:} The publisher finds a CODECHECKER to check the code. This is
 analogous to the publisher finding one or more reviewers to evaluate
-the paper. However, a crucial difference is that we suggest the CODECHECKER
-and the author can talk freely to each other directly (see next), independent
-of the openness of the scientific peer review process.
-
-\textbf{Step 3:} the CODECHECKER runs the code, based on instructions provided by
+the paper.
+However, a crucial difference is that we suggest the CODECHECKER and the author are able to talk freely and directly to each, regardless how open the scientific peer review process is.
+
+\textbf{Step 3:} The CODECHECKER runs the code, based on instructions provided by
 the author. They check if some or all of the results from the paper can be
@@
 This process iterates until either the CODECHECKER is successful,
-or the CODECHECKER concludes the code does not work.
-As part of this process, the CODECHECKER could work entirely
-locally on their own compute resource, or in the cloud, e.g.,~using the open
-MyBinder infrastructure \cite{jupyter_binder_2018} or alternatives, some of which
-are more tailored for scientific publications while others offer commercial
-options to, e.g.,~publishers (cf. \cite{konkol_publishing_2020}).
-Cloud-based 
-infrastructure allows for the CODECHECKER and author to collaboratively improve
-the code and enforces a complete definition of the computing environment,
-but, unless secure infrastructure is provided, e.g.,~by the PUBLISHER,
-requires code and data to be published openly online.
-
-Note that the task of the CODECHECKER is to check the ``mechanics'' of the
-workflow, not to evaluate its correctness. 
-Stodden et~al. \cite{stodden_setting_2013} distinguish between \emph{verification}
-and \emph{validation}
-in the context of Mathematics. Following their definition, a CODECHECK
-ensures a verification\footnote{Note this is a different use as the
-\emph{Verification Reports} 
-article type in the journal \emph{Cortex} \cite{chambers_verification_2020}},
-of computational results, i.e.,~checking that the code correctly produces the 
-output it claims to create, but not a validation, i.e.,~does the code implement
-the right algorithm to solve the problem that is researched.
-Nevertheless, the simple attempt to reproduce may highlight shortcomings in
-a submission to meet journal policy requirements
-(cf.~\cite{christian_journal_2020})
-and can effectively increase transparency and thereby improve practices
-(cf.~\cite{nosek_scientific_2012})
-even if the check goes not into all details.
-
-\textbf{Step 4:} the CODECHECKER writes a certificate stating how the code was
-run and includes a copy of outputs (figures, tables) that were
-independently generated. The certificate may include further recommendations to
-the author how to improve the material.
+or the CODECHECKER concludes the code can not be executed within the context of a CODECHECK.
+As part of this process, the CODECHECKER could work entirely locally, relying on their own computing resources, or in the cloud, e.g.,~using the open MyBinder infrastructure \cite{jupyter_binder_2018} or alternatives, some of which are more tailored to scientific publications while others offer commercial options for, e.g.,~publishers (cf.
+\cite{konkol_publishing_2020}).
+A cloud-based infrastructure allows for the CODECHECKER and author to collaboratively improve the code and enforces a complete definition of the computing environment; but, unless secure infrastructure is provided, e.g.,~by the PUBLISHER, this requires the code and data to be published openly online.
+Note that the task of the CODECHECKER is to check the ``mechanics'' of the workflow not to evaluate its correctness. 
+In the context of mathematics, Stodden et~al. \cite{stodden_setting_2013} distinguish between \emph{verification} and \emph{validation};
+following their definition, a CODECHECK ensures verification
+\footnote{Note this is a different use from that referred to as the \emph{Verification Reports} article type in the journal \emph{Cortex} \cite{chambers_verification_2020}.},
+of computational results, i.e.,~checking that the code correctly produces the output it claims to create, but not a validation, i.e.,~checking that the code implements the right algorithm to solve the specific research problem.
+Nevertheless, simply attempting to reproduce an output may highlight a submisison's shortcomings in meeting a journal's requirements (cf.~\cite{christian_journal_2020}) and may effectively increase transparency, thereby improving practices (cf.~\cite{nosek_scientific_2012}) even if the check does not go into every detail.
+
+\textbf{Step 4:} The CODECHECKER writes a certificate stating how the code was
+run and includes a copy of outputs (e.g.,~data, figures, or tables) that were
+independently generated.
+The certificate may include further recommendations to the author on how to improve the material.
 The free text in the certificate is the most flexible and feasible way to 
 make clear what was really checked, because each workflow is unique.
-Since no specific tool or platform should be required so that no authors are
-excluded, automation or fixed checklists are futile.
-
-\textbf{Step 5:} the certificate as well as auxiliary files created
+Since no specific tool or platform should be required, such that no authors are excluded, it would be futile for the CODECHECKER to use automation or fixed checklists.
+
+\textbf{Step 5:} The certificate as well as auxiliary files created
 during the check, e.g.,~a specification of a computing environment, data 
@@
 sensitivity apply.
-Currently, the CODECHECKERs deposit the material on Zenodo themselves, 
-while processes managed by the PUBLISHER may have them complete this step.
+Currently, the CODECHECKERs deposit the material on Zenodo themselves, but the PUBLISHER may complete this step after integrating CODECHECK into its review process.
 A badge or other visual aid may be added to the deposit and the paper and
 link to the certificate.
-Despite the simplification into a binary value and the risk of confusion of
-the extent of the check, especially across disciplines and methods,
-a badge can provide recognition value and creditation for the completed
-CODECHECK. The badge and the actual check are incentives for the efforts
-of providing a readily computationally reproducible workflows.
-
-\textbf{Step 6:} the PUBLISHER can, depending on the timing, provide the
-certificate to scientific reviewers or handling editors, or publish it
+Although a badge simplifies the CODECHECK into a binary value and risks introducing confusion regarding the extent of the check, especially across disciplines and methods, a badge can provide recognition value and acknowledgement of the completed CODECHECK.
+The badge and the actual check are incentives for undertaking the effort needed to provide a readily computationally reproducible workflow.
+
+\textbf{Step 6:} The PUBLISHER can, depending on the timing, provide the
+certificate to scientific reviewers or handling editors or publish it
 and link between the certificate, the paper, and any repositories 
 through metadata. Currently, the CODECHECKER creates these connections on
-Zenodo \footnote{
+Zenodo\footnote{
 See ``related identifiers'' and ``alternate identifiers'' in the right-hand
@@
 
-The example workflow leaves room for different
-implementations of this process. To elaborate on these options, we consider
-the several dimensions in a space of possible CODECHECK workflows, as shown in 
-Figure~\ref{fig:dimensions}, to represent relevant aspects for the involved
-stakeholders. These aspects touch on timing, responsibilities, and 
+The example workflow leaves room for this process to be implemented differently.
+To elaborate on these options, we consider several dimensions in a space of possible CODECHECK workflows, as shown in Figure~\ref{fig:dimensions}, to represent relevant aspects for the involved stakeholders.
+These aspects touch on timing, responsibilities, and 
 transparency and will be discussed in the following sections.
@@
 
-The point in time for a CODECHECK and the ascribed importance are closely
-connected, so the dimensions \emph{When} and \emph{Imporance} are described
-together.
-The earlier a CODECHECK happens in the process, the more the
-results can play a role in editorial decisions:
-is a paper published, sent back for revisions, or rejected.
-Even earlier checks, i.e.,~a CODECHECK of a preprint, may even help to 
-improve the workflow itself even before a publisher is involved. The 
-possibility for codechecking papers could be part of a preprint server's
-policy, or simply be initiated by interested readers and community members.
-
-A \textbf{strict prerequisite} for passing a check can reduce the workload
-of reviewers, and therefore seems out of place late in the review process.
-The later in a review process the check happens, the less challenging is it
-to allow a bidirectional communication between author and codechecker, e.g.,
-because the author might already be notified of the acceptance and may more
-freely share materials online closer to the paper publication date.
-
-A \textbf{pre-review} CODECHECK allows editors to use it as a filter to only
-send  a submission out to peer review if it passes the check.
-The certificate would  be included in the submission package provided
-to the reviewers, so they can include the CODECHECK's results in their 
-evaluation of the work.
-Reviewers may then judge the the relevance of the computations for the results
-of the work, or follow journal guidelines as how strict to apply the check's
-outcomes.
-
-A CODECHECK may also be conducted in \textbf{parallel} to the scientific review.
-This puts less burden on the turnaround time for the check, yet only makes the 
-outcomes available for the final consideration of the handling editor.
-The check could also be tasked after suggestion by a reviewer and remove the
-need for a pre-review screening for computational workflows. However, 
-soliciting such a ``specialist review'' is much less desirable than having
-a regular CODECHECK, without any special treatment for some submissions.
-In both cases, a CODECHECK could then be awarded \textbf{equal weight} as 
-the other reviews for the editor's decision.
-
-A \textbf{post acceptance} CODECHECK has the smallest impact on editorial 
-decisions and may simply provide an \textbf{extra merit} on top of the 
-acceptance of the submission. 
-This  is the least impactful solution while still evaluating all material
-and properly acknowledging the check results from the paper, because the 
-check can be completed  before the publication of the paper.
-The GIScience group of checks (see below) falls into this
-category: the AGILE conference highlights articles where the reproducibility
-was successfully reviewed with a badge on the volume and article landing
-pages. Similarly, the GIScience papers in collaborations with journals were 
-checked while the authors worked on revisions.
-
-A CODECHECK may also be conducted \textbf{post publication}, though it
-would required to update the article and article metadata later to reference 
-the check, otherwise readers would presumably not be aware of it. In general,
-publishers are hesitant to make such a revision of published articles.
-While this has the least impact on the current publishing practices and also
-enables open discourse,
-it also has the smallest effect on the original submission
-and does not acknowledge the importance of reproducible workflows for
-ensuring good scientific practice.
+The point in time at which a CODECHECK is done and its ascribed importance are closely connected, so we describe the dimensions \emph{When} and \emph{Importance} together.
+The earlier a CODECHECK happens in the publishing process, the more it can affect editorial decisions:
+Is a paper published, sent back for revisions, or rejected?
+Even earlier checks, i.e.,~a CODECHECK of a preprint, may even help to improve the workflow itself, even before a publisher is involved.
+As such, codechecking papers could be part of a preprint server's policy or simply be initiated by interested readers and community members.
+
+Publishers could introduce a CODECHECK as a \textbf{strict prerequisite}.
+As this can reduce the workload of reviewers, such a check should not occur late in the review process.
+Yet, the later in the review process the check happens, the easier is it to allow bidirectional communication between the author and codechecker, e.g., because the author might already be notified of the paper's acceptance and may more freely share materials online closer to the paper's publication date.
+A \textbf{pre-review} CODECHECK would allow editors to only send a submission out for peer review if it passes the check, or to include the certificate in the submission package provided to the reviewers.
+Reviewers may then judge the the relevance of the computations for the results of the work, or follow journal guidelines regarding how strictly to apply the check's outcomes in their evaluation.
+
+A CODECHECK may also be conducted in \textbf{parallel} to the scientific peer review.
+This puts less burden on the turnaround time for the check, yet it only makes the outcomes available during the final consideration by the handling editor.
+The check could also be assigned after suggestion by a reviewer, which would remove the need for submissions to undergo a pre-review screening.
+However, soliciting such a ``specialist review'' is much less desirable than having a regular CODECHECK, thus avoiding the situation in which some submissions get special treatment.
+In both cases, a CODECHECK could then be awarded \textbf{equal weight} compared to the other reviews for the editor's decision.
+
+A \textbf{post-acceptance} CODECHECK would have the smallest impact on editorial decisions and may simply provide \textbf{extra merit} on top of the submission's acceptance.
+This is the least impactful solution in which all material is still evaluated and the results of the check are properly acknowledged, because the check can be completed before the publication of the paper.
+The GIScience group of checks (see below) falls into this category:
+By displaying a badge on the volume and article landing pages, the AGILE conference highlights articles whose reproducibility was successfully reviewed.
+Similarly, in collaborations with journals, some GIScience have been checked while the authors worked on revisions.
+
+A CODECHECK may also be conducted \textbf{post-publication}, though it would require someone to update the article and article metadata to reference the check; otherwise, readers would presumably not be aware of it.
+In general, publishers hesite to make such revisions to published articles.
+While this option has the least impact on current publishing practices and enables open discourse, it also has the smallest effect on the original submission and does not acknowledge the importance of reproducible workflows for ensuring good scientific practice.
 
@@
 established or typical scholarly communication process.
-While more radical ideas to change the paradigms
-how researchers can share their work much more granually exist\footnote{
-For example Octopus, 
-\url{https://science-octopus.org/about}, or Hypergraph
-\url{https://www.libscie.org/hypergraph}.},
+While more radical ideas do exist for changing the paradigms of how researchers can more granually share their work much
+\footnote{For example Octopus, \url{https://science-octopus.org/about}, or Hypergraph, \url{https://www.libscie.org/hypergraph}.},
 % mimosa is another platform idea, but no real prototype AFAICT:
@@
 % https://docs.google.com/presentation/d/1V_K8hghgnvGfEtW7TdwtowTveZyy4Qs8rMemNyHx_Dg/edit#slide=id.p
-there is a value in an evolutionary approach and we do not suggest
-to completely remove the paper as a point in time to share work just yet.
-Enhancing existing processes with CODECHECKs allows communities to 
-transition towards more open practices at their own pace and ensure
-no one is left behind unfairly.
-When integrating a CODECHECK in existing review and publication processes,
-the \emph{turnaround time} is crucial. Depending on when and who conducts
-the check, it may be done faster, or add considerably to the overall
-time from submission to publication.
-In our first checks, we found that a codecheck may generally take from 2-5
-hours, with some outliers on the higher end. This time includes writing and
-publishing the certificate, but excludes the time for actual computations.
+an evolutionary approach has value, and we do not yet suggest to completely eliminating research papers as a way of sharing work.
+Enhancing existing processes with CODECHECKs allows communities to transition towards more open practices at their own pace and ensure that no one is unfaily left behind.
+When integrating a CODECHECK into existing review and publication processes, the \emph{turnaround time} is crucial.
+Depending on when and who conducts the check, it might be done quickly or it might add considerable time to how long it takes from submission to publication.
+In our first checks, we found that a CODECHECK generally takes from 2-5 hours, with some outliers on the higher end.
+This time includes writing and publishing the certificate but excludes the time for actual computations.
 These efforts are comparable to the time needed to peer review a submission,
-which concurs with the efforts some volunteer codecheckers are willing to
-spend.
-However, the time includes a considerable share of communicating about the
-process,  especially around who publishes which document when, so that proper
+which aligns with the efforts some volunteer codecheckers are willing to
+make.
+However, the time includes a considerable amount of communicating about the
+process, especially regarding who publishes which document when, so that proper
 cross-referencing between paper and certificate is ensured via persistent
 identifiers.
-When integrated into a peer review system, this handling of documents should
+When integrated into a peer review platform, this handling of documents should
 become much more effective and reduce organisational overhead.
 
-\subsubsection*{Openness, or: Who knows who?}\label{who-knows-who}
-
-The topic of anonymity is broadly discussed, not the least in the push towards
+\subsubsection*{Openness, or ``Who knows who?''}\label{who-knows-who}
+
+The topic of anonymity is broadly discussed, especially in the push towards
 open peer review as part of the Open Science movement 
 (cf. \cite{ross-hellauer_guidelines_2019}).
-Without taking a strong stance on this discussion, the general motivation 
+Without taking a strong stance on this topic, the general motivation 
 behind CODECHECK
@@
 review process.
-However, anonymity can also serve to protect individuals 
-\cite{tennant_limitations_2020}, e.g.~junior scientists, whereas it might 
-be acceptable to require revealing of names from tenured researchers.
-The possible negative effects of a non-anonymous review are reduced
-if a CODECHECK is not relevant for the decision to accept or reject, but that
-is of course not desirable. Instead, we argue that a CODECHECK is technical
-process that should in any case be fixable and not a question of opinion or
-a faulty approach.
+However, anonymity can also serve to protect individuals \cite{tennant_limitations_2020}, e.g.,~junior scientists, whereas revealing of names from tenured researchers might be more reasonable to require.
+The possible negative effects of a non-anonymous review are reduced if a CODECHECK is not relevant for a journal's decision to accept or reject, but that is, of course, not desirable when the goal is higher transparency and reproducibility.
+Instead, we argue that a CODECHECK is technical process that should in general find fixable problems; it is not aimed at giving an opinion or identifying a faulty approach.
 It is technically possible for a nonsense workflow to receive
@@
 If passing a CODECHECK becomes mandatory, the full transparency may have to be 
-revisited as the relations between authors and codecheckers fall under the
-same social and acommunity challenges as non-anonymous peer review
+revisited as the relations between authors and codecheckers would fall under the
+same social and community challenges as non-anonymous peer review
 (cf.~\cite{everythinghertz123}).
@@
 
-The technical nature of the check and the challenge of good enough 
-documentation is why we see great benefits in a effective bidirectional
-means of communication between author and codechecker. Instead of trying to
+The technical nature of the check and the challenge of providing sufficient documentation is why we see great benefits in allowing for bidirectional communication between author and codechecker.
+Instead of trying to
 fix problems or guess the next step, the codechecker can ask the author to 
-rework the docs  or the update code.
+rework the documentation or the update code.
+Instead of struggling to provide perfect instructions and as a result possibly not sharing any code or data, the author can make a best effort to document sufficiently.
 Of course, the codechecker may also provide a fix within the code instead
-of lengthily describing a small change if it is quickly done.
-Authors and readers can profit from codecheckers' experience and approach,
-as the latter may create useful and instructive files, e.g.,~a machine-readable
-computing environment specification, during the check.
+of lengthily describing a small change if it can be quickly done.
+Authors and readers can profit from a codecheckers' experience and approach, as during the check they may create useful and instructive files, e.g.,~a machine-readable computing environment specification.
 While communication between author and codechecker may be facilitated in an 
 anonymous way by the 
-publisher, it most likely only helps to protect the identity of the 
-codechecker, because, especially when of a certain size, not trivial, or 
-if very specialised, it is close to impossible to fully anonymise a codebase.
-Therefore, the most effective means with the most desirable variant for
-the stakeholders is a non-anonymous and collaborative CODECHECK.
+publisher, it most likely only helps to protect the identity of the codechecker, because, especially when codebases are of a certain size, not trivial, or if very specialised, they are almost impossible to fully anonymise.
+Therefore, the most effective and desirable situation for the stakeholders is to hold a non-anonymous and collaborative CODECHECK.
 The contributions by the codechecker may even be integrated into
 the code of the workflow and be acknowledged as code commits. This way, 
-proper credit is given within the research software development community.
-Nevertheless, depending on needs represented by other dimensions,
-the aspect of `Who knows who?' may be the most flexible one.
+proper credit can be given within the research software development community.
+Nevertheless, depending on the needs represented by other dimensions,
+the aspect of ``Who knows who?'' may be the most flexible.
 
@@
 
-The options for the person who can conduct a CODECHECK require a not quite
-simple matching of a combination of skills and availability. Ideally, the
-codechecker has a matching code \emph{and} domain expertise, for example to
-check a workflow based on Python code analysing neuroimaging datasets. However,
+The options for who can conduct a CODECHECK require a somewhat complex matching of a combination of skills and availability.
+Ideally, thecodechecker has a matching code \emph{and} domain expertise; for example, to check a workflow based on Python code analysing neuroimaging datasets, the codechecker should ideally be experienced in Python and neuroimaging analysis.
+However,
 a well-documented workflow should be executable by any person with a basic
@@
 codechecker has, the less time is needed to understand the goals and 
-the mechanics of an analysis. From our experiences, the priority should be
-given to matching technical expertise first, as the a lack of knowledge in
-setting  up a computing environment with a particular language or tool is 
-much more discouraging and the assessment of the outcome, e.g.,~comparing
-created figures with the original, is largely possible without 
-understanding the science.
-The time and expertise allotted to the check
-are the main drivers for the depth of checking, though in general, we 
-expect a CODECHECK not to evaluate performance, e.g.,~does a racing car
-have the best apex speed, but to assess operation free of failure, i.e.,~is
-it a car with 4 wheels and at least one door, and does it move forwards
-when accelerating. This depth can be seen independent from the power of
-the check (see above), as long as it is communicated clearly to all
-stakeholders, most importantly authors and readers.
-
-The concrete people to provide these expertises can range from researchers,
+the mechanics of an analysis.
+From our experiences, the priority should be given to matching technical expertise first, as lacking knowledge in setting up a computing environment with a particular language or tool is much more of a problem than assessing the outcome, e.g.,~comparing created figures with the original, without an in-depth understanding of the science.
+The depth of the check will mostly be driven by the time required and expertise of the checker, though in general, we expect a CODECHECK not to evaluate performance (e.g.,~it will not check whether a racing car has the best apex speed) but to assess operation free of failure (i.e.,~is it a car with 4 wheels and does it move forwards when accelerating).
+This depth can be seen independent of the check's importance, as long as it is communicated clearly to all stakeholders, most importantly authors and readers.
+
+The concrete people to provide such expertise can range from researchers,
 possibly drawn from the regular pool of volunteer
-\textbf{scientific reviewers} or from a special group of
-\textbf{reproducibility reviewers}, via specific roles
+\textbf{scientific reviewers}, or from a special group of
+\textbf{reproducibility reviewers} via specific roles
 such as \textbf{reproducibility editors}, to experts working as part of
@@
 We consider a \emph{single codechecker} to be enough, unlike peer review where
-multiple referees are common. Because of the technical focus, the 
-determination if the concrete outputs of author and codechecker, who can
-collaborate to solve any issues, actually match should be a very factual 
-process and not
-require much debate or opinion. Code usually makes systematic and repeatable
-mistakes is is thereby more reliable and auditable than processes controlled
-by humans \cite{tibav:42484}, e.g.,~in a laboratory.
+multiple referees are common.
+Because of the technical focus, determining whether the concrete outputs produced by the author and the codechecker, who can collaborate to solve any issues, actually match should be a very factual process and not require much debate or opinion.
+Code usually harbours systematic and repeatable mistakes and is thereby more reliable and auditable than processes controlled by humans \cite{tibav:42484}, e.g.,~in a laboratory.
 Depending on the importance of the CODECHECK
@@
 
-We see a great opportunity to especially involve
-\emph{early-career researchers}
-(ECRs) as codecheckers. They arguably have a high interest in learning
-about the latest tools and technologies as they are building up their own
-expertise and specialisation. The possibility to gain insights into latest
-science for ECRs and at the same time increase the appreciation of
-reproduction efforts is a sweet spot that CODECHECK shares
-with \emph{ReScience X}, a journal devoted to reproduction and replication 
-experiments \cite{roesch_new_2020}.
-In fact, ECRs are often much more familiar with latest technology and are 
-likely to be early adopters as creators, i.e.,~be authors of CODECHECK-ready manuscripts, too\footnote{Cf. the 77\% of 141 registered reports who were 
-submitted to leading Neuroscience journals by ECRs
-\cite{chambers_registered_2019}.}
-They are also introduced to peer review and 
-may transition into the role of scientific reviewer over time.
-Not to solve the general problems of peer review, which is largely an
-unsupervised process, i.e.,~rarely are reviewers taught how to do it, but
-a junior codechecker may be supported by a senior codechecker, too.
-We see a high potential in setting up the new processes for CODECHECKING with
-a clear commitment on openness and transparency, independent of the
-respective current peer review process (see \emph{Openness} dimension).
-
-A publisher's \textbf{editorial staff} represents the most controlled but also
-resource-intensive
-option. Hiring the required technical and or domain expertise puts a financial
-burden on the publisher, of course, but such a commitment also shows the 
-value given to publishing reproducibility and would positively impact the 
-submission experience, because of more controllability, and external 
-perception of the journal or conference. A staff codechecker reduces the 
-challenges when it comes to open communication and awarding credit, as
-in-house personnel may be more readily be given access to non-anonymised
-submission material and receives a proper salary instead of public credit
-as part of the scientific community.
-Keeping diverse specialists as staff to realise a codechecking process is 
-unlikely to be feasible, especially for small or independent publishers.
+We also see a great opportunity to involve \emph{early-career researchers} (ECRs) as codecheckers.
+ECRs arguably have a high interest in learning
+about the latest tools and technologies, as they are building up their own
+expertise and specialisation.
+CODECHECK offes a way for ECRs to gain insights into the latest science and at the same time can help increase the appreciation of reproduction efforts, a combination that is of particular interest.
+\emph{ReScience X}, a journal devoted to reproduction and replication experiments \cite{roesch_new_2020}, shares an interest in this combination.
+Further, ECRs are often much more familiar with latest technology and tend to be early adopters, thus also making them likely to author CODECHECK-ready manuscripts
+\footnote{Cf. the 77\% of 141 registered reports that were submitted to leading Neuroscience journals by ECRs \cite{chambers_registered_2019}.}.
+As ECRs are introduced to peer review as codecheckers, they may transition into the role of scientific reviewer over time.
+Another potential highlight is that during the codechecking process, a senior codechecker may support a junior codechecker, which would be the opposite of what happens during general peer review, which is largely an unsupervised process, i.e.,~rarely are reviewers taught how to do it.
+Overall, we see several opportunities and benefits to setting up a new process for CODECHECKING with a clear commitment to openness and transparency, independent of the current peer review process (see \emph{Openness} dimension).
+
+Having the codechecker be part of the publisher's \textbf{editorial staff} represents the most controlled but also
+resource-intensive option.
+Hiring the required technical and or domain expertise will put a financial burden on the publisher, of course, but such a commitment would also show that publishers are investing in reproducibility.
+Codecheckers on staff would positively impact the submission experience and the external perception, because having an in-house codechecker would make the process easier to control and would ensure the author gets an external, unbiased review.
+Furthermore, a staff codechecker would also reduce the challenges regarding open communication and awarding credit, as in-house personnel may more readily be given access to submission material would receive a salary instead of public credit as part of the scientific community.
+Yet, keeping diverse specialists on as staff to realise a codechecking process is 
+unlikely to be feasible for small or independent publishers.
 
 In contrast, for researchers it can be very important to be publicly credited
-with their activity as a reviewer.
+for their activity as a reviewer.
 A regular
@@
 \footnote{\href{https://publons.com/}{https://publons.com/}} or 
-ReviewerCredits\footnote{\href{https://www.reviewercredits.com/}{https://www.reviewercredits.com/}}).
-A codechecker should be listed just the same. The number of scientific reviews
-is of course a very coarse indicator for the community contribution by an 
+ReviewerCredits\footnote{\href{https://www.reviewercredits.com/}{https://www.reviewercredits.com/}});
+a codechecker should be similarly listed.
+The number of scientific reviews is of course only a coarse indicator of an individual's community contribution, yet it is unclear if a regular reviewer, who conductsa regular scientific review and in addition also conducts a CODECHECK, should be credited twice.
 In this case, the published CODECHECK certificate and possibly the published peer review can clearly describe the possibly large efforts put into both.
-conducts a CODECHECK to the scientific review, should be credited twice.
-
-The codecheckers community\footnote{\href{https://github.com/codecheckers/codecheckers/}{https://github.com/codecheckers/codecheckers/}}
-currently has over 20 individuals who signed up as volunteer codecheckers
-following talks given on the project in the last 12 months.
-That is even though no checks are actively elicited in collaboration
-with journals.
-Their motivations, mentioned in the sign-up issues\footnote{
-\url{https://github.com/codecheckers/codecheckers/labels/registration}},
-are:
-supporting reproducible research and open science,
-establishing a culture of open and reproducible science and being part
-of making it the norm,
-improving coding skills,
-giving back to the community,
-gaining experience in helping scientists with their code,
-reducing the amount of bad scientific code and encouraging a culture that
-reduces friction in passing on academic codebases,
-learning from other people's mistakes,
-and curiosity.
-We see a benefits in an open shared list of codecheckers
-across journals
-compared to a private in-house group, as it may better match expertise
-and share the workload.
-The overarching community is also a mean to ensure CODECHECK is a viable
-option not just for large publishers, independent of their financial model,
-but also for independent no-cost Open Access journals.
+
+The codechecker community\footnote{\href{https://github.com/codecheckers/codecheckers/}{https://github.com/codecheckers/codecheckers/}}
+currently has over 20 individuals who signed up in the last 12 months as volunteer codecheckers after talks given on the project, even though no checks are actively being requested in collaboration with journals.
+These volunteer's motivations, mentioned in the registration information\footnote{\url{https://github.com/codecheckers/codecheckers/labels/registration}}, are that they want to
+support reproducible research and Open Science,
+establish a culture of open and reproducible science,
+be part of making this culture the norm,
+improve coding skills,
+give back to the community,
+gain experience in helping scientists with their code,
+reduce the amount of bad scientific code, 
+encourage a culture that reduces friction in sharing or passing along on academic codebases,
+learn from other people's mistakes;
+many are also motivated simply by curiosity.
+We see benefits to an open shared list of codecheckers across journals rather than a private in-house group, as this may allow for better matches regarding expertise and more even sharing of the workload.
+Overall, this community will aim to ensure that CODECHECK is a viable option not just for large publishers, independent of their financial model, but also for independent no-cost Open Access journals.
 
@@
 
-The example workflow and variations just outlined above reflect our current
-views on how a codecheck should be performed. They are not set in
-stone, but we do believe the following core principles underpin our
-CODECHECK:
+The example workflow and variations outlined above reflect our current
+views on how a workflow execution and code checking should be performed as part of peer review.
+They are not set in stone, but we do believe the following core principles underpin our CODECHECK process:
 
@@
 the code. If any steps are unclear, or if code does not run correctly,
-the codechecker records this fact, and then communicates this
+the codechecker records this fact and then communicates this
 information to the author. We believe that the job of the codechecker is
-not to fix all these problems, but simply to report them, and await a
+not to fix all these problems but simply to report them and await a
 fix from the author.
 The level of documentation that is needed for third parties
-to reproduce a workflow is extremely hard to get right. Too often this 
-uncertainty leads to researchers giving up and not documenting at all.
+to reproduce a workflow is extremely hard to get right, and too often this 
+uncertainty leads researchers to give up and not document it at all.
 The conversation with a codechecker fixes this problem.
 
-\textbf{2. Communication between humans is key.} Some code may just
+\textbf{2. Communication between humans is key.} Some code may
 work without any interaction (e.g.,~CODECHECK certificate \texttt{2020-013}
-\cite{cert-2020-013}); however, more often there are hidden
-dependencies that need to be found, or revised software
-installations. Allowing the codechecker and human to communicate
+\cite{cert-2020-013}) but more often there are hidden
+dependencies that need to be found or software installations that need to be adjusted for a particular system.
+Allowing the (human) codechecker to communicate
 directly and openly with each other should make this process as
@@
 (possibly anonymously) through a publisher would slow things down and
-reduce the chances for community building as well as the lessons
-learned for both author and codecheckers.
-
-\textbf{3. Credit is given to codecheckers.} The value of contributing a
-CODECHECK is comparable to providing a scientific peer review. Individual
-variations aside, it may also require a similar amount of time. Therefore,
+reduce the chances for community building as well as obscure some of the lessons
+learned for both authors and codecheckers.
+
+\textbf{3. Credit is given to codecheckers.} The value of performing a
+CODECHECK is comparable to that of a scientific peer review, and
+it may also require a similar amount of time. Therefore,
 the codechecker's activity should be publicly recorded and mentioned in the
@@
 has at least enough material to validate the workflow outputs submitted by 
-the authors. Stark~\cite{stark_before_2018} calls this `preproducibility'
+the authors. Stark~\cite{stark_before_2018} calls this ``preproducibility''
 and the ICERM report \cite{stodden_setting_2013} defines the level
@@
 was given and could be investigated more deeply or extended in the future.
-Ideally, this is a "one~click" step, but achieving this requires particular 
-skills. It is also hard to create a sufficient level of documentation for 
-third parties. Furthermore automation may lead to people gaming the system
-or reliance on technology hiding important details. All such aspects can
-reduce the understandability of the material, which we estimate to be the
-best mean to ensure long-term transparency and usefulness.
-That is why the possibility to communicate is more important
-than perfected workflows, and why a CODECHECK is not automated.
+Ideally, this is a ``one~click'' step, but achieving this requires particular 
+skills and a sufficient level of documentation for 
+third parties. Furthermore, automation may lead to people gaming the system
+or reliance on technology, which can often hide important details.
+All such aspects can reduce the understandability of the material, so we estimate our approach to codechecking, done without automation and with open human communication, to be the best way to ensure long-term transparency and usefulness.
 We acknowledge that others have argued in favour of bitwise reproducibility
-because in the long run it can be automated\footnote{Konrad Hinsen on 
-Twitter: \url{https://twitter.com/khinsen/status/1242842759733665799}};
+because, in the long run, it can be automated\footnote{Konrad Hinsen on 
+Twitter: \url{https://twitter.com/khinsen/status/1242842759733665799}},
 but until then we need CODECHECK's approach.
 
-\textbf{5. Open by default and transitional by disposition.} % or "design"?
+\textbf{5. Open by default and transitional by disposition.}
 By default, unless there are strong
-reasons to the contrary (e.g.,~clinically-sensitive data), all code and
+reasons to the contrary (e.g.,~sensitive data on human subjects), all code and
 data, both from author and codechecker, will be made freely available, 
 i.e.,~under open licenses, before or when the certificate is published.
-Note that openness is not required for the paper itself as to accommodate 
-journals to transition to sustainable Open Access models.
+Note that openness is not required for the paper itself, as to accommodate 
+journals in their transition to sustainable Open Access models.
 The code and data publication should follow community good practices.
-By disposition, a CODECHECK workflow should be designed (a) as a recognition of
+By disposition, a CODECHECK workflow should be designed (a) with a recognition of
 shortcomings in asserting reproducibility of submissions and (b)
 with an understanding of the
-complex situation that lead to this. However, establishing a CODECHECK workflow
-should go hand in hand with initiatives and a perspective to improve education
-around reproducibility of methods and communication of comprehensive 
-computational workflows, with the long term goal for a specific community or
+complex situation that led to this. However, establishing a CODECHECK workflow
+should go hand in hand with initiatives aiming to improve education
+around reproducibility of methods and communication surrounding comprehensive 
+computational workflows.
+The long term goal is for a specific community or
 journal's target audience to reach a state where the level of verification 
-ensured by codecheckers becomes part of peer review and regular
+ensured by codecheckers becomes part of peer review and a regular
 merit of all accepted papers.
@@
 \url{https://codecheck.org.uk/register}.
-Table~\ref{tab:register} lists the certificates completed to date,
+Table~\ref{tab:register} lists the certificates completed to date
 along with citations to the original papers.
@@
 as as a conversation starter for potential collaborating journals.
-This intention worked out with the first CODECHECK (certificate number
-\texttt{2020-001}) before publication of an article for the journal
-\emph{GigaScience}, which successfully reproduced the visualisations
+This intention was brought to the first CODECHECK (certificate number
+\texttt{2020-001}) performed before publication on an article for the journal
+\emph{GigaScience}, which successfully reproduced the visualisations,
 though not the full machine learning process.
@@
 The second theme is rather opportunistic. As the COVID-19 pandemic swept
-over the whole Earth, we hoped to contribute a small part to science's answer
+over the globe, we hoped to contribute a small part to science's answer
 to the unknown virus. The checks were solicited through community interaction
 %Several COVID-19 models have been done during the time of peer review of these papers,
-and under our initiative rather than requested from journal, but some certificates
+or by our initiative rather than requested from journals, but some certificates
 have since then been acknowledged in reviewed papers
@@
 AGILE reproducibility reviews adhere to the CODECHECK principles.
-Furthermore, a small number of checks was completed as part of a peer reviews
+Furthermore, a small number of checks were completed as part of peer reviews
 for GIScience journals.
@@
 
-At the completion of each CODECHECK, the codechecker writes a certificate
-stating which outputs, i.e.,~numbers, figures or tables, from the 
-original article, could be reproduced by the codechecker.
+At the completion of each CODECHECK, the codechecker writes a certificate stating which outputs from the original article, i.e.,~numbers, figures or tables, could be reproduced.
 This certificate is made openly
 available so that all readers of the article can see which elements
-have been reproduced and what limitations or issues were faced.
+were reproduced and what limitations or issues were faced.
 The certificate also includes links to the code
@@
 See technical specification of the CODECHECK configuration file at
-\url{https://codecheck.org.uk/spec/config/latest/}}.
+\url{https://codecheck.org.uk/spec/config/latest/}.}.
 The configuration file enables current and future automation of workflows and
@@
 
-To demonstrate the typical contents of a certificate the the underlying
+To demonstrate the typical contents of a certificate and the underlying
 metadata, 
@@
 annotations.
-The referenced article \cite{unwin_report_2020} describes the modelling COVID-19
-across the USA.
+The referenced article \cite{unwin_report_2020} describes the modelling of COVID-19 across the USA.
 Figure~\ref{fig:cert}A shows the certificate number and its DOI,
@@
 Figure~\ref{fig:cert}B provides the key metadata as a table generated
-from the \texttt{codecheck.yml} --
-which paper was checked (title, DOI), who the authors and the codechecker 
-were, when it was checked, and in which repository related material is
+from the \texttt{codecheck.yml}; in particular it names the paper that
+was checked (title, DOI), the authors, the codechecker, when the check was performed,
+and in which repository related material is
 accessible.
-Figure~\ref{fig:cert}C is a textual summary of how the CODECHECK was
-performed, and any interesting findings.
-
+Figure~\ref{fig:cert}C shows a textual summary of how the CODECHECK was
+performed and any interesting findings.
 Figure~\ref{fig:cert}D (page 2 of the certificate) shows a table that
-lists the outputs that were generated from the check.
-This table is generated from the manifest of output files in the CODECHECK
-configuration file. The manifest lists output files with a short description.
+lists the outputs that were generated from the check based on the 
+manifest of output files in the CODECHECK.
 The certificate's file table shows the file name, the description stating to
-which figure/table each file should be compared to in the original paper, and
+which figure/table each file should be compared in the original paper, and
 the size of the file.
@@
 codechecker. In this particular example, the notes document what steps were 
-needed to initialize the environment and run the code, and that the execution
-took quite long to run -- about 17 hours.  Finally, on page 4 of
+needed to initialize the environment and run the code and that the execution
+took quite long to run -- about 17~hours. Finally, on page 4 of
 the certificate, we see the first output that was generated by the
@@
 for creating the certificate, even a regular word processor.
-This is a deliberate flexibility to
-accomodate different skill sets and unforseen technical advances or
-challenges.
+This flexibility is deliberate to accomodate different skill sets and unforseen technical advances or challenges.
 
@@
 The automation should ensure that more checks can
-be conducted in the future, though the management by a paid CODECHECK
+be conducted in the future, though having the process be managed by a paid CODECHECK
 editor or further automation, e.g.,~with a bot, could improve turnaround time.
 The only costs are the project's domain, which is negligible.
-This leaves the initiative's main resource -- the \textbf{humans}
+This means that the initiative's main resource are the \textbf{humans} needed
 for managing the project and processes and the codecheckers.
-All contributions are currently rely on (partly grant-based) public funding
+All contributions currently rely on (partly grant-based) public funding
 and volunteering.
@@
 The journal \emph{ACM Transactions on Mathematical Software (TOMS)} 
-established a "Replicated Computational Results" (RCR) review process
-\cite{heroux_editorial_2015}, using "replicable" with the meaning 
-as this work calls "reproducible".
+established a ``Replicated Computational Results'' (RCR) review process
+\cite{heroux_editorial_2015}, where ``replicable'' means the same thing as what we call ``reproducible''.
 A search on Crossref\footnote{\url{https://search.crossref.org/?q=\%22Replicated+Computations+Results+\%28RCR\%29+Report\%22&from_ui=yes&page=2}; search executed
-on 2020-12-10.} shows 15 RCR Reports to be published and the 
+on 2020-12-10.} shows that 15 RCR Reports have been published and the 
 % Number of found TOMAS RCR reports:
@@
 % 15: https://academic.microsoft.com/search?q=Replicated%20Computations%20Results%20%22(RCR)%20Report%22&f=&orderBy=0&skip=10&take=10
-workflow to be 
+workflow is being
 extended to the ACM journal \emph{Transactions on Modeling and Computer 
 Simulation (TOMACS)}\footnote{\url{https://dl.acm.org/journal/tomacs/author-guidelines}}.
-The CODECHECK principles were independently developed as we only later 
-learned about TOMS RCR, which we judge as fulfilling CODECHECK principles
-1 to 4.
-This work provides an extensive discussion of the possible variants, 
+The TOMS RCR follows the same principles as those used by CODECHECK (principles 1-4), which were independently developed as we only later learned about TOMS RCR.
+Here, we provide an extensive discussion of the possible variants of CODECHECK,
 while TOMS RCR realises one particular workflow for reproducibility 
@@
 \cite{heroux_editorial_2015} shares similar concerns about
-selection of reviewers as above. Unlike existing CODECHECK certificates, 
-the RCR reports undergo editorial review, but this could very well become
-part of CODECHECK processes at journals.
+selection of reviewers, as we discussed above. Unlike existing CODECHECK certificates, 
+the RCR reports, which are very similar to CODECHECK certificates, undergo editorial review, but this could eventually become
+part of the CODECHECK processes at journals.
 The publication of the RCR report is seen as a reward for the efforts of
-the reproducing person, while the possible conflict of interest of this 
-motivation is acknowledged. TOMS also recognises reviewer activity in a 
+the reproducing person, while the potential for this motive to be a conflict of interest is acknowledged.
+TOMS also recognises reviewer activity in a 
 partnership with Publons\footnote{
 \url{https://authors.acm.org/author-services/publons}},
-though it is unclear if the author of an RCR report receives both the report's
+though it is unclear whether the author of an RCR report receives both the report's
 authorship and the reviewer recognition. In any case, this is different
-from CODECHECK, which specifically aims at recognition as reviewers.
+from CODECHECK, which specifically aims to give codecheckers recognition as reviewers.
 In our view, the reviewer role removes the possible conflict of interest while
 keeping the public acknowledgement.
-Specific to the field of Mathematics, the RCR is also expected to apply a
+Specific to the field of mathematics, the RCR is also expected to apply a
 review of the software itself if the system it runs on cannot be evaluated by
@@
 in reviewer judgement over numerical bit-wise perfection.
-A key difference is that authors opt-in with an \emph{RCR Review Request} and
-the RCR reports, which are very similar to CODECHECK certificates, are
-published in the TOMS journal next to the actual papers.
-
-A number of journals provide guidelines for 
-ensuring reproducible publications and their verification,
-e.g.,~\emph{Information Systems} with its invited 
+A key difference is that for TOMS RCR, authors opt-in with an \emph{RCR Review Request} and
+the RCR reports are published in the TOMS journal next to the actual papers.
+
+A number of journals provide special article types for reproductions of papers
+published in the journal.
+\emph{Information Systems} has an invited 
 reproducibility paper track\footnote{
 \url{https://www.elsevier.com/journals/information-systems/0306-4379/guide-for-authors}.}.
-% first glimpse into "Code exeuction processes in scholarly publishing" ?
-% `https://docs.google.com/document/d/1fMWrFvBTwAYXg5J_bYIjgb-gY7DfQ0DOjx9Vlg_Rbmg/edit# ?
 \emph{Nature Machine Intelligence} recently introduced a new type of article,
@@
 \cite{noauthor_research_2020}. This answers the specific community's 
-challenges around computational reproducibility and also values these kind
+challenges around computational reproducibility and also values these kinds
 of contributions as independent publications, which goes beyond the goals 
@@
 The journal \emph{Cortex} has a special article type 
-\emph{Verification Reports}, which  actually are about replication of results 
+\emph{Verification Reports}, which are actually about replication of results
 but are very well designed/reasoned \cite{chambers_verification_2020}.
@@
 
-Going beyond article types, the journal \emph{ReScience C} publishes only
-replications\footnote{\url{https://rescience.github.io/faq/}} requiring
-open code and replication by a third party to. ReScience also relies on
+Going beyond article types and a specific journal, the journal \emph{ReScience C} publishes only
+replications\footnote{\url{https://rescience.github.io/faq/}}, also requiring
+open code and replication by a third party. ReScience also relies on
 free infrastructure (GitHub and Zenodo).
+
 For research with very high stakes, where reproduction would be too weak and
-post publication replication possibly too late because of policy impact,
+post-publication replication possibly too late because of policy impact,
 Benjamin-Chung et~al. \cite{benjamin-chung_internal_2020} propose
 \emph{internal replication}.
-A workflow that underwent internal
+A workflow that has undergone internal
 replication would very likely be of high quality and relatively easy to check.
@@
 \emph{Verifiable Result Identifier} would indeed provide excellent
-reproduciblity. However, the system is also complex and was not taken up by
-any publisher as far as we know; the system is indeed not open source itself.
+reproduciblity. However, the system is also complex and has not been adopted by
+any publisher as far as we know; also, the system is indeed not open source itself.
 In comparison, CODECHECK is less powerful but also much more flexible and
@@
 
-\emph{SciGen.Report} is a community-run platform to foster communication on
-reproducibility\footnote{\url{https://scigen.report/}}. People can report on
-fully, particially or un-successful reproductions post publication.
+Another platform for publishing results of reproductions is \emph{SciGen.Report}\footnote{\url{https://scigen.report/}}.
+It is a community-run independent platform to foster communication on reproducibility.
+People can report on fully, partiially, or un-successful reproductions of articles after publication.
 
@@
 intentionally designed to be picked up and implemented by many different 
-journals or events, and to build a community of codecheckers. CODECHECK shares
+journals or events and to build a community of codecheckers. CODECHECK shares
 its interdisciplinary nature with other community initiatives concerned with 
-reproducibility awareness, education, and supports, such as ReproHack
-\footnote{\url{https://reprohack.github.io/reprohack-hq/}}, Code Copilot
-\footnote{\url{https://twitter.com/Code_Copilot}}, or Papers with Code
-\footnote{\url{https://paperswithcode.com/about}}.
-The latter recently announced a collaborations with the preprint server
+reproducibility awareness, education, and support, such as
+ReproHack\footnote{\url{https://reprohack.github.io/reprohack-hq/}},
+Code Copilot\footnote{\url{https://twitter.com/Code_Copilot}}, or Papers with Code\footnote{\url{https://paperswithcode.com/about}}.
+The latter recently announced a collaboration with the preprint server
 \emph{arXiv} on providing data and code supplements for machine learning
-manuscripts\footnote{\url{https://medium.com/paperswithcode/papers-with-code-partners-with-arxiv-ecc362883167}} and runs a reproduciblity challenge
-\footnote{\url{https://paperswithcode.com/rc2020}}.
+manuscripts\footnote{\url{https://medium.com/paperswithcode/papers-with-code-partners-with-arxiv-ecc362883167}} and runs a reproduciblity challenge\footnote{\url{https://paperswithcode.com/rc2020}}.
 % Papers with Code now also does more disciplines than just ML.
 Likewise, different disciplines and journals provide reproducibility
-checklists, e.g.,~\cite{rosenberg_next_2020}, which naturally share some
+checklists, e.g.,~sciene and engineering \cite{rosenberg_next_2020} or GIScience \cite{nust_agile_2019}, which naturally share some
 aspects while addressing particularities as well as addressing researchers
 from different fields. Regarding the education and guidance for authors, 
-we see CODECHECK's role in referencing and linking educational efforts
-and helpful material, not as a creator and maintainer of such content.
+we see CODECHECK's role as referencing and linking educational efforts
+and helpful material, not as creating and maintaining such content.
 
@@
 \textbf{Isn't CODECHECK what peer review should be doing already?}
-On the surface, that is correct.
+On the surface, yes.
 However, the general observation is that the
 increasing number of researchers, submissions, and reviews to conduct
-under the pressure to publish, puts the scholarly publication system
-under a lot of strain.  
-Miyakawa described how simply requesting raw data from authors as an 
-editor lead to withdrawal or rejection of nearly all submissions where the
-request was necessary \cite{miyakawa_no_2020}. Such basic checks that
-sufficient data supports the results requires large efforts from editors,
-yet they can be based on availability of data files and the content of the
-paper. These checks can be enhanced by more complex CODECHECKs asking for
-the code, and, because code is not inert but needs to be run in an specific
-environment, executing the code.
-This might fall within the idealistic expectations for peer review,
+under the pressure to publish puts the scholarly publication system
+under significant strain.
+Miyakawa described how when he as an editor simply requested raw data from authors, this led to almost all of those papers in which data was requested to be withdrawn or rejected \cite{miyakawa_no_2020}.
+Such basic checks to show that
+sufficient data supports the results requires much effort from editors,
+yet they only rely on the availability of data files and the content of the
+paper.
+These availability checks can be enhanced by having more complex CODECHECKs request the code and then execute it, because code is not inert but needs to be run in a specific environment.
+This might fall within idealistic expectations of peer review,
 but is far from common in reality.
@@
 process is largely an acknowledgement that peer reviewing practices
-have faltered to adapt to the challenges of digitisation data science.
+have been unable to adapt to the challenges of digitisation of data-based science.
 The concept of a CODECHECK,
-just as the reproducibility of research and the openness of science,
-may be of transitional nature. If the activities described here as
+just as the concepts of reproducible research and Open Science,
+may be transitional by nature. If the activities described here as
 being part of a CODECHECK become the norm for scientific reviews and
@@
 eternally functional workflow and neither suggests a specific software stack
-or practical approaches, such as research compendia
-\footnote{\url{https://research-compendium.science/}},
-even if they are very valuable.
+or practical approach even if such suggestions, e.g.,~research compendia\footnote{\url{https://research-compendium.science/}}, are very valuable.
 Creating something that anyone can reproduce has been called a 
-fool's errand
-\footnote{\url{https://twitter.com/DougBlank/status/1135904909663068165?s=09}}
+fool's errand\footnote{\url{https://twitter.com/DougBlank/status/1135904909663068165?s=09}}
 and we tend to agree.
@@
 working analysis that greatly increases the likelihood of a successful 
-reproduction and possibility to extensions by third parties, with access
-to suitable resources and matching skill set, in the future.
+reproduction and the possibility that a workflow can be extended by third parties in the future, as long as they have access
+to suitable resources and matching skill set.
 Concrete implementations of CODECHECK workflows, especially for specific
-disciplines, may reify much more helpful guidelines for authors how to
+disciplines, may reify much more helpful guidelines for authors on how to
 create reproducibility packages.
@@
 as part of the widely accepted peer review concept, can provide may
-eventually allow to raise the bar as high as possible, e.g.,~with
-executable research compendia \cite{nust_opening_2017}
+eventually allow the bar to be raised as high as possible, e.g.,~with
+executable research compendia \cite{nust_opening_2017}, ``Whole Tales'' \cite{brinckman_computing_2018},
 or continuous analysis \cite{beaulieu-jones_reproducibility_2017-1}.
 However, considering that missing artefacts and lack of documentation
-have repeatedly been identified as main barriers to reproducibility
+have repeatedly been identified as key barriers to reproducibility
 (e.g.,~\cite{stagge_assessing_2019,nust_improving_2020}),
 we would not underestimate the power of a simple check.
-ModelDB curation policies ensure manually to reproduce one figure only 
-\cite{mcdougal_reproducibility_2016}, but that has not limited the usefulness
-nor success of the platform.
-
-It should be noted that a codechecker is not taking the same roles as the
-\emph{statistical reviewer}, as it is practices by some journals in the 
+For example, ModelDB curation policies ensure that only one figure is manually reproduced \cite{mcdougal_reproducibility_2016}, but that has not limited the usefulness nor success of the platform.
+
+It should be noted that a codechecker does not fulfil the same role as a
+\emph{statistical reviewer}, as it is applied by some journals in the 
 biomedical domain (cf.~\cite{petrovecki_role_2009,greenwood_how_2015}).
@@
 tests are applied \cite{petrovecki_role_2009}.
-The codechecker may go equally deep into the review, if they have, by 
-intention of by chance, the required expertise and the required time, 
+The codechecker may go equally deep into the review, if they have
+the required expertise and the required time, 
 but they are by no means expected to be able to provide
-such a level of scrutiny. We can imagine a tiered workflow where the 
-codechecker, just as the conventional reviewer, could recommend a detailed
-statistical review to the editor as they might come across upon different
-hints.
-%The CODECHECK provides a certain level of safety with regard to the inherent
-%problems that a computational workflow might have. We want to capture the
-%obvious errors, but ... https://www.quora.com/What-is-the-difference-between-safety-and-security
+such a level of scrutiny. We can imagine a tiered workflow where a 
+codechecker could, just as a conventional reviewer could, recommend a detailed
+statistical review to the editor if they come upon certain issues while
+examining the work from a different perspective.
 
 An important distinction is that a codechecker does not conduct a 
-\emph{code review}. Code reviews are without doubt extremely valuable in 
-terms of improving reproducibility and re-usability, and their proponents 
+\emph{code review}. Code reviews are without a doubt extremely valuable in 
+terms of improving reproducibility and reusability, and their proponents 
 even believe they can improve the science \cite{petre_code_2014}.
@@
 require even more resources. That said, a well-reviewed codebase is likely
-to be easier to codecheck, and the awareness for high-quality code raised
+to be easier to codecheck, and the awareness of high-quality code raised
 through CODECHECK may lead to more support for code reviewing.
-Initiatives and journals that conduct software reviews independently of 
+Initiatives and journals that conduct software reviews independent of 
 a specific publication or workflow include
@@
 We have no established process for the case
-that a reproduction fails. This may be the sign of a pre-selection bias
-of the examples, as so far all our CODECHECKs succeeded.
-In case of a journal adopting CODECHECK for all submissions, the open 
-question remains as what to do in this case will have to be answered.
-At the start, we doubt publicly reporting failures (i.e.,~the code 
-would not run) will increase overall reproducibility and recommend not
-to share the outcome publicly but to share the latest draft of the 
-certificate with the authors.
+in which a reproduction completey fails.
+This may be a sign that our pre-selection of the examples was biased, since so far all our CODECHECKs succeeded.
+In the case that a journal adopts CODECHECK for all submissions, the open 
+question remains as what to do if a check fails.
 A negative comment in a CODECHECK certificate or a failed check does
@@
 (cf. discussion on negative comments in \cite{everythinghertz123}).
-Therefore, one must be careful as to the impact of acceptance by authors 
-and the concerns of volunteers this approach, or a contrary one, has.
-Private feedback for failures and public acknowledgement/encouragement
-for successes seems to be the way to go.
+At the start, we doubt that publicly reporting failures (i.e.,~the code 
+would not run) will increase overall reproducibility, and we recommend not
+sharing that outcome publicly but sharing the latest draft of the 
+certificate with the authors.
+One must be careful how policies on disclosing failures may negatively impact the acceptance by authors, but also how efforts by volunteers may go unnoticed if no report is published.
+The most appropriate way forwards seems to be to give private feedback for failures and public acknowledgement and encouragement for successes.
 % eLife's new model (see https://osf.io/9cftx/) will publish reports evantually, e.g.,~after preprint has been accepted elsewhere
-\cite{Rosenthal2016b} discuss incentives for different actors around the
-implementation of reproducibility. We see CODECHECK as one mean for
+Rosenthal et~al. \cite{Rosenthal2016b} discuss such incentives for different actors around the
+implementation of reproducibility. We see CODECHECK as one way for
 organisations to invest in reproducibility by creating incentives
@@
 some disciplines, e.g MATLAB, this seems like a pragmatic choice.
-However, it prohibits to benefit from open infrastructure for reproducibility
+However, this prohibits publications from benefiting from open infrastructure for reproducibility
 (cf.~\cite{konkol_publishing_2020,perkel_make_2019})
@@
 
-\textbf{Can't someone cheat?} Yes. We don't check it is correct code or
+\textbf{Can't someone cheat?} Yes. We don't check that the code is correct or
 sound science,
-just simply that it runs. This `mechanical' test is indeed a very low bar.
+just simply that it runs. This ``mechanical'' test is indeed a very low bar.
 Yet, by having the code and (raw) data openly deposited, interested parties can
-then examine the code, and knowing that the code will be open we hope will
-be an incentive for authors to share. It will also give the opportunity 
+then examine the code, and we hope that knowing the code will be open will
+be an incentive for authors to ensure good quality.
+It will also give opportunities
 to future researchers, potentially with new methods, to look for errors.
-This is a much more effective means to avoid cheating than an arms race
-trying to catch the cheater with closed datasets and code. This approach
-is comparable to the storing of blood samples in sports to detect
-doping in the future (cf. \cite{everythinghertz97}).
-Another picture that helped us to define the scope of a CODECHECK is:
-the codechecker is the forensic photographer, making sure all the details
+This prospective evaluation is a much more effective way to capture cheaters than engaging in an arms race on tools, methods, or resources trying to catch malicious intent now with closed datasets and code; this approach is comparable to storing blood samples of sport champions today to possibly detect doping in the future with more advanced methods (cf.
+\cite{everythinghertz97}).
+Another comparison that helped us define the scope of a CODECHECK is was that we think of the codechecker as forensic photographer, making sure all the details
 are captured so that an investigator may, now or in the future, scrutinize
@@
 \textbf{Who's got time for more peer review?} Agree.
-Ignoring the complex community contract behind volunteered peer review,
-the time spent with CODECHECKs likely reduces the individual's time to work
+The time one spends performing CODECHECKs will likely reduce the time an individual has to work as a reviewer.
+However, a CODECHECK is different from peer review.
 First, the \textbf{technical nature} of a CODECHECK can provide a more 
-However, there are differences to peer review.
-First, the technical nature of a CODECHECK can provide a more 
 clearly defined scope and expectations and thereby time budget compared to conventional 
-peer review. The authors are told what to provide and the codechecker
-can be told when to stop, for example.
-Second, the specific skill set allows to introduce different groups
-into the peer review process. ECRs might be attracted to working with
-groups to learn more about recent methods and reproducibility practices.
-Research software experts (Research Software Engineers, RSE) who might
-not regularly be involved in writing or reviewing papers could be 
-interested to increase their connection with scholarly practices.
+peer review. 
+For example, authors are told what to provide and the codechecker can be told when to stop.
+The option for the codechecker to ask for clarification or more 
+material can greatly reduce the digging required and thereby
+increase \textbf{efficiency}.
+Second, the specific skill set of a codechecker allows for \textbf{different groups} to take part in the review process.
+ECRs might be attracted to learn more about recent methods, peer review, and reproducibility practices.
+Research software experts (Research Software Engineers, RSE\footnote{https://en.wikipedia.org/wiki/Research_software_engineering}) who might
+not regularly be involved in writing or reviewing papers might be interested in increasing their connection with scholarly practices.
 An extra codechecker may simplify the matchmaking an editor has to do
 when identifying suitable reviewers for a submission, as technical and
-topical expertise can be provided by different people, who must not even
-be traditional peer reviewers. In fact, the codechecker could even be
-a software professional employed by a publisher.
-Third, recall that we expect CODECHECKs to be not anonymous and always
-publicly deposited, which is common practice for some journals, but not
-broadly practiced. The possibility to ask for clarification or more 
-material/information can greatly reduce the digging required and thereby
-increase \textbf{efficiency}.
+topical expertise can be provided by different people who do not even have to be traditional peer reviewers.
+In fact, the codechecker could even be a software professional employed by a publisher.
+Third, recall that we expect CODECHECKs not to be anonymous and always
+publicly deposited, which is common practice for reviews of few journals only.
+With code and workflows, there also is a good chance that a codechecker's
+feedback can directly impact and improve the author's 
+work, which may make for a rewarding experience or even provide tangible contributions.
+The public certificates and contributions provide \textbf{peer recognition}.
 Fourth, we also found that the focus on the mechanics of a
 workflow and the ability to interact with the authors can make
-reproductions very \textbf{educational}. It's a different role and as such could
-be a welcome alternation/diversion for researchers to give back their 
-time to the community.
-With code and workflows, there also is a good chance that a codechecker's
-feedback can have direct impact and lead to improvements of the author's 
-work that comes with a rewarding experience or even tangible contributions.
-While such impact is also part of idealistic peer review, it is
-rather rare beyond paraphrased anonymous acknowledgement.
-
-\textbf{Checking multiple times?} If you check a preprint before peer
+reproductions very \textbf{educational}. It also is a different role and, as such, could
+be a welcome option for researchers to give back their time to the community.
+
+While such benefits are also part of idealistic peer review, they are mostly hidden behind paraphrased anonymous acknowledgement.
+
+\textbf{Do workflows need to be codechecked multiple times?}
+If you check a preprint before peer
 review or a submission to a journal with non-public review, it might
-need checking again as it is modified during peer review.  This is in
-part unavoidable, and as happened to us (e.g.,~certificate \texttt{2020-012}
-\cite{cert-2020-012} has been revised). This can also be desired, if
-the interaction between author, reviewer, and codechecker lead to
+need checking again as it is modified during peer review.
+This is somewhat unavoidable, and has happened to us (e.g.,~certificate \texttt{2020-012}
+\cite{cert-2020-012} has been revised). This can also be desired tough, if
+the interaction between author, reviewer, and codechecker led to
 improvements. Since checking the manuscript the second time is likely
@@
 open and regular communication. The higher the reproducibility and
-automation, the more aspects of a CODECHECK may eventually be
+automation, the more aspects of a CODECHECK that may eventually be
 supported by automation, though we believe the human codecheckers
-should remain the decision maker.  Automatically detecting if results
+should remain the decision makers. Automatically detecting whether results
 are ``the same'' is a question intrinsically linked with each unique
-process and quite hard to realise for common outputs, such as plots or
-maps.  Nevertheless, continuous integration (CI) infrastructure may
-help codecheckers, and, without much further work, use scientific
-workflows as detectors for when a piece of software breaks if a
-workflow is set up to use the latest versions of tools.
-
-As a principle, we do not require results to be exactly the same for a
-CODECHECK to pass, simply that the code runs and generates the output
+process and quite hard to realise for graphical outputs, such as plots or
+maps.
+That is why we do not require results to be exactly the same for a
+CODECHECK to pass but simply that the code runs and generates the output
 files that the author claims. Stochastic simulations mean that often
-we cannot get exactly the same, or even the same versions of libraries
+we cannot get exactly the same results, and that even the same versions of libraries
 can generate outputs that differ by operating system
@@
 
-\textbf{Not revolutionary enough?}
+\textbf{Shouldn't the next step be more revolutionary?}
 CODECHECK's approach is to acknowledge the shortcomings around 
-computational reproducibility and iteratively improve within the current
-system of scholarly communication through journals.
-It remains to be proved if this approach is welcomed broadly and if 
-involving the traditional system of publishers helps to push the 
-conversation along or not.
+computational reproducibility and to iteratively improve the current
+system of scholarly communication based on journals and conferences.
+It remains to be proven whether this approach is welcomed broadly and whether 
+involving the traditional publishing stakeholders helps to further the cause.
 We have discussed stronger more drastic rules at length, such as only 
@@
 
-We also have considered to require latest technologial solutions which can
-support reproducibility (cf.~\cite{konkol_publishing_2020}), but have 
-decided to focus on the human interface and the judgement of an experienced
-researchers/developer as a more sustainable approach.
-The structurelessness of the required data and software gives the needed
+We also have deliberated requiring the latest technologial solutions which can
+support reproducibility (cf.~\cite{konkol_publishing_2020}), but we have 
+decided to focus on the human interface and the judgement of experienced
+researchers and developers as a more sustainable and powerful approach.
+The structurelessness of the required data and code gives all types of research the freedom to adopt CODECHECK.
 CODECHECK could be complemented by automated scoring (e.g.,~\cite{menke_rigor_2020}), yet automation and metrics bear new risks.
@@
 The creation of the certificates and the interactions with authors and
-editors not only improved the principles and the workflow, but also
+editors not only improved the principles and the workflow but also
 confirmed the approach taken. This result corroborates findings from
@@
 shared principles and common name, through recognition value,
-allow researchers
+will allow researchers
 across publication venues to judge the level of scrutiny that results have
@@
 of reproductions or informal crediation and closed inner circles.
-Codechecking can be a worthwhile educational activity, especially for
+Forther, codechecking can be a worthwhile educational activity, especially for
 ECRs.
 
-Yet, CODECHECK underlies the same limitations as peer review
+Yet, CODECHECK harbours the same limitations as peer review
 in general and is closely connected to larger disruptions and challenges
@@
 \cite{eglen_recent_2018,tennant_ten_2019,fyfe_mission_2019}, 
-including the tensions between commercial publishing and reviewer's often
-free labour, and a global pandemic jumbling up scientific publishing
-and exposing a broader general audience to science, even preliminary 
+including the tensions between commercial publishing and reviewers' often
+free labour, and a global pandemic that has jumbled up scientific publishing
+and exposed a broader general audience to science, even preliminary 
 results \cite{munafo_what_2020}.
 It is clear that establishing CODECHECK-like systems can be 
-impeded by and must be seen in connection to much larger issues in 
-science, such as broken metrics or malpractices triggered by publication
+impeded by and must be seen as interconnected with much larger issues in 
+science, such as broken metrics or malpractice triggered by publication
 pressure \cite{piwowar_altmetrics:_2013,nosek_promoting_2015}.
@@
 
-While developed for the current "paper"-centric publication process and
+While developed for the current ``paper''-centric publication process and
 being compatible with different styles of peer review, the CODECHECK 
@@
 Public researcher profiles, such as ORCID, may consider different
-types of reviewer activity to capture the contributions of independent 
-code execution to science.
+types of reviewer activity to capture how independent code execution contributes to science.
 Notably, the discussed limitations are largely self-imposed for easier
 acceptance and evolutionary integration, as to not break the current 
-system and increase demands over time without leaving practitioners behind.
-
-If adopted today, a CODECHECK system, even if temporal in a sustainable 
+system and increase demands gradually without leaving practitioners behind.
+
+If adopted today, a CODECHECK system, even if temporary in a sustainable 
 transition towards more open publication and review practices, can contribute
@@
 touched on the procedural and technical implications of openness, be it 
-that OPR is neither applied nor well-known by stakeholders, or that process
+that OPR is neither applied by nor familiar to stakeholders, or that process
 management software would not support such open communication.
@@
 More reproducible practices initiated by CODECHECKs could lead
-communities/disciplines to reach a state where authors do provide good
-enough material and reviewers have acquired broady enough skills that
-the latter will generally conduct a CODECHECK-level of checking, and
+communities or whole disciplines to reach a state where authors provide 
+sufficient material and reviewers have acquired sufficient skills that
+the reviewers will generally conduct a CODECHECK-level of checking;
 only in especially sophisticated cases will a specialised codechecker
-be included.
-The main challenge for us remains to get journals to embrace the
-idea behind CODECHECK and realise processes conform to the principles,
-whether they use the actual name for it or not.
+be needed.
+The main challenge for us remains getting journals to embrace the
+idea behind CODECHECK and to realise processes that conform to the principles,
+whether or not they use CODECHECK by name.
 We would be keen to use the flexibility of the principles and cooperate
@@
 specific challenges -- Do CODECHECKs really work better with OPR?
+To facilitate the adoption, the CODECHECK badge is, intentionally, not branded beyond the checkmark
+and green colour and simply states ``code works''.
+
 Each rollout of CODECHECK should be accompanied by studies to ensure 
@@
 \cite{tennant_limitations_2020}
-and to ensure positive change within the respective peer review system.
-The CODECHECK badge is, intentionally, not branded beyond the checkmark
-and green colour and simply states ``code works''.
+and to ensure positive change within the review system.
 This \emph{cultural change}, however, is needed for the valuation of the
@@
 Multiple adopting journals will also help to answer relevant questions on
-the establishing process (What are crucial decisions or pain points? Can
-authors retract code/data once a CODECHECK has started?),
-what variants of CODECHECKs will be most common, and how open CODECHECKs
-may influence the scope and anonymity of conventional review over time.
+the establishing process: What are crucial decisions or pain points? Can
+authors retract code/data once a CODECHECK has started?
+What variants of CODECHECKs will be most common? How will open CODECHECKs
+influence or codevelop with the scope and anonymity of conventional review over time.
 
@@
 codecheckers will become relevant. We expect both a mentoring scheme within
-the CODECHECK community (experienced codecheckers provide on the job training
+the CODECHECK community (experienced codecheckers provide on-the-job training
 or serve as fallback advisors) as well as collaboration with reproducible
@@
 to be able to grow the pool of codecheckers as needed.
-The initial reaction of the scientific community and the spontaneously signing
-up volunteer codecheckers make us feel optimistic that the time is right for 
-putting scholarly peer review on the path to facilitate sharing and execution
+The initial reaction of the scientific community and the spontaneous signing
+up volunteer codecheckers make us optimistic that the time is right for 
+putting scholarly peer review on the path towards facilitating the sharing and execution
 of computational workflows.
@@
 
-SJE is on the editorial board at the journal \emph{Scientific Data}.
-DN is reproducibility chair at the Association of Geographic
-Information Laboratories in Europe's (AGILE) annual conference.
+SJE was on the editorial board at the journal \emph{Scientific Data} until January 2021.
+DN is reproducibility chair at the Association of Geographic Information Laboratories in Europe's (AGILE) annual conference.
 
@@
 all authors of auditable workflows for their participation.
+We thank Celeste R. Brennecka from the Scientific Editing Service, University of M\"unster, for her editorial review.
 

### 2021-02-01 codecheck-paper.tex
manual page breaks

@@
 certificate repository to the published paper or the original code repository.
+
+\newpage
 
@@
 
-\textbf{1. Codecheckers record but don't investigate or fix.} The
-codechecker follows the instructions provided by the author in running
+\textbf{1. Codecheckers record but don't investigate or fix.} \\
+The codechecker follows the instructions provided by the author in running
 the code. If any steps are unclear, or if code does not run correctly,
@@
 
-\textbf{2. Communication between humans is key.} Some code may
-work without any interaction (e.g.,~CODECHECK certificate \texttt{2020-013}
+\textbf{2. Communication between humans is key.} \\ 
+Some code may
+work without any interaction (e.g., CODECHECK certificate \texttt{2020-013}
 \cite{cert-2020-013}) but more often there are hidden
@@
 
-\textbf{3. Credit is given to codecheckers.} The value of performing a
+\textbf{3. Credit is given to codecheckers.} \\ 
+The value of performing a
 CODECHECK is comparable to that of a scientific peer review, and
@@
 
-\textbf{4. Workflows must be auditable.}
+\textbf{4. Workflows must be auditable.} \\ 
 This requires that the codechecker 
@@
 
-\textbf{5. Open by default and transitional by disposition.}
+\textbf{5. Open by default and transitional by disposition.} \\ 
 By default, unless there are strong
@@
 
-\textbf{Isn't CODECHECK what peer review should be doing already?}
+\textbf{Isn't CODECHECK what peer review should be doing already?} \\
 On the surface, yes.
@@
 
-\textbf{Aren't CODECHECK requirements not demanding enough?}
+\newpage
+
+\textbf{Should CODECHECK requirements not be more demanding?} \\
 CODECHECK by design does not require authors to provide (and sustain) an
@@
 
+\newpage
+
 \subsection*{Competing interests}

### 2021-02-02 codecheck-paper.tex
mention repro-packs alongside research compendia

@@
 eternally functional workflow and neither suggests a specific software stack
-or practical approach even if such suggestions, e.g.,~research compendia\footnote{\url{https://research-compendium.science/}}, are very valuable.
+or practical approach even if such suggestions, e.g.,~research compendia or ``repro-packs''\footnote{See \url{https://research-compendium.science/} respectively \cite{barba_praxis_2018}.}, are very valuable.
 Creating something that anyone can reproduce has been called a 

### 2021-02-04 codecheck-paper.tex

@@
 the work presented here: Andy Collings, Melissa Harrison, Giuliano
-Maciocci, Naomi Penfold (eLife), Rebecca Kirk (PLOS Computational Biology), Scott
+Maciocci, Naomi Penfold, Emmy Tsang (eLife), Rebecca Kirk (PLOS Computational Biology), Scott
 Edmunds (GigaScience), and Andrew Hufton (Scientific Data). Iain Davies and

### 2021-02-04 codecheck-paper.tex
about to read

@@
 ECRs might be attracted to learn more about recent methods, peer review, and reproducibility practices.
-Research software experts (Research Software Engineers, RSE\footnote{https://en.wikipedia.org/wiki/Research_software_engineering}) who might
+Research software experts (Research Software Engineers, RSE\footnote{\url{https://en.wikipedia.org/wiki/Research_software_engineering}}) who might
 not regularly be involved in writing or reviewing papers might be interested in increasing their connection with scholarly practices.
@@
 
-% question mark indicates we are waiting for person to respond.
 We are grateful to the following individuals for discussions regarding

### 2021-02-07 codecheck-paper.tex
finished up to "core principles" section

@@
 
+\usepackage{fontspec}
 \usepackage{fontawesome}
@@
 \affil[1]{Institute for Geoinformatics, University of M\"unster, Germany; \email{daniel.nuest@uni-muenster.de}}
-\affil[2]{Department of Applied Mathematics and Theoretical Physics, University of Cambridge, GB; \email{sje30@cam.ac.uk}}
+\affil[2]{Department of Applied Mathematics and Theoretical Physics, University of Cambridge, UK; \email{sje30@cam.ac.uk}}
 \affil[$\ast$]{Contributed equally}
@@
 \begin{abstract}
-  In its traditional format, the scientific paper falls short of effectively communicating computational research.
+  The traditional scientific paper falls short of effectively communicating computational research.
   Since this issue is difficult to address due to lacking incentives, guidance, benefits, recognitions, and reproducibility, we come at this problem from a different angle:
-  We propose to introduce, as part of peer review, a system by which the workflows underlying research articles are executed.
-  The CODECHECK system, which uses free intrastructure and open tools, relies on five principles and can be integrated into review and publication processes in multiple ways.
+  We propose a system by which the computational workflows underlying research articles are checked.
+  The CODECHECK system uses open intrastructure and tools and can be integrated into review and publication processes in multiple ways.
   We describe these integrations along multiple dimensions (importance, who, openness, when).
@@
   These CODECHECKs show that asking for reproducible workflows during a collaborative review can effectively improve executability.
-  While CODECHECK has clear limitations, it may represent a building block in the Open Science and publishing ecosystems for improving the reproducibility, appreciation, and, potentially, the quality of non-textual research artefacts.
+  While CODECHECK has clear limitations, it may represent a building block in Open Science and publishing ecosystems for improving the reproducibility, appreciation, and, potentially, the quality of non-textual research artefacts.
 \end{abstract}
 
+\clearpage
+
 \section*{Introduction}\label{introduction}
 
-Many areas of scientific research use computations to either simulate or analyse their data.
-These computations are becoming increasingly complex and are difficult to explain coherently in a paper \citep{marwick_how_2015}.
+Many areas of scientific research use computations to simulate or analyse their data. These computations are complex and are difficult to explain coherently in a paper \citep{marwick_how_2015}.
 To complement the traditional route of sharing research by writing papers,
@@
 
-So, assuming that researchers begin to share more artefacts, how might these artefacts be examined to ensure that they do what they claim?
+If researchers start sharing more artefacts, how might these artefacts be examined to ensure that they do what they claim?
 For example,
-most scientific journals now require a data sharing statement that
-outlines what data the authors have (or will) share.
-Yet, each journal can implement this requirement differently.
+although  scientific journals now require a data sharing statement that
+outlines what data the authors have (or will) share, journals  implement this differently.
 At one end of the spectrum, journals have been created to accept ``data papers''
@@
 the data is validated according to standards in each field.
-At the other end of the spectrum, many journals still simply allow authors to state ``Data available upon reasonable request''.
-Authors, while possibly well intentioned at the time of writing the article, often cannot provide the data when readers ask for it, as data disappears over time \cite{Vines2014-hf}.
+At the other end of the spectrum, many journals still allow authors to state ``Data available upon reasonable request''.
+Authors, while possibly well intentioned at the time of writing the article, often cannot provide data when requested as data disappears over time \cite{Vines2014-hf}.
 
 Given that no clear standards yet exist for sharing data, what hope might there be for sharing computer programs?
-Both data and software together are required to validate outcomes of a computational analysis, but they differ in that data can be seen as static and inert,
-while code requires an environment to be run in and interacted with.
+Both data and software are required to validate a
+computational analysis; data can be seen as inert whereas
+code requires an environment to be run in.
 This makes software harder to share.
-Our anectodal experiences on this matter suggest that researchers often give a variety of reasons for why code cannot be shared, e.g.,~``there is no documentation'', ``I cannot maintain it'', or ``I do not want to give away my code to competitors''.
-Our view, rooted in Open~Science principles\footnote{\url{https://en.wikipedia.org/wiki/Open_science}}, is that sharing code, wherever possible, is good for the community and the individual \cite{Barnes2010-iv,markowetz_five_2015}.
-Having code and data openly available and archived provides a valuable resource for others to learn from, even if the code no longer runs or if there is no documentation.
-However, with a little effort, we believe that if an independent person can execute an analysis, this is worth documenting as early as possible and that such a focused approach can greatly reduce the barrier to evaluating non-text research materials.
-Just as data journals' validations of data and all journals' peer review can provide a ``baseline reassurance'', i.e.,~that a paper has been checked by someone with an understanding of the topic \cite{fyfe_mission_2019}, we think the same baseline should be provided for the workflow underlying a paper.
-With this in mind, we have developed a set of principles and an example workflow that provides what we hope is a pragmatic way of checking that a paper's code works.
-
-As this work's main contribution, we offer a thorough description of a process and its variations to integrate a much needed evaluation of computational reproducibility\footnote{Following the Claerbout/Donoho/Peng terminology, see \cite{barba_terminologies_2018}.} into peer review, and we demonstrate its feasibility by means of 25 reproductions across different scientific disciplines.
-We call this system CODECHECK as outlined next.
+Our experience is that researchers offer several reasons for why code
+is not shared, e.g.,~``there is no documentation'', ``I cannot maintain it'', or ``I do not want to give away my code to competitors''.
+Our view is that sharing code, wherever possible, is good for the community and the individual \cite{Barnes2010-iv,markowetz_five_2015}.
+Having code and data openly available and archived provides a valuable
+resource for others to learn from, even if the code is broken or lacks documentation.
+However, with a little effort, we believe that if an independent person can execute an analysis, this is worth documenting and that this  reduces the barrier to evaluating non-text research materials.
+Just as data journals' validations of data and all journals' peer review  provides a ``baseline reassurance'', i.e.,~that a paper has been checked by someone with an understanding of the topic \cite{fyfe_mission_2019}, the same baseline could be provided for the workflow underlying a paper.
+With this in mind, we have developed a set of principles and an example workflow that provides a pragmatic way of checking that a paper's code works.
+
+Here we offer a thorough description of a process and its variations to integrate a much-needed evaluation of computational reproducibility\footnote{Following the Claerbout/Donoho/Peng terminology, see \cite{barba_terminologies_2018}.} into peer review, and we demonstrate its feasibility by means of 25 reproductions across scientific disciplines.
+We call this system CODECHECK.
 
@@
 (3) the CODECHECKER, who checks that the authors's code works.
-The workflow we have refined is shown in Figure~\ref{fig:workflow}
-and consists of six steps.
+The six-step workflow we have refined is shown in Figure~\ref{fig:workflow}.
 
@@
       \includegraphics[width=\textwidth]{figs/codecheck_overview.pdf}
-  \caption{The CODECHECK example process implementation. Codechecker act as a detectives:
-  They investigate and record, but do not fix issues.}
+      \caption{The CODECHECK example process implementation.
+        Codechecker act as a detectives: They investigate and record,
+        but do not fix issues.  Numbers in bold refer to the six steps
+        outlined in the text.}
   \label{fig:workflow}
@@
 
-\textbf{Step 2:} The publisher finds a CODECHECKER to check the code. This is
-analogous to the publisher finding one or more reviewers to evaluate
-the paper.
-However, a crucial difference is that we suggest the CODECHECKER and the author are able to talk freely and directly to each, regardless how open the scientific peer review process is.
+\textbf{Step 2:} The publisher finds a CODECHECKER to check the
+code. This is analogous to the publisher finding one or more
+peer-reviewers to evaluate the paper, except we suggest that the
+CODECHECKER and the author talk directly to each other.
 
@@
 
-\textbf{Step 4:} The CODECHECKER writes a certificate stating how the code was
-run and includes a copy of outputs (e.g.,~data, figures, or tables) that were
-independently generated.
-The certificate may include further recommendations to the author on how to improve the material.
-The free text in the certificate is the most flexible and feasible way to 
-make clear what was really checked, because each workflow is unique.
-Since no specific tool or platform should be required, such that no authors are excluded, it would be futile for the CODECHECKER to use automation or fixed checklists.
-
-\textbf{Step 5:} The certificate as well as auxiliary files created
+\textbf{Step 4:} The CODECHECKER writes a certificate stating how the
+code was run and includes a copy of outputs (e.g.,~data, figures, or
+tables) that were independently generated.  The certificate may
+include further recommendations on how to improve the material.  The
+free text in the certificate can describe what was really checked,
+because each workflow is unique.  Since no specific tool or platform
+should be required, such that no authors are excluded, it would be
+futile for the CODECHECKER to use automation or fixed checklists.
+
+\textbf{Step 5:} The certificate and auxiliary files created
 during the check, e.g.,~a specification of a computing environment, data 
@@
 \textbf{Step 6:} The PUBLISHER can, depending on the timing, provide the
-certificate to scientific reviewers or handling editors or publish it
-and link between the certificate, the paper, and any repositories 
+certificate to peer-reviewers or handling editors or publish it
+and link between  certificate, paper, and any repositories 
 through metadata. Currently, the CODECHECKER creates these connections on
@@
 sidebar of certificate \texttt{2020-025} \cite{cert-2020-025}.}.
-The PUBLISHER also gives credit to the CODECHECKER for their work by depositing
+The PUBLISHER also credits the CODECHECKER for their work by depositing
 the activity appropriately in scholarly profiles, such as ORCID\footnote{
@@
 
-The example workflow leaves room for this process to be implemented differently.
-To elaborate on these options, we consider several dimensions in a space of possible CODECHECK workflows, as shown in Figure~\ref{fig:dimensions}, to represent relevant aspects for the involved stakeholders.
+The example workflow leaves room for variations to this process.
+Here we consider several dimensions in a space of possible CODECHECK workflows(Figure~\ref{fig:dimensions}), to represent relevant aspects for stakeholders.
 These aspects touch on timing, responsibilities, and 
@@
 
-The point in time at which a CODECHECK is done and its ascribed importance are closely connected, so we describe the dimensions \emph{When} and \emph{Importance} together.
+The time at which a CODECHECK is done and its ascribed importance are closely connected, so we describe the dimensions \emph{When} and \emph{Importance} together.
 The earlier a CODECHECK happens in the publishing process, the more it can affect editorial decisions:
 Is a paper published, sent back for revisions, or rejected?
-Even earlier checks, i.e.,~a CODECHECK of a preprint, may even help to improve the workflow itself, even before a publisher is involved.
-As such, codechecking papers could be part of a preprint server's policy or simply be initiated by interested readers and community members.
+Even earlier checks, i.e.,~a CODECHECK of a preprint, may help to improve the workflow itself, even before a publisher is involved.
+As such, codechecking papers could be part of a preprint server's
+policy or initiated by interested authors.
 
 Publishers could introduce a CODECHECK as a \textbf{strict prerequisite}.
-As this can reduce the workload of reviewers, such a check should not occur late in the review process.
+As this can reduce the workload of reviewers, such a check should occur early in the review process.
 Yet, the later in the review process the check happens, the easier is it to allow bidirectional communication between the author and codechecker, e.g., because the author might already be notified of the paper's acceptance and may more freely share materials online closer to the paper's publication date.
-A \textbf{pre-review} CODECHECK would allow editors to only send a submission out for peer review if it passes the check, or to include the certificate in the submission package provided to the reviewers.
-Reviewers may then judge the the relevance of the computations for the results of the work, or follow journal guidelines regarding how strictly to apply the check's outcomes in their evaluation.
+A \textbf{pre-review} CODECHECK means editors could send a submission
+for peer review only if it passes the check, or include the certificate in the submission package provided to peer-reviewers.
+Reviewers may then judge the relevance of the computations for the results of the work, or follow journal guidelines regarding how strictly to apply the check's outcomes in their evaluation.
 
 A CODECHECK may also be conducted in \textbf{parallel} to the scientific peer review.
-This puts less burden on the turnaround time for the check, yet it only makes the outcomes available during the final consideration by the handling editor.
+This puts less burden on the turnaround time for the CODECHECK, yet it only makes the outcomes available during the final consideration by the handling editor.
 The check could also be assigned after suggestion by a reviewer, which would remove the need for submissions to undergo a pre-review screening.
-However, soliciting such a ``specialist review'' is much less desirable than having a regular CODECHECK, thus avoiding the situation in which some submissions get special treatment.
-In both cases, a CODECHECK could then be awarded \textbf{equal weight} compared to the other reviews for the editor's decision.
-
-A \textbf{post-acceptance} CODECHECK would have the smallest impact on editorial decisions and may simply provide \textbf{extra merit} on top of the submission's acceptance.
-This is the least impactful solution in which all material is still evaluated and the results of the check are properly acknowledged, because the check can be completed before the publication of the paper.
-The GIScience group of checks (see below) falls into this category:
-By displaying a badge on the volume and article landing pages, the AGILE conference highlights articles whose reproducibility was successfully reviewed.
-Similarly, in collaborations with journals, some GIScience have been checked while the authors worked on revisions.
-
-A CODECHECK may also be conducted \textbf{post-publication}, though it would require someone to update the article and article metadata to reference the check; otherwise, readers would presumably not be aware of it.
-In general, publishers hesite to make such revisions to published articles.
-While this option has the least impact on current publishing practices and enables open discourse, it also has the smallest effect on the original submission and does not acknowledge the importance of reproducible workflows for ensuring good scientific practice.
-
-In general, the timing of a CODECHECK needs to be connected with 
-established or typical scholarly communication process.
-While more radical ideas do exist for changing the paradigms of how researchers can more granually share their work much
-\footnote{For example Octopus, \url{https://science-octopus.org/about}, or Hypergraph, \url{https://www.libscie.org/hypergraph}.},
-% mimosa is another platform idea, but no real prototype AFAICT:
-% https://projects.invisionapp.com/share/E9Z3RKE7W3P#/screens/436978924
-% https://docs.google.com/presentation/d/1V_K8hghgnvGfEtW7TdwtowTveZyy4Qs8rMemNyHx_Dg/edit#slide=id.p
-an evolutionary approach has value, and we do not yet suggest to completely eliminating research papers as a way of sharing work.
-Enhancing existing processes with CODECHECKs allows communities to transition towards more open practices at their own pace and ensure that no one is unfaily left behind.
+However, soliciting such a ``specialist review'' is much less
+desirable than having a regular CODECHECK, thus avoiding the situation
+in which some submissions get special treatment. In both cases, the
+editor's decision could be based both on CODECHECK and peer-review reports.
+
+A \textbf{post-acceptance} CODECHECK would have the smallest impact on
+editorial decisions and may simply provide \textbf{extra merit} on top
+of the submission's acceptance.  This is the least impactful solution
+in which all material is still evaluated and the results of the check
+are properly acknowledged, because the check can be completed before
+publication of the paper.  The GIScience checks (see
+below) falls into this category: By displaying a badge on the volume
+and article landing pages, the AGILE conference highlights articles
+whose reproducibility was confirmed.  Similarly, in
+collaborations with journals, some GIScience articles were
+checked whilst authors worked on revisions.
+
+A CODECHECK may also be conducted \textbf{post-publication}, though
+this requires an update to the article and article metadata to
+reference the check so that readers can find the CODECHECK.  In
+general, publishers hesitate to make such revisions to published
+articles.  We do not prefer this option as it has the least impact on
+current publishing practices and downplays the importance
+of reproducible workflows for ensuring good scientific practice.
+
+%The timing of a CODECHECK needs connecting to typical scholarly communication process.
+%While ideas exist for radically changing how researchers can share their work
+% \footnote{For example Octopus, \url{https://science-octopus.org/about}, or Hypergraph, \url{https://www.libscie.org/hypergraph}.},
+% % mimosa is another platform idea, but no real prototype AFAICT:
+% % https://projects.invisionapp.com/share/E9Z3RKE7W3P#/screens/436978924
+% % https://docs.google.com/presentation/d/1V_K8hghgnvGfEtW7TdwtowTveZyy4Qs8rMemNyHx_Dg/edit#slide=id.p
+% an evolutionary approach has value, and we do not yet suggest to completely eliminating research papers as a way of sharing work.
+Enhancing existing processes with CODECHECKs allows communities to
+gradually transition towards more open practices.
 When integrating a CODECHECK into existing review and publication processes, the \emph{turnaround time} is crucial.
-Depending on when and who conducts the check, it might be done quickly or it might add considerable time to how long it takes from submission to publication.
-In our first checks, we found that a CODECHECK generally takes from 2-5 hours, with some outliers on the higher end.
-This time includes writing and publishing the certificate but excludes the time for actual computations.
+Depending on when and who conducts the check, it might be done quickly
+or it might delay publication.
+We found that a CODECHECK generally takes  2--5 hours, with some outliers on the higher end.
+This time includes writing and publishing the certificate but excludes
+actual computation time, some of which took days.
 These efforts are comparable to the time needed to peer review a submission,
@@
 make.
-However, the time includes a considerable amount of communicating about the
+Currently, there is considerable amount of communicating about the
 process, especially regarding who publishes which document when, so that proper
 cross-referencing between paper and certificate is ensured via persistent
-identifiers.
-When integrated into a peer review platform, this handling of documents should
-become much more effective and reduce organisational overhead.
+identifiers.  When integrated into a peer review platform, this
+handling of documents should become much more streamlined.
 
@@
 
-The topic of anonymity is broadly discussed, especially in the push towards
+Anonymity is broadly discussed, especially in the push towards
 open peer review as part of the Open Science movement 
 (cf. \cite{ross-hellauer_guidelines_2019}).
-Without taking a strong stance on this topic, the general motivation 
+Without taking a strong stance on this topic, our motivation 
 behind CODECHECK
@@
 review process.
-However, anonymity can also serve to protect individuals \cite{tennant_limitations_2020}, e.g.,~junior scientists, whereas revealing of names from tenured researchers might be more reasonable to require.
-The possible negative effects of a non-anonymous review are reduced if a CODECHECK is not relevant for a journal's decision to accept or reject, but that is, of course, not desirable when the goal is higher transparency and reproducibility.
-Instead, we argue that a CODECHECK is technical process that should in general find fixable problems; it is not aimed at giving an opinion or identifying a faulty approach.
-It is technically possible for a nonsense workflow to receive
-a CODECHECK certificate.
-If passing a CODECHECK becomes mandatory, the full transparency may have to be 
-revisited as the relations between authors and codecheckers would fall under the
-same social and community challenges as non-anonymous peer review
+However, anonymity can protect individuals
+\cite{tennant_limitations_2020}, e.g.,~junior scientists.  
+The negative effects of a signed review may be reduced if a CODECHECK is not relevant for a journal's decision to accept or reject, but that is, of course, not desirable when the goal is higher transparency and reproducibility.
+Instead, CODECHECK is a technical process that should generally find fixable problems; it is not aimed at giving an opinion or identifying a faulty approach.
+If passing a CODECHECK becomes mandatory, full transparency may need
+revisiting as the relations between authors and codecheckers would fall under the
+same social and community challenges as open peer review
 (cf.~\cite{everythinghertz123}).
@@
 
-The technical nature of the check and the challenge of providing sufficient documentation is why we see great benefits in allowing for bidirectional communication between author and codechecker.
+The technical nature of the check and the challenge of providing sufficient documentation is why we see great benefits in bidirectional communication between author and codechecker.
 Instead of trying to
 fix problems or guess the next step, the codechecker can ask the author to 
-rework the documentation or the update code.
+rework the documentation or update code.
 Instead of struggling to provide perfect instructions and as a result possibly not sharing any code or data, the author can make a best effort to document sufficiently.
-Of course, the codechecker may also provide a fix within the code instead
-of lengthily describing a small change if it can be quickly done.
 Authors and readers can profit from a codecheckers' experience and approach, as during the check they may create useful and instructive files, e.g.,~a machine-readable computing environment specification.
-While communication between author and codechecker may be facilitated in an 
-anonymous way by the 
-publisher, it most likely only helps to protect the identity of the codechecker, because, especially when codebases are of a certain size, not trivial, or if very specialised, they are almost impossible to fully anonymise.
-Therefore, the most effective and desirable situation for the stakeholders is to hold a non-anonymous and collaborative CODECHECK.
+While communication between author and codechecker may be anonymised via
+the publisher, it most likely only helps to protect the identity of
+the codechecker, because code is hard to anonymise.
+Therefore, the most effective and desirable situation for the
+stakeholders is to hold a open and collaborative CODECHECK.
 The contributions by the codechecker may even be integrated into
@@
 proper credit can be given within the research software development community.
-Nevertheless, depending on the needs represented by other dimensions,
-the aspect of ``Who knows who?'' may be the most flexible.
 
@@
 
-The options for who can conduct a CODECHECK require a somewhat complex matching of a combination of skills and availability.
-Ideally, thecodechecker has a matching code \emph{and} domain expertise; for example, to check a workflow based on Python code analysing neuroimaging datasets, the codechecker should ideally be experienced in Python and neuroimaging analysis.
-However,
-a well-documented workflow should be executable by any person with a basic
-understanding of computers. Naturally, the more prerequisite knowledge the
-codechecker has, the less time is needed to understand the goals and 
-the mechanics of an analysis.
-From our experiences, the priority should be given to matching technical expertise first, as lacking knowledge in setting up a computing environment with a particular language or tool is much more of a problem than assessing the outcome, e.g.,~comparing created figures with the original, without an in-depth understanding of the science.
-The depth of the check will mostly be driven by the time required and expertise of the checker, though in general, we expect a CODECHECK not to evaluate performance (e.g.,~it will not check whether a racing car has the best apex speed) but to assess operation free of failure (i.e.,~is it a car with 4 wheels and does it move forwards when accelerating).
-This depth can be seen independent of the check's importance, as long as it is communicated clearly to all stakeholders, most importantly authors and readers.
-
-The concrete people to provide such expertise can range from researchers,
-possibly drawn from the regular pool of volunteer
+Just as with peer-reviewers, a potential codechecker should have the
+right skills and availability to do the work.  Ideally, the
+codechecker has a matching code \emph{and} domain expertise to the
+paper, although a well-documented workflow should be executable by any
+computationally-competent person. Naturally, the more
+prerequisite knowledge the codechecker has, the less time is needed to
+understand the goals and the mechanics of an analysis.  From our
+experiences, the priority should be given to matching technical
+expertise first, as lacking knowledge in setting up a computing
+environment with a particular language or tool is much more of a
+problem than assessing the outcome, e.g.,~comparing created figures
+with the original, without an in-depth understanding of the science.
+The depth of the check will mostly be driven by the time required and
+expertise of the checker, though in general, we expect a CODECHECK 
+to consider reproducbility of the results above performance of the code.
+
+Codecheckers could be drawn from a regular pool of volunteer
 \textbf{scientific reviewers}, or from a special group of
-\textbf{reproducibility reviewers} via specific roles
-such as \textbf{reproducibility editors}, to experts working as part of
-\textbf{editorial staff} with a publisher. If the regular reviewers
-are relied upon, the editor's task to find suitable ones becomes more
-tedious; extra information such as software familiarity would have to be
-modelled in review platforms.
-
-We consider a \emph{single codechecker} to be enough, unlike peer review where
-multiple referees are common.
-Because of the technical focus, determining whether the concrete outputs produced by the author and the codechecker, who can collaborate to solve any issues, actually match should be a very factual process and not require much debate or opinion.
-Code usually harbours systematic and repeatable mistakes and is thereby more reliable and auditable than processes controlled by humans \cite{tibav:42484}, e.g.,~in a laboratory.
-Depending on the importance of the CODECHECK
-(see above), a mediator, e.g.,~the editor, or a
-second opinion might be warranted in rare cases.
+\textbf{reproducibility reviewers} via specific roles such as
+\textbf{reproducibility editors}, or \textbf{editorial staff} with a
+publisher.  A \emph{single codechecker} is sufficient to verify the
+workflow since it is mostly a factual process and not require much
+debate or opinion.  Code usually harbours systematic and repeatable
+mistakes and is thereby more reliable and auditable than processes
+controlled by humans \cite{tibav:42484}, e.g.,~in a laboratory.  If
+however publication of the paper depends on the CODECHECK, a second
+opinion may be required.
 
@@
 ECRs arguably have a high interest in learning
-about the latest tools and technologies, as they are building up their own
+about the latest tools and technologies, as they build up their own
 expertise and specialisation.
-CODECHECK offes a way for ECRs to gain insights into the latest science and at the same time can help increase the appreciation of reproduction efforts, a combination that is of particular interest.
+CODECHECK offes a way for ECRs to gain insights into new research and
+highlight the importance of reproduction.  
 \emph{ReScience X}, a journal devoted to reproduction and replication experiments \cite{roesch_new_2020}, shares an interest in this combination.
-Further, ECRs are often much more familiar with latest technology and tend to be early adopters, thus also making them likely to author CODECHECK-ready manuscripts
+ECRs are also often familiar with new technologies, thus also making them likely to author CODECHECK-ready manuscripts
 \footnote{Cf. the 77\% of 141 registered reports that were submitted to leading Neuroscience journals by ECRs \cite{chambers_registered_2019}.}.
 As ECRs are introduced to peer review as codecheckers, they may transition into the role of scientific reviewer over time.
-Another potential highlight is that during the codechecking process, a senior codechecker may support a junior codechecker, which would be the opposite of what happens during general peer review, which is largely an unsupervised process, i.e.,~rarely are reviewers taught how to do it.
 Overall, we see several opportunities and benefits to setting up a new process for CODECHECKING with a clear commitment to openness and transparency, independent of the current peer review process (see \emph{Openness} dimension).
 
-Having the codechecker be part of the publisher's \textbf{editorial staff} represents the most controlled but also
-resource-intensive option.
-Hiring the required technical and or domain expertise will put a financial burden on the publisher, of course, but such a commitment would also show that publishers are investing in reproducibility.
-Codecheckers on staff would positively impact the submission experience and the external perception, because having an in-house codechecker would make the process easier to control and would ensure the author gets an external, unbiased review.
-Furthermore, a staff codechecker would also reduce the challenges regarding open communication and awarding credit, as in-house personnel may more readily be given access to submission material would receive a salary instead of public credit as part of the scientific community.
-Yet, keeping diverse specialists on as staff to realise a codechecking process is 
-unlikely to be feasible for small or independent publishers.
-
-In contrast, for researchers it can be very important to be publicly credited
-for their activity as a reviewer.
-A regular
-review may be listed in public databases (e.g.,~ORCID
-\footnote{\href{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}} or commercial offerings such as Publons
+The codechecker could be a member of \textbf{editorial staff}; this is
+the most controlled but also resource-intensive option.  Such a
+resource commitment would show that publishers are investing in
+reproducibility, yet this commitment may be hard for small publishers.
+These codecheckers could be fully integrated into internal workflows.
+Credit for doing the codecheck is also achieved, as it is part of
+their duties.  By contrast, it is useful for researchers to be
+publicly credited for their reviewing activity.  A regular review
+may be listed in public databases (e.g.,~ORCID
+\footnote{\href{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}}
+or commercial offerings such as Publons
 \footnote{\href{https://publons.com/}{https://publons.com/}} or
 ReviewerCredits\footnote{\href{https://www.reviewercredits.com/}{https://www.reviewercredits.com/}});
-a codechecker should be similarly listed.
-The number of scientific reviews is of course only a coarse indicator of an individual's community contribution, yet it is unclear if a regular reviewer, who conductsa regular scientific review and in addition also conducts a CODECHECK, should be credited twice.
-In this case, the published CODECHECK certificate and possibly the published peer review can clearly describe the possibly large efforts put into both.
-
-The codechecker community\footnote{\href{https://github.com/codecheckers/codecheckers/}{https://github.com/codecheckers/codecheckers/}}
-currently has over 20 individuals who signed up in the last 12 months as volunteer codecheckers after talks given on the project, even though no checks are actively being requested in collaboration with journals.
-These volunteer's motivations, mentioned in the registration information\footnote{\url{https://github.com/codecheckers/codecheckers/labels/registration}}, are that they want to
-support reproducible research and Open Science,
-establish a culture of open and reproducible science,
-be part of making this culture the norm,
-improve coding skills,
-give back to the community,
-gain experience in helping scientists with their code,
-reduce the amount of bad scientific code, 
-encourage a culture that reduces friction in sharing or passing along on academic codebases,
-learn from other people's mistakes;
-many are also motivated simply by curiosity.
-We see benefits to an open shared list of codecheckers across journals rather than a private in-house group, as this may allow for better matches regarding expertise and more even sharing of the workload.
-Overall, this community will aim to ensure that CODECHECK is a viable option not just for large publishers, independent of their financial model, but also for independent no-cost Open Access journals.
+a codechecker could be similarly listed.
+% The number of scientific reviews is of course only a coarse indicator of an individual's community contribution, yet it is unclear if a regular reviewer, who conductsa regular scientific review and in addition also conducts a CODECHECK, should be credited twice.
+% In this case, the published CODECHECK certificate and possibly the published peer review can clearly describe the possibly large efforts put into both.
+
+The codechecker
+community\footnote{\href{https://github.com/codecheckers/codecheckers/}{https://github.com/codecheckers/codecheckers/}}
+currently has over 20 individuals who volunteered in the last 12
+years.  These volunteer's motivations, mentioned in the registration
+information\footnote{\url{https://github.com/codecheckers/codecheckers/labels/registration}},
+include: supporting reproducible research and Open Science, improve
+coding skills, gaining experience in helping scientists with their
+code, encouraging a sharing culture, and learning from other people's
+mistakes; many are also motivated simply by curiosity.  We see
+benefits to an open shared list of codecheckers across journals rather
+than a private in-house group, as this may allow for better matches
+regarding expertise and more even sharing of the workload.  Overall,
+this community will aim to ensure that CODECHECK is a viable option
+for independent no-cost Open Access journals.
 

### 2021-02-07 codecheck-paper.tex
edits to principles section

@@
 community\footnote{\href{https://github.com/codecheckers/codecheckers/}{https://github.com/codecheckers/codecheckers/}}
-currently has over 20 individuals who volunteered in the last 12
-years.  These volunteer's motivations, mentioned in the registration
+has over 20 volunteers who signed up in the last year.  Their
+motivations, mentioned in the registration
 information\footnote{\url{https://github.com/codecheckers/codecheckers/labels/registration}},
@@
 
-The example workflow and variations outlined above reflect our current
-views on how a workflow execution and code checking should be performed as part of peer review.
-They are not set in stone, but we do believe the following core principles underpin our CODECHECK process:
+The workflow and variations outlined describe our current
+views on how code could be checked.  They are not immutable, but
+we believe the following core principles underpin our CODECHECK
+process:
 
 \textbf{1. Codecheckers record but don't investigate or fix.} \\
-The codechecker follows the instructions provided by the author in running
-the code. If any steps are unclear, or if code does not run correctly,
-the codechecker records this fact and then communicates this
-information to the author. We believe that the job of the codechecker is
-not to fix all these problems but simply to report them and await a
-fix from the author.
-The level of documentation that is needed for third parties
-to reproduce a workflow is extremely hard to get right, and too often this 
-uncertainty leads researchers to give up and not document it at all.
-The conversation with a codechecker fixes this problem.
+The codechecker follows the author's instructions to run the code. If
+instructions are unclear, or if code does not run, the codechecker
+tells the author. We believe that the job of the codechecker is not to
+fix all these problems but simply to report them and await a fix from
+the author.  The level of documentation that is needed for third
+parties to reproduce a workflow is hard to get right, and
+too often this uncertainty leads researchers to give up and not
+document it at all.  The conversation with a codechecker fixes this
+problem.
 
@@
 \cite{cert-2020-013}) but more often there are hidden
-dependencies that need to be found or software installations that need to be adjusted for a particular system.
+dependencies or installations that need adjusting for a particular system.
 Allowing the (human) codechecker to communicate
@@
 constructive as possible. We feel that routing this conversation
-(possibly anonymously) through a publisher would slow things down and
-reduce the chances for community building as well as obscure some of the lessons
-learned for both authors and codecheckers.
+(possibly anonymously) through a publisher would introduce delays and
+reduce the chances for community building.
 
 \textbf{3. Credit is given to codecheckers.} \\
-The value of performing a
-CODECHECK is comparable to that of a scientific peer review, and
-it may also require a similar amount of time. Therefore,
-the codechecker's activity should be publicly recorded and mentioned in the
-published paper.
-The public record can be realised by publishing the certificate in a 
-citable form (i.e.,~with a DOI, as conducted in the community CODECHECK
-process), by listing codecheckers on the journal's website or, ideally, by
-publishing the checks alongside peer review activities in public databases.
+The value of performing a CODECHECK is comparable to that of a peer
+review, and it may also require a similar amount of time. Therefore,
+the codechecker's activity should be recorded, ideally in the
+published paper.  The public record can be realised by publishing the
+certificate in a citable form (i.e.,~with a DOI, as conducted in the
+community CODECHECK process), by listing codecheckers on the journal's
+website or, ideally, by publishing the checks alongside peer review
+activities in public databases.
 
 \textbf{4. Workflows must be auditable.} \\
-This requires that the codechecker 
-has at least enough material to validate the workflow outputs submitted by 
-the authors. Stark~\cite{stark_before_2018} calls this ``preproducibility''
-and the ICERM report \cite{stodden_setting_2013} defines the level
-``Auditable Research'' similarly.
-Every community may establish their own good practices or adapt generic
-concepts and practical tools, such as publishing all building blocks
-of science in a research compendium
-\footnote{\url{https://research-compendium.science/}}.
-A completed check means that code could be executed at least once without
-critical errors or
-warnings using the provided instructions, and, therefore, all code and data 
-was given and could be investigated more deeply or extended in the future.
-Ideally, this is a ``one~click'' step, but achieving this requires particular 
-skills and a sufficient level of documentation for 
-third parties. Furthermore, automation may lead to people gaming the system
+The codechecker should have sufficient material to validate the
+workflow outputs submitted by the
+authors. Stark~\cite{stark_before_2018} calls this
+``preproducibility'' and the ICERM report \cite{stodden_setting_2013}
+defines the level ``Auditable Research'' similarly.  Every community
+may establish their own good practices or adapt generic concepts and
+practical tools, such as publishing all building blocks of science in
+a research compendium
+\footnote{\url{https://research-compendium.science/}}.  A completed
+check means that code could be executed at least once using the
+provided instructions, and, therefore, all code and data was given and
+could be investigated more deeply or extended in the future.  Ideally,
+this is a ``one~click'' step, but achieving this requires particular
+skills and a sufficient level of documentation for third
+parties. Furthermore, automation may lead to people gaming the system
 or reliance on technology, which can often hide important details.
-All such aspects can reduce the understandability of the material, so we estimate our approach to codechecking, done without automation and with open human communication, to be the best way to ensure long-term transparency and usefulness.
-We acknowledge that others have argued in favour of bitwise reproducibility
-because, in the long run, it can be automated\footnote{Konrad Hinsen on 
-Twitter: \url{https://twitter.com/khinsen/status/1242842759733665799}},
-but until then we need CODECHECK's approach.
+All such aspects can reduce the understandability of the material, so
+we estimate our approach to codechecking, done without automation and
+with open human communication, to be a simple way to ensure long-term
+transparency and usefulness.  We acknowledge that others have argued
+in favour of bitwise reproducibility because, in the long run, it can
+be automated\footnote{Konrad Hinsen on Twitter:
+  \url{https://twitter.com/khinsen/status/1242842759733665799}}, but
+until then we need CODECHECK's approach.
 
 \textbf{5. Open by default and transitional by disposition.} \\
-By default, unless there are strong
-reasons to the contrary (e.g.,~sensitive data on human subjects), all code and
-data, both from author and codechecker, will be made freely available, 
-i.e.,~under open licenses, before or when the certificate is published.
-Note that openness is not required for the paper itself, as to accommodate 
-journals in their transition to sustainable Open Access models.
-The code and data publication should follow community good practices.
-By disposition, a CODECHECK workflow should be designed (a) with a recognition of
-shortcomings in asserting reproducibility of submissions and (b)
-with an understanding of the
-complex situation that led to this. However, establishing a CODECHECK workflow
-should go hand in hand with initiatives aiming to improve education
-around reproducibility of methods and communication surrounding comprehensive 
-computational workflows.
-The long term goal is for a specific community or
-journal's target audience to reach a state where the level of verification 
-ensured by codecheckers becomes part of peer review and a regular
-merit of all accepted papers.
+By default, unless there are strong reasons to the contrary
+(e.g.,~sensitive data on human subjects), all code and data, both from
+author and codechecker, will be made freely available before or when
+the certificate is published.  Openness is not required for the paper
+itself, to accommodate journals in their transition to sustainable
+Open Access models.  The code and data publication should follow
+community good practices.  Ultimately we may find that CODECHECKS are
+subsumed within peer review.
+%% By disposition, a CODECHECK workflow should
+%% be designed (a) recognising shortcomings in asserting reproducibility
+%% of submissions and (b) with an understanding of the complex situation
+%% that led to this. However, establishing a CODECHECK workflow should
+%% mirror initiatives to improve education around reproducibility of
+%% methods.  The long term goal is for a specific community to reach a
+%% state where the level of verification ensured by codecheckers becomes
+%% part of peer review and a regular merit of all accepted papers.
 
@@
 % LocalWords:  CODECHECK CODECHECKER Zenodo ORCID ECRs codecheck codechecker
+
+% Local Variables:
+% TeX-engine: xetex
+% End:

### 2021-02-08 codecheck-paper.tex
edited annotated cert

@@
 \url{https://codecheck.org.uk/register}.
-Table~\ref{tab:register} lists the certificates completed to date
+Table~\ref{tab:register} lists the certificates
 along with citations to the original papers.
@@
 
-The first theme was the initial set of papers used to explore the concept
-of CODECHECK. 
-The idea was to take very well known historical articles from a 
-specific domain, in this case the work area of SJE, possibly 
-as as a conversation starter for potential collaborating journals.
-This intention was brought to the first CODECHECK (certificate number
-\texttt{2020-001}) performed before publication on an article for the journal
-\emph{GigaScience}, which successfully reproduced the visualisations,
-though not the full machine learning process.
-
-The second theme is rather opportunistic. As the COVID-19 pandemic swept
-over the globe, we hoped to contribute a small part to science's answer
-to the unknown virus. The checks were solicited through community interaction
-%Several COVID-19 models have been done during the time of peer review of these papers,
-or by our initiative rather than requested from journals, but some certificates
-have since then been acknowledged in reviewed papers
-\cite{Davies2020-vj,kucharski_effectiveness_2020}. Unlike many claims in
-popular media at the time,
-% reference missing here TODO SJE, you had a note about some Noorden news piece
-we found the results of Report 9, a model of the UK lockdown procedures,
-to be reproducible (certificate number \texttt{2020-10}).
-
-The third theme represents DN's service as a Reproducibility Reviewer at the AGILE conference series, where the \emph{Reproducible AGILE} Initiative \cite{reproducible_agile} established a process for reproducing workflows at the AGILE conference series independent of CODECHECK \cite{nust_improving_2020}.
-While using slightly different terms and infrastructure (``reproducibility
-reports'' are published on the Open Science Framework (OSF) instead of certificates on Zenodo) the 
-AGILE reproducibility reviews adhere to the CODECHECK principles.
-Furthermore, a small number of checks were completed as part of peer reviews
+The first theme was the initial set of papers used to explore the
+concept of CODECHECK.  The idea was to take very well known historical
+articles from a domains of interest to one of the co-authors (SJE).
+Our first CODECHECK (certificate number \texttt{2020-001}) was
+performed before publication on an article for the journal
+\emph{GigaScience}, which visusalized the outputs from a family of
+supervised classification algorithms.
+
+The second theme was a response to the COVID-19 pandemic, choosing
+modelling papers to predict outcomes. The checks were solicited
+through community interaction or by our initiative rather than
+requested from journals.  Some certificates were since acknowledged in
+the accepted papers
+\cite{Davies2020-vj,kucharski_effectiveness_2020}. In particular, we
+codechecked the well-known Imperial college model of UK lockdown
+procedures from March 2020(certificate \texttt{2020-10}).  Contrary to
+opinion in the mainstream media, the model results were reproducible
+(CITE NOORDEN).
+
+The third theme represents co-author DN's service as a Reproducibility
+Reviewer at the AGILE conference series, where the \emph{Reproducible
+  AGILE} Initiative \cite{reproducible_agile} independently
+established a process for reproducing workflows at the AGILE
+conference series \cite{nust_improving_2020}.  While using slightly
+different terms and infrastructure (``reproducibility reports'' are
+published on the Open Science Framework (OSF) instead of certificates
+on Zenodo) AGILE reproducibility reviews adhere to CODECHECK
+principles.  A few checks were also completed as part of peer reviews
 for GIScience journals.
@@
 
-At the completion of each CODECHECK, the codechecker writes a certificate stating which outputs from the original article, i.e.,~numbers, figures or tables, could be reproduced.
-This certificate is made openly
-available so that all readers of the article can see which elements
-were reproduced and what limitations or issues were faced.
-The certificate also includes links to the code
-and data used by the codechecker, allowing others to build on the
-work.
-The format of the certificates has evolved slightly during the
-project, as we have learnt to automate different aspects of the
-certification, such as inserting figures into certificate documents
-and managing core metadata surrounding a check.
-The metadata is stored in a machine-readable structured file in YAML,
-the CODECHECK configuration file \texttt{codecheck.yml}\footnote{
-See technical specification of the CODECHECK configuration file at
-\url{https://codecheck.org.uk/spec/config/latest/}.}.
-The configuration file enables current and future automation of workflows and
-meta-analyses.
-
-To demonstrate the typical contents of a certificate and the underlying
-metadata, 
-Figure~\ref{fig:cert} shows the first four out of ten
-pages of the certificate \texttt{2020-012} \cite{cert-2020-012} with
-annotations.
-The referenced article \cite{unwin_report_2020} describes the modelling of COVID-19 across the USA.
-Figure~\ref{fig:cert}A shows the certificate number and its DOI,
-which points to the certificate and any supplemental files on Zenodo.
-The CODECHECK logo is
-added for recognition and to denote successful reproduction.
-Figure~\ref{fig:cert}B provides the key metadata as a table generated
-from the \texttt{codecheck.yml}; in particular it names the paper that
-was checked (title, DOI), the authors, the codechecker, when the check was performed,
-and in which repository related material is
-accessible.
-Figure~\ref{fig:cert}C shows a textual summary of how the CODECHECK was
-performed and any interesting findings.
-Figure~\ref{fig:cert}D (page 2 of the certificate) shows a table that
-lists the outputs that were generated from the check based on the 
-manifest of output files in the CODECHECK.
-The certificate's file table shows the file name, the description stating to
-which figure/table each file should be compared in the original paper, and
-the size of the file.
-Page 3 of the certificate,
-as Figure~\ref{fig:cert}E shows, provides detailed notes from the
-codechecker. In this particular example, the notes document what steps were 
-needed to initialize the environment and run the code and that the execution
-took quite long to run -- about 17~hours. Finally, on page 4 of
-the certificate, we see the first output that was generated by the
-CODECHECK (Figure~\ref{fig:cert}F). In this case, the figure matches
-figure~4 of \cite{unwin_report_2020}.
-Subsequent pages of the certificate show other outputs and the computing
-environment in which the certificate itself was created (not shown here).
+At the end of each CODECHECK, the codechecker writes a certificate
+stating which outputs from the original article, i.e.,~numbers,
+figures or tables, could be reproduced.  This certificate is made
+openly available so that all readers of the article can see which
+elements were reproduced and what limitations or issues were found.
+The certificate links to code and data used by the codechecker,
+allowing others to build on the work.  The format of the certificates
+evolved during the project, as we have learnt to automate different
+aspects of the certification, such as inserting figures into
+certificate documents and managing metadata of a check.  The metadata
+is stored in a machine-readable structured file in YAML, the CODECHECK
+configuration file \texttt{codecheck.yml}\footnote{ See technical
+  specification of the CODECHECK configuration file at
+  \url{https://codecheck.org.uk/spec/config/latest/}.}.  The
+configuration file enables current and future automation of workflows
+and meta-analyses.
+
+Figure~\ref{fig:cert} shows pages 1--4 (of 10) of a typical
+certificate \texttt{2020-012} \cite{cert-2020-012}.  The referenced
+article \cite{unwin_report_2020} described models of COVID-19 across
+the USA.  Figure~\ref{fig:cert}A shows the certificate number and its
+DOI, which points to the certificate and any supplemental files on
+Zenodo.  The CODECHECK logo is added for recognition and to denote
+successful reproduction.  Figure~\ref{fig:cert}B provides the key
+metadata as a table generated from \texttt{codecheck.yml}; it names
+the paper that was checked (title, DOI), the authors, the codechecker,
+when the check was performed, and in which repository related material
+is accessible.  Figure~\ref{fig:cert}C shows a textual summary of how
+the CODECHECK was performed and any interesting findings.
+Figure~\ref{fig:cert}D (page 2 of the certificate) shows the outputs
+that were generated based on the MANIFEST of output files in the
+CODECHECK.  It shows the file name (Output), the description stating
+to which figure/table each file should be compared in the original
+paper (Comment), and the file size.  Page 3 of the certificate,
+(Figure~\ref{fig:cert}E) gives detailed notes from the codechecker,
+here documenting what steps were needed to run the code and that the
+code took about 17~hours to complete. Finally, page 4 of the
+certificate shows the first output generated by the CODECHECK
+(Figure~\ref{fig:cert}F). In this case, the figure matched figure~4 of
+\cite{unwin_report_2020}.  The remaining pages of the certificate show
+other outputs and the computing environment in which the certificate
+itself was created (not shown here).
 

### 2021-02-08 codecheck-paper.tex
tools and resources

@@
 
-To run the system, we rely on freely available infrastructure, more 
-specifically GitHub and Zenodo.
-The \texttt{codecheckers} \textbf{GitHub} organisation\footnote{
-\url{https://github.com/codecheckers}} contains projects for managing the
-project website, the community of codecheckers, community discussions,
-forks of checked code repositories, and the main register of 
+We use freely available infrastructure, GitHub and Zenodo, to run our
+system.  The \texttt{codecheckers} \textbf{GitHub}
+organisation\footnote{ \url{https://github.com/codecheckers}} contains
+projects for managing the project website, the codecheckers community
+and its discussions, code repositories, and the main register of
 CODECHECKs. Both the website\footnote{
@@
   \url{https://codecheck.org.uk/register}} are hosted as GitHub
-pages. The register is a single table in CSV format that
-connects the certificate identifier with the repository associated with
-a specific conducted CODECHECK. Each of these repositories, which 
-currently can be hosted on GitHub or Open Science Framework (OSF), 
-contains the CODECHECK metadata file \texttt{codecheck.yml}. The register
-further contains a column for the type of check, e.g.,~community, journal,
-or conference, and the respective GitHub issue where communications and 
+pages. The register is a single table in CSV format that connects the
+certificate identifier with the repository associated with a specific
+conducted CODECHECK. Each of these repositories, which currently can
+be hosted on GitHub or Open Science Framework (OSF), contains the
+CODECHECK metadata file \texttt{codecheck.yml}. The register further
+contains a column for the type of check, e.g.,~community, journal, or
+conference, and the respective GitHub issue where communications and
 assignments around a specific check are organised. No information is
-duplicated between the register and the metadata files. When a new
-certificate is entered into the register, an automated process using the
-continuous integration infrastructure of GitHub, GitHub Actions, retrieves
-all metadata files and builds different representations of the register
-for viewing (HTML) and for integration with other services (CSV, JSON).
-\textbf{Zenodo} is an open repository for scientific data and documents. It
-mints DOIs for deposits and ensures long-term availability of all digital
-artefacts related to the project. The CODECHECK community on Zenodo
+duplicated between the register and the metadata files. The continuous
+integration infrastructure of GitHub, GitHub Actions, is used to
+automate the register for viewing (HTML) and for integration with
+other services (CSV, JSON).  \textbf{Zenodo} is our preferred open repository for
+storing certificates. It mints DOIs for deposits and ensures
+long-term availability of all digital artefacts related to the
+project. The CODECHECK community on Zenodo
 \footnote{\url{https://zenodo.org/communities/codecheck/}} includes
@@
 A \textbf{custom R package}, \texttt{codecheck}\footnote{
-\url{https://github.com/codecheckers/codecheck}}, is used to automate repetitive
-tasks around the authoring of certificates and the management of the register.
-For example, it includes scripts for the deposition of certificates and related 
-files to Zenodo using the R package \texttt{zen4R} \cite{zen4r} and for the
-register update process outlined above.
-Codecheckers may choose to not use the package and rely on their own tools
-for creating the certificate, even a regular word processor.
-This flexibility is deliberate to accomodate different skill sets and unforseen technical advances or challenges.
+  \url{https://github.com/codecheckers/codecheck}}, automates
+repetitive tasks around authoring certificates and managing the
+register.  It includes scripts to deposit certificates and related
+files to Zenodo using the R package \texttt{zen4R} \cite{zen4r} and
+for the register update process outlined above.  Codecheckers can
+ignore this package, and use their own tools for creating
+and depositing the certificate.  This flexibility  accommodates
+different skill sets and unforseen technical advances or challenges.
 
@@
 managed without infrastructure costs on freely available platforms.
-The automation should ensure that more checks can
-be conducted in the future, though having the process be managed by a paid CODECHECK
-editor or further automation, e.g.,~with a bot, could improve turnaround time.
-The only costs are the project's domain, which is negligible.
-This means that the initiative's main resource are the \textbf{humans} needed
-for managing the project and processes and the codecheckers.
-All contributions currently rely on (partly grant-based) public funding
-and volunteering.
-Naturally, similar set-ups are conceivable, such as using GitLab.com or
-a hosted GitLab instance and the GitLab continuous integration.
+The automation should ensure that more checks can be conducted in the
+future, though having the process be managed by a paid CODECHECK
+editor or further automation, e.g.,~with a bot, to reduce turnaround
+time. Our main resource requirements are the \textbf{humans} needed for managing
+the project and processes and the codecheckers.  All contributions
+currently rely on (partly grant-based) public funding and
+volunteering.
 

### 2021-02-08 codecheck-paper.tex
related work section
(I think this will need further tweaking)

@@
 \footnote{\url{https://zenodo.org/communities/codecheck/}} includes
-certificates as well as the regularly archived register
+certificates and the regularly archived register
 \cite{codecheck_register_jan2021}.
@@
 The journal \emph{ACM Transactions on Mathematical Software (TOMS)} 
-established a ``Replicated Computational Results'' (RCR) review process
-\cite{heroux_editorial_2015}, where ``replicable'' means the same thing as what we call ``reproducible''.
-A search on Crossref\footnote{\url{https://search.crossref.org/?q=\%22Replicated+Computations+Results+\%28RCR\%29+Report\%22&from_ui=yes&page=2}; search executed
-on 2020-12-10.} shows that 15 RCR Reports have been published and the 
-% Number of found TOMAS RCR reports:
-% 8: https://scholar.google.de/scholar?start=10&q=Replicated+Computations+Results+%22(RCR)+Report%22&hl=de&as_sdt=0,5
-% 15: https://search.crossref.org/?q=Replicated+Computations+Results+%22%28RCR%29+Report%22&from_ui=yes&publication=ACM+Transactions+on+Mathematical+Software
+recently established a ``Replicated Computational Results'' (RCR) review process
+\cite{heroux_editorial_2015}, where ``replicable'' is the same as our
+use of ``reproducible''.
+
+Fifteen RCR Reports have been published so far
+\footnote{\url{https://search.crossref.org/?q=\%22Replicated+Computations+Results+\%28RCR\%29+Report\%22&from_ui=yes&page=2};
+  search executed on 2020-12-10.}
 % 15: https://academic.microsoft.com/search?q=Replicated%20Computations%20Results%20%22(RCR)%20Report%22&f=&orderBy=0&skip=10&take=10
-workflow is being
-extended to the ACM journal \emph{Transactions on Modeling and Computer 
-Simulation (TOMACS)}\footnote{\url{https://dl.acm.org/journal/tomacs/author-guidelines}}.
-The TOMS RCR follows the same principles as those used by CODECHECK (principles 1-4), which were independently developed as we only later learned about TOMS RCR.
-Here, we provide an extensive discussion of the possible variants of CODECHECK,
-while TOMS RCR realises one particular workflow for reproducibility 
-reviews for one particular journal. The TOMS editorial
-\cite{heroux_editorial_2015} shares similar concerns about
-selection of reviewers, as we discussed above. Unlike existing CODECHECK certificates, 
-the RCR reports, which are very similar to CODECHECK certificates, undergo editorial review, but this could eventually become
-part of the CODECHECK processes at journals.
-The publication of the RCR report is seen as a reward for the efforts of
-the reproducing person, while the potential for this motive to be a conflict of interest is acknowledged.
-TOMS also recognises reviewer activity in a 
-partnership with Publons\footnote{
-\url{https://authors.acm.org/author-services/publons}},
-though it is unclear whether the author of an RCR report receives both the report's
-authorship and the reviewer recognition. In any case, this is different
-from CODECHECK, which specifically aims to give codecheckers recognition as reviewers.
-In our view, the reviewer role removes the possible conflict of interest while
-keeping the public acknowledgement.
-Specific to the field of mathematics, the RCR is also expected to apply a
-review of the software itself if the system it runs on cannot be evaluated by
-an independent party.
-The TOMS RCR creators concur with the importance of communication, expect
-collaboration between author and RCR reviewers,
-share the considerations around reviewer selection, and also put trust
-in reviewer judgement over numerical bit-wise perfection.
-A key difference is that for TOMS RCR, authors opt-in with an \emph{RCR Review Request} and
-the RCR reports are published in the TOMS journal next to the actual papers.
-
-A number of journals provide special article types for reproductions of papers
-published in the journal.
-\emph{Information Systems} has an invited 
-reproducibility paper track\footnote{
+and the process is being extended extended to the ACM journal
+\emph{Transactions on Modeling and Computer Simulation
+  (TOMACS)}\footnote{\url{https://dl.acm.org/journal/tomacs/author-guidelines}}.
+The TOMS RCR follows CODECHECK principles 1-4, although our work was
+independently developed of theirs.  Here, we discuss variants of
+CODECHECK, while TOMS RCR realises one particular workflow for
+reproducibility reviews for one particular journal. The TOMS editorial
+\cite{heroux_editorial_2015} shares similar concerns about selection
+of reviewers, as we discussed above. Unlike existing CODECHECK
+certificates, the RCR reports, which are similar to CODECHECK
+certificates, undergo editorial review, but this could eventually
+become part of the CODECHECK processes at journals.  The publication
+of the RCR report is seen as a reward for the efforts of the
+reproducing person, while the potential for this motive to be a
+conflict of interest is acknowledged.  TOMS also recognises reviewer
+activity in a partnership with Publons\footnote{
+  \url{https://authors.acm.org/author-services/publons}}.
+%, but it is
+unclear if the author of an RCR report receives both the report's
+%authorship and the reviewer recognition. 
+CODECHECK however specifically aims to give codecheckers
+recognition as reviewers.  In our view, the reviewer role removes the
+possible conflict of interest while keeping the public
+acknowledgement.  Specific to the field of mathematics, the RCR is
+also expected to apply a review of the software itself if the system
+it runs on cannot be evaluated by an independent party.  The TOMS RCR
+creators concur with the importance of communication, expect
+collaboration between author and RCR reviewers, share the
+considerations around reviewer selection, and also put trust in
+reviewer judgement over numerical bit-wise perfection.  A key
+difference is that for TOMS RCR, authors opt-in with an \emph{RCR
+  Review Request} and the RCR reports are published in the TOMS
+journal next to the actual papers.
+%% SJE: think this could be tightened a bit, and some of it I don't
+%% fully understand.
+
+Several journals provide special article types for reproductions of
+published papers.  \emph{Information Systems} has an
+invited reproducibility paper track\footnote{
   \url{https://www.elsevier.com/journals/information-systems/0306-4379/guide-for-authors}.}.
-\emph{Nature Machine Intelligence} recently introduced a new type of article,
-the reusability report \cite{noauthor_research_2020}.
+\emph{Nature Machine Intelligence} recently introduced a new type of
+article, the reusability report \cite{noauthor_research_2020}.
 Inspired by the detailed and nuanced submissions to a reproducibility
-challenge, the reusability report focuses on the exploration of robustness
-and generalizability of the original paper's claims
+challenge, the reusability report focuses on the exploration of
+robustness and generalizability of the original paper's claims
 \cite{noauthor_research_2020}. This answers the specific community's
-challenges around computational reproducibility and also values these kinds
-of contributions as independent publications, which goes beyond the goals 
-of CODECHECK.
-The journal \emph{Cortex} has a special article type 
-\emph{Verification Reports}, which are actually about replication of results
-but are very well designed/reasoned \cite{chambers_verification_2020}.
-In a similar vein, the CODECHECK certificates
-could also be published as a special article type within journals.
-
-Going beyond article types and a specific journal, the journal \emph{ReScience C} publishes only
-replications\footnote{\url{https://rescience.github.io/faq/}}, also requiring
-open code and replication by a third party. ReScience also relies on
-free infrastructure (GitHub and Zenodo).
+challenges around computational reproducibility and also values these
+kinds of contributions as independent publications, which goes beyond
+the goals of CODECHECK.  The journal \emph{Cortex} has a special
+article type \emph{Verification Reports}, which are actually about
+replication of results and are very well designed/reasoned
+\cite{chambers_verification_2020}.  In a similar vein, the CODECHECK
+certificates could also be published as a special article type within
+journals.
+
+Going beyond article types and a specific journal, the journal
+\emph{ReScience C} publishes only
+replications\footnote{\url{https://rescience.github.io/faq/}}, also
+requiring open code and replication by a third party. ReScience also
+relies on free infrastructure (GitHub and Zenodo).
 
@@
 
-Gavis and Donoho \cite{gavish_universal_2011} propose a new discipline and 
-infrastructure for reproducible computational research with its own tools and
-API. Their specific packaging format, provenance record, and cryptographic
-\emph{Verifiable Result Identifier} would indeed provide excellent
-reproduciblity. However, the system is also complex and has not been adopted by
-any publisher as far as we know; also, the system is indeed not open source itself.
-In comparison, CODECHECK is less powerful but also much more flexible and
-less dependent on specific tools or infrastructure. If data and code are 
-deposited properly, i.e.,~very unlikely to disappear, then the certificate's 
-DOI is practically close to the cryptographic identifier.
-
-Another platform for publishing results of reproductions is \emph{SciGen.Report}\footnote{\url{https://scigen.report/}}.
-It is a community-run independent platform to foster communication on reproducibility.
-People can report on fully, partiially, or un-successful reproductions of articles after publication.
+Gavis and Donoho \cite{gavish_universal_2011} propose a new discipline
+and infrastructure for reproducible computational research with its
+own tools and API. Their specific packaging format, provenance record,
+and cryptographic \emph{Verifiable Result Identifier} would indeed
+provide excellent reproduciblity. However, the system is also complex
+and since its creation in 2011 we are not aware of any publisher using
+it; also, the system is not open source itself.  In comparison,
+CODECHECK is less powerful but also much more flexible and less
+dependent on specific tools or infrastructure. If data and code are
+deposited properly, i.e.,~very unlikely to disappear, then the
+certificate's DOI is practically close to the cryptographic
+identifier.
+
+Another platform for publishing results of reproductions is
+\emph{SciGen.Report}\footnote{\url{https://scigen.report/}}.  It is a
+community-run independent platform to foster communication on
+reproducibility.  People can report on fully, partiially, or
+un-successful reproductions of articles after publication.
 
 No approach besides CODECHECK spans across journals and is
-intentionally designed to be picked up and implemented by many different 
-journals or events and to build a community of codecheckers. CODECHECK shares
-its interdisciplinary nature with other community initiatives concerned with 
-reproducibility awareness, education, and support, such as
+intentionally designed to be adopted across journals or events and to
+build a community of codecheckers. CODECHECK shares its
+interdisciplinary nature with other community initiatives concerned
+with reproducibility awareness, education, and support, such as
 ReproHack\footnote{\url{https://reprohack.github.io/reprohack-hq/}},
-Code Copilot\footnote{\url{https://twitter.com/Code_Copilot}}, or Papers with Code\footnote{\url{https://paperswithcode.com/about}}.
+Code Copilot\footnote{\url{https://twitter.com/Code_Copilot}}, or
+Papers with Code\footnote{\url{https://paperswithcode.com/about}}.
 The latter recently announced a collaboration with the preprint server
-\emph{arXiv} on providing data and code supplements for machine learning
-manuscripts\footnote{\url{https://medium.com/paperswithcode/papers-with-code-partners-with-arxiv-ecc362883167}} and runs a reproduciblity challenge\footnote{\url{https://paperswithcode.com/rc2020}}.
+\emph{arXiv} on providing data and code supplements for machine
+learning
+manuscripts\footnote{\url{https://medium.com/paperswithcode/papers-with-code-partners-with-arxiv-ecc362883167}}
+and runs a reproduciblity
+challenge\footnote{\url{https://paperswithcode.com/rc2020}}.
 % Papers with Code now also does more disciplines than just ML.
 Likewise, different disciplines and journals provide reproducibility
-checklists, e.g.,~sciene and engineering \cite{rosenberg_next_2020} or GIScience \cite{nust_agile_2019}, which naturally share some
-aspects while addressing particularities as well as addressing researchers
-from different fields. Regarding the education and guidance for authors, 
-we see CODECHECK's role as referencing and linking educational efforts
-and helpful material, not as creating and maintaining such content.
+checklists, e.g.,~sciene and engineering \cite{rosenberg_next_2020} or
+GIScience \cite{nust_agile_2019}, which naturally share some aspects
+while addressing particularities as well as addressing researchers
+from different fields. Regarding the education and guidance for
+authors, we see CODECHECK's role as referencing and linking
+educational efforts and helpful material, not as creating and
+maintaining such content.
 

### 2021-02-08 codecheck-paper.tex
done

@@
 
-No approach besides CODECHECK spans across journals and is
-intentionally designed to be adopted across journals or events and to
-build a community of codecheckers. CODECHECK shares its
+CODECHECK is uniquely designed to be adopted across journals or events
+and to build a community of codecheckers. CODECHECK shares its
 interdisciplinary nature with other community initiatives concerned
@@
 \textbf{Isn't CODECHECK what peer review should be doing already?} \\
-On the surface, yes.
-However, the general observation is that the
-increasing number of researchers, submissions, and reviews to conduct
-under the pressure to publish puts the scholarly publication system
-under significant strain.
-Miyakawa described how when he as an editor simply requested raw data from authors, this led to almost all of those papers in which data was requested to be withdrawn or rejected \cite{miyakawa_no_2020}.
-Such basic checks to show that
-sufficient data supports the results requires much effort from editors,
-yet they only rely on the availability of data files and the content of the
-paper.
-These availability checks can be enhanced by having more complex CODECHECKs request the code and then execute it, because code is not inert but needs to be run in a specific environment.
-This might fall within idealistic expectations of peer review,
-but is far from common in reality.
-Establishing a CODECHECK
-process is largely an acknowledgement that peer reviewing practices
-have been unable to adapt to the challenges of digitisation of data-based science.
-The concept of a CODECHECK,
-just as the concepts of reproducible research and Open Science,
-may be transitional by nature. If the activities described here as
-being part of a CODECHECK become the norm for scientific reviews and
+On the surface, yes, but peer reviewers are overburdened enough and
+asking them to do more work usring peer review is not likely to
+suceed.  Miyakawa described how when he as an editor simply requested
+raw data from authors during submission, this led to almost all of
+those papers in which data was requested to be withdrawn or rejected
+after peer review \cite{miyakawa_no_2020}.  Such basic checks to show
+that sufficient data supports the results requires effort from
+editors, yet they only rely on the availability of data files and the
+content of the paper.  These availability checks can be enhanced by
+having more complex CODECHECKs request the code and then execute it.
+This might fall within idealistic expectations of peer review, but is
+rare.  Establishing a CODECHECK process is largely an acknowledgement
+that peer reviewing practices have been unable to adapt to the
+challenges of computational papers.  The concept of a CODECHECK, just
+as the concepts of reproducible research and Open Science, may be
+transitional by nature. If the activities described here as being part
+of a CODECHECK become the norm for scientific reviews and
 publications, then the initiative will have succeeded. However, the
@@
 
-\textbf{Should CODECHECK requirements not be more demanding?} \\
-CODECHECK by design does not require authors to provide (and sustain) an
-eternally functional workflow and neither suggests a specific software stack
-or practical approach even if such suggestions, e.g.,~research compendia or ``repro-packs''\footnote{See \url{https://research-compendium.science/} respectively \cite{barba_praxis_2018}.}, are very valuable.
-Creating something that anyone can reproduce has been called a 
-fool's errand\footnote{\url{https://twitter.com/DougBlank/status/1135904909663068165?s=09}}
-and we tend to agree.
-However, the package of data, code, and documentation
-collaboratively created by authors and codecheckers is a snapshot of a 
-working analysis that greatly increases the likelihood of a successful 
-reproduction and the possibility that a workflow can be extended by third parties in the future, as long as they have access
-to suitable resources and matching skill set.
-Concrete implementations of CODECHECK workflows, especially for specific
-disciplines, may reify much more helpful guidelines for authors on how to
-create reproducibility packages.
-The author-friendly ``low bar'' should not stay low forever, but cultural
-change takes time and the encouragement and guidance that CODECHECK,
-as part of the widely accepted peer review concept, can provide may
-eventually allow the bar to be raised as high as possible, e.g.,~with
-executable research compendia \cite{nust_opening_2017}, ``Whole Tales'' \cite{brinckman_computing_2018},
-or continuous analysis \cite{beaulieu-jones_reproducibility_2017-1}.
-However, considering that missing artefacts and lack of documentation
-have repeatedly been identified as key barriers to reproducibility
-(e.g.,~\cite{stagge_assessing_2019,nust_improving_2020}),
-we would not underestimate the power of a simple check.
-For example, ModelDB curation policies ensure that only one figure is manually reproduced \cite{mcdougal_reproducibility_2016}, but that has not limited the usefulness nor success of the platform.
-
-It should be noted that a codechecker does not fulfil the same role as a
-\emph{statistical reviewer}, as it is applied by some journals in the 
-biomedical domain (cf.~\cite{petrovecki_role_2009,greenwood_how_2015}).
-The statistical reviewer actually evaluates the appropriateness of
-statistical methods \cite{greenwood_how_2015} and can support topical
-reviewers if, e.g.,~complex methods or sophisticated variants of statistical
-tests are applied \cite{petrovecki_role_2009}.
-The codechecker may go equally deep into the review, if they have
-the required expertise and the required time, 
-but they are by no means expected to be able to provide
-such a level of scrutiny. We can imagine a tiered workflow where a 
-codechecker could, just as a conventional reviewer could, recommend a detailed
-statistical review to the editor if they come upon certain issues while
-examining the work from a different perspective.
-
-An important distinction is that a codechecker does not conduct a 
-\emph{code review}. Code reviews are without a doubt extremely valuable in 
-terms of improving reproducibility and reusability, and their proponents 
-even believe they can improve the science \cite{petre_code_2014}.
-Code reviews, however, have quite different structural challenges and 
-require even more resources. That said, a well-reviewed codebase is likely
-to be easier to codecheck, and the awareness of high-quality code raised
-through CODECHECK may lead to more support for code reviewing.
-Initiatives and journals that conduct software reviews independent of 
-a specific publication or workflow include
-ROpenSci\footnote{\url{https://ropensci.org/}},
-PyOpenSci\footnote{\url{https://www.pyopensci.org/}},
-and JOSS\footnote{\url{https://joss.theoj.org/}}.
-Furthermore, the codechecker's task list is intentionally not
-overloaded with related concerns, albeit potentially important ones,
-such as ensuring proper citation of data and software or depositing
-data and software in suitable repositories. Nevertheless,
-we expect codecheckers to often recognise these issues and highlight them.
+\textbf{Should CODECHECK requirements be more demanding?} \\
+CODECHECK by design does not require authors to provide (and sustain)
+an eternally functional workflow nor suggests a specific software
+stack or practical approach even if such suggestions, e.g.,~research
+compendia or ``repro-packs''\footnote{See
+  \url{https://research-compendium.science/} respectively
+  \cite{barba_praxis_2018}.}, are valuable.  Creating something
+that anyone can reproduce has been called a fool's
+errand\footnote{\url{https://twitter.com/DougBlank/status/1135904909663068165?s=09}}
+and we tend to agree.  However, the package of data, code, and
+documentation collaboratively created by authors and codecheckers is a
+snapshot of a working analysis that greatly increases the likelihood
+of a successful reproduction and the possibility that a workflow can
+be extended by third parties in the future, as long as they have
+access to suitable resources and matching skill set.  Concrete
+implementations of CODECHECK workflows, especially for specific
+disciplines, may reify much more helpful guidelines for authors on how
+to create reproducibility packages.  Our author-friendly ``low bar''
+should not stay low forever, but cultural change takes time and the
+encouragement and guidance that CODECHECK, as part of the widely
+accepted peer review concept, can provide may eventually allow the bar
+to be raised much higher, e.g.,~with executable research
+compendia \cite{nust_opening_2017}, ``Whole Tales''
+\cite{brinckman_computing_2018}, or continuous analysis
+\cite{beaulieu-jones_reproducibility_2017-1}.  However, considering
+that missing artefacts and lack of documentation have repeatedly been
+identified as key barriers to reproducibility
+(e.g.,~\cite{stagge_assessing_2019,nust_improving_2020}), we would not
+underestimate the power of a simple check.  For example, ModelDB
+curation policies require that only one figure need be manually reproduced
+\cite{mcdougal_reproducibility_2016}, but that has not limited the
+usefulness nor success of the platform.
+
+A codechecker does not fulfil the same role as a \emph{statistical
+  reviewer}, as it is applied by some journals in the biomedical
+domain (cf.~\cite{petrovecki_role_2009,greenwood_how_2015}).  The
+statistical reviewer evaluates the appropriateness of statistical
+methods \cite{greenwood_how_2015} and can support topical reviewers
+if, e.g.,~complex methods or sophisticated variants of statistical
+tests are applied \cite{petrovecki_role_2009}.  The codechecker may go
+equally deep into the review, but only if they have the expertise and
+time. We can imagine a tiered workflow where a codechecker could, just
+as a conventional reviewer could, recommend a detailed code review
+(see next paragraph) to the editor if they come upon certain issues while
+examining the work.
+
+A codechecker does not conduct a \emph{code review}. Code reviews are
+valuable to improve reproducibility and reusability, and their
+proponents even believe they can improve the science
+\cite{petre_code_2014}.  Code reviews, however, have quite different
+structural challenges and require even more resources. That said, a
+well-reviewed codebase is likely to be easier to codecheck, and the
+awareness of high-quality code raised through CODECHECK may lead to
+more support for code reviewing.  Initiatives and journals that
+conduct software reviews independent of a specific publication or
+workflow include ROpenSci\footnote{\url{https://ropensci.org/}},
+PyOpenSci\footnote{\url{https://www.pyopensci.org/}}, and
+JOSS\footnote{\url{https://joss.theoj.org/}}.  Furthermore, the
+codechecker's task list is intentionally not overloaded with related
+concerns, albeit potentially important ones, such as ensuring proper
+citation of data and software or depositing data and software in
+suitable repositories. Nevertheless, we expect codecheckers to often
+recognise these issues and highlight them.
 
 \textbf{How are failures during checks handled?}
-We have no established process for the case
-in which a reproduction completey fails.
-This may be a sign that our pre-selection of the examples was biased, since so far all our CODECHECKs succeeded.
-In the case that a journal adopts CODECHECK for all submissions, the open 
-question remains as what to do if a check fails.
-A negative comment in a CODECHECK certificate or a failed check does
-not necessarily mean the paper or the science is bad
-(cf. discussion on negative comments in \cite{everythinghertz123}).
-At the start, we doubt that publicly reporting failures (i.e.,~the code 
-would not run) will increase overall reproducibility, and we recommend not
-sharing that outcome publicly but sharing the latest draft of the 
-certificate with the authors.
-One must be careful how policies on disclosing failures may negatively impact the acceptance by authors, but also how efforts by volunteers may go unnoticed if no report is published.
-The most appropriate way forwards seems to be to give private feedback for failures and public acknowledgement and encouragement for successes.
+
+We have not yet established a process for if a reproduction fails.  This
+may simply reflect our biases in pre-selection of case studies.  In
+the case that a journal adopts CODECHECK for all submissions, the open
+question remains as what to do if a check fails, after exhausting
+efforts between author and codechecker to reproduce the workflow.  A
+negative comment in a CODECHECK certificate or a failed check does not
+necessarily mean the paper or the science is bad (cf. discussion on
+negative comments in \cite{everythinghertz123}).  We doubt that
+publicly reporting failures (i.e.,~the code would not run) will
+increase overall reproducibility, and we recommend sharing the latest
+draft of the certificate only with the authors.  One must be careful
+how policies on disclosing failures may \textbf{negatively impact the
+  acceptance by authors}
+%% SJE: what does this mean?
+, and also how volunteer efforts may go unnoticed if no report is
+published.  The most appropriate way forwards seems to be to give
+private feedback for failures and public acknowledgement and
+encouragement for successes.
 % eLife's new model (see https://osf.io/9cftx/) will publish reports evantually, e.g.,~after preprint has been accepted elsewhere
@@
 
-\textbf{Who will pay for the compute time?}
-For those papers that take significant compute
-time (days, not minutes), it is unclear who will pay for it.
-The openness of CODECHECK certificates could help to establish community
-peer pressure, or recognition, as submitting authors can be evaluated on their engagement
-as codecheckers. This may lead to a balance of the shared load to reproduce
-computations.
-However, one must carfully consider the sustainability of rerunning computations and the
-environmental impact large calculations, such as training machine learning models, have.
-The challenges of computational resources can also quickly be reduced when
-authors provide a small sample or a synthetic dataset together with the
-expected outputs for this small example, which can equally well demonstrate the 
-usefulness and validity of their workflow.
-
-\textbf{What about my proprietary software and sensitive data?}
-Authors can currently provide code that requires
-proprietary software. Given the prevalence of proprietary software in
-some disciplines, e.g MATLAB, this seems like a pragmatic choice.
-However, this prohibits publications from benefiting from open infrastructure for reproducibility
-(cf.~\cite{konkol_publishing_2020,perkel_make_2019})
-and requires the codechecker to have access to that particular software.
-Non-open software also considerably hampers reuse, especially by researchers
-from the global south. Therefore, allowing proprietary software is a compromise
-that each implementation of CODECHECK should critically reconsider.
-
-Solutions for proprietary and sensitive data exists.
-Authors can provide synthetic data (cf. \cite{shannon_opening_2018}), some
-data can effectively be redacted \cite{oloughlin_data_2015}, and publishers or
-independent entities can provide infrastructure for sharing data and workflows
-confidentially \cite{perignon_certify_2019} or with access to derived results
-but not raw data \cite{shannon_opening_2018},
-i.e.,~data enclaves \cite{foster_research_2018},
-or domains of reproducibility \cite{harris_more_2017})
-
-\textbf{Can't someone cheat?} Yes. We don't check that the code is correct or
-sound science,
-just simply that it runs. This ``mechanical'' test is indeed a very low bar.
-Yet, by having the code and (raw) data openly deposited, interested parties can
-then examine the code, and we hope that knowing the code will be open will
-be an incentive for authors to ensure good quality.
-It will also give opportunities
-to future researchers, potentially with new methods, to look for errors.
-This prospective evaluation is a much more effective way to capture cheaters than engaging in an arms race on tools, methods, or resources trying to catch malicious intent now with closed datasets and code; this approach is comparable to storing blood samples of sport champions today to possibly detect doping in the future with more advanced methods (cf.
-\cite{everythinghertz97}).
-Another comparison that helped us define the scope of a CODECHECK is was that we think of the codechecker as forensic photographer, making sure all the details
-are captured so that an investigator may, now or in the future, scrutinize
-them.
-
-\textbf{Who's got time for more peer review?} Agree.
-The time one spends performing CODECHECKs will likely reduce the time an individual has to work as a reviewer.
-However, a CODECHECK is different from peer review.
-First, the \textbf{technical nature} of a CODECHECK can provide a more 
-clearly defined scope and expectations and thereby time budget compared to conventional 
-peer review. 
-For example, authors are told what to provide and the codechecker can be told when to stop.
-The option for the codechecker to ask for clarification or more 
-material can greatly reduce the digging required and thereby
-increase \textbf{efficiency}.
-Second, the specific skill set of a codechecker allows for \textbf{different groups} to take part in the review process.
-ECRs might be attracted to learn more about recent methods, peer review, and reproducibility practices.
-Research software experts (Research Software Engineers, RSE\footnote{\url{https://en.wikipedia.org/wiki/Research_software_engineering}}) who might
-not regularly be involved in writing or reviewing papers might be interested in increasing their connection with scholarly practices.
-An extra codechecker may simplify the matchmaking an editor has to do
-when identifying suitable reviewers for a submission, as technical and
-topical expertise can be provided by different people who do not even have to be traditional peer reviewers.
-In fact, the codechecker could even be a software professional employed by a publisher.
-Third, recall that we expect CODECHECKs not to be anonymous and always
-publicly deposited, which is common practice for reviews of few journals only.
-With code and workflows, there also is a good chance that a codechecker's
-feedback can directly impact and improve the author's 
-work, which may make for a rewarding experience or even provide tangible contributions.
-The public certificates and contributions provide \textbf{peer recognition}.
-Fourth, we also found that the focus on the mechanics of a
-workflow and the ability to interact with the authors can make
-reproductions very \textbf{educational}. It also is a different role and, as such, could
-be a welcome option for researchers to give back their time to the community.
-
-While such benefits are also part of idealistic peer review, they are mostly hidden behind paraphrased anonymous acknowledgement.
-
-\textbf{Do workflows need to be codechecked multiple times?}
-If you check a preprint before peer
-review or a submission to a journal with non-public review, it might
-need checking again as it is modified during peer review.
-This is somewhat unavoidable, and has happened to us (e.g.,~certificate \texttt{2020-012}
-\cite{cert-2020-012} has been revised). This can also be desired tough, if
-the interaction between author, reviewer, and codechecker led to
-improvements. Since checking the manuscript the second time is likely
-to be much less work than the first time, this is largely mitigated by
-open and regular communication. The higher the reproducibility and
-automation, the more aspects of a CODECHECK that may eventually be
-supported by automation, though we believe the human codecheckers
-should remain the decision makers. Automatically detecting whether results
-are ``the same'' is a question intrinsically linked with each unique
-process and quite hard to realise for graphical outputs, such as plots or
-maps.
-That is why we do not require results to be exactly the same for a
-CODECHECK to pass but simply that the code runs and generates the output
-files that the author claims. Stochastic simulations mean that often
-we cannot get exactly the same results, and that even the same versions of libraries
+\textbf{Who will pay for the compute time?}  For papers that take
+significant compute time (days, not minutes), it is unclear who will
+pay for it. One must carefully consider the sustainability of
+rerunning computations and the environmental impact large
+calculations, such as training machine learning models, have.  A
+pragmatic workaround is to request that authors provide a ``toy''
+example, or small dataset that can be quickly analysed to demonstrate
+that the workflow runs correctly.
+
+\textbf{What about my proprietary software and sensitive data?}  Given
+the prevalence of proprietary software, e.g MATLAB, we pragmatically
+decided that we should accept code as long as we could find a machine
+with suitable licences to run it.  However, this prohibits us from
+using open infrastructure for reproducibility
+(cf.~\cite{konkol_publishing_2020,perkel_make_2019}) and requires the
+codechecker to have access to that particular software.  Non-open
+software also considerably hampers reuse, especially by researchers
+from the global south. Therefore, allowing proprietary software is a
+compromise that should be reconsidered.
+
+Solutions for proprietary and sensitive data exists.  Authors can
+provide synthetic data (cf. \cite{shannon_opening_2018}), some data
+can effectively be redacted \cite{oloughlin_data_2015}, and publishers
+or independent entities can provide infrastructure for sharing data
+and workflows confidentially \cite{perignon_certify_2019} or with
+access to derived results but not raw data
+\cite{shannon_opening_2018}, i.e.,~data enclaves
+\cite{foster_research_2018}, or domains of reproducibility
+\cite{harris_more_2017}).
+
+\textbf{Can't someone cheat?} Yes. We simply check that the code runs,
+not that is correct or sound science. This ``mechanical'' test is
+indeed a low bar.  By having the code and data openly deposited, third
+parties can later examine the code, and we hope that knowing the code
+will be open ensures that authors will not cheat.  It also allows
+researchers, potentially with new methods, to look for errors.  This
+is more effective than engaging in an arms race on building methods to
+detect malicious intent now with closed datasets and code.  This is
+analogous to storing blood samples of sport champions today to
+possibly detect doping in the future with more sensitive methods
+(cf. \cite{everythinghertz97}).  Another comparison that helped us
+define the scope of a CODECHECK is that we think of the codechecker as
+forensic photographer, capturing details so that an investigator may
+later scrutinize them.
+
+\textbf{Who's got time for more peer review?} Agree; codechecking
+takes time that could otherwise be used for traditional peer review.
+However, a CODECHECK is different from peer review.  First, the
+\textbf{technical nature} of a CODECHECK sets clear expectations and
+thereby time budget compared to conventional peer review.  For
+example, authors are told what to provide and the codechecker can be
+told when to stop.  Codecheckers can always directly ask the author
+when clarification is required, thereby increasing
+\textbf{efficiency}.  Second, the specific skill set of a codechecker
+allows for \textbf{different groups} to participate in the review
+process.  ECRs might be attracted to learn more about recent methods,
+peer review, and reproducibility practices.  Research software experts
+(Research Software Engineers,
+RSE\footnote{\url{https://en.wikipedia.org/wiki/Research_software_engineering}})
+who might not regularly be involved in writing or reviewing papers
+might be interested in increasing their connection with scholarly
+practices.  An extra codechecker may simplify the matchmaking an
+editor does when identifying suitable reviewers for a submission,
+as technical and topical expertise can be provided by different
+people.  Third, recall that CODECHECKs should always be publicly
+available, unlike peer review reports.  With code and workflows, the
+codechecker's feedback may directly impact and improve the author's
+work.  The public certificates and contributions provide \textbf{peer
+  recognition} for the codechecker.  Fourth, we found that focusing on
+workflow mechanics and interacting with the author makes reproductions
+\textbf{educational}. It also is a different role and, as such, could
+be a welcome option for researchers to give back their time to the
+community.
+
+While such benefits are also part of idealistic peer review, they are
+mostly hidden behind paraphrased anonymous acknowledgement.
+
+\textbf{Do workflows need to be codechecked multiple times?}  If a
+paper is checked at the start of peer review, it might need
+re-checking if the paper is modified during peer review.  This is
+somewhat unavoidable, and has happened to us (e.g.,~certificate
+\texttt{2020-012} \cite{cert-2020-012}). This can
+also be desired though, if interactions between author, reviewer,
+and codechecker led to improvements.  Checking the manuscript the
+second time is likely to be much less work than the first time.
+
+\textbf{What does it mean for a figure to be reproducible?}
+
+Automatically detecting if a codechecker's results are ``the same'' as
+an author's is more challenging than it might appear.  That is why we
+do not require results to be \textit{identical} for a CODECHECK to pass
+but simply that the code runs and generates output files that the
+author claims. Stochastic simulations mean that often we cannot get
+exactly the same results, and that even the same versions of libraries
 can generate outputs that differ by operating system
-\cite{Gronenschild2012-pp}.
-While reproducibility practices can mitigate some of these problems, e.g., by using a seed, the flexibility of the human judgement is still needed.
-
-\textbf{Shouldn't the next step be more revolutionary?}
-CODECHECK's approach is to acknowledge the shortcomings around 
-computational reproducibility and to iteratively improve the current
-system of scholarly communication based on journals and conferences.
-It remains to be proven whether this approach is welcomed broadly and whether 
-involving the traditional publishing stakeholders helps to further the cause.
-We have discussed stronger more drastic rules at length, such as only 
-considering fully free and open source software, diamond Open Access journals or community-led no-fee journals,
-and journals with collaborative/open peer review, but we eventually decided
-against them on the level of the principles.
-For the CODECHECK community process\footnote{\url{https://codecheck.org.uk/guide/community-process}}
-and the volunteer codechecker community, these requirements are reconsidered.
-
-We also have deliberated requiring the latest technologial solutions which can
-support reproducibility (cf.~\cite{konkol_publishing_2020}), but we have 
-decided to focus on the human interface and the judgement of experienced
-researchers and developers as a more sustainable and powerful approach.
-The structurelessness of the required data and code gives all types of research the freedom to adopt CODECHECK.
-CODECHECK could be complemented by automated scoring (e.g.,~\cite{menke_rigor_2020}), yet automation and metrics bear new risks.
-Note that the focus of the CODECHECK principles on code execution allows journals
-and publishers to innovate on the financial model and peer review practices
-at their own pace.
+\cite{Gronenschild2012-pp}.  While reproducibility practices can
+mitigate some of these problems, e.g., by using a seed, the
+flexibility of the human judgement is still needed, rather than
+bitwise reproducibility.  The codechecker is free to comment on
+visible differences in outputs in their report.
+
+\textbf{Shouldn't the next step be more revolutionary?}  CODECHECK's
+approach is to acknowledge shortcomings around computational
+reproducibility and to iteratively improve the current system.  It
+remains to be proven whether this approach is welcomed broadly and if
+involving publishing stakeholders helps to further the cause.  We have
+discussed more stringent rules at length, e.g. only considering fully
+free and open source software, diamond Open Access journals, but we
+eventually decided against them on the level of the principles.  For
+the CODECHECK community
+process\footnote{\url{https://codecheck.org.uk/guide/community-process}}
+and the volunteer codechecker community, these requirements are
+reconsidered.
+
+We also have deliberated requiring modern technological solutions to
+support reproducibility (cf.~\cite{konkol_publishing_2020}), but we
+decided to focus on the human interface and the judgement of
+experienced researchers and developers as a more sustainable and
+flexible approach.  All types of research can adopt CODECHECK due to
+its flexible design.  CODECHECK could include automated scoring
+(e.g.,~\cite{menke_rigor_2020}), yet automation and metrics bear new
+risks.  The focus of the CODECHECK principles on code execution allows
+journals and publishers to innovate on financial  models and peer
+review practices at their own pace.
 
@@
 
-CODECHECK works!
-We have created a considerable number of certificates to demonstrate it.
-The creation of the certificates and the interactions with authors and
-editors not only improved the principles and the workflow but also
-confirmed the approach taken. This result corroborates findings from
-similar evaluations of reproducible computational research in journals and
-at conferences.
-CODECHECKs increase transparency of the checked papers and can contribute
-to building trust in scientific results.
+CODECHECK works --- we have created a considerable number of
+certificates to demonstrate it.  The creation of certificates and
+interactions with authors and editors shaped the principles and the
+workflow and also confirmed the approach taken. This result
+corroborates findings from similar evaluations of reproducible
+computational research in journals and conferences.  CODECHECKs
+increase transparency of the checked papers and can contribute to
+building trust in scientific results.
 % "science earns trust through transparency" > https://twitter.com/briannosek/status/1290240244101521408
-The set of 
-shared principles and common name, through recognition value,
-will allow researchers
-across publication venues to judge the level of scrutiny that results have
-faced. CODECHECK requires direct acknowledgement of the contribution to the
-scientific community in the role of codechecker, not indirectly via citations 
-of reproductions or informal crediation and closed inner circles.
-Forther, codechecking can be a worthwhile educational activity, especially for
-ECRs.
-
-Yet, CODECHECK harbours the same limitations as peer review
-in general and is closely connected to larger disruptions and challenges
+The set of shared principles and common name, through recognition
+value, will allow researchers to judge the level of scrutiny that
+results have faced. CODECHECK requires direct acknowledgement of the
+codechecker's contributions, not indirectly via citations of
+reproductions or informal crediation and closed inner circles.
+
+CODECHECK however harbours the same limitations as peer review in
+general and is closely connected to larger disruptions and challenges
 in scholarly communication
-\cite{eglen_recent_2018,tennant_ten_2019,fyfe_mission_2019}, 
-including the tensions between commercial publishing and reviewers' often
-free labour, and a global pandemic that has jumbled up scientific publishing
-and exposed a broader general audience to science, even preliminary 
-results \cite{munafo_what_2020}.
-It is clear that establishing CODECHECK-like systems can be 
-impeded by and must be seen as interconnected with much larger issues in 
-science, such as broken metrics or malpractice triggered by publication
-pressure \cite{piwowar_altmetrics:_2013,nosek_promoting_2015}.
-We certainly do not want the binary attribute of ``code works''
-to become a factor in bibliometric approaches for performance assesments.
+\cite{eglen_recent_2018,tennant_ten_2019,fyfe_mission_2019}, including
+the tensions between commercial publishing and reviewers' often free
+labour, and a global pandemic that has jumbled up scientific
+publishing and exposed a broader general audience to science, even
+preliminary results \cite{munafo_what_2020}.  Establishing CODECHECK
+workflows must be seen as interconnected with much larger issues in
+science, such as broken metrics or malpractice triggered by
+publication pressure
+\cite{piwowar_altmetrics:_2013,nosek_promoting_2015}.  We certainly do
+not want the binary attribute of ``code works'' to become a factor in
+bibliometric approaches for performance assesments.
 % Mittermaier, Peer Review and Bibliometrics: http://hdl.handle.net/2128/22745
 
-While developed for the current ``paper''-centric publication process and
-being compatible with different styles of peer review, the CODECHECK 
-principles would also work well with novel publication paradigms, e.g.,
-peer-reviewed computational notebooks \cite{earthcube_new_2020},
-iterative and granular communication of scientific outputs, 
-articles with live-code \cite{perkel_pioneering_2019-1}
-\footnote{For example, eLife's ERA, see 
+While developed for the current ``paper''-centric publication process,
+the CODECHECK principles would also work well with novel publication
+paradigms, e.g., peer-reviewed computational notebooks
+\cite{earthcube_new_2020}, iterative and granular communication of
+scientific outputs, articles with live-code
+\cite{perkel_pioneering_2019-1} \footnote{For example, eLife's ERA,
+  see
   \url{https://elifesciences.org/labs/dc5acbde/welcome-to-a-new-era-of-reproducible-publishing}.},
 decentralized infrastructure and public reviewer reputation systems
-\cite{tenorio-fornes_towards_2019},
-and completely new visions for scholarly communication and peer 
-review\footnote{Cf., for example, \emph{A modern vision for peer review} by 
-Amy J. Ko: \url{https://medium.com/bits-and-behavior/a-modern-vision-for-peer-review-d5f73f0fae07}}.
+\cite{tenorio-fornes_towards_2019}, and completely new visions for
+scholarly communication and peer review\footnote{Cf., for example,
+  \emph{A modern vision for peer review} by Amy J. Ko:
+  \url{https://medium.com/bits-and-behavior/a-modern-vision-for-peer-review-d5f73f0fae07}}.
 An explicit segmentation of research steps could even make the focus
-of a CODECHECK easier by only checking the ``analysis'' sub-publication.
-The discovery of CODECHECKs could be increased by depositing the reports
-into public databases of reproductions, such as \emph{SciGen.Report}.
-Public researcher profiles, such as ORCID, may consider different
-types of reviewer activity to capture how independent code execution contributes to science.
-Notably, the discussed limitations are largely self-imposed for easier
-acceptance and evolutionary integration, as to not break the current 
-system and increase demands gradually without leaving practitioners behind.
-
-If adopted today, a CODECHECK system, even if temporary in a sustainable 
-transition towards more open publication and review practices, can contribute
-to increased trust in the results of research. Introducing CODECHECK
-should be informed by lessons learned from (introducing) open peer review
-(OPR) \cite{ross-hellauer_guidelines_2019}.
-All our conversations with publishers and editors at some point
-touched on the procedural and technical implications of openness, be it 
-that OPR is neither applied by nor familiar to stakeholders, or that process
-management software would not support such open communication.
+of a CODECHECK easier by only checking the ``analysis''
+sub-publication.  The discovery of CODECHECKs could be increased by
+depositing the reports into public databases of reproductions, such as
+\emph{SciGen.Report}.  Public researcher profiles, such as ORCID, may
+consider different types of reviewer activity to capture how
+independent code execution contributes to science.  Notably, the
+discussed limitations are largely self-imposed for easier acceptance
+and evolutionary integration, as to not break the current system and
+increase demands gradually without leaving practitioners behind.
+
+A CODECHECK system, even if temporarily adopted as a sustainable
+transition towards more open publication and review practices, can
+contribute to increased trust in research outputs. Introducing
+CODECHECK should be informed by lessons learned from (introducing)
+open peer review (OPR) \cite{ross-hellauer_guidelines_2019}.  Our
+conversations with publishers and editors included procedural and
+technical implications of openness, be it that OPR is neither applied
+by nor familiar to stakeholders, or that process management software
+would not support such open communication.
 
 More reproducible practices initiated by CODECHECKs could lead
-communities or whole disciplines to reach a state where authors provide 
-sufficient material and reviewers have acquired sufficient skills that
-the reviewers will generally conduct a CODECHECK-level of checking;
-only in especially sophisticated cases will a specialised codechecker
-be needed.
-The main challenge for us remains getting journals to embrace the
-idea behind CODECHECK and to realise processes that conform to the principles,
-whether or not they use CODECHECK by name.
-We would be keen to use the flexibility of the principles and cooperate
-with journals to learn more about the advantages and yet unclear
-specific challenges -- Do CODECHECKs really work better with OPR?
-To facilitate the adoption, the CODECHECK badge is, intentionally, not branded beyond the checkmark
-and green colour and simply states ``code works''.
-
-Each rollout of CODECHECK should be accompanied by studies to ensure 
-codechecking does not fall into the same gaps as peer review did
-\cite{tennant_limitations_2020}
-and to ensure positive change within the review system.
-This \emph{cultural change}, however, is needed for the valuation of the
-efforts that go into proper evaluation of papers.
-Multiple adopting journals will also help to answer relevant questions on
-the establishing process: What are crucial decisions or pain points? Can
-authors retract code/data once a CODECHECK has started?
-What variants of CODECHECKs will be most common? How will open CODECHECKs
-influence or codevelop with the scope and anonymity of conventional review over time.
-
-With the implementation of CODECHECK workflows, the question of training
-codecheckers will become relevant. We expect both a mentoring scheme within
-the CODECHECK community (experienced codecheckers provide on-the-job training
-or serve as fallback advisors) as well as collaboration with reproducible
-research initiatives such as
+communities to reach a state where authors provide sufficient material
+and reviewers have acquired sufficient skills that peer reviewers will
+generally conduct a CODECHECK-level of checking; only in especially
+sophisticated cases will a specialised codechecker be needed.  The
+main challenge for us remains getting journals to embrace the idea
+behind CODECHECK and to realise processes that conform to the
+principles, whether or not they use CODECHECK by name.  We would be
+keen to use the flexibility of the principles and cooperate with
+journals to learn more about the advantages and yet unclear specific
+challenges -- e.g do CODECHECKs really work better with OPR?  To
+facilitate the adoption, the CODECHECK badge is, intentionally, not
+branded beyond the checkmark and green colour and simply states ``code
+works''.
+
+Future CODECHECK versions may be accompanied by studies to ensure
+codechecking does not fall into the same traps as peer review did
+\cite{tennant_limitations_2020} and to ensure positive change within
+the review system.  This \emph{cultural change}, however, is needed
+for the valuation of the efforts that go into proper evaluation of
+papers.  Journals can help us to answer open questions in our system: What are
+crucial decisions or pain points? Can authors retract code/data once a
+CODECHECK has started?  What variants of CODECHECKs will be most
+common? How will open CODECHECKs influence or codevelop with the scope
+and anonymity of conventional review over time?
+
+The question of training codecheckers is also relevant. We expect a
+mentoring scheme within the CODECHECK community (experienced
+codecheckers will provide on-the-job training or serve as fallback
+advisors). Codecheckers may also be found by collaborating with
+reproducible research initiatives such as
 ReproHack\footnote{\url{https://reprohack.github.io/reprohack-hq/}},
 ReproducibiliTea\footnote{\url{https://reproducibilitea.org/}}
-\cite{fitzgibbon_brewing_2020},
-and Repro4Everyone\footnote{\url{https://repro4everyone.org/}}
-\cite{auer_reproducibility_2020},
-to be able to grow the pool of codecheckers as needed.
-The initial reaction of the scientific community and the spontaneous signing
-up volunteer codecheckers make us optimistic that the time is right for 
-putting scholarly peer review on the path towards facilitating the sharing and execution
-of computational workflows.
+\cite{fitzgibbon_brewing_2020}, and
+Repro4Everyone\footnote{\url{https://repro4everyone.org/}}
+\cite{auer_reproducibility_2020}.  The initial reaction of the
+scientific community to these ideas shows that scholarly peer review
+should continue on the path towards facilitating the sharing and
+execution of computational workflows.
 
@@
 
-SJE was on the editorial board at the journal \emph{Scientific Data} until January 2021.
-DN is reproducibility chair at the Association of Geographic Information Laboratories in Europe's (AGILE) annual conference.
+%% SJE: check again in April! s/is/was
+SJE is on the editorial board at the journal \emph{Scientific Data}
+until March 2021.  DN is reproducibility chair at the Association of
+Geographic Information Laboratories in Europe's (AGILE) annual
+conference.
 
@@
 the work presented here: Andy Collings, Melissa Harrison, Giuliano
-Maciocci, Naomi Penfold, Emmy Tsang (eLife), Rebecca Kirk (PLOS Computational Biology), Scott
-Edmunds (GigaScience), and Andrew Hufton (Scientific Data). Iain Davies and
-Yuhao (Sebastian) Wang developed code and example certificates.
-We thank Antonio P\'{a}ez (Journal of Geographical Systems) for enabling
-CODECHECKs, Carlos Granell and Frank Ostermann for contributing
-certificates as reproducibility reviewers at the AGILE conference, and 
-all authors of auditable workflows for their participation.
-We thank Celeste R. Brennecka from the Scientific Editing Service, University of M\"unster, for her editorial review.
+Maciocci, Naomi Penfold, Emmy Tsang (eLife), Rebecca Kirk (PLOS
+Computational Biology), Scott Edmunds (GigaScience), and Andrew Hufton
+(Scientific Data). Iain Davies and Yuhao (Sebastian) Wang developed
+code and example certificates.  We thank Antonio P\'{a}ez (Journal of
+Geographical Systems) for enabling CODECHECKs, Carlos Granell and
+Frank Ostermann for contributing certificates as reproducibility
+reviewers at the AGILE conference, and all authors of auditable
+workflows for their participation.  We thank Celeste R. Brennecka from
+the Scientific Editing Service, University of M\"unster, for her
+editorial review.
 

### 2021-02-09 codecheck-paper.tex
comply with submission guidelines
- include license and citation for R package
- keywords
- data availability

@@
 \affil[2]{Department of Applied Mathematics and Theoretical Physics, University of Cambridge, UK; \email{sje30@cam.ac.uk}}
-\affil[$\ast$]{Contributed equally}
+\affil[$\ast$]{Corresponding authors}
 \affil[$$]{Public project for this article is at \url{https://github.com/codecheckers/paper}.}
@@
 \begin{abstract}
+%\textbf{Background:}
+%\textbf{Methods:}
+%\textbf{Results:}
+%\textbf{Conclusions:}
   The traditional scientific paper falls short of effectively communicating computational research.
@@
 \end{abstract}
+
+\section*{Keywords}
+
+reproducible research, Open Science, peer review, reproducibility, code sharing, data sharing, quality control, scholarly publishing
 
@@
 
-A \textbf{custom R package}, \texttt{codecheck}\footnote{
-  \url{https://github.com/codecheckers/codecheck}}, automates
+A \textbf{custom R package}, \texttt{codecheck}, automates
 repetitive tasks around authoring certificates and managing the
-register.  It includes scripts to deposit certificates and related
+register. The package is published at
+\url{https://github.com/codecheckers/codecheck} under MIT license
+\cite{stephen_eglen_codecheckerscodecheck_2021}.
+It includes scripts to deposit certificates and related
 files to Zenodo using the R package \texttt{zen4R} \cite{zen4r} and
@@
 
+\subsection{Data availability}
+% https://f1000research.com/for-authors/data-guidelines#addstatement
+
+Repository: Daniel Nüst, Stephen Eglen, & Iain Davies.
+(2021).
+codecheckers/register: CODECHECK Register Deposit January 2021 (Version 2021-01) [Data set].
+Zenodo.
+http://doi.org/10.5281/zenodo.4486559
+
+This project contains the following underlying data:
+
+\begin{itemize}
+  \item `register.csv`. List of all CODECHECK certificates with references to repositories and reports.
+\end{itemize}
+
+Data are available under the terms of the \href{https://creativecommons.org/licenses/by-sa/4.0/legalcode}{Creative Commons Attribution Share Alike license} (CC-BY-SA 4.0 International).
+
 {\small\bibliographystyle{unsrtnat}

### 2021-02-09 codecheck-paper.tex
typos

@@
   We propose a system by which the computational workflows underlying research articles are checked.
-  The CODECHECK system uses open intrastructure and tools and can be integrated into review and publication processes in multiple ways.
+  The CODECHECK system uses open infrastructure and tools and can be integrated into review and publication processes in multiple ways.
   We describe these integrations along multiple dimensions (importance, who, openness, when).
@@
 (2) the PUBLISHER of a journal interested in publishing the author's paper, and
-(3) the CODECHECKER, who checks that the authors's code works.
+(3) the CODECHECKER, who checks that the author's code works.
 The six-step workflow we have refined is shown in Figure~\ref{fig:workflow}.
@@
 of computational results, i.e.,~checking that the code correctly produces the output it claims to create, but not a validation, i.e.,~checking that the code implements the right algorithm to solve the specific research problem.
-Nevertheless, simply attempting to reproduce an output may highlight a submisison's shortcomings in meeting a journal's requirements (cf.~\cite{christian_journal_2020}) and may effectively increase transparency, thereby improving practices (cf.~\cite{nosek_scientific_2012}) even if the check does not go into every detail.
+Nevertheless, simply attempting to reproduce an output may highlight a submission's shortcomings in meeting a journal's requirements (cf.~\cite{christian_journal_2020}) and may effectively increase transparency, thereby improving practices (cf.~\cite{nosek_scientific_2012}) even if the check does not go into every detail.
 
@@
 (cf.~\cite{everythinghertz123}).
-% previous sentenc inspired by https://osf.io/9cftx/
+% previous sentence inspired by https://osf.io/9cftx/
 
@@
 expertise of the checker, though in general, we expect a CODECHECK 
-to consider reproducbility of the results above performance of the code.
+to consider reproducibility of the results above performance of the code.
 
@@
 publisher.  A \emph{single codechecker} is sufficient to verify the
-workflow since it is mostly a factual process and not require much
+workflow since it is mostly a factual process and does not require much
 debate or opinion.  Code usually harbours systematic and repeatable
@@
 expertise and specialisation.
-CODECHECK offes a way for ECRs to gain insights into new research and
+CODECHECK offers a way for ECRs to gain insights into new research and
 highlight the importance of reproduction.  
@@
 a codechecker could be similarly listed.
-% The number of scientific reviews is of course only a coarse indicator of an individual's community contribution, yet it is unclear if a regular reviewer, who conductsa regular scientific review and in addition also conducts a CODECHECK, should be credited twice.
+% The number of scientific reviews is of course only a coarse indicator of an individual's community contribution, yet it is unclear if a regular reviewer, who conducts a regular scientific review and in addition also conducts a CODECHECK, should be credited twice.
 % In this case, the published CODECHECK certificate and possibly the published peer review can clearly describe the possibly large efforts put into both.
@@
     \texttt{2020-009}  \cite{cert-2020-009} & COVID-19 & Code for analysis of effectiveness of measures to reduce transmission of SARS-CoV-2 checked as preprint \cite{kucharski-preprint-2020} and later published \cite{kucharski_effectiveness_2020}.
-\\ % R, check mentiond in acks, but not linked
+\\ % R, check mentioned in acks, but not linked
     \texttt{2020-010}  \cite{cert-2020-010} & COVID-19 & Code for analysis of non-pharmaceutical interventions (Report 9) checked as a preprint \cite{ferguson_report_2020}. \\ % CovidSim
@@
 \\
-    \texttt{2020-013}  \cite{cert-2020-013} & Neuroscience & Code for analsis of rest-activity patterns in people wihout con-mediated vision was checked as a preprint \cite{Spitschan2020.06.02.129502} after direct contact with the authors.
+    \texttt{2020-013}  \cite{cert-2020-013} & Neuroscience & Code for analysis of rest-activity patterns in people without con-mediated vision was checked as a preprint \cite{Spitschan2020.06.02.129502} after direct contact with the authors.
 \\ % MATLAB
-    \texttt{2020-014}  \cite{cert-2020-014} & Neuroscience & Code for analysis of pertubation patterns of neural activity was checked after publication as part of publisher collaboration \cite{Sadeh2020}.
+    \texttt{2020-014}  \cite{cert-2020-014} & Neuroscience & Code for analysis of perturbation patterns of neural activity was checked after publication as part of publisher collaboration \cite{Sadeh2020}.
 \\ % MATLAB
@@
 and depositing the certificate.  This flexibility  accommodates
-different skill sets and unforseen technical advances or challenges.
+different skill sets and unforeseen technical advances or challenges.
 
@@
 and cryptographic \emph{Verifiable Result Identifier} would indeed
-provide excellent reproduciblity. However, the system is also complex
+provide excellent reproducibility. However, the system is also complex
 and since its creation in 2011 we are not aware of any publisher using
@@
 community-run independent platform to foster communication on
-reproducibility.  People can report on fully, partiially, or
+reproducibility.  People can report on fully, partially, or
 un-successful reproductions of articles after publication.
@@
 manuscripts\footnote{\url{https://medium.com/paperswithcode/papers-with-code-partners-with-arxiv-ecc362883167}}
-and runs a reproduciblity
+and runs a reproducibility
 challenge\footnote{\url{https://paperswithcode.com/rc2020}}.
@@
 Likewise, different disciplines and journals provide reproducibility
-checklists, e.g.,~sciene and engineering \cite{rosenberg_next_2020} or
+checklists, e.g.,~science and engineering \cite{rosenberg_next_2020} or
 GIScience \cite{nust_agile_2019}, which naturally share some aspects
@@
 On the surface, yes, but peer reviewers are overburdened enough and
-asking them to do more work usring peer review is not likely to
-suceed.  Miyakawa described how when he as an editor simply requested
+asking them to do more work around peer review is not likely to
+succeed.  Miyakawa described how when he as an editor simply requested
 raw data from authors during submission, this led to almost all of
@@
 
-A codechecker does not fulfil the same role as a \emph{statistical
+A codechecker does not fulfill the same role as a \emph{statistical
   reviewer}, as it is applied by some journals in the biomedical
@@
 published.  The most appropriate way forwards seems to be to give
-private feedback for failures and public acknowledgement and
+private feedback for failures and public acknowledgment and
 encouragement for successes.
-% eLife's new model (see https://osf.io/9cftx/) will publish reports evantually, e.g.,~after preprint has been accepted elsewhere
+% eLife's new model (see https://osf.io/9cftx/) will publish reports eventually, e.g.,~after preprint has been accepted elsewhere
 Rosenthal et~al. \cite{Rosenthal2016b} discuss such incentives for different actors around the
@@
 not want the binary attribute of ``code works'' to become a factor in
-bibliometric approaches for performance assesments.
+bibliometric approaches for performance assessments.
 % Mittermaier, Peer Review and Bibliometrics: http://hdl.handle.net/2128/22745

### 2021-02-09 codecheck-paper.tex
bye bye footnotes, remove manual page breaks

@@
 Just as data journals' validations of data and all journals' peer review  provides a ``baseline reassurance'', i.e.,~that a paper has been checked by someone with an understanding of the topic \cite{fyfe_mission_2019}, the same baseline could be provided for the workflow underlying a paper.
-With this in mind, we have developed a set of principles and an example workflow that provides a pragmatic way of checking that a paper's code works.
-
-Here we offer a thorough description of a process and its variations to integrate a much-needed evaluation of computational reproducibility\footnote{Following the Claerbout/Donoho/Peng terminology, see \cite{barba_terminologies_2018}.} into peer review, and we demonstrate its feasibility by means of 25 reproductions across scientific disciplines.
+With this in mind, we have developed a set of principles and an example workflow that provides a pragmatic way of checking that a paper's code works, i.e., it is reproducible following the Claerbout/Donoho/Peng terminology \cite{barba_terminologies_2018}.
+
+Here we offer a thorough description of a process and its variations to integrate a much-needed evaluation of computational reproducibility into peer review, and we demonstrate its feasibility by means of 25 reproductions across scientific disciplines.
 We call this system CODECHECK.
@@
 following their definition, a CODECHECK ensures verification
-\footnote{Note this is a different use from that referred to as the \emph{Verification Reports} article type in the journal \emph{Cortex} \cite{chambers_verification_2020}.},
+%\footnote{Note this is a different use from that referred to as the \emph{Verification Reports} article type in the journal \emph{Cortex} \cite{chambers_verification_2020}.},
 of computational results, i.e.,~checking that the code correctly produces the output it claims to create, but not a validation, i.e.,~checking that the code implements the right algorithm to solve the specific research problem.
@@
 through metadata. Currently, the CODECHECKER creates these connections on
-Zenodo\footnote{
-See ``related identifiers'' and ``alternate identifiers'' in the right-hand
-sidebar of certificate \texttt{2020-025} \cite{cert-2020-025}.}.
+Zenodo. They appear as links with a relationship type on the Zenodo
+landing page for a record in the right-hand sidebar, e.g., the 
+``related identifiers'' and ``alternate identifiers'' 
+of certificate \texttt{2020-025} \cite{cert-2020-025}.
 The PUBLISHER also credits the CODECHECKER for their work by depositing
-the activity appropriately in scholarly profiles, such as ORCID\footnote{
-See peer review contributions in ORCID records at
-\url{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}.}.
+the activity appropriately in scholarly profiles, such as ORCID
+(see peer review contributions in ORCID records,
+\url{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}).
 The PUBLISHER also ensures proper publication metadata, e.g.,~links from the 
 certificate repository to the published paper or the original code repository.
-
-\newpage
 
@@
 \emph{ReScience X}, a journal devoted to reproduction and replication experiments \cite{roesch_new_2020}, shares an interest in this combination.
-ECRs are also often familiar with new technologies, thus also making them likely to author CODECHECK-ready manuscripts
-\footnote{Cf. the 77\% of 141 registered reports that were submitted to leading Neuroscience journals by ECRs \cite{chambers_registered_2019}.}.
+ECRs are also often familiar with new technologies, thus also making them likely to author CODECHECK-ready manuscripts.
+A supporting data point for ECRs as early adopters is that they are responsible for 77\% of 141 registered reports that were submitted to leading Neuroscience \cite{chambers_registered_2019}.
 As ECRs are introduced to peer review as codecheckers, they may transition into the role of scientific reviewer over time.
@@
 publicly credited for their reviewing activity.  A regular review
-may be listed in public databases (e.g.,~ORCID
-\footnote{\href{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}}
-or commercial offerings such as Publons
-\footnote{\href{https://publons.com/}{https://publons.com/}} or
-ReviewerCredits\footnote{\href{https://www.reviewercredits.com/}{https://www.reviewercredits.com/}});
+may be listed in public databases (e.g.,~ORCID, see Step~6 above,
+or commercial offerings such as Publons, \url{https://publons.com/}, and
+ReviewerCredits, \url{https://www.reviewercredits.com/});
 a codechecker could be similarly listed.
@@
 
-The codechecker
-community\footnote{\href{https://github.com/codecheckers/codecheckers/}{https://github.com/codecheckers/codecheckers/}}
-has over 20 volunteers who signed up in the last year.  Their
+The codechecker community has over 20 volunteers who signed up in the last
+year, see \url{https://github.com/codecheckers/codecheckers/}.  Their
 motivations, mentioned in the registration
-information\footnote{\url{https://github.com/codecheckers/codecheckers/labels/registration}},
+information at \url{https://github.com/codecheckers/codecheckers/labels/registration},
 include: supporting reproducible research and Open Science, improve
@@
 practical tools, such as publishing all building blocks of science in
-a research compendium
-\footnote{\url{https://research-compendium.science/}}.  A completed
+a research compendium (cf.~\url{https://research-compendium.science/})
+or ``repro-pack'' \cite{barba_praxis_2018}.
+A completed
 check means that code could be executed at least once using the
@@
 in favour of bitwise reproducibility because, in the long run, it can
-be automated\footnote{Konrad Hinsen on Twitter:
-  \url{https://twitter.com/khinsen/status/1242842759733665799}}, but
+be automated (e.g., Konrad Hinsen on Twitter,
+\url{https://twitter.com/khinsen/status/1242842759733665799}), but
 until then we need CODECHECK's approach.
@@
 is stored in a machine-readable structured file in YAML, the CODECHECK
-configuration file \texttt{codecheck.yml}\footnote{ See technical
-  specification of the CODECHECK configuration file at
-  \url{https://codecheck.org.uk/spec/config/latest/}.}.  The
+configuration file \texttt{codecheck.yml}.
+The technical specification of the CODECHECK configuration file is
+published at \url{https://codecheck.org.uk/spec/config/latest/}.  The
 configuration file enables current and future automation of workflows
@@
 system.  The \texttt{codecheckers} \textbf{GitHub}
-organisation\footnote{ \url{https://github.com/codecheckers}} contains
+organisation at \url{https://github.com/codecheckers} contains
 projects for managing the project website, the codecheckers community
 and its discussions, code repositories, and the main register of
-CODECHECKs. Both the website\footnote{
-  \url{https://codecheck.org.uk/}} and the register\footnote{
-  \url{https://codecheck.org.uk/register}} are hosted as GitHub
-pages. The register is a single table in CSV format that connects the
+CODECHECKs. Both the project website \url{https://codecheck.org.uk/}}
+and the register at \url{https://codecheck.org.uk/register}
+are hosted as GitHub pages.
+The register database is a single table in CSV format that connects the
 certificate identifier with the repository associated with a specific
@@
 long-term availability of all digital artefacts related to the
-project. The CODECHECK community on Zenodo
-\footnote{\url{https://zenodo.org/communities/codecheck/}} includes
-certificates and the regularly archived register
-\cite{codecheck_register_jan2021}.
+project. The CODECHECK community on Zenodo is available at
+\url{https://zenodo.org/communities/codecheck/}. It holds
+certificates, the regularly archived register
+\cite{codecheck_register_jan2021}, and other material related to
+CODECHECK.
 
@@
 
-Fifteen RCR Reports have been published so far
-\footnote{\url{https://search.crossref.org/?q=\%22Replicated+Computations+Results+\%28RCR\%29+Report\%22&from_ui=yes&page=2};
-  search executed on 2020-12-10.}
+Fifteen RCR Reports have been published so far (search on 
+\url{https://search.crossref.org/} with the term
+\texttt{"Replicated Computations Results (RCR) Report"} on 2020-12-10).
 % 15: https://academic.microsoft.com/search?q=Replicated%20Computations%20Results%20%22(RCR)%20Report%22&f=&orderBy=0&skip=10&take=10
@@
 \emph{Transactions on Modeling and Computer Simulation
-  (TOMACS)}\footnote{\url{https://dl.acm.org/journal/tomacs/author-guidelines}}.
+  (TOMACS)} (cf. \url{https://dl.acm.org/journal/tomacs/author-guidelines}).
 The TOMS RCR follows CODECHECK principles 1-4, although our work was
@@
 conflict of interest is acknowledged.  TOMS also recognises reviewer
-activity in a partnership with Publons\footnote{
-  \url{https://authors.acm.org/author-services/publons}}.
-%, but it is
-unclear if the author of an RCR report receives both the report's
-%authorship and the reviewer recognition. 
-CODECHECK however specifically aims to give codecheckers
+activity in a partnership with Publons
+(see \url{https://authors.acm.org/author-services/publons}).
+It is unclear if the author of an RCR report receives both recognitions.
+CODECHECK, however, specifically aims to give codecheckers
 recognition as reviewers.  In our view, the reviewer role removes the
@@
 Several journals provide special article types for reproductions of
-published papers.  \emph{Information Systems} has an
-invited reproducibility paper track\footnote{
-  \url{https://www.elsevier.com/journals/information-systems/0306-4379/guide-for-authors}.}.
+published papers.  \emph{Information Systems} has an invitation only
+Reproducibility Section for articles describing the reproducibility 
+efforts of published articles, which are co-authored by the original
+authors and the reproducibility reviewer(s) (see
+\url{https://www.elsevier.com/journals/information-systems/0306-4379/guide-for-authors}).
 \emph{Nature Machine Intelligence} recently introduced a new type of
@@
 Going beyond article types and a specific journal, the journal
-\emph{ReScience C} publishes only
-replications\footnote{\url{https://rescience.github.io/faq/}}, also
+\emph{ReScience C} (\url{https://rescience.github.io/}) publishes only
+replications, also
 requiring open code and replication by a third party. ReScience also
@@
 Such internal checks are professionalised in local reproduction services,
-such as CISER R-squared\footnote{
-\url{https://ciser.cornell.edu/research/results-reproduction-r-squared-service/}} (established/permanent),
-%Uni Münster's Reproducible Research Support Service\footnote{
-%\url{https://go.wwu.de/r2s2}} (pilot),
-Oxford's code review network (OxCRN)\footnote{
-\url{https://github.com/OxfordCodeReviewNet/forum}} (community, broader scope), or
-Yale's YARD\footnote{
-\url{https://isps.yale.edu/research/data/approach}} (staff for replications).
+such as CISER R-squared
+(\url{https://ciser.cornell.edu/research/results-reproduction-r-squared-service/})
+or YARD (\url{https://isps.yale.edu/research/data/approach}),
+or in communities such as Oxford's code review network (OxCRN,
+\url{https://github.com/OxfordCodeReviewNet/forum}).
 
@@
 Another platform for publishing results of reproductions is
-\emph{SciGen.Report}\footnote{\url{https://scigen.report/}}.  It is a
+\emph{SciGen.Report} (\url{https://scigen.report/}).  It is a
 community-run independent platform to foster communication on
@@
 with reproducibility awareness, education, and support, such as
-ReproHack\footnote{\url{https://reprohack.github.io/reprohack-hq/}},
-Code Copilot\footnote{\url{https://twitter.com/Code_Copilot}}, or
-Papers with Code\footnote{\url{https://paperswithcode.com/about}}.
+ReproHack (\url{https://reprohack.github.io/reprohack-hq/}),
+Code Copilot (\url{https://twitter.com/Code_Copilot}), or
+Papers with Code (\url{https://paperswithcode.com/about}).
 The latter recently announced a collaboration with the preprint server
 \emph{arXiv} on providing data and code supplements for machine
-learning
-manuscripts\footnote{\url{https://medium.com/paperswithcode/papers-with-code-partners-with-arxiv-ecc362883167}}
-and runs a reproducibility
-challenge\footnote{\url{https://paperswithcode.com/rc2020}}.
+learning manuscripts
+%\footnote{\url{https://medium.com/paperswithcode/papers-with-code-partners-with-arxiv-ecc362883167}}
+and runs a reproducibility challenge.
+%\footnote{\url{https://paperswithcode.com/rc2020}}.
 % Papers with Code now also does more disciplines than just ML.
@@
 
-\newpage
-
 \textbf{Should CODECHECK requirements be more demanding?} \\
@@
 an eternally functional workflow nor suggests a specific software
-stack or practical approach even if such suggestions, e.g.,~research
-compendia or ``repro-packs''\footnote{See
-  \url{https://research-compendium.science/} respectively
-  \cite{barba_praxis_2018}.}, are valuable.  Creating something
+stack or practical approach.
+Creating something
 that anyone can reproduce has been called a fool's
-errand\footnote{\url{https://twitter.com/DougBlank/status/1135904909663068165?s=09}}
+errand (Douglas Blank on Twitter,
+\url{https://twitter.com/DougBlank/status/1135904909663068165})
 and we tend to agree.  However, the package of data, code, and
@@
 conduct software reviews independent of a specific publication or
-workflow include ROpenSci\footnote{\url{https://ropensci.org/}},
-PyOpenSci\footnote{\url{https://www.pyopensci.org/}}, and
-JOSS\footnote{\url{https://joss.theoj.org/}}.  Furthermore, the
+workflow include ROpenSci (\url{https://ropensci.org/}),
+PyOpenSci (\url{https://www.pyopensci.org/}), and
+JOSS (\url{https://joss.theoj.org/}).  Furthermore, the
 codechecker's task list is intentionally not overloaded with related
@@
 (Research Software Engineers,
-RSE\footnote{\url{https://en.wikipedia.org/wiki/Research_software_engineering}})
+RSE, see \url{https://en.wikipedia.org/wiki/Research_software_engineering})
 who might not regularly be involved in writing or reviewing papers
@@
 the CODECHECK community
-process\footnote{\url{https://codecheck.org.uk/guide/community-process}}
+process, which is documented online at
+\url{https://codecheck.org.uk/guide/community-process},
 and the volunteer codechecker community, these requirements are
@@
 scientific outputs, articles with live-code
-\cite{perkel_pioneering_2019-1} \footnote{For example, eLife's ERA,
-  see
-  \url{https://elifesciences.org/labs/dc5acbde/welcome-to-a-new-era-of-reproducible-publishing}.},
+\cite{perkel_pioneering_2019-1} such as eLife's ERA
+(see \url{https://elifesciences.org/labs/dc5acbde/welcome-to-a-new-era-of-reproducible-publishing}),
 decentralized infrastructure and public reviewer reputation systems
 \cite{tenorio-fornes_towards_2019}, and completely new visions for
-scholarly communication and peer review\footnote{Cf., for example,
-  \emph{A modern vision for peer review} by Amy J. Ko:
-  \url{https://medium.com/bits-and-behavior/a-modern-vision-for-peer-review-d5f73f0fae07}}.
+scholarly communication and peer review such as described by Amy J. Ko
+in \emph{A modern vision for peer review}
+(\url{https://medium.com/bits-and-behavior/a-modern-vision-for-peer-review-d5f73f0fae07}).
 An explicit segmentation of research steps could even make the focus
@@
 reproducible research initiatives such as
-ReproHack\footnote{\url{https://reprohack.github.io/reprohack-hq/}},
-ReproducibiliTea\footnote{\url{https://reproducibilitea.org/}}
-\cite{fitzgibbon_brewing_2020}, and
-Repro4Everyone\footnote{\url{https://repro4everyone.org/}}
-\cite{auer_reproducibility_2020}.  The initial reaction of the
+ReproHack (\url{https://reprohack.github.io/reprohack-hq/}),
+ReproducibiliTea (\url{https://reproducibilitea.org/},
+\cite{fitzgibbon_brewing_2020}), and
+Repro4Everyone (\url{https://repro4everyone.org/},
+\cite{auer_reproducibility_2020}).  The initial reaction of the
 scientific community to these ideas shows that scholarly peer review
@@
 execution of computational workflows.
-
-\newpage
 
@@
 \begin{itemize}
-  \item `register.csv`. List of all CODECHECK certificates with references to repositories and reports.
+  \item \texttt{register.csv}. List of all CODECHECK certificates with references to repositories and reports.
 \end{itemize}

### 2021-02-09 codecheck-paper.tex
switch CI to latex image

@@
 and its discussions, code repositories, and the main register of
-CODECHECKs. Both the project website \url{https://codecheck.org.uk/}}
+CODECHECKs. Both the project website \url{https://codecheck.org.uk/}
 and the register at \url{https://codecheck.org.uk/register}
@@
 
-Repository: Daniel Nüst, Stephen Eglen, & Iain Davies.
+Repository: Daniel Nüst, Stephen Eglen, \& Iain Davies.
 (2021).

### 2021-02-09 codecheck-paper.tex
consistent formatting

@@
 
-\textbf{Isn't CODECHECK what peer review should be doing already?} \\
+\textbf{Isn't CODECHECK what peer review should be doing already?}
 On the surface, yes, but peer reviewers are overburdened enough and
@@
 
-\textbf{Should CODECHECK requirements be more demanding?} \\
+\textbf{Should CODECHECK requirements be more demanding?}
 CODECHECK by design does not require authors to provide (and sustain)
@@
 \textbf{How are failures during checks handled?}
-
 We have not yet established a process for if a reproduction fails.  This
@@
 \textbf{What does it mean for a figure to be reproducible?}
-
 Automatically detecting if a codechecker's results are ``the same'' as

### 2021-02-09 codecheck-paper.tex
try to clarify

@@
 publicly reporting failures (i.e.,~the code would not run) will
-increase overall reproducibility, and we recommend sharing the latest
-draft of the certificate only with the authors.  One must be careful
-how policies on disclosing failures may \textbf{negatively impact the
-  acceptance by authors}
-%% SJE: what does this mean?
-, and also how volunteer efforts may go unnoticed if no report is
+increase overall reproducibility, and it may prohibit authors from
+sharing their best efforts, which is always more desirable than
+nothing shared. Therefore, we recommend sharing the latest
+draft of the certificate only with the authors, even if that
+means that volunteer efforts may go unnoticed if no report is
 published.  The most appropriate way forwards seems to be to give
-private feedback for failures and public acknowledgment and
-encouragement for successes.
+private feedback for failures, public acknowledgment and
+encouragement for successes, and recognise codecheckers as reviewers.
 % eLife's new model (see https://osf.io/9cftx/) will publish reports eventually, e.g.,~after preprint has been accepted elsewhere

### 2021-02-23 codecheck-paper.tex
interim progress

@@
 \begin{document}
-%\title{CODECHECK: An open-science initiative for the independent execution of computations underlying research articles}
-\title{CODECHECK: An Open Science initiative for the independent execution of computations underlying research articles during peer review improves reproducibility}
+\title{CODECHECK: An Open Science initiative for the independent
+  execution of computations underlying research articles during peer
+  review to  improve reproducibility}
 \author[1,$\ast$]{Daniel~N\"{u}st \orcid{0000-0002-0024-5046} \github{nuest} \twitter{nordholmen}}
@@
   The traditional scientific paper falls short of effectively communicating computational research.
-  Since this issue is difficult to address due to lacking incentives, guidance, benefits, recognitions, and reproducibility, we come at this problem from a different angle:
-  We propose a system by which the computational workflows underlying research articles are checked.
+ To help improve this situation, we propose a system by which the computational workflows underlying research articles are checked.
   The CODECHECK system uses open infrastructure and tools and can be integrated into review and publication processes in multiple ways.
@@
 
-Many areas of scientific research use computations to simulate or analyse their data. These computations are complex and are difficult to explain coherently in a paper \citep{marwick_how_2015}.
-To complement the traditional route of sharing research by writing papers,
-there is a growing demand to share the underlying artefacts, notably 
-code and datasets, so that others can inspect, reproduce or expand that work
-(see Figure~\ref{fig:inverse}).
-Some of the earliest proponents of this initiative were Buckheit and Donoho \cite{buckheit_wavelab_1995}, who coined what has been called \emph{Claerbout's claim} (extending on \citet{claerbout_electronic_1992}):
-\emph{``An article about computational science in a scientific publication 
-is \text{not} the scholarship itself, it is merely \textbf{advertising} of
-the scholarship. The actual scholarship is the complete software development
-environment and the complete set of instructions which generated the 
-figures.''}
-
-If researchers start sharing more artefacts, how might these artefacts be examined to ensure that they do what they claim?
-For example,
+Many areas of scientific research use computations to simulate or
+analyse their data. These complex computations are difficult to
+explain coherently in a paper \citep{marwick_how_2015}.  To complement
+the traditional route of sharing research by writing papers, there is
+a growing demand to share the underlying artefacts, notably code and
+datasets, so that others can inspect, reproduce or expand that work
+(see Figure~\ref{fig:inverse}).  Early proponents of this initiative
+were Buckheit and Donoho
+\cite{buckheit_wavelab_1995,claerbout_electronic_1992}, who noted:
+\emph{``An article about computational science in a scientific
+  publication is \text{not} the scholarship itself, it is merely
+  \textbf{advertising} of the scholarship. The actual scholarship is
+  the complete software development environment and the complete set
+  of instructions which generated the figures.''}
+
+If researchers start sharing more artefacts, how might these artefacts
+be examined to ensure that they do what they claim?  For example,
 although scientific journals now require a data sharing statement that
-outlines what data the authors have (or will) share, journals  implement this differently.
-At one end of the spectrum, journals have been created to accept ``data papers''
+outlines what data the authors have (or will) share, journals
+implement this differently.  On one hand, journals have
+been created to accept ``data papers''
 (e.g.,~\href{https://www.nature.com/sdata/}{\emph{Scientific~Data}},
 \href{https://essd.copernicus.org/}{\emph{Earth System Science Data}},
-\href{https://rmets.onlinelibrary.wiley.com/journal/20496060}{\emph{Geoscience Data Journal}},
-\href{https://bdj.pensoft.net/}{\emph{Biodiversity Data Journal}},
-\href{https://openpsychologydata.metajnl.com/}{\emph{Journal of Open Psychology Data}},
-\href{https://odjar.org/}{\emph{Open Data Journal for Agricultural Research}},
-\href{https://openhealthdata.metajnl.com}{\emph{Journal of Open Health Data}});
-these journals have established rigorous procedures by which
-the data is validated according to standards in each field.
-At the other end of the spectrum, many journals still allow authors to state ``Data available upon reasonable request''.
-Authors, while possibly well intentioned at the time of writing the article, often cannot provide data when requested as data disappears over time \cite{Vines2014-hf}.
-
-Given that no clear standards yet exist for sharing data, what hope might there be for sharing computer programs?
-Both data and software are required to validate a
-computational analysis; data can be seen as inert whereas
-code requires an environment to be run in.
-This makes software harder to share.
-Our experience is that researchers offer several reasons for why code
-is not shared, e.g.,~``there is no documentation'', ``I cannot maintain it'', or ``I do not want to give away my code to competitors''.
-Our view is that sharing code, wherever possible, is good for the community and the individual \cite{Barnes2010-iv,markowetz_five_2015}.
-Having code and data openly available and archived provides a valuable
-resource for others to learn from, even if the code is broken or lacks documentation.
-However, with a little effort, we believe that if an independent person can execute an analysis, this is worth documenting and that this  reduces the barrier to evaluating non-text research materials.
-Just as data journals' validations of data and all journals' peer review  provides a ``baseline reassurance'', i.e.,~that a paper has been checked by someone with an understanding of the topic \cite{fyfe_mission_2019}, the same baseline could be provided for the workflow underlying a paper.
-With this in mind, we have developed a set of principles and an example workflow that provides a pragmatic way of checking that a paper's code works, i.e., it is reproducible following the Claerbout/Donoho/Peng terminology \cite{barba_terminologies_2018}.
-
-Here we offer a thorough description of a process and its variations to integrate a much-needed evaluation of computational reproducibility into peer review, and we demonstrate its feasibility by means of 25 reproductions across scientific disciplines.
-We call this system CODECHECK.
+\href{https://rmets.onlinelibrary.wiley.com/journal/20496060}{\emph{Geoscience
+    Data Journal}}, \href{https://bdj.pensoft.net/}{\emph{Biodiversity
+    Data Journal}},
+\href{https://openpsychologydata.metajnl.com/}{\emph{Journal of Open
+    Psychology Data}}, \href{https://odjar.org/}{\emph{Open Data
+    Journal for Agricultural Research}},
+\href{https://openhealthdata.metajnl.com}{\emph{Journal of Open Health
+    Data}}); these journals have established rigorous procedures by
+which data are validated according to standards in each field.  On
+the other hand, many journals still allow authors to
+state ``Data available upon reasonable request''.  Authors, while
+possibly well intentioned at the time of writing the article, often
+cannot provide data when requested as data disappears over time
+\cite{Vines2014-hf}.
+
+Given that data are not routinely shared, what hope might there be for
+sharing computer programs?  Both data and software are required to
+validate a computational analysis; data can be seen as inert whereas
+code requires an environment to be run in.  This makes software harder
+to share.  Our experience is that researchers offer several reasons
+for why code is not shared, e.g.,~``there is no documentation'', ``I
+cannot maintain it'', or ``I do not want to give away my code to
+competitors''.  Our view is that sharing code, wherever possible, is
+good for the community and the individual
+\cite{Barnes2010-iv,markowetz_five_2015}.  Having code and data openly
+available, and archived, provides a valuable resource for others to
+learn from, even if the code is broken or lacks documentation.
+However, with a little effort, we believe that if an independent
+person can re-run the programs, this is worth documenting and that
+this reduces the barrier to evaluating non-text research materials.
+Just as data journals' validations of data and all journals' peer
+review provides a ``baseline reassurance'', i.e.,~that a paper has
+been checked by someone with an understanding of the topic
+\cite{fyfe_mission_2019}, the same baseline could be provided for the
+workflow underlying a paper.  With this in mind, we have developed a
+set of principles and an example workflow that provides a pragmatic
+way of checking that a paper's code works, i.e., it is reproducible
+following the Claerbout/Donoho/Peng terminology
+\cite{barba_terminologies_2018}.
+
+Here we offer a thorough description of a process and its variations
+to integrate a much-needed evaluation of computational reproducibility
+into peer review, and we demonstrate its feasibility by means of 25
+reproductions across scientific disciplines.  We call this system
+CODECHECK.
 
@@
   \caption{The inverse problem in reproducible research. The left
-  half of the diagram shows the diverse range of materials used
+  half of the diagram shows a diverse range of materials used
   within a laboratory. These materials are often then
@@
   ineffective for science as a whole. By sharing the
-  materials on the left, others outside the lab can reproduce
-  or build on this work.}
+  materials on the left, others outside the lab can enhance this work.}
   \label{fig:inverse}
@@
 groups of people:
-(1) the AUTHOR of a paper providing the code to be checked,
-(2) the PUBLISHER of a journal interested in publishing the author's paper, and
-(3) the CODECHECKER, who checks that the author's code works.
-The six-step workflow we have refined is shown in Figure~\ref{fig:workflow}.
+(1) the \textbf{author} of a paper providing the code to be checked,
+(2) the \textbf{publisher} of a journal interested in publishing the author's paper, and
+(3) the \textbf{codechecker}, who checks that the author's code works.
+The six-step workflow we have refined is shown in
+Figure~\ref{fig:workflow}.  In this article, we also refer to a
+\textbf{peer-reviewer} who is independent of this process, and
+performs the traditional academic review of the content of an article.
 
@@
       \caption{The CODECHECK example process implementation.
-        Codechecker act as a detectives: They investigate and record,
-        but do not fix issues.  Numbers in bold refer to the six steps
+        Codecheckers act as detectives: They investigate and record,
+        but do not fix issues.  Numbers in bold refer to steps
         outlined in the text.}
@@
 
-\textbf{Step 1:} The author submits their manuscript along with the code and data to the PUBLISHER.
-The code and data need not be openly available at this point.
-However, in many cases the code and data may be published on a code hosting platform,
-such as GitHub or GitLab. Ideally, the AUTHOR is expecting the CODECHECK and
-prepares for it, e.g.,~by asking a colleague to attempt a reproduction or by 
-following the PUBLISHER's guidelines for sharing computational workflows.
+\textbf{Step 1:} The author submits their manuscript along with the
+code and data to the PUBLISHER.  The code and data need not be openly
+available at this point.  However, in many cases the code and data may
+be published on a code hosting platform, such as GitHub or
+GitLab. Ideally, the AUTHOR is expecting the CODECHECK and prepares
+for it, e.g.,~by asking a colleague to attempt a reproduction, and
+providing a set of instructions on how to re-run the workflow.
 
@@
 This process iterates until either the CODECHECKER is successful,
-or the CODECHECKER concludes the code can not be executed within the context of a CODECHECK.
+or the CODECHECKER concludes the workflow is not reproducible.
 As part of this process, the CODECHECKER could work entirely locally, relying on their own computing resources, or in the cloud, e.g.,~using the open MyBinder infrastructure \cite{jupyter_binder_2018} or alternatives, some of which are more tailored to scientific publications while others offer commercial options for, e.g.,~publishers (cf.
@@
 %\footnote{Note this is a different use from that referred to as the \emph{Verification Reports} article type in the journal \emph{Cortex} \cite{chambers_verification_2020}.},
-of computational results, i.e.,~checking that the code correctly produces the output it claims to create, but not a validation, i.e.,~checking that the code implements the right algorithm to solve the specific research problem.
-Nevertheless, simply attempting to reproduce an output may highlight a submission's shortcomings in meeting a journal's requirements (cf.~\cite{christian_journal_2020}) and may effectively increase transparency, thereby improving practices (cf.~\cite{nosek_scientific_2012}) even if the check does not go into every detail.
+of computational results, i.e.,~checking that code generates the
+output it claims to create, but not a validation, i.e.,~checking that
+the code implements the right algorithm to solve the specific research
+problem.  Nevertheless, simply attempting to reproduce an output may
+highlight a submission's shortcomings in meeting a journal's
+requirements (cf.~\cite{christian_journal_2020}) and may effectively
+increase transparency, thereby improving practices
+(cf.~\cite{nosek_scientific_2012}) even if the check does not go into
+every detail.
 
@@
 subsets or helper scripts, and the original code and data get deposited in
-an open archive unless restrictions regarding data size, license or 
-sensitivity apply.
+an open archive unless restrictions (data size, license or 
+sensitivity) apply.
 Currently, the CODECHECKERs deposit the material on Zenodo themselves, but the PUBLISHER may complete this step after integrating CODECHECK into its review process.
@@
 link to the certificate.
-Although a badge simplifies the CODECHECK into a binary value and risks introducing confusion regarding the extent of the check, especially across disciplines and methods, a badge can provide recognition value and acknowledgement of the completed CODECHECK.
-The badge and the actual check are incentives for undertaking the effort needed to provide a readily computationally reproducible workflow.
+Although a badge simplifies the CODECHECK into a binary value and risks introducing confusion regarding the extent of the check, especially across disciplines and methods, a badge can provide recognition value and acknowledges the completed CODECHECK.
+The badge and the actual check are incentives for undertaking the effort needed to provide a reproducible workflow.
 
 \textbf{Step 6:} The PUBLISHER can, depending on the timing, provide the
-certificate to peer-reviewers or handling editors or publish it
+certificate to peer-reviewers or editors or publish it
 and link between  certificate, paper, and any repositories 
@@
 The PUBLISHER also credits the CODECHECKER for their work by depositing
-the activity appropriately in scholarly profiles, such as ORCID
+the activity in scholarly profiles, such as ORCID
 (see peer review contributions in ORCID records,
@@
 
-The example workflow leaves room for variations to this process.
-Here we consider several dimensions in a space of possible CODECHECK workflows(Figure~\ref{fig:dimensions}), to represent relevant aspects for stakeholders.
-These aspects touch on timing, responsibilities, and 
-transparency and will be discussed in the following sections.
-The options on each dimension have their own pros and cons and can,
-for the most part, be freely combined.
+Our workflow is just one of many possibilities of a CODECHECK
+workflow.  Here we consider several dimensions in a space of possible
+CODECHECK workflows(Figure~\ref{fig:dimensions}).  These aspects touch
+on timing, responsibilities, and transparency.
 
@@
 As this can reduce the workload of reviewers, such a check should occur early in the review process.
-Yet, the later in the review process the check happens, the easier is it to allow bidirectional communication between the author and codechecker, e.g., because the author might already be notified of the paper's acceptance and may more freely share materials online closer to the paper's publication date.
-A \textbf{pre-review} CODECHECK means editors could send a submission
+Yet, the later in the review process the check happens, the easier is
+it to allow bidirectional communication between the author and
+codechecker, e.g., because the author might already be notified of the
+paper's acceptance and may be more willing to share materials online closer to the paper's publication date.
+A \textbf{pre-review} CODECHECK means editors would send a submission
 for peer review only if it passes the check, or include the certificate in the submission package provided to peer-reviewers.
-Reviewers may then judge the relevance of the computations for the results of the work, or follow journal guidelines regarding how strictly to apply the check's outcomes in their evaluation.
-
-A CODECHECK may also be conducted in \textbf{parallel} to the scientific peer review.
-This puts less burden on the turnaround time for the CODECHECK, yet it only makes the outcomes available during the final consideration by the handling editor.
-The check could also be assigned after suggestion by a reviewer, which would remove the need for submissions to undergo a pre-review screening.
-However, soliciting such a ``specialist review'' is much less
-desirable than having a regular CODECHECK, thus avoiding the situation
-in which some submissions get special treatment. In both cases, the
-editor's decision could be based both on CODECHECK and peer-review reports.
+Peer-reviewers may then judge the relevance of the computations for the results of the work.
+
+A CODECHECK may also be conducted in \textbf{parallel} to the academic
+peer review.  This puts less burden on the turnaround time for the
+CODECHECK, yet it only makes the outcomes available during the final
+consideration by the handling editor.  The check could also be
+assigned after suggestion by a reviewer, which would remove the need
+for submissions to undergo a pre-review screening.  However,
+soliciting such a ``specialist review'' is much less desirable than
+having a regular CODECHECK, thus avoiding the situation in which some
+submissions get special treatment. In both cases, the editor's
+decision could be based both on CODECHECK and peer-review reports.
 
@@
 computationally-competent person. Naturally, the more
-prerequisite knowledge the codechecker has, the less time is needed to
-understand the goals and the mechanics of an analysis.  From our
+prerequisite knowledge the codechecker has, the quicker they can
+understand the goals and mechanics of an analysis.  From our
 experiences, the priority should be given to matching technical
@@
 problem than assessing the outcome, e.g.,~comparing created figures
-with the original, without an in-depth understanding of the science.
+with the original, without an in-depth understanding of the domain.
 The depth of the check will mostly be driven by the time required and
@@
 
-We also see a great opportunity to involve \emph{early-career researchers} (ECRs) as codecheckers.
-ECRs arguably have a high interest in learning
-about the latest tools and technologies, as they build up their own
-expertise and specialisation.
-CODECHECK offers a way for ECRs to gain insights into new research and
-highlight the importance of reproduction.  
-\emph{ReScience X}, a journal devoted to reproduction and replication experiments \cite{roesch_new_2020}, shares an interest in this combination.
-ECRs are also often familiar with new technologies, thus also making them likely to author CODECHECK-ready manuscripts.
-A supporting data point for ECRs as early adopters is that they are responsible for 77\% of 141 registered reports that were submitted to leading Neuroscience \cite{chambers_registered_2019}.
-As ECRs are introduced to peer review as codecheckers, they may transition into the role of scientific reviewer over time.
-Overall, we see several opportunities and benefits to setting up a new process for CODECHECKING with a clear commitment to openness and transparency, independent of the current peer review process (see \emph{Openness} dimension).
+We also see a great opportunity to involve \emph{early-career
+  researchers} (ECRs) as codecheckers.  ECRs arguably have a high
+interest in learning about new tools and technologies, to build up
+their own expertise.  CODECHECK offers a way for ECRs to gain insights
+into new research and highlight the importance of reproduction.
+\emph{ReScience X}, a journal devoted to reproduction and replication
+experiments \cite{roesch_new_2020}, shares an interest in this
+combination.  ECRs are also often familiar with new technologies, thus
+also making them likely to author CODECHECK-ready manuscripts.  A
+supporting data point for ECRs as early adopters is that they are
+responsible for 77\% of 141 registered reports that were submitted
+\cite{chambers_registered_2019}.  As ECRs are introduced to peer
+review as codecheckers, they may transition into the role of
+peer-reviewer over time.  Overall, we see several opportunities
+and benefits to setting up a new process for CODECHECKING with a clear
+commitment to openness and transparency, independent of the current
+peer review process (see \emph{Openness} dimension).
 
@@
 than a private in-house group, as this may allow for better matches
-regarding expertise and more even sharing of the workload.  Overall,
-this community will aim to ensure that CODECHECK is a viable option
+regarding expertise and workload sharing.  
+This community can establish CODECHECK as a viable option
 for independent no-cost Open Access journals.

### 2021-02-23 codecheck-paper.tex
about to tackle the TOMS bit

@@
 tells the author. We believe that the job of the codechecker is not to
-fix all these problems but simply to report them and await a fix from
-the author.  The level of documentation that is needed for third
-parties to reproduce a workflow is hard to get right, and
-too often this uncertainty leads researchers to give up and not
-document it at all.  The conversation with a codechecker fixes this
-problem.
+fix these problems but simply to report them to the author and await a
+fix.  The level of documentation required for third parties to
+reproduce a workflow is hard to get right, and too often this
+uncertainty leads researchers to give up and not document it at all.
+The conversation with a codechecker fixes this problem.
 
 \textbf{2. Communication between humans is key.} \\
-Some code may
-work without any interaction (e.g., CODECHECK certificate \texttt{2020-013}
-\cite{cert-2020-013}) but more often there are hidden
-dependencies or installations that need adjusting for a particular system.
-Allowing the (human) codechecker to communicate
-directly and openly with each other should make this process as
-constructive as possible. We feel that routing this conversation
-(possibly anonymously) through a publisher would introduce delays and
-reduce the chances for community building.
+Some code may work without any interaction, e.g. \cite{cert-2020-013},
+but often there are hidden dependencies that need adjusting for a
+particular system.  Allowing the codechecker to communicate directly
+and openly with the author make this process as constructive as
+possible; routing this conversation (possibly anonymously) through a
+publisher would introduce delays and inhibit community building.
 
@@
 The value of performing a CODECHECK is comparable to that of a peer
-review, and it may also require a similar amount of time. Therefore,
-the codechecker's activity should be recorded, ideally in the
-published paper.  The public record can be realised by publishing the
-certificate in a citable form (i.e.,~with a DOI, as conducted in the
-community CODECHECK process), by listing codecheckers on the journal's
-website or, ideally, by publishing the checks alongside peer review
-activities in public databases.
+review, and it may require a similar amount of time. Therefore, the
+codechecker's activity should be recorded, ideally in the published
+paper.  The public record can be realised by publishing the
+certificate in a citable form (i.e.,~with a DOI), by
+listing codecheckers on the journal's website or, ideally, by
+publishing the checks alongside peer review activities in public
+databases.
 
@@
 ``preproducibility'' and the ICERM report \cite{stodden_setting_2013}
-defines the level ``Auditable Research'' similarly.  Every community
-may establish their own good practices or adapt generic concepts and
+defines the level ``Auditable Research'' similarly.  Communities can
+establish their own good practices or adapt generic concepts and
 practical tools, such as publishing all building blocks of science in
 a research compendium (cf.~\url{https://research-compendium.science/})
-or ``repro-pack'' \cite{barba_praxis_2018}.
-A completed
-check means that code could be executed at least once using the
-provided instructions, and, therefore, all code and data was given and
-could be investigated more deeply or extended in the future.  Ideally,
-this is a ``one~click'' step, but achieving this requires particular
-skills and a sufficient level of documentation for third
+or ``repro-pack'' \cite{barba_praxis_2018}.  A completed check means
+that code could be executed at least once using the provided
+instructions, and, therefore, all code and data was given and could be
+investigated more deeply or extended in the future.  Ideally, this is
+a ``one~click'' step, but achieving this requires particular skills
+and a sufficient level of documentation for third
 parties. Furthermore, automation may lead to people gaming the system
@@
 in favour of bitwise reproducibility because, in the long run, it can
-be automated (e.g., Konrad Hinsen on Twitter,
+be automated (e.g.,
 \url{https://twitter.com/khinsen/status/1242842759733665799}), but
@@
 \textbf{5. Open by default and transitional by disposition.} \\
-By default, unless there are strong reasons to the contrary
+Unless there are strong reasons to the contrary
 (e.g.,~sensitive data on human subjects), all code and data, both from
-author and codechecker, will be made freely available before or when
+author and codechecker, will be made freely available when
 the certificate is published.  Openness is not required for the paper
-itself, to accommodate journals in their transition to sustainable
+itself, to accommodate journals in their transition to
 Open Access models.  The code and data publication should follow
@@
 
-We have a curated list of 25 certificates available at
-\url{https://codecheck.org.uk/register}.
-Table~\ref{tab:register} lists the certificates
-along with citations to the original papers.
-These CODECHECKs fall into three themes:
-(1) classic and current papers from computational neuroscience,
-(2) COVID-19 modelling preprints, and
-(3) GIScience.
-
-The first theme was the initial set of papers used to explore the
-concept of CODECHECK.  The idea was to take very well known historical
-articles from a domains of interest to one of the co-authors (SJE).
-Our first CODECHECK (certificate number \texttt{2020-001}) was
-performed before publication on an article for the journal
-\emph{GigaScience}, which visusalized the outputs from a family of
-supervised classification algorithms.
-
-The second theme was a response to the COVID-19 pandemic, choosing
-modelling papers to predict outcomes. The checks were solicited
-through community interaction or by our initiative rather than
-requested from journals.  Some certificates were since acknowledged in
-the accepted papers
-\cite{Davies2020-vj,kucharski_effectiveness_2020}. In particular, we
-codechecked the well-known Imperial college model of UK lockdown
-procedures from March 2020(certificate \texttt{2020-10}).  Contrary to
-opinion in the mainstream media, the model results were reproducible
-(CITE NOORDEN).
+We have a curated list of 25 certificates (Table~\ref{tab:register}).
+These CODECHECKs fall into three themes: (1) classic and current
+papers from computational neuroscience, (2) COVID-19 modelling
+preprints, and (3) GIScience.
+
+The first theme was an initial set of papers used to explore the
+concept of CODECHECK.  The idea was to take well-known articles from a
+domain of interest (Neuroscience).  Our first CODECHECK (certificate
+number \texttt{2020-001}) was performed before publication on an
+article for the journal \emph{GigaScience}, which visusalized the
+outputs from a family of supervised classification algorithms.
+
+The second theme was a response to the COVID-19 pandemic, selecting
+papers that predicted outcomes. The checks were solicited through
+community interaction or by our initiative rather than requested from
+journals.  Some certificates were since acknowledged in the accepted
+papers \cite{Davies2020-vj,kucharski_effectiveness_2020}. In
+particular, we codechecked the well-known Imperial college model of UK
+lockdown procedures from March 2020,
+demonstrating that the model results were reproducible
+\cite{Chawla2020-hi,cert-2020-010}.
 
@@
 
-At the end of each CODECHECK, the codechecker writes a certificate
+After running the workflow, the codechecker writes a certificate
 stating which outputs from the original article, i.e.,~numbers,
 figures or tables, could be reproduced.  This certificate is made
-openly available so that all readers of the article can see which
+openly available so that everyone can see which
 elements were reproduced and what limitations or issues were found.
@@
 allowing others to build on the work.  The format of the certificates
-evolved during the project, as we have learnt to automate different
-aspects of the certification, such as inserting figures into
-certificate documents and managing metadata of a check.  The metadata
+evolved during the project, as we learnt to automate different
+aspects of the certification.  The metadata
 is stored in a machine-readable structured file in YAML, the CODECHECK
@@
 
-Figure~\ref{fig:cert} shows pages 1--4 (of 10) of a typical
-certificate \texttt{2020-012} \cite{cert-2020-012}.  The referenced
-article \cite{unwin_report_2020} described models of COVID-19 across
-the USA.  Figure~\ref{fig:cert}A shows the certificate number and its
-DOI, which points to the certificate and any supplemental files on
-Zenodo.  The CODECHECK logo is added for recognition and to denote
-successful reproduction.  Figure~\ref{fig:cert}B provides the key
-metadata as a table generated from \texttt{codecheck.yml}; it names
-the paper that was checked (title, DOI), the authors, the codechecker,
-when the check was performed, and in which repository related material
-is accessible.  Figure~\ref{fig:cert}C shows a textual summary of how
-the CODECHECK was performed and any interesting findings.
-Figure~\ref{fig:cert}D (page 2 of the certificate) shows the outputs
-that were generated based on the MANIFEST of output files in the
-CODECHECK.  It shows the file name (Output), the description stating
-to which figure/table each file should be compared in the original
-paper (Comment), and the file size.  Page 3 of the certificate,
-(Figure~\ref{fig:cert}E) gives detailed notes from the codechecker,
-here documenting what steps were needed to run the code and that the
-code took about 17~hours to complete. Finally, page 4 of the
-certificate shows the first output generated by the CODECHECK
-(Figure~\ref{fig:cert}F). In this case, the figure matched figure~4 of
-\cite{unwin_report_2020}.  The remaining pages of the certificate show
-other outputs and the computing environment in which the certificate
-itself was created (not shown here).
+Figure~\ref{fig:cert} shows pages 1--4 (of 10) of an example
+certificate to check predictions of COVID-19 spread across the USA
+\cite{cert-2020-012,unwin_report_2020}.  Figure~\ref{fig:cert}A shows
+the certificate number and its DOI, which points to the certificate
+and any supplemental files on Zenodo.  The CODECHECK logo is added for
+recognition and to denote successful reproduction.
+Figure~\ref{fig:cert}B provides the key metadata extracted from
+\texttt{codecheck.yml}; it names the paper that was checked (title,
+DOI), the authors, the codechecker, when the check was performed, and
+where code/data are available.
+Figure~\ref{fig:cert}C shows a textual summary of how the CODECHECK
+was performed and key findings.  Figure~\ref{fig:cert}D
+(page 2 of the certificate) shows the outputs that were generated
+based on the MANIFEST of output files in the CODECHECK.  It shows the
+file name (Output), the description stating to which figure/table each
+file should be compared in the original paper (Comment), and the file
+size.  Page 3 of the certificate, (Figure~\ref{fig:cert}E) gives
+detailed notes from the codechecker, here documenting what steps were
+needed to run the code and that the code took about 17~hours to
+complete. Finally, page 4 of the certificate shows the first output
+generated by the CODECHECK (Figure~\ref{fig:cert}F). In this case, the
+figure matched figure~4 of \cite{unwin_report_2020}.  The remaining
+pages of the certificate show other outputs and the computing
+environment in which the certificate itself was created (not shown
+here).
 
@@
 We use freely available infrastructure, GitHub and Zenodo, to run our
-system.  The \texttt{codecheckers} \textbf{GitHub}
-organisation at \url{https://github.com/codecheckers} contains
-projects for managing the project website, the codecheckers community
-and its discussions, code repositories, and the main register of
-CODECHECKs. Both the project website \url{https://codecheck.org.uk/}
-and the register at \url{https://codecheck.org.uk/register}
-are hosted as GitHub pages.
-The register database is a single table in CSV format that connects the
-certificate identifier with the repository associated with a specific
-conducted CODECHECK. Each of these repositories, which currently can
-be hosted on GitHub or Open Science Framework (OSF), contains the
-CODECHECK metadata file \texttt{codecheck.yml}. The register further
-contains a column for the type of check, e.g.,~community, journal, or
-conference, and the respective GitHub issue where communications and
-assignments around a specific check are organised. No information is
-duplicated between the register and the metadata files. The continuous
+system.  The \texttt{codecheckers} \textbf{GitHub} organisation at
+\url{https://github.com/codecheckers} contains projects for managing
+the project website, the codecheckers community and its discussions,
+code repositories, and the main register of CODECHECKs. Both the
+project website \url{https://codecheck.org.uk/} and the register at
+\url{https://codecheck.org.uk/register} are hosted as GitHub pages.
+The register database is a single table in CSV format that connects
+the certificate identifier with the repository associated with a
+CODECHECK. Each of these repositories, which currently can be hosted
+on GitHub or Open Science Framework (OSF), contains the CODECHECK
+metadata file \texttt{codecheck.yml}. The register further contains a
+column for the type of check, e.g.,~community, journal, or conference,
+and the respective GitHub issue where communications and assignments
+around a specific check are organised. No information is duplicated
+between the register and the metadata files. The continuous
 integration infrastructure of GitHub, GitHub Actions, is used to
 automate the register for viewing (HTML) and for integration with
-other services (CSV, JSON).  \textbf{Zenodo} is our preferred open repository for
-storing certificates. It mints DOIs for deposits and ensures
-long-term availability of all digital artefacts related to the
+other services (CSV, JSON).  \textbf{Zenodo} is our preferred open
+repository for storing certificates. It mints DOIs for deposits and
+ensures long-term availability of all digital artefacts related to the
 project. The CODECHECK community on Zenodo is available at
@@
 managed without infrastructure costs on freely available platforms.
-The automation should ensure that more checks can be conducted in the
-future, though having the process be managed by a paid CODECHECK
-editor or further automation, e.g.,~with a bot, to reduce turnaround
-time. Our main resource requirements are the \textbf{humans} needed for managing
-the project and processes and the codecheckers.  All contributions
+Automation of some aspects may improve turnaround time. Our main
+resource requirements are the \textbf{humans} needed for managing the
+project and processes and the codecheckers.  All contributions
 currently rely on (partly grant-based) public funding and

### 2021-02-23 codecheck-paper.tex
another round of edits completed.

@@
 \textbf{Step 1:} The author submits their manuscript along with the
-code and data to the PUBLISHER.  The code and data need not be openly
+code and data to the publisher.  The code and data need not be openly
 available at this point.  However, in many cases the code and data may
 be published on a code hosting platform, such as GitHub or
-GitLab. Ideally, the AUTHOR is expecting the CODECHECK and prepares
+GitLab. Ideally, the author is expecting the CODECHECK and prepares
 for it, e.g.,~by asking a colleague to attempt a reproduction, and
@@
 
-\textbf{Step 2:} The publisher finds a CODECHECKER to check the
+\textbf{Step 2:} The publisher finds a codechecker to check the
 code. This is analogous to the publisher finding one or more
 peer-reviewers to evaluate the paper, except we suggest that the
-CODECHECKER and the author talk directly to each other.
-
-\textbf{Step 3:} The CODECHECKER runs the code, based on instructions provided by
-the author. They check if some or all of the results from the paper can be
-reproduced. If there are
-any problems running the code, the CODECHECKER asks the AUTHOR for help,
-updates, or further documentation.
-The burden to provide reproducible material lies with the author.
-The CODECHECKER then tries to run the code again.
-This process iterates until either the CODECHECKER is successful,
-or the CODECHECKER concludes the workflow is not reproducible.
-As part of this process, the CODECHECKER could work entirely locally, relying on their own computing resources, or in the cloud, e.g.,~using the open MyBinder infrastructure \cite{jupyter_binder_2018} or alternatives, some of which are more tailored to scientific publications while others offer commercial options for, e.g.,~publishers (cf.
-\cite{konkol_publishing_2020}).
-A cloud-based infrastructure allows for the CODECHECKER and author to collaboratively improve the code and enforces a complete definition of the computing environment; but, unless secure infrastructure is provided, e.g.,~by the PUBLISHER, this requires the code and data to be published openly online.
-Note that the task of the CODECHECKER is to check the ``mechanics'' of the workflow not to evaluate its correctness. 
-In the context of mathematics, Stodden et~al. \cite{stodden_setting_2013} distinguish between \emph{verification} and \emph{validation};
-following their definition, a CODECHECK ensures verification
+codechecker and the author talk directly to each other.
+
+\textbf{Step 3:} The codechecker runs the code, based on instructions
+provided by the author. They check if some or all of the results from
+the paper can be reproduced. If there are any problems running the
+code, the codechecker asks the author for help, updates, or further
+documentation.  The burden to provide reproducible material lies with
+the author.  The codechecker then tries to run the code again.  This
+process iterates until either the codechecker is successful, or the
+codechecker concludes the workflow is not reproducible.  As part of
+this process, the codechecker could work entirely locally, relying on
+their own computing resources, or in the cloud, e.g.,~using the open
+MyBinder infrastructure \cite{jupyter_binder_2018} or alternatives,
+some of which are more tailored to scientific publications while
+others offer commercial options for, e.g.,~publishers
+(cf. \cite{konkol_publishing_2020}).  A cloud-based infrastructure
+allows for the codechecker and author to collaboratively improve the
+code and enforces a complete definition of the computing environment;
+but, unless secure infrastructure is provided, e.g.,~by the publisher,
+this requires the code and data to be published openly online.  Note
+that the task of the codechecker is to check the ``mechanics'' of the
+workflow not to evaluate its correctness.  In the context of
+mathematics, Stodden et~al. \cite{stodden_setting_2013} distinguish
+between \emph{verification} and \emph{validation}; following their
+definition, a CODECHECK ensures verification
 %\footnote{Note this is a different use from that referred to as the \emph{Verification Reports} article type in the journal \emph{Cortex} \cite{chambers_verification_2020}.},
@@
 
-\textbf{Step 4:} The CODECHECKER writes a certificate stating how the
+\textbf{Step 4:} The codechecker writes a certificate stating how the
 code was run and includes a copy of outputs (e.g.,~data, figures, or
@@
 should be required, such that no authors are excluded, it would be
-futile for the CODECHECKER to use automation or fixed checklists.
-
-\textbf{Step 5:} The certificate and auxiliary files created
-during the check, e.g.,~a specification of a computing environment, data 
-subsets or helper scripts, and the original code and data get deposited in
-an open archive unless restrictions (data size, license or 
-sensitivity) apply.
-Currently, the CODECHECKERs deposit the material on Zenodo themselves, but the PUBLISHER may complete this step after integrating CODECHECK into its review process.
-A badge or other visual aid may be added to the deposit and the paper and
-link to the certificate.
-Although a badge simplifies the CODECHECK into a binary value and risks introducing confusion regarding the extent of the check, especially across disciplines and methods, a badge can provide recognition value and acknowledges the completed CODECHECK.
-The badge and the actual check are incentives for undertaking the effort needed to provide a reproducible workflow.
-
-\textbf{Step 6:} The PUBLISHER can, depending on the timing, provide the
-certificate to peer-reviewers or editors or publish it
-and link between  certificate, paper, and any repositories 
-through metadata. Currently, the CODECHECKER creates these connections on
+futile for the codechecker to use automation or fixed checklists.
+
+\textbf{Step 5:} The certificate and auxiliary files created during
+the check, e.g.,~a specification of a computing environment, data
+subsets or helper scripts, and the original code and data get
+deposited in an open archive unless restrictions (data size, license
+or sensitivity) apply.  Currently, the codecheckers deposit the
+material on Zenodo themselves, but the publisher may complete this
+step after integrating CODECHECK into its review process.  A badge or
+other visual aid may be added to the deposit and the paper and link to
+the certificate.  Although a badge simplifies the CODECHECK into a
+binary value and risks introducing confusion regarding the extent of
+the check, especially across disciplines and methods, a badge can
+provide recognition value and acknowledges the completed CODECHECK.
+The badge and the actual check are incentives for undertaking the
+effort needed to provide a reproducible workflow.
+
+\textbf{Step 6:} The publisher can, depending on the timing, provide
+the certificate to peer-reviewers or editors or publish it and link
+between certificate, paper, and any repositories through
+metadata. Currently, the codechecker creates these connections on
 Zenodo. They appear as links with a relationship type on the Zenodo
 landing page for a record in the right-hand sidebar, e.g., the
-``related identifiers'' and ``alternate identifiers'' 
-of certificate \texttt{2020-025} \cite{cert-2020-025}.
-The PUBLISHER also credits the CODECHECKER for their work by depositing
-the activity in scholarly profiles, such as ORCID
-(see peer review contributions in ORCID records,
+``related identifiers'' and ``alternate identifiers'' of certificate
+\texttt{2020-025} \cite{cert-2020-025}.  The publisher also credits
+the codechecker for their work by depositing the activity in scholarly
+profiles, such as ORCID (see peer review contributions in ORCID
+records,
 \url{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}).
-The PUBLISHER also ensures proper publication metadata, e.g.,~links from the 
-certificate repository to the published paper or the original code repository.
+The publisher also ensures proper publication metadata, e.g.,~links
+from the certificate repository to the published paper or the original
+code repository.
 
@@
 The journal \emph{ACM Transactions on Mathematical Software (TOMS)}
-recently established a ``Replicated Computational Results'' (RCR) review process
-\cite{heroux_editorial_2015}, where ``replicable'' is the same as our
-use of ``reproducible''.
-
-Fifteen RCR Reports have been published so far (search on 
-\url{https://search.crossref.org/} with the term
-\texttt{"Replicated Computations Results (RCR) Report"} on 2020-12-10).
+recently established a ``Replicated Computational Results'' (RCR)
+review process \cite{heroux_editorial_2015}, where ``replicable'' is
+the same as our use of ``reproducible''.  Fifteen RCR Reports have
+been published so far (search on \url{https://search.crossref.org/}
+with the term \texttt{"Replicated Computations Results (RCR) Report"}
+on 2020-12-10).
 % 15: https://academic.microsoft.com/search?q=Replicated%20Computations%20Results%20%22(RCR)%20Report%22&f=&orderBy=0&skip=10&take=10
 and the process is being extended extended to the ACM journal
-\emph{Transactions on Modeling and Computer Simulation
-  (TOMACS)} (cf. \url{https://dl.acm.org/journal/tomacs/author-guidelines}).
+\emph{Transactions on Modeling and Computer Simulation (TOMACS)}.
+%% (cf. \url{https://dl.acm.org/journal/tomacs/author-guidelines}).
 The TOMS RCR follows CODECHECK principles 1-4, although our work was
-independently developed of theirs.  Here, we discuss variants of
-CODECHECK, while TOMS RCR realises one particular workflow for
-reproducibility reviews for one particular journal. The TOMS editorial
+independently developed of theirs.  The TOMS editorial
 \cite{heroux_editorial_2015} shares similar concerns about selection
 of reviewers, as we discussed above. Unlike existing CODECHECK
-certificates, the RCR reports, which are similar to CODECHECK
-certificates, undergo editorial review, but this could eventually
-become part of the CODECHECK processes at journals.  The publication
-of the RCR report is seen as a reward for the efforts of the
-reproducing person, while the potential for this motive to be a
-conflict of interest is acknowledged.  TOMS also recognises reviewer
-activity in a partnership with Publons
-(see \url{https://authors.acm.org/author-services/publons}).
-It is unclear if the author of an RCR report receives both recognitions.
-CODECHECK, however, specifically aims to give codecheckers
-recognition as reviewers.  In our view, the reviewer role removes the
-possible conflict of interest while keeping the public
-acknowledgement.  Specific to the field of mathematics, the RCR is
-also expected to apply a review of the software itself if the system
-it runs on cannot be evaluated by an independent party.  The TOMS RCR
-creators concur with the importance of communication, expect
-collaboration between author and RCR reviewers, share the
-considerations around reviewer selection, and also put trust in
-reviewer judgement over numerical bit-wise perfection.  A key
-difference is that for TOMS RCR, authors opt-in with an \emph{RCR
-  Review Request} and the RCR reports are published in the TOMS
-journal next to the actual papers.
-%% SJE: think this could be tightened a bit, and some of it I don't
-%% fully understand.
+certificates, the RCR reports undergo editorial review.  Publication
+of the RCR report recognises the efforts of the reproducing person,
+while the potential for this motive to be a conflict of interest is
+acknowledged.  TOMS also recognises reviewer activity in a partnership
+with Publons (see
+\url{https://authors.acm.org/author-services/publons}).
+
+This activity in the ACM journals can be seen as one possible workflow
+within a CODECHECK system, and clearly shares much in spirit.
+CODECHECK, however, specifically aims to give codecheckers recognition
+as reviewers.  In our view, the reviewer role removes the possible
+conflict of interest while keeping the public acknowledgement.
+Specific to the field of mathematics, the RCR is also expected to
+apply a review of the software itself if the system it runs on cannot
+be evaluated by an independent party.  The TOMS RCR creators concur
+with the importance of communication, expect collaboration between
+author and RCR reviewers, share the considerations around reviewer
+selection, and also put trust in reviewer judgement over numerical
+bit-wise perfection.  A key difference is that for TOMS RCR, authors
+opt-in with an \emph{RCR Review Request} and the RCR reports are
+published in the TOMS journal next to the actual papers.
 
@@
 
-Going beyond article types and a specific journal, the journal
-\emph{ReScience C} (\url{https://rescience.github.io/}) publishes only
-replications, also
+Going beyond individual articles, the journal \emph{ReScience C}
+(\url{https://rescience.github.io/}) publishes only replications, also
 requiring open code and replication by a third party. ReScience also
@@
 
-For research with very high stakes, where reproduction would be too weak and
+For research with high stakes, where reproduction would be too weak and
 post-publication replication possibly too late because of policy impact,
@@
 A workflow that has undergone internal
-replication would very likely be of high quality and relatively easy to check.
+replication would likely be of high quality and relatively easy to check.
 Similarly, internal CODECHECKs may be used, with the same limitations such
@@
 and since its creation in 2011 we are not aware of any publisher using
-it; also, the system is not open source itself.  In comparison,
+it; also, the system is not open source.  In comparison,
 CODECHECK is less powerful but also much more flexible and less
@@
 
-\textbf{Isn't CODECHECK what peer review should be doing already?}
-On the surface, yes, but peer reviewers are overburdened enough and
+\textbf{Isn't CODECHECK what peer review should be doing already?}  On
+the surface, yes, but peer reviewers are overburdened enough and
 asking them to do more work around peer review is not likely to
-succeed.  Miyakawa described how when he as an editor simply requested
-raw data from authors during submission, this led to almost all of
-those papers in which data was requested to be withdrawn or rejected
-after peer review \cite{miyakawa_no_2020}.  Such basic checks to show
-that sufficient data supports the results requires effort from
+succeed.  When an editor (Tsuyoshi Miyakawa) requested raw data from
+n=41 authors before reviewing, 21 authors withdrew their manuscripts;
+19 of the 20 remaining articles were rejected after peer review
+\cite{miyakawa_no_2020}.  Such basic checks require effort from
 editors, yet they only rely on the availability of data files and the
@@
 This might fall within idealistic expectations of peer review, but is
-rare.  Establishing a CODECHECK process is largely an acknowledgement
-that peer reviewing practices have been unable to adapt to the
-challenges of computational papers.  The concept of a CODECHECK, just
-as the concepts of reproducible research and Open Science, may be
+rare.  Establishing a CODECHECK process acknowledges that peer
+reviewing practices have been unable to adapt to the challenges of
+computational papers.  The concept of a CODECHECK, just as the
+concepts of reproducible research and Open Science, may be
 transitional by nature. If the activities described here as being part
-of a CODECHECK become the norm for scientific reviews and
-publications, then the initiative will have succeeded. However, the
-rapidly changing technology used in research leads us to believe that
-a separate role of codechecker and the expertise it brings will be
-beneficial in the long term.
+of a CODECHECK are integrated into the publication process the
+initiative will have succeeded.
 
@@
 that anyone can reproduce has been called a fool's
-errand (Douglas Blank on Twitter,
-\url{https://twitter.com/DougBlank/status/1135904909663068165})
+errand (\url{https://twitter.com/DougBlank/status/1135904909663068165})
 and we tend to agree.  However, the package of data, code, and
@@
 of a successful reproduction and the possibility that a workflow can
-be extended by third parties in the future, as long as they have
+be extended by third parties in the future, if they have
 access to suitable resources and matching skill set.  Concrete
@@
 valuable to improve reproducibility and reusability, and their
-proponents even believe they can improve the science
+proponents even believe they can improve the research
 \cite{petre_code_2014}.  Code reviews, however, have quite different
@@
 conduct software reviews independent of a specific publication or
-workflow include ROpenSci (\url{https://ropensci.org/}),
-PyOpenSci (\url{https://www.pyopensci.org/}), and
-JOSS (\url{https://joss.theoj.org/}).  Furthermore, the
-codechecker's task list is intentionally not overloaded with related
-concerns, albeit potentially important ones, such as ensuring proper
-citation of data and software or depositing data and software in
-suitable repositories. Nevertheless, we expect codecheckers to often
-recognise these issues and highlight them.
+workflow include ROpenSci (\url{https://ropensci.org/}), PyOpenSci
+(\url{https://www.pyopensci.org/}), and JOSS
+(\url{https://joss.theoj.org/}).  Furthermore, the codechecker's task
+list is intentionally not overloaded with related issues such as
+ensuring proper citation of data and software or depositing material
+in suitable repositories. Nevertheless, codecheckers are free to
+highlight these issues.
 
 \textbf{How are failures during checks handled?}
-We have not yet established a process for if a reproduction fails.  This
-may simply reflect our biases in pre-selection of case studies.  In
-the case that a journal adopts CODECHECK for all submissions, the open
-question remains as what to do if a check fails, after exhausting
-efforts between author and codechecker to reproduce the workflow.  A
-negative comment in a CODECHECK certificate or a failed check does not
-necessarily mean the paper or the science is bad (cf. discussion on
-negative comments in \cite{everythinghertz123}).  We doubt that
-publicly reporting failures (i.e.,~the code would not run) will
-increase overall reproducibility, and it may prohibit authors from
-sharing their best efforts, which is always more desirable than
-nothing shared. Therefore, we recommend sharing the latest
-draft of the certificate only with the authors, even if that
-means that volunteer efforts may go unnoticed if no report is
-published.  The most appropriate way forwards seems to be to give
-private feedback for failures, public acknowledgment and
-encouragement for successes, and recognise codecheckers as reviewers.
+
+We do not yet have a process for denoting if a reproduction fails, as
+our case-studies were all successful.  In the case that a journal
+adopts CODECHECK for all submissions, the question remains as what to
+do if a check fails, after exhausting efforts between author and
+codechecker to reproduce the workflow.  A negative comment in a
+CODECHECK certificate or a failed check does not necessarily mean the
+paper or research is bad (cf. discussion on negative comments in
+\cite{everythinghertz123}).  We doubt that publicly reporting failures
+(i.e.,~the code would not run) will increase overall reproducibility,
+and may prohibit authors from sharing their work, which is always more
+desirable than nothing shared. Therefore, we recommend sharing interim
+reproduction efforts only with the authors, even if that means that volunteer
+efforts may go unnoticed if no certificate is published.  
 % eLife's new model (see https://osf.io/9cftx/) will publish reports eventually, e.g.,~after preprint has been accepted elsewhere
-Rosenthal et~al. \cite{Rosenthal2016b} discuss such incentives for different actors around the
-implementation of reproducibility. We see CODECHECK as one way for
-organisations to invest in reproducibility by creating incentives
-until reproducible computations become the norm.
+Rosenthal et~al. \cite{Rosenthal2016b} discuss such incentives for
+different actors around the implementation of reproducibility. We see
+CODECHECK as one way for organisations to invest in reproducibility by
+creating incentives until reproducible computations become the norm.
 
@@
 not that is correct or sound science. This ``mechanical'' test is
-indeed a low bar.  By having the code and data openly deposited, third
+indeed a low bar.  By having code and data openly deposited, third
 parties can later examine the code, and we hope that knowing the code
@@
 process.  ECRs might be attracted to learn more about recent methods,
-peer review, and reproducibility practices.  Research software experts
-(Research Software Engineers,
-RSE, see \url{https://en.wikipedia.org/wiki/Research_software_engineering})
+peer review, and reproducibility practices.  Research Software
+Engineers (RSE) %% , see \url{https://en.wikipedia.org/wiki/Research_software_engineering})
 who might not regularly be involved in writing or reviewing papers
@@
 re-checking if the paper is modified during peer review.  This is
-somewhat unavoidable, and has happened to us (e.g.,~certificate
-\texttt{2020-012} \cite{cert-2020-012}). This can
+inevitable, and happened to us \cite{cert-2020-012}. This can
 also be desired though, if interactions between author, reviewer,
@@
 an author's is more challenging than it might appear.  That is why we
-do not require results to be \textit{identical} for a CODECHECK to pass
-but simply that the code runs and generates output files that the
-author claims. Stochastic simulations mean that often we cannot get
-exactly the same results, and that even the same versions of libraries
-can generate outputs that differ by operating system
+do not require results to be \textit{identical} for a CODECHECK to
+pass but simply that the code runs and generates output files that the
+author claims. Stochastic simulations mean that often we will get
+different results, and even the same versions of libraries can
+generate outputs that differ by operating system
 \cite{Gronenschild2012-pp}.  While reproducibility practices can
@@
 the CODECHECK community
-process, which is documented online at
+process, documented at
 \url{https://codecheck.org.uk/guide/community-process},
-and the volunteer codechecker community, these requirements are
+and the volunteer codechecker community, these requirements can be 
 reconsidered.
 
-We also have deliberated requiring modern technological solutions to
-support reproducibility (cf.~\cite{konkol_publishing_2020}), but we
-decided to focus on the human interface and the judgement of
-experienced researchers and developers as a more sustainable and
-flexible approach.  All types of research can adopt CODECHECK due to
-its flexible design.  CODECHECK could include automated scoring
+We have deliberated requiring modern technologies to support
+reproducibility (cf.~\cite{konkol_publishing_2020}), 
+focusing instead on the human interface and the judgement of experienced
+researchers and developers as a more sustainable and flexible
+approach.  All types of research can adopt CODECHECK due to its
+flexible design.  CODECHECK could include automated scoring
 (e.g.,~\cite{menke_rigor_2020}), yet automation and metrics bear new
@@
 the tensions between commercial publishing and reviewers' often free
-labour, and a global pandemic that has jumbled up scientific
-publishing and exposed a broader general audience to science, even
-preliminary results \cite{munafo_what_2020}.  Establishing CODECHECK
+labour, and a global pandemic that has jumbled up academic
+publishing and exposed a broader general audience to preprints
+\cite{munafo_what_2020}.  Establishing CODECHECK
 workflows must be seen as interconnected with much larger issues in
-science, such as broken metrics or malpractice triggered by
+research, such as broken metrics or malpractice triggered by
 publication pressure
@@
 sub-publication.  The discovery of CODECHECKs could be increased by
-depositing the reports into public databases of reproductions, such as
+depositing certificates into public databases of reproductions, such as
 \emph{SciGen.Report}.  Public researcher profiles, such as ORCID, may
@@
 CODECHECK should be informed by lessons learned from (introducing)
-open peer review (OPR) \cite{ross-hellauer_guidelines_2019}.  Our
+open peer review (OPR) \cite{ross-hellauer_guidelines_2019}.  \textbf{Our
 conversations with publishers and editors included procedural and
@@
 by nor familiar to stakeholders, or that process management software
-would not support such open communication.
+would not support such open communication. WHAT DOES THIS SENTENCE
+MEAN??  JUST DELETE?}
 
@@
 journals to learn more about the advantages and yet unclear specific
-challenges -- e.g do CODECHECKs really work better with OPR?  To
+challenges -- e.g do CODECHECKs really work better with open peer review?  To
 facilitate the adoption, the CODECHECK badge is, intentionally, not
@@
 
-% LocalWords:  CODECHECK CODECHECKER Zenodo ORCID ECRs codecheck codechecker
+% LocalWords:  CODECHECK codechecker Zenodo ORCID ECRs codecheck codechecker
 

### 2021-02-23 codecheck-paper.tex
spell-checked; removed a couple of abbrevs

@@
 
-The time at which a CODECHECK is done and its ascribed importance are closely connected, so we describe the dimensions \emph{When} and \emph{Importance} together.
-The earlier a CODECHECK happens in the publishing process, the more it can affect editorial decisions:
-Is a paper published, sent back for revisions, or rejected?
-Even earlier checks, i.e.,~a CODECHECK of a preprint, may help to improve the workflow itself, even before a publisher is involved.
-As such, codechecking papers could be part of a preprint server's
-policy or initiated by interested authors.
+The time at which a CODECHECK is done and its ascribed importance are
+closely connected, so we describe the dimensions \emph{When} and
+\emph{Importance} together.  The earlier a CODECHECK happens in the
+publishing process, the more it can affect editorial decisions: Is a
+paper published, sent back for revisions, or rejected?  Even earlier
+checks, i.e.,~a CODECHECK of a preprint, may help to improve the
+workflow itself, even before a publisher is involved.  As such,
+codechecking papers could be part of a preprint server's policy or
+initiated by interested authors.
 
@@
 peer-reviewer over time.  Overall, we see several opportunities
-and benefits to setting up a new process for CODECHECKING with a clear
+and benefits to setting up a new process for codechecking with a clear
 commitment to openness and transparency, independent of the current
@@
 Open Access models.  The code and data publication should follow
-community good practices.  Ultimately we may find that CODECHECKS are
+community good practices.  Ultimately we may find that CODECHECK activities are
 subsumed within peer review.
@@
 different terms and infrastructure (``reproducibility reports'' are
-published on the Open Science Framework (OSF) instead of certificates
+published on the Open Science Framework instead of certificates
 on Zenodo) AGILE reproducibility reviews adhere to CODECHECK
@@
 CODECHECK. Each of these repositories, which currently can be hosted
-on GitHub or Open Science Framework (OSF), contains the CODECHECK
+on GitHub or Open Science Framework, contains the CODECHECK
 metadata file \texttt{codecheck.yml}. The register further contains a
@@
 or YARD (\url{https://isps.yale.edu/research/data/approach}),
-or in communities such as Oxford's code review network (OxCRN,
-\url{https://github.com/OxfordCodeReviewNet/forum}).
+or in communities such as Oxford's code review network (\url{https://github.com/OxfordCodeReviewNet/forum}).
 
@@
 reproducibility.  People can report on fully, partially, or
-un-successful reproductions of articles after publication.
+failed reproductions of articles after publication.
 
@@
 
-A codechecker does not fulfill the same role as a \emph{statistical
+A codechecker does not fulfil the same role as a \emph{statistical
   reviewer}, as it is applied by some journals in the biomedical
@@
 forensic photographer, capturing details so that an investigator may
-later scrutinize them.
+later scrutinise them.
 
@@
 peer review, and reproducibility practices.  Research Software
-Engineers (RSE) %% , see \url{https://en.wikipedia.org/wiki/Research_software_engineering})
+Engineers  %% (RSE), see \url{https://en.wikipedia.org/wiki/Research_software_engineering})
 who might not regularly be involved in writing or reviewing papers
@@
 codechecker's contributions, not indirectly via citations of
-reproductions or informal crediation and closed inner circles.
+reproductions or informal credit.
 
@@
 CODECHECK should be informed by lessons learned from (introducing)
-open peer review (OPR) \cite{ross-hellauer_guidelines_2019}.  \textbf{Our
+open peer review \cite{ross-hellauer_guidelines_2019}.  \textbf{Our
 conversations with publishers and editors included procedural and
-technical implications of openness, be it that OPR is neither applied
+technical implications of openness, be it that open peer review is neither applied
 by nor familiar to stakeholders, or that process management software
@@
 % End:
+% LocalWords:  preprint reproducibility compendia

### 2021-02-23 codecheck-paper.tex
few more abbrevs

@@
 
-\section*{Keywords}
+\subsection*{Keywords}
 
@@
 
+\subsection*{Abbreviations}
+
+ACM: Association for Computing Machinery; 
+ECRs: Early Career Researchers;
+RCR: Replicated Computational Results;
+TOMS: Transactions on Mathematical Software;
 \clearpage
@@
 this requires the code and data to be published openly online.  Note
-that the task of the codechecker is to check the ``mechanics'' of the
-workflow not to evaluate its correctness.  In the context of
+that the task of the codechecker is to check only the ``mechanics'' of the
+workflow.  In the context of
 mathematics, Stodden et~al. \cite{stodden_setting_2013} distinguish
@@
 \textbf{Step 4:} The codechecker writes a certificate stating how the
-code was run and includes a copy of outputs (e.g.,~data, figures, or
-tables) that were independently generated.  The certificate may
-include further recommendations on how to improve the material.  The
-free text in the certificate can describe what was really checked,
-because each workflow is unique.  Since no specific tool or platform
-should be required, such that no authors are excluded, it would be
-futile for the codechecker to use automation or fixed checklists.
+code was run and includes a copy of outputs (figures or tables) that
+were independently generated.  The certificate may include
+recommendations on how to improve the material.  The free text in the
+certificate can describe exactly what was checked, because each
+workflow is unique.  Since no specific tool or platform is required,
+such that no authors are excluded, it is futile for the
+codechecker to use automation or fixed checklists.
 
@@
 deposited in an open archive unless restrictions (data size, license
-or sensitivity) apply.  Currently, the codecheckers deposit the
-material on Zenodo themselves, but the publisher may complete this
-step after integrating CODECHECK into its review process.  A badge or
-other visual aid may be added to the deposit and the paper and link to
-the certificate.  Although a badge simplifies the CODECHECK into a
-binary value and risks introducing confusion regarding the extent of
-the check, especially across disciplines and methods, a badge can
-provide recognition value and acknowledges the completed CODECHECK.
-The badge and the actual check are incentives for undertaking the
-effort needed to provide a reproducible workflow.
+or sensitivity) apply.  Currently, codecheckers deposit the material
+on Zenodo themselves, but a publisher may complete this step after
+integrating CODECHECK into its review process.  A badge or other
+visual aid may be added to the deposit and the paper and link to the
+certificate.  Although a badge simplifies the CODECHECK into a binary
+value and risks introducing confusion regarding the extent of the
+check, a badge provides recognition value and acknowledges the
+completed CODECHECK.  The badge and the actual check are incentives
+for undertaking the effort needed to provide a reproducible workflow.
 
@@
 the certificate to peer-reviewers or editors or publish it and link
-between certificate, paper, and any repositories through
-metadata. Currently, the codechecker creates these connections on
-Zenodo. They appear as links with a relationship type on the Zenodo
-landing page for a record in the right-hand sidebar, e.g., the
-``related identifiers'' and ``alternate identifiers'' of certificate
-\texttt{2020-025} \cite{cert-2020-025}.  The publisher also credits
-the codechecker for their work by depositing the activity in scholarly
-profiles, such as ORCID (see peer review contributions in ORCID
-records,
+between certificate, paper, and any repositories. Currently, the
+codechecker creates these connections on Zenodo. They appear as links
+with a relationship type on the Zenodo landing page for a certificate,
+e.g., the ``related identifiers'' and ``alternate identifiers'' of
+certificate \texttt{2020-025} \cite{cert-2020-025}.  The publisher
+also credits the codechecker's work by depositing the activity
+in scholarly profiles, such as ORCID (see peer review contributions in
+ORCID records,
 \url{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}).
@@
 
-We have a curated list of 25 certificates (Table~\ref{tab:register}).
-These CODECHECKs fall into three themes: (1) classic and current
+To date we have created 25 certificates (Table~\ref{tab:register}) 
+falling into three broad themes: (1) classic and current
 papers from computational neuroscience, (2) COVID-19 modelling
@@
 integration infrastructure of GitHub, GitHub Actions, is used to
-automate the register for viewing (HTML) and for integration with
-other services (CSV, JSON).  \textbf{Zenodo} is our preferred open
+automate generation of the register.  \textbf{Zenodo} is our preferred open
 repository for storing certificates. It mints DOIs for deposits and
@@
 These tools and resources demonstrate that a CODECHECK process can be
-managed without infrastructure costs on freely available platforms.
+managed on freely available platforms.
 Automation of some aspects may improve turnaround time. Our main
@@
 and the process is being extended extended to the ACM journal
-\emph{Transactions on Modeling and Computer Simulation (TOMACS)}.
+\emph{Transactions on Modeling and Computer Simulation}.
 %% (cf. \url{https://dl.acm.org/journal/tomacs/author-guidelines}).
@@
 Gavis and Donoho \cite{gavish_universal_2011} propose a new discipline
-and infrastructure for reproducible computational research with its
-own tools and API. Their specific packaging format, provenance record,
-and cryptographic \emph{Verifiable Result Identifier} would indeed
-provide excellent reproducibility. However, the system is also complex
-and since its creation in 2011 we are not aware of any publisher using
-it; also, the system is not open source.  In comparison,
-CODECHECK is less powerful but also much more flexible and less
-dependent on specific tools or infrastructure. If data and code are
-deposited properly, i.e.,~very unlikely to disappear, then the
-certificate's DOI is practically close to the cryptographic
-identifier.
+and infrastructure for reproducible computational research. Their
+specific packaging format, provenance record, and cryptographic
+\emph{Verifiable Result Identifier} would indeed provide excellent
+reproducibility. However, the system is also complex and since its
+creation in 2011 we are not aware of any publisher using it; also, the
+system is not open source.  In comparison, CODECHECK is less powerful
+but also much more flexible and less dependent on specific tools or
+infrastructure. If data and code are deposited properly, i.e.,~very
+unlikely to disappear, then the certificate's DOI is practically close
+to the cryptographic identifier.
 
@@
 Repro4Everyone (\url{https://repro4everyone.org/},
-\cite{auer_reproducibility_2020}).  The initial reaction of the
-scientific community to these ideas shows that scholarly peer review
-should continue on the path towards facilitating the sharing and
+\cite{auer_reproducibility_2020}).  The initial reaction of 
+researchers to these ideas shows that scholarly peer review
+should continue on the path towards facilitating sharing and
 execution of computational workflows.

### 2021-02-23 codecheck-paper.tex
I'm done for now...

@@
 
-Codecheckers could be drawn from a regular pool of volunteer
-\textbf{scientific reviewers}, or from a special group of
+Codecheckers could be drawn from a regular pool of
+\textbf{peer-reviewers}, or from a special group of
 \textbf{reproducibility reviewers} via specific roles such as
 \textbf{reproducibility editors}, or \textbf{editorial staff} with a
-publisher.  A \emph{single codechecker} is sufficient to verify the
-workflow since it is mostly a factual process and does not require much
-debate or opinion.  Code usually harbours systematic and repeatable
-mistakes and is thereby more reliable and auditable than processes
-controlled by humans \cite{tibav:42484}, e.g.,~in a laboratory.  If
-however publication of the paper depends on the CODECHECK, a second
-opinion may be required.
+publisher.  One codechecker is sufficient to verify the workflow since
+it is mostly a factual process.  Code usually harbours systematic and
+repeatable mistakes and is thereby more reliable and auditable than
+processes controlled by humans \cite{tibav:42484}, e.g.,~in a
+laboratory.  If however publication of the paper depends on the
+CODECHECK, a second opinion may be required.
 
@@
 re-checking if the paper is modified during peer review.  This is
-inevitable, and happened to us \cite{cert-2020-012}. This can
-also be desired though, if interactions between author, reviewer,
-and codechecker led to improvements.  Checking the manuscript the
-second time is likely to be much less work than the first time.
+inevitable, and happened to us \cite{cert-2020-012}. This is desirable
+though, if interactions between author, reviewer, and codechecker led
+to improvements.  Checking the manuscript the second time is likely to
+be much less work than the first time.
 
@@
 increase transparency of the checked papers and can contribute to
-building trust in scientific results.
+building trust in research findings.
 % "science earns trust through transparency" > https://twitter.com/briannosek/status/1290240244101521408
@@
 \cite{earthcube_new_2020}, iterative and granular communication of
-scientific outputs, articles with live-code
+research outputs, articles with live-code
 \cite{perkel_pioneering_2019-1} such as eLife's ERA

### 2021-02-24 codecheck-paper.tex
fix

@@
 CODECHECK should be informed by lessons learned from (introducing)
-open peer review \cite{ross-hellauer_guidelines_2019}.  \textbf{Our
-conversations with publishers and editors included procedural and
-technical implications of openness, be it that open peer review is neither applied
-by nor familiar to stakeholders, or that process management software
-would not support such open communication. WHAT DOES THIS SENTENCE
-MEAN??  JUST DELETE?}
+open peer review \cite{ross-hellauer_guidelines_2019}.  Our
+conversations with publishers and editors indicate a willingness to
+adopt open practices like these, but that it is hard to innovate with
+legacy infrastructure and established practices.
 

### 2021-02-24 codecheck-paper.tex
final fixes

@@
 RCR: Replicated Computational Results;
-TOMS: Transactions on Mathematical Software;
+TOMS: Transactions on Mathematical Software.
 \clearpage

### 2021-05-27 reviews/response-round1.md
rename

<https://doi.org/10.5256/f1000research.54932.r82470>

We thank the reviewers for their comments on our manuscript.  The
comments are below.  Indented are our responses.

## Reviewer 1

In this article, authors propose to implement a procedure to check for
the code accompanying a submission to a journal. To do so, they
describe a pipeline made of 6 steps that ultimately lead to the
delivery of a code check certificate meaning that someone external to
the author's lab has managed to re-run the code. At this point, no
checking that the results are correct is necessary. The authors
already issued several codecheck certificates in different
disciplines. I find the idea really nice and certainly necessary but
I've a few questions (even though some of them are already addressed
in the "limitations" section). Given the structure of the paper, I'll
just list my questions here:

1. How does CODECHECK compare to ACM Artifact reviews badges?
(https://www.acm.org/publications/policies/artifact-review-and-badging-current)

    These badges, introduced in August 2020, show whether code is
    available (different levels) and reproduces same results.  In
    principle CODECHECKER could award these badges (artifacts
    evaluated, functional).  We have made a note to this effect in the
    manuscript (Related work, end of paragraph 1.)

2. What would be the incentive for someone to code check the code? Being
aware of the increasing difficulty in finding reviewers, I don't think
it would be easy to recruit people to perform a task that can rapidly
become very technical and time consuming.

    This was addressed in our section "Who's got time for more peer
    review?" We would however note that we have a pool of about 20
    volunteers currently willing to do codechecks.

3. How do you handle the case when specific hardware is necessary
(e.g. NVidia GPU)? Is it documented somewhere such that code-checkers
might first verify if they have the necessary hardware to run the
code?

    This was handled in the limitation "What about my proprietary
    software and sensitive data." but we now mention hardware too in
    the first paragraph of that section.

4. How do you establish a check has failed? For example, what happens if
a code-checker gets a segfault (for some unknown reason) and the
author is unable to help. Is it deemed failed?

    We hope that codechecker and author can resolve problems, but in
    the end there may be problems that cannot be solved.  Open
    infrastructure could help as both author and codechecker can work
    together in the same environment to minimise these failures.
    Ultimately however, there may be failures, which are noted in the
    section "How are failures during checks handled?".

5. Who will pay for the computing resources needed to run heavy
simulations and/or to acquire necessary software such as e.g. Matlab?
When a simulation consumes a lot of resources, it might wise to give
the checker access to computing resources. This could be paid for by
the journal.

    In the section "Who will pay for compute time?" we mention this
	problem, and that toy examples might alleviate the need to re-run
	resource-intensive computations.  We agree that one model might be
	that a journal provide some resource for this service.  Likewise,
	in the following paragraph, we describe that our pragmatic
	approach for now is to find codecheckers that have access to
	particular software, e.g. MATLAB.

6. I did not see in the report example a description of the environment
necessary to run the software. How did you solve the "dependency
hell"? Since the code might break at some point in the future because
of incompatibility in some libraries or environments, it would be
necessary to have a mechanism describing the running environment such
that it can be re-run later.

    The short answer is "we didn't.".  In the paragraph "Should
    CODECHECK requirements be more demanding?" we note our low bar of
    simply getting a codecheck to run once.  We do, however, encourage
    CODECHECKERS to describe the environment in free text form in
    their report.  Moving towards machine-readable descriptions would
    be a natural extension.

7. What do you recommend if the reviews are both excellents but the code
check failed? Does this mean the paper is blocked until code check
passes or rejected or else?

   This is up to the editor of the journal -- see the "Importance"
   dimension of Figure 3.  At one end, it could indeed be a "strict
   requirement" to get a codecheck certificate for the paper to be
   accepted.  On the other hand, it could be entirely optional.

8. The code check proposal is close to some extents to the Journal of
Open Science Software where each reviewer is assigned a list of things
to check during the review. Do authors consider this pipeline when
establishing their own pipeline?

   We have not considered this pipeline, nor do we have an explicit
   idea.  We now note this reviewer list  at the end of the third paragraph
   of "Related Work".

9. To what extent the codecheck certificate can be updated automatically
via some kind of "manual continuous-integration"? I mean that when
reading a paper online, would it be possible to click a button to test
if the code still runs considering the latest versions of libraries?
(for example, the certificate has been issued for Python 2 but I want
to know if this is usable with Python 3).

    To follow on from point 6, this would make a natural extension, but for
    now we are still considering one point in time, and keeping the
    requirements as close to the authors as we can.

10. When you look at journals advertising open data policies, it is
unfortunately not rare to find articles in these same journals without
the actual data. Do you have some suggestion for educating editors to
actually enforce the code check a journal adopt it?

    We share this concern, and unfortunately have no simple
    suggestions for helping editors.  At this early stage, we think
    the approach should be one of encouraging uptake, rather than
    mandating it.  We also hope that having specific in-house
    experience, e.g. editorial staff to examine for code and data
    availability, can note this.  But at the end of the day, this
    again is dependent on the journal's workflow.

Some suggestions:

11. The badge that is delivered would need some time information since the
check is valid at one point in time (with a given software stack) and
does not guarantee future runs.

    Great idea. we could add the certificate number to the URL, or add
    the certificate number.  We will try to implement this when
    revising our workflows.

12. For specialized journals, you could consider to offer a common generic
environment where a code could be first tested. It this fails, then
you would need only to slightly modify the environment to add missing
dependencies. For example, in neuroscience, a Neuro Debian would
probably suit the needs of a large number of models.

    Yes. We will certainly bear this in mind in future work.

13. As editor-in-chief of ReScience C, I would like to inform authors
that the journal now accepts "reproduction report". The idea it to try
to re-run the code accompanying a published article and to report if
it succeeded or failed. Our own procedure to check for reproduction is
not standardized and we'll certainly benefit from the code check
initiative.

    Thank you for noting this.  We now mention the reproduction report
    in the manuscript where we describe Rescience C.

Overall, it's nice to have a clean description of a pipeline to check
for code even though some questions need to be addressed. Also, I'm
not too confident that journals will adopt it immediately and I'm
afraid such initiative will take time to be generalized. But we have
to start somewhere.

    We share your realistic assessment that (a) journals may be slow
    to adopt but that (b) we should start somewhere.

## Reviewer 2

1. Is there any community feedback on the transparency and
reusability of research published with CODECHECK certificates?

This is perhaps too big of an ask this early in the
initiative as research reuse and citations are independent factors of
the publication and peer review of this specific paper - but I’d still
be interested in any insights the authors have to offer on this topic.

    This is an excellent question.  It is perhaps too early for us to
	assess this given most certificates are under a year old, but we
	are not aware of any reuse yet of the codecheck-deposited
	material.  However, we certainly think it interesting to try and
	monitor this over a longer (3-5 year) timescale if possible.   As
	well as looking for citations of certificates, we could also check
	for forks of our repositories, and download statistics from
	Zenodo.

### 2021-05-27 reviews/response-round1.md
URL for reviewer 2

@@
-<https://doi.org/10.5256/f1000research.54932.r82470>
-
-We thank the reviewers for their comments on our manuscript.  The
-comments are below.  Indented are our responses.
+We thank the reviewers for their comments on our manuscript.  Their
+key comments, requiring responses, are below.  Indented are our responses.
 
@@
 
-In this article, authors propose to implement a procedure to check for
-the code accompanying a submission to a journal. To do so, they
-describe a pipeline made of 6 steps that ultimately lead to the
-delivery of a code check certificate meaning that someone external to
-the author's lab has managed to re-run the code. At this point, no
-checking that the results are correct is necessary. The authors
-already issued several codecheck certificates in different
-disciplines. I find the idea really nice and certainly necessary but
-I've a few questions (even though some of them are already addressed
-in the "limitations" section). Given the structure of the paper, I'll
-just list my questions here:
+<https://doi.org/10.5256/f1000research.54932.r82470>
 
@@
 ## Reviewer 2
+<https://f1000research.com/articles/10-253/v1#referee-response-82472>
 

### 2021-05-27 reviews/response-round1.md
added text to address the reusability angle.
include URL for response doc

@@
+# Prelim for F1000R
+
+We have written a detailed 'response to reviewers' document which is
+available at
+<https://github.com/codecheckers/paper/blob/master/reviews/response-round1.md>
+
+However, in brief, we have added text to the manuscript, to the
+"limitations" and "future work" to address key issues made by the
+reviewers.  We
+
+# Response to reviewers
+
 We thank the reviewers for their comments on our manuscript.  Their
@@
 	for forks of our repositories, and download statistics from
-	Zenodo.
+	Zenodo.  We note this as a good closing point for our article.

### 2021-05-27 reviews/response-round1.md
expand.
test quote syntax

@@
 "limitations" and "future work" to address key issues made by the
-reviewers.  We
+reviewers.  These edits have resulted in a few additional sentences in
+our article; no changes to figures/tables or references were required.
 
@@
 
-    These badges, introduced in August 2020, show whether code is
-    available (different levels) and reproduces same results.  In
-    principle CODECHECKER could award these badges (artifacts
-    evaluated, functional).  We have made a note to this effect in the
-    manuscript (Related work, end of paragraph 1.)
+> These badges, introduced in August 2020, show whether code is
+> available (different levels) and reproduces same results.  In
+> principle CODECHECKER could award these badges (artifacts evaluated,
+> functional).  We have made a note to this effect in the manuscript
+> (Related work, end of paragraph 1.)
 

### 2021-05-27 reviews/response-round1.md
used quotes for reviewer comments
(seems more natural this way; looks like email)

--- 2021-05-27 reviews/response-round1.md
@@
 However, in brief, we have added text to the manuscript, to the
-"limitations" and "future work" to address key issues made by the
-reviewers.  We
+"related work", "limitations" and "future work" to address key issues
+made by the reviewers.  These edits have resulted in a few additional
+sentences in our article; no changes to figures/tables or references
+were required.
 
@@
 
-We thank the reviewers for their comments on our manuscript.  Their
-key comments, requiring responses, are below.  Indented are our responses.
+We thank the reviewers for their comments on our manuscript.  
+key comments, requiring responses, are below, quoted, followed by our
+answers.  (We have added numbers to the key points.)
 
@@
 
-1. How does CODECHECK compare to ACM Artifact reviews badges?
-(https://www.acm.org/publications/policies/artifact-review-and-badging-current)
+> 1. How does CODECHECK compare to ACM Artifact reviews badges?
+> (https://www.acm.org/publications/policies/artifact-review-and-badging-current)
 
@@
 available (different levels) and reproduces same results.  In
-    principle CODECHECKER could award these badges (artifacts
-    evaluated, functional).  We have made a note to this effect in the
-    manuscript (Related work, end of paragraph 1.)
+principle CODECHECKER could award these badges (artifacts evaluated,
+functional).  We have made a note to this effect in the manuscript
+(Related work, end of paragraph 1.)
 
-2. What would be the incentive for someone to code check the code? Being
-aware of the increasing difficulty in finding reviewers, I don't think
-it would be easy to recruit people to perform a task that can rapidly
-become very technical and time consuming.
+> 2. What would be the incentive for someone to code check the code?
+> Being aware of the increasing difficulty in finding reviewers, I
+> don't think it would be easy to recruit people to perform a task
+> that can rapidly become very technical and time consuming.
 
@@
 
-3. How do you handle the case when specific hardware is necessary
-(e.g. NVidia GPU)? Is it documented somewhere such that code-checkers
-might first verify if they have the necessary hardware to run the
-code?
+> 3. How do you handle the case when specific hardware is necessary
+> (e.g. NVidia GPU)? Is it documented somewhere such that
+> code-checkers might first verify if they have the necessary hardware
+> to run the code?
 
-    This was handled in the limitation "What about my proprietary
-    software and sensitive data." but we now mention hardware too in
-    the first paragraph of that section.
+This was handled in the limitation "What about my proprietary software
+and sensitive data." but we now mention hardware too in the first
+paragraph of that section.
 
-4. How do you establish a check has failed? For example, what happens if
-a code-checker gets a segfault (for some unknown reason) and the
-author is unable to help. Is it deemed failed?
+> 4. How do you establish a check has failed? For example, what
+> happens if a code-checker gets a segfault (for some unknown reason)
+> and the author is unable to help. Is it deemed failed?
 
@@
 
-5. Who will pay for the computing resources needed to run heavy
-simulations and/or to acquire necessary software such as e.g. Matlab?
-When a simulation consumes a lot of resources, it might wise to give
-the checker access to computing resources. This could be paid for by
-the journal.
+> 5. Who will pay for the computing resources needed to run heavy
+> simulations and/or to acquire necessary software such as
+> e.g. Matlab?  When a simulation consumes a lot of resources, it
+> might wise to give the checker access to computing resources. This
+> could be paid for by the journal.
 
@@
 
-6. I did not see in the report example a description of the environment
-necessary to run the software. How did you solve the "dependency
-hell"? Since the code might break at some point in the future because
-of incompatibility in some libraries or environments, it would be
-necessary to have a mechanism describing the running environment such
-that it can be re-run later.
+> 6. I did not see in the report example a description of the
+> environment necessary to run the software. How did you solve the
+> "dependency hell"? Since the code might break at some point in the
+> future because of incompatibility in some libraries or environments,
+> it would be necessary to have a mechanism describing the running
+> environment such that it can be re-run later.
 
-    The short answer is "we didn't.".  In the paragraph "Should
-    CODECHECK requirements be more demanding?" we note our low bar of
-    simply getting a codecheck to run once.  We do, however, encourage
-    CODECHECKERS to describe the environment in free text form in
-    their report.  Moving towards machine-readable descriptions would
-    be a natural extension.
+The short answer is "we didn't".  In the paragraph "Should CODECHECK
+requirements be more demanding?" we note our low bar of simply getting
+a codecheck to run once.  We do, however, encourage CODECHECKERS to
+describe the environment in free text form in their report.  Moving
+towards machine-readable descriptions would be a natural extension.
 
-7. What do you recommend if the reviews are both excellents but the code
-check failed? Does this mean the paper is blocked until code check
-passes or rejected or else?
+> 7. What do you recommend if the reviews are both excellents but the code
+> check failed? Does this mean the paper is blocked until code check
+> passes or rejected or else?
 
@@
 
-8. The code check proposal is close to some extents to the Journal of
-Open Science Software where each reviewer is assigned a list of things
-to check during the review. Do authors consider this pipeline when
-establishing their own pipeline?
+> 8. The code check proposal is close to some extents to the Journal
+> of Open Science Software where each reviewer is assigned a list of
+> things to check during the review. Do authors consider this pipeline
+> when establishing their own pipeline?
 
-   We have not considered this pipeline, nor do we have an explicit
-   idea.  We now note this reviewer list  at the end of the third paragraph
-   of "Related Work".
+We have not considered this pipeline, nor do we have an explicit idea.
+We now note this reviewer list at the end of the third paragraph of
+"Related Work".
 
-9. To what extent the codecheck certificate can be updated automatically
-via some kind of "manual continuous-integration"? I mean that when
-reading a paper online, would it be possible to click a button to test
-if the code still runs considering the latest versions of libraries?
-(for example, the certificate has been issued for Python 2 but I want
-to know if this is usable with Python 3).
+> 9. To what extent the codecheck certificate can be updated
+> automatically via some kind of "manual continuous-integration"? I
+> mean that when reading a paper online, would it be possible to click
+> a button to test if the code still runs considering the latest
+> versions of libraries?  (for example, the certificate has been
+> issued for Python 2 but I want to know if this is usable with Python
+> 3).
 
@@
 
-10. When you look at journals advertising open data policies, it is
-unfortunately not rare to find articles in these same journals without
-the actual data. Do you have some suggestion for educating editors to
-actually enforce the code check a journal adopt it?
+> 10. When you look at journals advertising open data policies, it is
+> unfortunately not rare to find articles in these same journals
+> without the actual data. Do you have some suggestion for educating
+> editors to actually enforce the code check a journal adopt it?
 
@@
 
-Some suggestions:
-
-11. The badge that is delivered would need some time information since the
-check is valid at one point in time (with a given software stack) and
-does not guarantee future runs.
+> 11. The badge that is delivered would need some time information
+> since the check is valid at one point in time (with a given software
+> stack) and does not guarantee future runs.
 
@@
 
-12. For specialized journals, you could consider to offer a common generic
-environment where a code could be first tested. It this fails, then
-you would need only to slightly modify the environment to add missing
-dependencies. For example, in neuroscience, a Neuro Debian would
-probably suit the needs of a large number of models.
+> 12. For specialized journals, you could consider to offer a common
+> generic environment where a code could be first tested. It this
+> fails, then you would need only to slightly modify the environment
+> to add missing dependencies. For example, in neuroscience, a Neuro
+> Debian would probably suit the needs of a large number of models.
 
@@
 
-13. As editor-in-chief of ReScience C, I would like to inform authors
-that the journal now accepts "reproduction report". The idea it to try
-to re-run the code accompanying a published article and to report if
-it succeeded or failed. Our own procedure to check for reproduction is
-not standardized and we'll certainly benefit from the code check
-initiative.
+> 13. As editor-in-chief of ReScience C, I would like to inform
+> authors that the journal now accepts "reproduction report". The idea
+> it to try to re-run the code accompanying a published article and to
+> report if it succeeded or failed. Our own procedure to check for
+> reproduction is not standardized and we'll certainly benefit from
+> the code check initiative.
 
-    Thank you for noting this.  We now mention the reproduction report
-    in the manuscript where we describe Rescience C.
+Thank you for noting this.  We now mention the reproduction report in
+the manuscript where we describe Rescience C.
 
-Overall, it's nice to have a clean description of a pipeline to check
-for code even though some questions need to be addressed. Also, I'm
-not too confident that journals will adopt it immediately and I'm
-afraid such initiative will take time to be generalized. But we have
-to start somewhere.
+> Overall, it's nice to have a clean description of a pipeline to
+> check for code even though some questions need to be
+> addressed. Also, I'm not too confident that journals will adopt it
+> immediately and I'm afraid such initiative will take time to be
+> generalized. But we have to start somewhere.
 
@@
 
-1. Is there any community feedback on the transparency and
-reusability of research published with CODECHECK certificates?
+> 1. Is there any community feedback on the transparency and
+> reusability of research published with CODECHECK certificates?
 
-This is perhaps too big of an ask this early in the
-initiative as research reuse and citations are independent factors of
-the publication and peer review of this specific paper - but I’d still
-be interested in any insights the authors have to offer on this topic.
+> This is perhaps too big of an ask this early in the initiative as
+> research reuse and citations are independent factors of the
+> publication and peer review of this specific paper - but I’d still
+> be interested in any insights the authors have to offer on this
+> topic.
 

### 2021-06-02 reviews/response-round1.md
small edits to response letter

@@
 "related work", "limitations" and "future work" to address key issues
-made by the reviewers.  These edits have resulted in a few additional
+made by the reviewers and by one public comment.
+These edits have resulted in a few additional
 sentences in our article; no changes to figures/tables or references
 were required.
+
+Furthermore, we noticed an inconsistent use of the terms "workflow" 
+and "process", and decided to use "CODECHECK workflow", "computational workflow", and 
+"publication/review process" more consistently in the revision of the manuscript.
 
@@
 the certificate number.  We will try to implement this when
-revising our workflows.
+revising our workflows.  Nevertheless, the point in time and software
+stack should be documented via the certificate already now.
 
@@
 
-Yes. We will certainly bear this in mind in future work.
+Yes. We will certainly bear this in mind in future work, especially
+for author guidelines.
 

### 2021-06-15 codecheck-paper.tex
revision after first round of reviews
Transferred from https://www.overleaf.com/project/609a7da53dfe13bf1a4b42b9

@@
 
-\usepackage{fontspec}
+\newcommand{\rev}[1]{#1}
+%\newcommand{\rev}[1]{\textit{#1}}
+%\newcommand{\rev}[1]{\textcolor{blue}{#1}}
+
 \usepackage{fontawesome}
@@
 \begin{document}
-\title{CODECHECK: An Open Science initiative for the independent
+\title{CODECHECK: an Open Science initiative for the independent
   execution of computations underlying research articles during peer
   review to  improve reproducibility}
-\author[1,$\ast$]{Daniel~N\"{u}st \orcid{0000-0002-0024-5046} \github{nuest} \twitter{nordholmen}}
-\author[2,$\ast$]{Stephen~J~Eglen \orcid{0000-0001-8607-8025} \github{sje30} \twitter{StephenEglen}}
+\author[1,$\ast$]{Daniel~N\"{u}st \orcid{0000-0002-0024-5046}}
+\author[2,$\ast$]{Stephen~J~Eglen \orcid{0000-0001-8607-8025}}
 \affil[1]{Institute for Geoinformatics, University of M\"unster, Germany; \email{daniel.nuest@uni-muenster.de}}
@@
 \affil[$\ast$]{Corresponding authors}
-\affil[$$]{Public project for this article is at \url{https://github.com/codecheckers/paper}.}
 \maketitle
+
 \begin{abstract}
-%\textbf{Background:}
-%\textbf{Methods:}
-%\textbf{Results:}
-%\textbf{Conclusions:}
  The traditional scientific paper falls short of effectively communicating computational research.
@@
 While CODECHECK has clear limitations, it may represent a building block in Open Science and publishing ecosystems for improving the reproducibility, appreciation, and, potentially, the quality of non-textual research artefacts.
+The CODECHECK website can be accessed here: \url{https://codecheck.org.uk/}.
 \end{abstract}
-
-\subsection*{Keywords}
-
-reproducible research, Open Science, peer review, reproducibility, code sharing, data sharing, quality control, scholarly publishing
 
@@
 \cite{fyfe_mission_2019}, the same baseline could be provided for the
-workflow underlying a paper.  With this in mind, we have developed a
-set of principles and an example workflow that provides a pragmatic
+\rev{computational} workflow underlying a paper.  With this in mind, we have developed a
+set of principles and an example \rev{CODECHECK} workflow that provides a pragmatic
 way of checking that a paper's code works, i.e., it is reproducible
@@
 
-\subsection*{Workflow and people}\label{workflow-people}
+\subsection*{\rev{CODECHECK w}orkflow and people}\label{workflow-people}
 
 CODECHECK is best demonstrated by way of our example workflow, and later
-we expand on the underlying principles. The workflow involves three
+we expand on the underlying principles. The \rev{CODECHECK} workflow involves three
 groups of people:
@@
 (3) the \textbf{codechecker}, who checks that the author's code works.
-The six-step workflow we have refined is shown in
+The six-step \rev{CODECHECK} workflow we have refined is shown in
 Figure~\ref{fig:workflow}.  In this article, we also refer to a
@@
       \includegraphics[width=\textwidth]{figs/codecheck_overview.pdf}
-      \caption{The CODECHECK example process implementation.
+      \caption{The CODECHECK example \rev{workflow} implementation.
         Codecheckers act as detectives: They investigate and record,
@@
 for it, e.g.,~by asking a colleague to attempt a reproduction, and
-providing a set of instructions on how to re-run the workflow.
+providing a set of instructions on how to re-run the \rev{computational} workflow.
 
@@
 process iterates until either the codechecker is successful, or the
-codechecker concludes the workflow is not reproducible.  As part of
+codechecker concludes the \rev{paper's} workflow is not reproducible.  As part of
 this process, the codechecker could work entirely locally, relying on
@@
 that the task of the codechecker is to check only the ``mechanics'' of the
-workflow.  In the context of
+\rev{computational} workflow.  In the context of
 mathematics, Stodden et~al. \cite{stodden_setting_2013} distinguish
@@
 certificate can describe exactly what was checked, because each
-workflow is unique.  Since no specific tool or platform is required,
+\rev{computational} workflow is unique.  Since no specific tool or platform is required,
 such that no authors are excluded, it is futile for the
@@
 in scholarly profiles, such as ORCID (see peer review contributions in
-ORCID records,
-\url{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}).
+\href{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}{ORCID records}).
 The publisher also ensures proper publication metadata, e.g.,~links
@@
       \includegraphics[width=0.8\textwidth]{figs/codecheck_dimensions.pdf}
-  \caption{The dimensions of implementing a CODECHECK process.}
+  \caption{The dimensions of implementing a CODECHECK \rev{workflow}.}
   \label{fig:dimensions}
@@
 checks, i.e.,~a CODECHECK of a preprint, may help to improve the
-workflow itself, even before a publisher is involved.  As such,
+\rev{computational} workflow itself, even before a publisher is involved.  As such,
 codechecking papers could be part of a preprint server's policy or
@@
 publication of the paper.  The GIScience checks (see
-below) falls into this category: By displaying a badge on the volume
+below) falls into this category: by displaying a badge on the volume
 and article landing pages, the AGILE conference highlights articles
@@
 % an evolutionary approach has value, and we do not yet suggest to completely eliminating research papers as a way of sharing work.
-Enhancing existing processes with CODECHECKs allows communities to
+Enhancing existing \rev{review and publication} processes with CODECHECKs allows communities to
 gradually transition towards more open practices.
@@
 make.
-Currently, there is considerable amount of communicating about the
-process, especially regarding who publishes which document when, so that proper
+Currently, there is considerable amount of communicating about the \rev{CODECHECK workflow}, especially regarding who publishes which document when, so that proper
 cross-referencing between paper and certificate is ensured via persistent
@@
 The contributions by the codechecker may even be integrated into
-the code of the workflow and be acknowledged as code commits. This way, 
+the code of the \rev{paper's} workflow and be acknowledged as code commits. This way, 
 proper credit can be given within the research software development community.
@@
 codechecker has a matching code \emph{and} domain expertise to the
-paper, although a well-documented workflow should be executable by any
+paper, although a well-documented \rev{analysis} should be executable by any
 computationally-competent person. Naturally, the more
@@
 \textbf{reproducibility editors}, or \textbf{editorial staff} with a
-publisher.  One codechecker is sufficient to verify the workflow since
+publisher.  One codechecker is sufficient to verify the \rev{paper's} workflow since
 it is mostly a factual process.  Code usually harbours systematic and
@@
 reproducibility, yet this commitment may be hard for small publishers.
-These codecheckers could be fully integrated into internal workflows.
+These codecheckers could be fully integrated into the internal \rev{publication process}.
 Credit for doing the codecheck is also achieved, as it is part of
@@
 may be listed in public databases (e.g.,~ORCID, see Step~6 above,
-or commercial offerings such as Publons, \url{https://publons.com/}, and
-ReviewerCredits, \url{https://www.reviewercredits.com/});
+or commercial offerings such as \href{https://publons.com/about/home/}{Publons}, and
+\href{https://www.reviewercredits.com/}{ReviewerCredits});
 a codechecker could be similarly listed.
@@
 year, see \url{https://github.com/codecheckers/codecheckers/}.  Their
-motivations, mentioned in the registration
-information at \url{https://github.com/codecheckers/codecheckers/labels/registration},
+motivations, mentioned in the \href{https://github.com/codecheckers/codecheckers/labels/registration}{registration
+information},
 include: supporting reproducible research and Open Science, improve
@@
 
-The workflow and variations outlined describe our current
+The \rev{CODECHECK} workflow and variations outlined describe our current
 views on how code could be checked.  They are not immutable, but
 we believe the following core principles underpin our CODECHECK
-process:
+\rev{workflow}:
 
@@
 fix.  The level of documentation required for third parties to
-reproduce a workflow is hard to get right, and too often this
+reproduce a \rev{computational} workflow is hard to get right, and too often this
 uncertainty leads researchers to give up and not document it at all.
@@
 
-\textbf{4. Workflows must be auditable.} \\
+\textbf{4. \rev{Computational} workflows must be auditable.} \\
 The codechecker should have sufficient material to validate the
-workflow outputs submitted by the
+\rev{computational} workflow outputs submitted by the
 authors. Stark~\cite{stark_before_2018} calls this
@@
 in favour of bitwise reproducibility because, in the long run, it can
-be automated (e.g.,
+\rev{help to automate checking by comparing outputs algorithmically} (e.g.,
 \url{https://twitter.com/khinsen/status/1242842759733665799}), but
-until then we need CODECHECK's approach.
+until \rev{such an ideal is achievable} we need CODECHECK's approach.
 
@@
   AGILE} Initiative \cite{reproducible_agile} independently
-established a process for reproducing workflows at the AGILE
+established a process for reproducing \rev{computational} workflows at the AGILE
 conference series \cite{nust_improving_2020}.  While using slightly
@@
 
-After running the workflow, the codechecker writes a certificate
+After running the \rev{paper's} workflow, the codechecker writes a certificate
 stating which outputs from the original article, i.e.,~numbers,
@@
 published at \url{https://codecheck.org.uk/spec/config/latest/}.  The
-configuration file enables current and future automation of workflows
+configuration file enables current and future automation of \rev{CODECHECK} workflows
 and meta-analyses.
@@
 file should be compared in the original paper (Comment), and the file
-size.  Page 3 of the certificate, (Figure~\ref{fig:cert}E) gives
+size.  Page 3 of the certificate, Figure~\ref{fig:cert}E gives
 detailed notes from the codechecker, here documenting what steps were
@@
 complete. Finally, page 4 of the certificate shows the first output
-generated by the CODECHECK (Figure~\ref{fig:cert}F). In this case, the
+generated by the CODECHECK Figure~\ref{fig:cert}F. In this case, the
 figure matched figure~4 of \cite{unwin_report_2020}.  The remaining
@@
 
-These tools and resources demonstrate that a CODECHECK process can be
+These tools and resources demonstrate that a CODECHECK \rev{workflow} can be
 managed on freely available platforms.
@@
 with Publons (see
-\url{https://authors.acm.org/author-services/publons}).
-
-This activity in the ACM journals can be seen as one possible workflow
+\url{https://authors.acm.org/author-services/publons}).  \rev{As well
+  as this, ACM provides several badges to indicate what kind of artifact review or reproduction a paper submitted to an ACM journal completed (\url{https://www.acm.org/publications/policies/artifact-review-and-badging-current}), but does not provide nor require a specific review process.
+In principle, these badges could be awarded by a codechecker, too, though the different levels and even partial replacement of artifacts required  to achieve a \emph{Results Reproduced} go beyond a CODECHECK's scope.
+A completed check certainly warrants the ACM badge \emph{Artifacts Evaluated - Functional} and possibly \emph{Artifacts Evaluated - Reusable} and likely \emph{Artifacts Available}, depending on additional requirements by implementing journals.
+However, we do not require codecheckers to evaluate code quality or ensuring proper archival of artifacts though, in our experience, they are likely to encounter or comment on these topics.}
+This activity in the ACM journals can be seen as one possible \rev{process}
 within a CODECHECK system, and clearly shares much in spirit.
@@
 replication of results and are very well designed/reasoned
-\cite{chambers_verification_2020}.  In a similar vein, the CODECHECK
+\cite{chambers_verification_2020}.
+\rev{The \emph{Journal of Water Resources Planning and Management}'s 
+policy recognises reproducible papers in a special collection and 
+incentivises authors with waived or reduced fees
+\cite{rosenberg_reproducible_2021}.}
+In a similar vein, the CODECHECK
 certificates could also be published as a special article type within
-journals.
-
-Going beyond individual articles, the journal \emph{ReScience C}
-(\url{https://rescience.github.io/}) publishes only replications, also
-requiring open code and replication by a third party. ReScience also
-relies on free infrastructure (GitHub and Zenodo).
+journals.  \rev{Finally, the \textit{Journal of Open Source Software}
+provides its reviewers with a checklist of items to check during
+review,
+\url{https://joss.readthedocs.io/en/latest/review_checklist.html\#software-paper},
+effectively providing a much more detailed form of check for scientific
+software that could complement CODECHECKs, too.}
+
+Going beyond individual articles, the journal
+\href{https://rescience.github.io}{\emph{ReScience~C}} publishes only
+replications, also requiring open code and replication by a third
+party.  \rev{The journal now accepts ``Reproduction reports'' that
+  describe if some code accompanying a published article can (or can
+  not) reproduce the same results as shown in the article.}  ReScience~C
+also relies on free infrastructure (GitHub and Zenodo).
 
@@
 \emph{internal replication}.
-A workflow that has undergone internal
+A \rev{computational} workflow that has undergone internal
 replication would likely be of high quality and relatively easy to check.
@@
 Such internal checks are professionalised in local reproduction services,
-such as CISER R-squared
-(\url{https://ciser.cornell.edu/research/results-reproduction-r-squared-service/})
-or YARD (\url{https://isps.yale.edu/research/data/approach}),
-or in communities such as Oxford's code review network (\url{https://github.com/OxfordCodeReviewNet/forum}).
+such as \href{https://ciser.cornell.edu/research/results-reproduction-r-squared-service/}{CISER R-squared} or \href{https://isps.yale.edu/research/data/approach}{YARD},
+or in communities such as \href{https://github.com/OxfordCodeReviewNet/forum}{Oxford's code review network}.
 
@@
 Another platform for publishing results of reproductions is
-\emph{SciGen.Report} (\url{https://scigen.report/}).  It is a
+\href{https://scigen.report}{\emph{SciGen.Report}}.  It is a
 community-run independent platform to foster communication on
@@
 with reproducibility awareness, education, and support, such as
-ReproHack (\url{https://reprohack.github.io/reprohack-hq/}),
-Code Copilot (\url{https://twitter.com/Code_Copilot}), or
-Papers with Code (\url{https://paperswithcode.com/about}).
+\href{https://reprohack.github.io/reprohack-hq/}{ReproHack},
+\href{https://twitter.com/Code_Copilot}{Code Copilot}, or
+\href{https://paperswithcode.com/about}{Papers with Code}.
 The latter recently announced a collaboration with the preprint server
@@
 This might fall within idealistic expectations of peer review, but is
-rare.  Establishing a CODECHECK process acknowledges that peer
+rare.  Establishing a CODECHECK \rev{workflow} acknowledges that peer
 reviewing practices have been unable to adapt to the challenges of
@@
 CODECHECK by design does not require authors to provide (and sustain)
-an eternally functional workflow nor suggests a specific software
+an eternally functional \rev{computational} workflow nor suggests a specific software
 stack or practical approach.
 Creating something
-that anyone can reproduce has been called a fool's
-errand (\url{https://twitter.com/DougBlank/status/1135904909663068165})
+that anyone can reproduce \href{https://twitter.com/DougBlank/status/1135904909663068165}{has been called a fool's
+errand}
 and we tend to agree.  However, the package of data, code, and
@@
 snapshot of a working analysis that greatly increases the likelihood
-of a successful reproduction and the possibility that a workflow can
+of a successful reproduction and the possibility that a \rev{computational} workflow can
 be extended by third parties in the future, if they have
-access to suitable resources and matching skill set.  Concrete
+access to suitable resources and matching skill set.
+\rev{The CODECHECK principles help to make very clear what a CODECHECK
+badge on a paper means and also ensure a minimum standard that 
+other processes or badges may not have, e.g., only superficially 
+checked self-awarded badges (
+\url{https://www.cambridge.org/core/journals/environmental-data-science/information/instructions-for-authors}).}
+
+Concrete
 implementations of CODECHECK workflows, especially for specific
@@
 equally deep into the review, but only if they have the expertise and
-time. We can imagine a tiered workflow where a codechecker could, just
+time. We can imagine a tiered \rev{CODECHECK} workflow where a codechecker could, just
 as a conventional reviewer could, recommend a detailed code review
@@
 conduct software reviews independent of a specific publication or
-workflow include ROpenSci (\url{https://ropensci.org/}), PyOpenSci
-(\url{https://www.pyopensci.org/}), and JOSS
-(\url{https://joss.theoj.org/}).  Furthermore, the codechecker's task
+\rev{venue} include \href{https://ropensci.org/}{ROpenSci}, \href{https://www.pyopensci.org/}{PyOpenSci}, and \href{https://joss.theoj.org}{JOSS}.  Furthermore, the codechecker's task
 list is intentionally not overloaded with related issues such as
@@
 
-\textbf{How are failures during checks handled?}
-
-We do not yet have a process for denoting if a reproduction fails, as
-our case-studies were all successful.  In the case that a journal
+\textbf{How are failures during checks handled?} We do not yet have a process for denoting if a reproduction fails, as our case-studies were all successful.  In the case that a journal
 adopts CODECHECK for all submissions, the question remains as what to
 do if a check fails, after exhausting efforts between author and
-codechecker to reproduce the workflow.  A negative comment in a
+codechecker to reproduce the \rev{computational} workflow.  A negative comment in a
 CODECHECK certificate or a failed check does not necessarily mean the
@@
 example, or small dataset that can be quickly analysed to demonstrate
-that the workflow runs correctly.
+that the \rev{paper's} workflow runs correctly.
 
 \textbf{What about my proprietary software and sensitive data?}  Given
-the prevalence of proprietary software, e.g MATLAB, we pragmatically
+the prevalence of proprietary software, e.g MATLAB, in some disciplines
+we pragmatically
 decided that we should accept code as long as we could find a machine
@@
 software also considerably hampers reuse, especially by researchers
-from the global south. Therefore, allowing proprietary software is a
-compromise that should be reconsidered.
-
-Solutions for proprietary and sensitive data exists.  Authors can
+from the global south.  \rev{Likewise, if a research
+requires specific hardware, e.g. GPU, we are reliant on the
+codechecker having access to similar hardware. Both licenses and costs
+can be barriers to a CODECHECK, but the focus on the codechecker's
+assessment provides options to overcome these barriers if needed.}  \rev{Therefore, allowing proprietary software and
+specialised hardware are compromises that should be reconsidered.
+In any case, authors must make such requirements clear and the 
+opportunity to answer them must be documented for codecheckers.}
+
+Solutions for proprietary and sensitive data exist.  Authors can
 provide synthetic data (cf. \cite{shannon_opening_2018}), some data
@@
 or independent entities can provide infrastructure for sharing data
-and workflows confidentially \cite{perignon_certify_2019} or with
+and \rev{computational} workflows confidentially \cite{perignon_certify_2019} or with
 access to derived results but not raw data
@@
 \cite{foster_research_2018}, or domains of reproducibility
-\cite{harris_more_2017}).
+\cite{harris_more_2017}.
 
@@
 as technical and topical expertise can be provided by different
-people.  Third, recall that CODECHECKs should always be publicly
-available, unlike peer review reports.  With code and workflows, the
+people \rev{(cf. segmentation of multidisciplinary
+works \cite{dinakaran_proposal_2021})}.
+Third, recall that CODECHECKs should always be publicly
+available, unlike peer review reports.  With code and \rev{computational} workflows, the
 codechecker's feedback may directly impact and improve the author's
@@
   recognition} for the codechecker.  Fourth, we found that focusing on
-workflow mechanics and interacting with the author makes reproductions
+\rev{the computational} workflow\rev{'s} mechanics and interacting with the author makes reproductions
 \textbf{educational}. It also is a different role and, as such, could
@@
 
-\textbf{Do workflows need to be codechecked multiple times?}  If a
+\textbf{Do \rev{computational} workflows need to be codechecked multiple times?}  If a
 paper is checked at the start of peer review, it might need
@@
 the CODECHECK community
-process, documented at
+\rev{workflow}, documented at
 \url{https://codecheck.org.uk/guide/community-process},
@@
 
-CODECHECK works --- we have created a considerable number of
-certificates to demonstrate it.  The creation of certificates and
+CODECHECK works --- we have \rev{reproduced} a considerable number of
+\rev{computational workflows across multiple disciplines, software stacks, and review
+processes, and we have documented all results transparently in CODECHECK}
+certificates.  The creation of certificates and
 interactions with authors and editors shaped the principles and the
-workflow and also confirmed the approach taken. This result
+\rev{CODECHECK} workflow and also confirmed the approach taken. This result
 corroborates findings from similar evaluations of reproducible
@@
 research outputs, articles with live-code
-\cite{perkel_pioneering_2019-1} such as eLife's ERA
-(see \url{https://elifesciences.org/labs/dc5acbde/welcome-to-a-new-era-of-reproducible-publishing}),
+\cite{perkel_pioneering_2019-1} such as \href{https://elifesciences.org/labs/dc5acbde/welcome-to-a-new-era-of-reproducible-publishing}{eLife's ERA},
 decentralized infrastructure and public reviewer reputation systems
 \cite{tenorio-fornes_towards_2019}, and completely new visions for
-scholarly communication and peer review such as described by Amy J. Ko
-in \emph{A modern vision for peer review}
-(\url{https://medium.com/bits-and-behavior/a-modern-vision-for-peer-review-d5f73f0fae07}).
+scholarly communication and peer review, such as described by Amy J. Ko
+in \href{https://medium.com/bits-and-behavior/a-modern-vision-for-peer-review-d5f73f0fae07}{\emph{A modern vision for peer review}}.
+\rev{A CODECHECK's impact on the published research outputs and the required
+infrastructure would also support answering needs for better integration of research 
+outputs and more openness \cite{eosc2020}.}
 An explicit segmentation of research steps could even make the focus
@@
 The question of training codecheckers is also relevant. We expect a
-mentoring scheme within the CODECHECK community (experienced
+mentoring scheme within the CODECHECK community\rev{, in which} experienced
 codecheckers will provide on-the-job training or serve as fallback
-advisors). Codecheckers may also be found by collaborating with
+advisors, \rev{would be most suitable. Given the difficulty to document 
+solutions for the unique problems every check has, practical experience
+in the craft of codechecking is paramount}.
+Codecheckers may also be found by collaborating with
 reproducible research initiatives such as
-ReproHack (\url{https://reprohack.github.io/reprohack-hq/}),
-ReproducibiliTea (\url{https://reproducibilitea.org/},
-\cite{fitzgibbon_brewing_2020}), and
-Repro4Everyone (\url{https://repro4everyone.org/},
-\cite{auer_reproducibility_2020}).  The initial reaction of 
-researchers to these ideas shows that scholarly peer review
-should continue on the path towards facilitating sharing and
-execution of computational workflows.
-
-\subsection*{Competing interests}
-
-%% SJE: check again in April! s/is/was
-SJE is on the editorial board at the journal \emph{Scientific Data}
-until March 2021.  DN is reproducibility chair at the Association of
-Geographic Information Laboratories in Europe's (AGILE) annual
-conference.
-
-\subsection*{Grant information}
-
-This work was financially supported by the UK Software
-Sustainability Institute and a Mozilla Science mini grant.
-DN is supported by grant
-\href{https://gepris.dfg.de/gepris/projekt/415851837}{PE~1632/17-1}
-from the German Research Foundation (DFG).
+\href{https://reprohack.github.io/reprohack-hq/}{ReproHack},
+\href{https://reproducibilitea.org/}{ReproducibiliTea},
+\cite{fitzgibbon_brewing_2020}, and
+\href{https://repro4everyone.org/}{Repro4Everyone},
+\cite{auer_reproducibility_2020}.  The initial reaction of researchers
+to these ideas shows that scholarly peer review should continue on the
+path towards facilitating sharing and execution of computational
+workflows.  \rev{It is perhaps too soon to see if CODECHECK increases
+  reuse of code and data, and we would certainly value a longer-term
+  critical assessment of the impact of material that has been checked.}
+
+\section*{Data availability}
+Zenodo: codecheckers/register: CODECHECK Register Deposit January 2021 \url{http://doi.org/10.5281/zenodo.4486559} [117].
+
+This project contains the following underlying data:
+
+\begin{itemize}
+  \item \texttt{register.csv}. List of all CODECHECK certificates with references to repositories and reports.
+\end{itemize}
+
+Data are available under the terms of the \href{https://creativecommons.org/licenses/by-sa/4.0/legalcode}{Creative Commons Attribution Share Alike license} (CC-BY-SA 4.0 International).
+
+\section*{Software availability}
+Codecheckers GitHub organisation: \url{https://github.com/codecheckers}\\
+CODECHECK community on Zenodo: \url{https://zenodo.org/communities/codecheck}\\
+
+\texttt{codecheck} R package: \url{https://github.com/codecheckers/codecheck}\\
+Archived R package as at time of publication: \url{http://doi.org/10.5281/zenodo.4522507} [79]\\
+License: MIT
 
 \subsection*{Acknowledgements}\label{acknowledgements}
-%Collaborators and potential collaborators:
-% someone from SpringerNature
-
 We are grateful to the following individuals for discussions regarding
@@
 reviewers at the AGILE conference, and all authors of auditable
-workflows for their participation.  We thank Celeste R. Brennecka from
+\rev{computational} workflows for their participation.  We thank Celeste R. Brennecka from
 the Scientific Editing Service, University of M\"unster, for her

### 2021-06-16 reviews/response-round1.md
Document \rev{}
reformat slightly for uploading

@@
 
-However, in brief, we have added text to the manuscript, to the
-"related work", "limitations" and "future work" to address key issues
-made by the reviewers and by one public comment.
-These edits have resulted in a few additional
-sentences in our article; no changes to figures/tables or references
-were required.
+However, in brief, we have added text to the manuscript, to the "related work", "limitations" and "future work" to address key issues made by the reviewers and by one public comment.
+These edits have resulted in a few additional sentences in our article; no changes to figures/tables or references were required.
 
-Furthermore, we noticed an inconsistent use of the terms "workflow" 
-and "process", and decided to use "CODECHECK workflow", "computational workflow", and 
-"publication/review process" more consistently in the revision of the manuscript.
+Furthermore, we noticed an inconsistent use of the terms "workflow" and "process", and decided to use "CODECHECK workflow", "computational workflow", and "publication/review process" more consistently in the revision of the manuscript.
+
+We have used a macro, \rev{...} to record our edits in this revision.  If you wish to see the edits in italics, please change the definition in the preamable of the latex file.
 
@@
 
-We thank the reviewers for their comments on our manuscript.  
-key comments, requiring responses, are below, quoted, followed by our
-answers.  (We have added numbers to the key points.)
+We thank the reviewers for their comments on our manuscript.  Key comments, requiring responses, are below, quoted, followed by our answers.  (We have added numbers to the key points.)
 
@@
 
-These badges, introduced in August 2020, show whether code is
-available (different levels) and reproduces same results.  In
-principle CODECHECKER could award these badges (artifacts evaluated,
-functional).  We have made a note to this effect in the manuscript
-(Related work, end of paragraph 1.)
+These badges, introduced in August 2020, show whether code is available (different levels) and reproduces same results.
+In principle CODECHECKER could award these badges (artifacts evaluated, functional).
+We have made a note to this effect in the manuscript (Related work, end of paragraph 1.)
 
@@
 
-This was addressed in our section "Who's got time for more peer
-review?" We would however note that we have a pool of about 20
-volunteers currently willing to do codechecks.
+This was addressed in our section "Who's got time for more peer review?" We would however note that we have a pool of about 20 volunteers currently willing to do codechecks.
 
@@
 
-This was handled in the limitation "What about my proprietary software
-and sensitive data." but we now mention hardware too in the first
-paragraph of that section.
+This was handled in the limitation "What about my proprietary software and sensitive data." but we now mention hardware too in the first paragraph of that section.
 
@@
 
-We hope that codechecker and author can resolve problems, but in
-the end there may be problems that cannot be solved.  Open
-infrastructure could help as both author and codechecker can work
-together in the same environment to minimise these failures.
-Ultimately however, there may be failures, which are noted in the
-section "How are failures during checks handled?".
+We hope that codechecker and author can resolve problems, but in the end there may be problems that cannot be solved.
+Open infrastructure could help as both author and codechecker can work together in the same environment to minimise these failures.
+Ultimately however, there may be failures, which are noted in the section "How are failures during checks handled?".
 
@@
 
-In the section "Who will pay for compute time?" we mention this
-problem, and that toy examples might alleviate the need to re-run
-resource-intensive computations.  We agree that one model might be
-that a journal provide some resource for this service.  Likewise,
-in the following paragraph, we describe that our pragmatic
-approach for now is to find codecheckers that have access to
-particular software, e.g. MATLAB.
+In the section "Who will pay for compute time?" we mention this problem, and that toy examples might alleviate the need to re-run resource-intensive computations.
+We agree that one model might be that a journal provide some resource for this service.
+Likewise, in the following paragraph, we describe that our pragmatic approach for now is to find codecheckers that have access to particular software, e.g.
+MATLAB.
 
@@
 
-The short answer is "we didn't".  In the paragraph "Should CODECHECK
-requirements be more demanding?" we note our low bar of simply getting
-a codecheck to run once.  We do, however, encourage CODECHECKERS to
-describe the environment in free text form in their report.  Moving
-towards machine-readable descriptions would be a natural extension.
+The short answer is "we didn't".
+In the paragraph "Should CODECHECK requirements be more demanding?" we note our low bar of simply getting a codecheck to run once.
+We do, however, encourage CODECHECKERS to describe the environment in free text form in their report.
+Moving towards machine-readable descriptions would be a natural extension.
 
@@
 
-This is up to the editor of the journal -- see the "Importance"
-dimension of Figure 3.  At one end, it could indeed be a "strict
-requirement" to get a codecheck certificate for the paper to be
-accepted.  On the other hand, it could be entirely optional.
+This is up to the editor of the journal -- see the "Importance" dimension of Figure 3.
+At one end, it could indeed be a "strict requirement" to get a codecheck certificate for the paper to be accepted.
+On the other hand, it could be entirely optional.
 
@@
 
-We have not considered this pipeline, nor do we have an explicit idea.
-We now note this reviewer list at the end of the third paragraph of
-"Related Work".
+We have not considered this pipeline, nor do we have an explicit idea.  We now note this reviewer list at the end of the third paragraph of "Related Work".
 
@@
 
-To follow on from point 6, this would make a natural extension, but for
-now we are still considering one point in time, and keeping the
-requirements as close to the authors as we can.
+To follow on from point 6, this would make a natural extension, but for now we are still considering one point in time, and keeping the requirements as close to the authors as we can.
 
@@
 
-We share this concern, and unfortunately have no simple
-suggestions for helping editors.  At this early stage, we think
-the approach should be one of encouraging uptake, rather than
-mandating it.  We also hope that having specific in-house
-experience, e.g. editorial staff to examine for code and data
-availability, can note this.  But at the end of the day, this
-again is dependent on the journal's workflow.
+We share this concern, and unfortunately have no simple suggestions for helping editors.
+At this early stage, we think the approach should be one of encouraging uptake, rather than mandating it.
+We also hope that having specific in-house experience, e.g.
+editorial staff to examine for code and data availability, can note this.
+But at the end of the day, this again is dependent on the journal's workflow.
 
@@
 
-Great idea. we could add the certificate number to the URL, or add
-the certificate number.  We will try to implement this when
-revising our workflows.  Nevertheless, the point in time and software
-stack should be documented via the certificate already now.
+Great idea.
+we could add the certificate number to the URL, or add the certificate number.
+We will try to implement this when revising our workflows.
+Nevertheless, the point in time and software stack should be documented via the certificate already now.
 
@@
 
-Yes. We will certainly bear this in mind in future work, especially
-for author guidelines.
+Yes. We will certainly bear this in mind in future work, especially for author guidelines.
 
@@
 
-Thank you for noting this.  We now mention the reproduction report in
-the manuscript where we describe Rescience C.
+Thank you for noting this.  We now mention the reproduction report in the manuscript where we describe Rescience C.
 
@@
 
-We share your realistic assessment that (a) journals may be slow
-to adopt but that (b) we should start somewhere.
+We share your realistic assessment that (a) journals may be slow to adopt but that (b) we should start somewhere.
 
@@
 
-This is an excellent question.  It is perhaps too early for us to
-assess this given most certificates are under a year old, but we
-are not aware of any reuse yet of the codecheck-deposited
-material.  However, we certainly think it interesting to try and
-monitor this over a longer (3-5 year) timescale if possible.   As
-well as looking for citations of certificates, we could also check
-for forks of our repositories, and download statistics from
-Zenodo.  We note this as a good closing point for our article.
+This is an excellent question.
+It is perhaps too early for us to assess this given most certificates are under a year old, but we are not aware of any reuse yet of the codecheck-deposited material.
+However, we certainly think it interesting to try and monitor this over a longer (3-5 year) timescale if possible.
+As well as looking for citations of certificates, we could also check for forks of our repositories, and download statistics from Zenodo.
+We note this as a good closing point for our article.
